### 2014-03-03 doc/LinksToDAG.tex
the paper so far and the bibtex file. Not very much on it, but I hope I can fill it out and write quickly in the next few days...

%
% File acl2014.tex
%
% Contact: koller@ling.uni-potsdam.de, yusuke@nii.ac.jp
%%
%% Based on the style files for ACL-2013, which were, in turn,
%% Based on the style files for ACL-2012, which were, in turn,
%% based on the style files for ACL-2011, which were, in turn, 
%% based on the style files for ACL-2010, which were, in turn, 
%% based on the style files for ACL-IJCNLP-2009, which were, in turn,
%% based on the style files for EACL-2009 and IJCNLP-2008...
%% Based on the style files for EACL 2006 by 
%%e.agirre@ehu.es or Sergi.Balari@uab.es
%% and that of ACL 08 by Joakim Nivre and Noah Smith

\documentclass[11pt]{article}
\usepackage{acl2014}
\usepackage{times}
\usepackage{url}
\usepackage{latexsym}

\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage{soul}
\newcommand{\Note}[1]{}
\renewcommand{\Note}[1]{\hl{[#1]}}  % comment out this definition to suppress all Notes                                                                                
\newcommand{\TODO}[1]{\Note{TODO: #1}}
\newcommand{\NoteSigned}[3]{{\sethlcolor{#2}\Note{#1: #3}}}
\newcommand{\NoteJE}[1]{\NoteSigned{JE}{LightBlue}{#1}}
\newcommand{\NoteJH}[1]{\NoteSigned{JH}{YellowGreen}{#1}}

\title{Link Parses to Directed Acyclic Graphs}

\author{Juneki Hong \\
  Department of Computer Science \\
  Johns Hopkins University \\
  Baltimore, MD 21218, USA \\ 
  {\tt juneki@cs.jhu.edu} \\\And
  Jason Eisner \\
  Department of Computer Science \\
  Johns Hopkins University \\
  Baltimore, MD 21218, USA \\
  {\tt jason@cs.jhu.edu} \\
}

\date{}

\begin{document}
\maketitle

\begin{abstract}

We will show that any legal link grammar parse is a connected DAG, with all of the nodes reachable from the root. Every undirected edge in a link grammar parse can be assigned directionality, by specifying the task as an ILP problem given the constraints of acyclicity, connectedness, and the link grammar rules described from a dictionary.

\end{abstract}

\section{Introduction}

\TODO 
% Why would someone be interested in reading this paper?

This can be solved as an Integer Linear program. Encoded in the Zimpl little language \cite{Koch2004}.

\section{Previous Work}

\TODO

\section{Link Grammars}

Link Grammars are a formalism that describes the links, or relationships between the constituents \cite{SleatorTemperly91}. They are represented as

\bibliographystyle{acl}
\bibliography{LinksToDAG}

\end{document}

### 2014-05-12 src/tools/experiments/conll_analysis.py
wrote conll_analysis.py. Updated trial.sh to run it when called.

@@ -1,0 +1,276 @@
+#!/usr/bin/python
+
+import os
+from pprint import pprint
+
+CONLL_DIR = "data/"
+SOL_DIR = "sol/"
+
+CONLL_LOC = CONLL_DIR+"english_bnews_train.conll"
+SOL_LOC = SOL_DIR+"links.conll"
+
+ANALYSIS_DIR = "sol/conll_analysis/"
+if not os.path.exists(ANALYSIS_DIR):
+    os.makedirs(ANALYSIS_DIR)
+ANALYSIS_FILE = ANALYSIS_DIR+"conll_analysis.txt"
+
+
+# The result maps will map a sentence to its conll representation.
+# We will use this to query conll results from both our conll and link-conll files
+conll_results = {}
+link_results = {}
+
+# Populating the conll_results and link_results
+f = open(CONLL_LOC)
+f = f.readlines()
+sentence = []
+conll = []
+for line in f:
+    line = line.strip()
+    if not line:
+        conll_results[" ".join(sentence)] = tuple(conll)
+        sentence = []
+        conll = []
+    else:
+        conll.append(line)
+        word = line.split()[1]
+        sentence.append(word)
+
+# Populating the link_results
+f = open(SOL_LOC)
+f = f.readlines()
+sentence = []
+conll = []
+for line in f:
+    line = line.strip()
+    if not line:
+        link_results[" ".join(sentence)] = tuple(conll)
+        sentence = []
+        conll = []
+    else:
+        conll.append(line)
+        word = line.split()[1]
+        
+        # Get rid of the [] brackets around words that the link-parser could not attach.
+        word = word.strip("[]")
+        sentence.append(word)
+
+
+
+def analysis(conlls, links):
+    # How many links match the conll corpus
+    matches = 0
+    mismatches = 0
+    extras = 0
+    total = 0
+
+    label_match = {}
+    label_match_count = {}
+
+    label_extra_count = {}
+
+    # Directionality mismatch
+    dir_miss = 0
+    dir_miss_labels = {}
+    dir_miss_labels_count = {}
+
+    for conll,link in zip(conlls, links):
+        conll = conll.split()
+        link = link.split()
+
+        conll_head = conll[6]
+        link_heads = link[6].split(",")
+        conll_label = conll[7]
+        link_labels = link[7].split(",")
+
+        if conll_head not in link_heads:
+            head = int(conll[6])-1
+
+            candidate = links[head]
+            candidate = candidate.split()
[…]

### 2014-05-12 src/tools/experiments/trial.sh

@@ -13,3 +13,8 @@ tail --lines 1 output_error >> /tmp/LinksToDAG_times
 
 cp sol/allowedLinks.txt /tmp/LinksToDAG_trial_sol/allowedLinks_$ITERATIONS.txt
 rm -f output_error
+
+
+python src/tools/experiments/conll_analysis.py
+mv sol/conll_analysis/conll_analysis.txt sol/conll_analysis/conll_analysis_$ITERATIONS.txt
+

### 2014-05-12 src/link_parser.py
fixed some divide-by-zero bugs. renamed a file to something more descriptive. modified the trials.sh script.

@@ -11,7 +11,7 @@ class linkParser:
     """ Will call the link-parser on a sentence and output the result"""
     
     def __init__(self):
-        self.cache = cache()
+        self.cache = cache("/tmp/LinksToDAG_linkparses_cache.p")
 
     def parse_all(self,sentences, outputFile = "/tmp/LinksToDAG_linkparses.txt"):
         output = open(outputFile, "wb")

### 2014-05-12 src/tools/experiments/conll_analysis.py

@@ -174,6 +174,15 @@ for linkSentence in link_results:
 
 
 
+match_percent = str(0)
+mismatch_percent = str(0)
+if final_total != 0:
+    match_percent = str(float(match_total)/final_total)
+    mismatch_percent = str(float(mismatch_total)/final_total)
+
+directional_percent = str(0)
+if mismatch_total != 0:
+    directional_percent = str(float(dir_mismatch_total)/mismatch_total)
 
 
 result= """------------------------------------------------------------
@@ -182,15 +191,15 @@ TOTALS
 
 How many conll arcs match a link. In both attachment and directionality.
 Match Total:\t\t\t"""+ str(match_total)+""" 
-Percent of all arcs:\t\t"""+ str(float(match_total)/final_total)+"""
+Percent of all arcs:\t\t"""+ match_percent +"""
 
 How many conll arcs do not match a link. In either attachment or directionality.
 Mismatch Total:\t\t\t"""+str(mismatch_total)+"""\t
-Percent of all arcs:\t\t"""+ str(float(mismatch_total)/final_total)+"""
+Percent of all arcs:\t\t"""+ mismatch_percent +"""
 
 How many conll arcs do not match a link in only directionality?
 Directional Mismatch Total:\t"""+str(dir_mismatch_total)+"""
-Percent of all mismatches:\t"""+str(float(dir_mismatch_total)/mismatch_total)+"""
+Percent of all mismatches:\t"""+ directional_percent +"""
 
 
 How many links attach to a word when there is already a matching link to the conll data. How many "extra" arcs.

### 2014-05-12 src/tools/experiments/trial.sh

@@ -5,14 +5,14 @@ ITERATIONS=$2
 
 mkdir -p /tmp/LinksToDAG_trial_sol/
 
-/usr/bin/time -f "%e" ./LinksToDAG $DATA $ITERATIONS 2> output_error
+/usr/bin/time -f "%e" -o runtime ./LinksToDAG $DATA $ITERATIONS #2> output_error
 
 echo -n $ITERATIONS " " >> /tmp/LinksToDAG_times
-tail --lines 1 output_error >> /tmp/LinksToDAG_times
+tail --lines 1 runtime >> /tmp/LinksToDAG_times
 
 
 cp sol/allowedLinks.txt /tmp/LinksToDAG_trial_sol/allowedLinks_$ITERATIONS.txt
-rm -f output_error
+rm -f runtime
 
 
 python src/tools/experiments/conll_analysis.py

### 2014-05-12 trials.sh

@@ -20,17 +20,26 @@ rm -rf /tmp/LinksToDAG_trial_sol/
 
 typeset -i i END
 
-#END=30
-END=60030
-INCREMENT=1
+#END=10000
+#END=60030
+END=10000
+INCREMENT=1000
 
 for ((i=1;i<=$END;i=$((i*2))));
 do
     bash src/tools/experiments/trial.sh data/english_bnews_train.sentences $i
     #INCREMENT=$((INCREMENT+1))
 done
-bash src/tools/experiments/trial.sh data/english_bnews_train.sentences $END
 
+
+# Lets sample the last section more densely.
+i=$((i/2))
+for ((i=$((i/2));i<=$END;i=$((i+$INCREMENT))));
+do
+    bash src/tools/experiments/trial.sh data/english_bnews_train.sentences $i
+done
+
+bash src/tools/experiments/trial.sh data/english_bnews_train.sentences $END
 
 
 python src/tools/plotter.py -x "Sentences" -y "Runtime (seconds)" -o $SOL_RUNTIME $RUNTIMES
@@ -38,7 +47,7 @@ python src/tools/plotter.py -x "Sentences" -y "Runtime (seconds)" -o $SOL_RUNTIM
 
 
 
-python src/tools/experiments/analysis.py
+python src/tools/experiments/type_analysis.py
 python src/tools/plotter.py -x "Sentences" -y "Precision" -o $SOL_PRECISION $TXT_PRECISION
 python src/tools/plotter.py -x "Sentences" -y "Recall" -o $SOL_RECALL $TXT_RECALL
 

### 2014-05-12 doc/LinksToDAG.tex
merged trials.sh

@@
 \documentclass[11pt]{article}
-\usepackage{acl2014}
+\usepackage{LinksToDAG}
 \usepackage{times}
@@
 
-\author{Juneki Hong \\
+\author{Juneki Hong and Jason Eisner\\
   Department of Computer Science \\
@@
   Baltimore, MD 21218, USA \\ 
-  {\tt juneki@cs.jhu.edu} \\\And
-  Jason Eisner \\
-  Department of Computer Science \\
-  Johns Hopkins University \\
-  Baltimore, MD 21218, USA \\
-  {\tt jason@cs.jhu.edu} \\
+  {\tt \{juneki,jason\}@cs.jhu.edu} \\
 }
@@
 
-\bibliographystyle{acl}
+\section{ILP Encoding}
+We ignored the link parses that the link parser could not find suitable attachments and returned a disconnected graph.
+
+\section{Conll-Link analysis}
+
+\bibliographystyle{LinksToDAG}
 \bibliography{LinksToDAG}

### 2014-05-12 trials.sh

@@ -20,10 +20,13 @@ rm -rf /tmp/LinksToDAG_trial_sol/
 
 typeset -i i END
 
+
 #END=10000
 #END=60030
 END=10000
 INCREMENT=1000
+
+
 
 for ((i=1;i<=$END;i=$((i*2))));
 do

### 2014-05-12 trials.sh
comment out the linear for loop after the exponential one

@@ -15,6 +15,7 @@ TYPE_ANALYSIS=$TYPE_AGREEMENT"type_analysis.txt"
 
 
 
+
 rm -f $RUNTIMES $TXT_RUNTIME $SOL_RUNTIME $TXT_PRECISION $SOL_PRECISION $SOL_RECALL
 rm -rf /tmp/LinksToDAG_trial_sol/
 
@@ -26,21 +27,21 @@ typeset -i i END
 END=10000
 INCREMENT=1000
 
-
+CONLL_ANALYSIS="sol/conll_analysis/conll_analysis_"$END".txt"
 
 for ((i=1;i<=$END;i=$((i*2))));
 do
     bash src/tools/experiments/trial.sh data/english_bnews_train.sentences $i
-    #INCREMENT=$((INCREMENT+1))
 done
 
-
-# Lets sample the last section more densely.
+<<BLOCK_COMMENT
+# Sampling the last section more densely.
 i=$((i/2))
 for ((i=$((i/2));i<=$END;i=$((i+$INCREMENT))));
 do
     bash src/tools/experiments/trial.sh data/english_bnews_train.sentences $i
 done
+BLOCK_COMMENT
 
 bash src/tools/experiments/trial.sh data/english_bnews_train.sentences $END
 
@@ -56,4 +57,4 @@ python src/tools/plotter.py -x "Sentences" -y "Recall" -o $SOL_RECALL $TXT_RECAL
 
 
 
-echo $RUNTIMES $END | mutt -s $END -a $RUNTIMES -a $TXT_PRECISION -a $TXT_RECALL -a $TYPE_ANALYSIS -- junekihong@gmail.com
+echo $RUNTIMES $END | mutt -s $END -a $RUNTIMES -a $TXT_PRECISION -a $TXT_RECALL -a $TYPE_ANALYSIS -a $CONLL_ANALYSIS -- junekihong@gmail.com

### 2014-05-13 src/link_parser.py
handled merge for trials.sh

@@ -10,8 +10,9 @@ import subprocess, sys, os
 class linkParser:
     """ Will call the link-parser on a sentence and output the result"""
     
-    def __init__(self):
+    def __init__(self, lang = "en"):
         self.cache = cache("/tmp/LinksToDAG_linkparses_cache.p")
+        self.language = lang
 
     def parse_all(self,sentences, outputFile = "/tmp/LinksToDAG_linkparses.txt"):
         output = open(outputFile, "wb")
@@ -20,13 +21,20 @@ class linkParser:
         for x in iterview(xrange(len(sentences)), msg='Sentences Processed',every=1):
             sentence = sentences[x].strip()
 
+            if self.language != "en":
+                ID = sentence + "#" + self.language
+            else:
+                ID = sentence
+
             # In the cache
-            if self.cache.check(sentence): 
-                parse = self.cache.get(sentence)
+            if self.cache.check(ID): 
+                parse = self.cache.get(ID)
             # Not in the cache
             else:
                 parse = self.parse(sentence)
-                self.cache.store(sentence,parse)
+                self.cache.store(ID,parse)
+
+            print parse
 
             output.write(parse)
         output.close()
@@ -36,9 +44,8 @@ class linkParser:
         self.cache.save()
 
     def parse(self,sentence):
-
         echo_process = Popen(["echo",sentence], stdout=PIPE)
-        linkparser_process = Popen(["link-parser", "-!graphics=0", "-!links=0", "-!echo=1", "-!postscript=1", "-!panic=0"], stdin=echo_process.stdout, stdout=PIPE, stderr=subprocess.PIPE)
+        linkparser_process = Popen(["link-parser", self.language, "-!graphics=0", "-!links=0", "-!echo=1", "-!postscript=1", "-!panic=0"], stdin=echo_process.stdout, stdout=PIPE, stderr=subprocess.PIPE)
         echo_process.stdout.close()
 
         output, err = linkparser_process.communicate()        
@@ -47,18 +54,26 @@ class linkParser:
 
 
 if __name__ == "__main__":
-    parser = linkParser()
+    if len(sys.argv) > 3:
+        language = sys.argv[3]
+        parser = linkParser(language)
+    else:
+        parser = linkParser()
+
     inputfile = open(sys.argv[1], "rb")
     sample = -1
+    
 
     if len(sys.argv) > 2:
         sample = int(sys.argv[2])
-    
+
     if sample == -1:
[…]

### 2014-05-13 src/tools/experiments/conll_analysis.py

@@ -57,23 +57,16 @@ for line in f:
 
 
 
-def analysis(conlls, links):
-    # How many links match the conll corpus
+# Analysis on the matches.
+def analysis_match(conlls, links):
     matches = 0
-    mismatches = 0
     extras = 0
     total = 0
-
     label_match = {}
     label_match_count = {}
-
     label_extra_count = {}
 
-    # Directionality mismatch
-    dir_miss = 0
-    dir_miss_labels = {}
-    dir_miss_labels_count = {}
-
+    total = len(conlls)
     for conll,link in zip(conlls, links):
         conll = conll.split()
         link = link.split()
@@ -84,45 +77,97 @@ def analysis(conlls, links):
         link_labels = link[7].split(",")
 
         if conll_head not in link_heads:
-            head = int(conll[6])-1
+            continue
 
-            candidate = links[head]
-            candidate = candidate.split()
-            candidate_heads = candidate[6].split(",")
-            candidate_labels = candidate[7].split(",")
-
-            index = conll[0]
-            
-            for head,label in zip(candidate_heads,candidate_labels):
-                if head == index:
-                    dir_miss += 1
-                    dir_miss_labels[conll_label] = dir_miss_labels.get(conll_label,set([])).union(set([label]))
-                    dir_miss_labels_count[(conll_label,label)] = dir_miss_labels_count.get((conll_label,label),0)+1
-
-            mismatches += 1
-        else:
-            for head,label in zip(link_heads, link_labels):
-                if head == conll_head:
-                    #print conll[0], conll[1], head
-                    matches += 1 
-                    label_match[conll_label] = label_match.get(conll_label, set([])).union(set([label]))
-                    label_match_count[(conll_label,label)] = label_match_count.get((conll_label,label), 0) + 1
-                else:
-                    extras += 1
-                    #label_extra = label_extra.union(set([label]))
-                    label_extra_count[label] = label_extra_count.get(label, 0) + 1
-
-        total += 1
+        for head,label in zip(link_heads, link_labels):
+            if head == conll_head:
+                matches += 1 
+                label_match[conll_label] = label_match.get(conll_label, set([])).union(set([label]))
[…]

### 2014-05-13 trials.sh

@@ -34,9 +34,9 @@ do
     bash src/tools/experiments/trial.sh data/english_bnews_train.sentences $i
 done
 
+
 <<BLOCK_COMMENT
-# Sampling the last section more densely.
-i=$((i/2))
+# Lets sample the last section more densely.
 for ((i=$((i/2));i<=$END;i=$((i+$INCREMENT))));
 do
     bash src/tools/experiments/trial.sh data/english_bnews_train.sentences $i

### 2014-05-16 trials.sh
merged trials.sh

@@ -4,8 +4,8 @@ typeset -i i END
 
 
 #END=10000
-#END=60030
-END=100
+END=60030
+#END=100
 #INCREMENT=1
 
 SENTENCES="data/english_bnews_train.sentences"
@@ -40,7 +40,6 @@ TYPE_ANALYSIS=$TYPE_AGREEMENT"type_analysis.txt"
 
 rm -f $RUNTIMES $TXT_RUNTIME $SOL_RUNTIME $TXT_PRECISION $SOL_PRECISION $SOL_RECALL
 rm -rf /tmp/LinksToDAG_trial_sol/
-
 
 CONLL_ANALYSIS="sol/conll_analysis/conll_analysis_"$END".txt"
 

### 2014-05-16 src/encoder.py
try to adjust encoder.py to handle the output of different versions of linkparser.

@@ -9,7 +9,7 @@ def getDataFromLinkParse(lines):
     #data format: (index1, index2, layer, label)
     links = []
 
-    #pprint(lines)
+    #pprint(lines) 
 
     # get rid of the beginning statements
     i = 2
@@ -21,6 +21,14 @@ def getDataFromLinkParse(lines):
     lines = lines[i:]
 
 
+
+    # Get rid of the last couple of lines because its just junk to us.
+    # We don't want to process the "Press RETURN for the next linkage" message
+    endString = "Press RETURN for the next linkage."
+    if endString in lines:
+        index = lines.index(endString)
+        lines = lines[:index-1]
+    
 
     sentence = ""
     processedSentence = []
@@ -38,6 +46,9 @@ def getDataFromLinkParse(lines):
             else:
                 continue
         
+
+        
+
         # extract the original sentence. Populate the corpus.
         if line.find("linkparser>") == 0:
             sentence = (line[len("linkparser>"):]).strip()

### 2014-05-17 doc/LinksToDAG.tex
working on the paper. Added a little bit of outline and some paragraphs.

@@
 \usepackage{latexsym}
+\usepackage{amsmath}
+
+\usepackage{algorithm}
+\usepackage[noend]{algpseudocode}
+\usepackage{graphicx}
 
@@
 
-\title{Link Parses to Directed Acyclic Graphs}
+\title{Deriving Multi-headed Planar Dependency Parses from Link Grammar Parses}
 
@@
 
-We will show that any legal link grammar parse is a connected DAG, with all of the nodes reachable from the root.
-Every undirected edge in a link grammar parse can be assigned directionality, by specifying the task as an ILP problem given the constraints of acyclicity, connectedness, and the link grammar rules described from a dictionary.
+% Copied from the google doc
+There has been recent theoretical work on multi-headed versions of dependency grammar (cite).
+Such formalisms can be more syntactically and semantically expressive.
+However, there aren't any corpora at present (check this) except for Buch-Kromann's work on Danish (cite), so it is hard to investigate the benefit of such parsers or to work on making them faster or more accurate.
+To fill this gap, we observe that link grammar produces parses that are similar to multi-headed planar dependency parses except that the links are undirected.
+We use Integer Linear Programming to assign consistent directions to the links in a corpus of NNN parses produced by the Link Grammar Parser, which has broad-coverage hand-written grammars of English, Arabic, and ...
+(what statistics does it use?).
+We find that such directions can indeed be consistently assigned in a way that yields valid multi-headed dependency parses.
+The resulting parses in English appear linguistically plausible, although they are not in general consistent with CoNLL-style parses of the same sentences; we discuss the differences.
+We also report ...
 
@@
 
-\TODO 
-% Why would someone be interested in reading this paper?
-
-This can be solved as an Integer Linear program. Encoded in the Zimpl little language \cite{Koch2004}.
-
-\section{Previous Work}
-
-\TODO
+Dependency Parsing is the task of mapping a sentence to a projective (not always projective?) directed acyclic tree.
+Link Parsing in contrast produces a planar graph with undirected edges, where every edge has a label.
+The labels describe the relationship between two constituents in a parse, and in this paper we will analyze whether this includes dependencies as well.
+In order to learn the directional dependencies within the Link edge labels we will use Integer Linear Programming, encoding the problem in the Zimpl little language \cite{Koch2004}.
 
@@
 
-Link Grammars are a formalism that describes the links, or relationships between the constituents \cite{SleatorTemperly91}. They are represented as
+Link Grammars are a formalism that describes the links, or relationships between the constituents \cite{SleatorTemperly91}.
 
-\section{ILP Encoding}
-We ignored the link parses that the link parser could not find suitable attachments and returned a disconnected graph.
+We ran the link parser on the english bnews corpus sentences \footnote{We ignored the link parses that the link parser could not find suitable attachments and returned a disconnected graph.
+This happened for roughly NNN\% of the corpus.}
+
+The link grammar on which the link-parser makes its linking decisions based on a set of handwritten dictionaries.
+Instead of going through these dictionaries, we learned the link grammar using an ILP.
+This approach allows us to analyze any link grammar dictionary other than English.
+
+\section{Integer Linear Programming}
+
+We formulated the ILP to assign directional dependencies to a link grammar parse such that it is a connected DAG, with all of the nodes reachable from the root.
+
+The ILP would solve it
+... given certain constraints.
+
+\subsection{Constraints}
+
+The constraints were
+acyclicity and connectedness
+
+\subsection{Slack Hierarchy}
+
+In order to encourage the same link types to align similarly with each other,
+we came up with a slack hierarchy.
+
+\begin{algorithm}
+\caption{ILP Encoding}\label{ILP}
+\begin{algorithmic}[1]
+%\Procedure{MyProcedure}{}
+
+\State $\text{min} \sum_{l \in labels} allowed_L(l) + allowed_R(l)$
+\State $\text{subject to}$
+
+%\EndProcedure
+\end{algorithmic}
+\end{algorithm}
+
+\subsection{Stability of Results}
+
+\section{Other languages}
 
 \section{Conll-Link analysis}
+
+\section{Related Work}
+
+\TODO {abc} def
+
+\section{Results}
+
+\section{Discussion}
 

### 2014-05-18 src/tools/plotter.py
plotter.py plots the x-axis range up to the maximum data point. Instead of padding out the xrange to a nicer number

@@ -45,7 +45,11 @@ for line in lines:
     
     Xs.append(x)
     Ys.append(y)
-    
+
+
+# Get the minimum, maximum X values
+xmin = 0
+xmax = max(Xs)
 
 
 plt.figure()
@@ -54,7 +58,7 @@ ax.set_xlabel(args.xLabel)
 ax.set_ylabel(args.yLabel)
 
 plt.plot(Xs,Ys)
-plt.xlim(xmin=0)
+plt.xlim(xmin, xmax)
 plt.ylim(ymin=0)
 
 

### 2014-05-18 trials.sh
updated trials.sh to sort the precision.txt and recall.txt data files before plotting them.

@@ -58,6 +58,9 @@ python src/tools/plotter.py -x "Sentences" -y "Runtime (seconds)" -o $SOL_RUNTIM
 
 
 python src/tools/experiments/type_analysis.py 
+cat $TXT_PRECISION | sort -V > temp; mv temp $TXT_PRECISION
+cat $TXT_RECALL | sort -V > temp; mv temp $TXT_RECALL
+
 python src/tools/plotter.py -x "Sentences" -y "Precision" -o $SOL_PRECISION $TXT_PRECISION
 python src/tools/plotter.py -x "Sentences" -y "Recall" -o $SOL_RECALL $TXT_RECALL
 

### 2014-05-18 trials.sh
handled a merge

@@ -52,6 +52,10 @@ if [ $i -ne $END -a $((i/2)) -ne $END ]; then
     bash src/tools/experiments/trial.sh $SENTENCES $CONLL $END $LANGUAGE
 fi
 
+
+
+
+
 python src/tools/plotter.py -x "Sentences" -y "Runtime (seconds)" -o $SOL_RUNTIME $RUNTIMES
 
 
@@ -66,4 +70,4 @@ python src/tools/plotter.py -x "Sentences" -y "Recall" -o $SOL_RECALL $TXT_RECAL
 
 
 
-#echo $RUNTIMES $END | mutt -s $END -a $RUNTIMES -a $TXT_PRECISION -a $TXT_RECALL -a $TYPE_ANALYSIS -a $CONLL_ANALYSIS -- junekihong@gmail.com
+echo $RUNTIMES $END | mutt -s $END -a $RUNTIMES -a $TXT_PRECISION -a $TXT_RECALL -a $TYPE_ANALYSIS -a $CONLL_ANALYSIS -- junekihong@gmail.com

### 2014-05-18 src/tools/plotter.py
updated plotter to handle multiple files. Plotting multiple lines on the same graph.

@@ -1,10 +1,8 @@
 #!/usr/bin/env python
 
-import math
+import math, argparse, sys
 import matplotlib.pyplot as plt
-import argparse
-import sys
-
+from pprint import pprint
 
 parser = argparse.ArgumentParser(description="Plots a list of x,y points.")
 parser.add_argument("-o", "--output", dest="output", default="sol/runtimes.png",
@@ -18,38 +16,58 @@ parser.add_argument("strings", nargs="*")
 args = parser.parse_args()
 
 
-lines = []
-for f in args.strings:
-    f = open(f)
+labels = []
+file_lines = []
+for filename in args.strings:
+    label = filename.split(".")[0]
+    label = label.split("/")[-1]
+    label = label.split("_")[0]
+    labels.append(label)
+    
+    lines = []
+    f = open(filename)
     f = f.readlines()
     for line in f:
         line = line.strip()
         lines.append(line)
+    file_lines.append(lines)
 
 
 
-Xs = []
-Ys = []
-
-
-for line in lines:
-    if not line:
-        continue
-    if line[0] == "#":
-        continue
-
-
-    line = line.split()
-    x = float(line[0])
-    y = float(line[1])
+file_coord = []
+for lines in file_lines:
+    Xs = []
+    Ys = []
     
-    Xs.append(x)
-    Ys.append(y)
+    coord = []
+
+    for line in lines:
+        if not line:
+            continue
+        if line[0] == "#":
+            continue
+
+        line = line.split()
+        x = float(line[0])
+        y = float(line[1])
+    
+        coord.append((x,y))
+    file_coord.append(coord)
 
 
 # Get the minimum, maximum X values
+xmax = None
 xmin = 0
-xmax = max(Xs)
+for coord in file_coord:
+    temp_max = max(coord)
+    if xmax == None or xmax < temp_max:
+        xmax = temp_max
+
+    temp_min = min(coord)
+    if xmin > temp_min:
+        xmin = temp_min
+xmax = xmax[0]
+
 
 
 plt.figure()
@@ -57,7 +75,15 @@ ax = plt.subplot(111)
 ax.set_xlabel(args.xLabel)
 ax.set_ylabel(args.yLabel)
 
-plt.plot(Xs,Ys)
+
+for l,coord in zip(labels,file_coord):
+    Xs = [elem[0] for elem in coord]
[…]

### 2014-05-18 doc/LinksToDAG.tex
merged trials.sh. Updated the email script at the end to tell me what to run to regenerate the graphs.

@@
 
-The ILP would solve it
-... given certain constraints.
+The ILP chooses whether to set a link type to be assigned left, right, or both.
+For this, two label\_direction variables were made for each link type that appears in the parses, corresponding to whether all of the $l$ link edges were to be set to go left, right, or both.
+Because we assign a directionality to every link type, at least one of these variables must be set to true.
 
-\subsection{Constraints}
-
-The constraints were
-acyclicity and connectedness
+For example, setting the variables $$l_{LEFT} = TRUE, l_{RIGHT} = FALSE$$ would mean that the directional dependency of all links with label $l$ would be set to $LEFT$.
+If both variables were set to true, then the link type is allowed to go in either direction.
+Our ILP objective is to minimize the number of label\_direction variables set to true, while satisfying the other constraints.
 
@@
 
-In order to encourage the same link types to align similarly with each other,
-we came up with a slack hierarchy.
+We introduced slack on the directional dependency variable tokens such that link types were allowed to deviate from the majority up to (1\% ?) of the time before the ILP would assign both directional dependency variables to true.
+This addressed noise in the link parser's label assignments, while still allowing for the possibility that both directions could still be assigned.
+
+\TODO{} We also introduced slack on the link types such that link types with the same coarse grained label would try and align the same way as the majority in the group, where the preceding capital letters of the link type denote the coarse grained label, while the subscript letters denote further information.
+This slack places a prior on rare or never-before seen link types to be assigned in the same way as other similar link types.
+
+This slack hierarchy gave the ILP the flexibility to handle noise and novel link types while still trying to learn the overall link grammar.
 
@@
 \subsection{Stability of Results}
+We tested to see how the solution changed as we increased the number of processed sentences.
+Taking the largest run as the answer key, we analyzed how much the previous runs deviated from it.
+We measured the precision of whether the assignments in the smaller runs matched the assignments of the largest run, and we measured the recall of whether the assignments in the largest run could be found in the smaller runs.
+
+\includegraphics[width=\linewidth, keepaspectratio=true]{figure/precision_recall.png}
 

### 2014-05-18 trials.sh

@@ -5,7 +5,7 @@ typeset -i i END
 
 #END=10000
 END=60030
-#END=100
+#END=10
 #INCREMENT=1
 
 SENTENCES="data/english_bnews_train.sentences"
@@ -29,8 +29,7 @@ SOL_RUNTIME="sol/runtimes.png"
 PRECISION_RECALL="sol/precision_recall/"
 TXT_PRECISION=$PRECISION_RECALL"precision.txt"
 TXT_RECALL=$PRECISION_RECALL"recall.txt"
-SOL_PRECISION=$PRECISION_RECALL"precision.png"
-SOL_RECALL=$PRECISION_RECALL"recall.png"
+SOL_PRECISION_RECALL=$PRECISION_RECALL"precision_recall.png"
 
 TYPE_AGREEMENT="sol/type_agreement/"
 TYPE_ANALYSIS=$TYPE_AGREEMENT"type_analysis.txt"
@@ -65,9 +64,9 @@ python src/tools/experiments/type_analysis.py
 cat $TXT_PRECISION | sort -V > temp; mv temp $TXT_PRECISION
 cat $TXT_RECALL | sort -V > temp; mv temp $TXT_RECALL
 
-python src/tools/plotter.py -x "Sentences" -y "Precision" -o $SOL_PRECISION $TXT_PRECISION
-python src/tools/plotter.py -x "Sentences" -y "Recall" -o $SOL_RECALL $TXT_RECALL
+python src/tools/plotter.py -x "Sentences" -y "Percent" -o $SOL_PRECISION_RECALL $TXT_PRECISION $TXT_RECALL
 
 
 
-echo $RUNTIMES $END | mutt -s $END -a $RUNTIMES -a $TXT_PRECISION -a $TXT_RECALL -a $TYPE_ANALYSIS -a $CONLL_ANALYSIS -- junekihong@gmail.com
+echo $RUNTIMES $END $'\n'The CLSP machines cant run plotter. Please download files and run: $'\n'python src/tools/plotter.py -x "Sentences" -y "Percent" -o $SOL_PRECISION_RECALL $TXT_PRECISION $TXT_RECALL | mutt -s $END -a $RUNTIMES -a $TXT_PRECISION -a $TXT_RECALL -a $TYPE_ANALYSIS -a $CONLL_ANALYSIS -- junekihong@gmail.com
+

### 2014-05-18 src/decoder.py
updated decoder because of a small change made to the zimpl_template

@@ -19,7 +19,7 @@ def decodeSCIPsolution(links, solutionFile, getAllowedLabels=False):
 
 
     for line in lines:
-        if line.find("rlink#") == 0:
+        if line.find("rLink#") == 0:
             line = line.split()
             ID = line[0]
             direction = int(round(float(line[1])))

### 2014-05-18 doc/compile.sh
Wrote more paragraphs of paper. The latex compile script regenerates the precision_recall graph using the makefile before generating the paper.

@@ -1,5 +1,8 @@
 #!/bin/sh
 
+cd ..; make graph_precision_recall; cd doc;
+
+
 if [ $1 ];
 then
 pdflatex $1 && bibtex *.aux && pdflatex $1 && pdflatex $1

### 2014-05-18 doc/LinksToDAG.tex
Wrote more paragraphs of paper. The latex compile script regenerates the precision_recall graph using the makefile before generating the paper.
paper writing

@@
 To fill this gap, we observe that link grammar produces parses that are similar to multi-headed planar dependency parses except that the links are undirected.
-We use Integer Linear Programming to assign consistent directions to the links in a corpus of NNN parses produced by the Link Grammar Parser, which has broad-coverage hand-written grammars of English, Arabic, and ...
+We use Integer Linear Programming to assign consistent directions to the links in a corpus of NNN parses produced by the Link Grammar Parser, which has broad-coverage hand-written grammars of English, Russian, and more.
 (what statistics does it use?).
@@
 
-Dependency Parsing is the task of mapping a sentence to a projective (not always projective?) directed acyclic tree.
-Link Parsing in contrast produces a planar graph with undirected edges, where every edge has a label.
-The labels describe the relationship between two constituents in a parse, and in this paper we will analyze whether this includes dependencies as well.
-In order to learn the directional dependencies within the Link edge labels we will use Integer Linear Programming, encoding the problem in the Zimpl little language \cite{Koch2004}.
+Dependency parsing is the task of mapping a sentence to a projective (not always projective?) directed acyclic tree.
+Link parsing in contrast produces a multiheaded planar graph with undirected edges, where every edge has a label describing the relationship between two constituents in a parse.
+In this paper we explore whether these relationships include dependencies.
+To determine the directional dependencies within the link edge labels we will use integer linear programming, encoding the problem in the Zimpl little language \cite{Koch2004}.
+It turns out that the link parses roughly only match half of the conll dependency corpus.
+However this is because \TODO{}.
+
+%Despite this, the link parses offer additional links in addition to dependencies...
 
@@
 
-Link Grammars are a formalism that describes the links, or relationships between the constituents \cite{SleatorTemperly91}.
+Link grammars describe the links, or relationships between constituents in a parse \cite{SleatorTemperly91}.
 
@@
 
+\TODO{cite model minimization papers}
+
 \subsection{Slack Hierarchy}
@@
 \State $\text{subject to}$
+\State $\text{Acyclicity}$
+\State $\text{All nodes reachable from the root.}$
 
@@
 \subsection{Stability of Results}
-We tested to see how the solution changed as we increased the number of processed sentences.
-Taking the largest run as the answer key, we analyzed how much the previous runs deviated from it.
+We tested to see how the solution changed over increasing numbers of processed sentences. 
+Taking the solution to the largest run as the answer key, we measured how much the previous runs deviated from it.
 We measured the precision of whether the assignments in the smaller runs matched the assignments of the largest run, and we measured the recall of whether the assignments in the largest run could be found in the smaller runs.
 
+\begin{figure}[ht!]
+  \centering
   \includegraphics[width=\linewidth, keepaspectratio=true]{figure/precision_recall.png}
+  \caption{Precision and Recall of different runs on English}
+\end{figure}
+
+In the end, the precision values converged to 100\% while the recall values continued to grow throughout our data.
+This indicates that the solutions of the previous runs were consistent with the largest run.
+However as the data grew, we would continue to encounter novel link types.
 
@@
 
-\section{Conll-Link analysis}
+\section{Discussion}
 
 \section{Related Work}
-
-\TODO {abc} def
-
-\section{Results}
-
-\section{Discussion}
 

### 2014-05-19 doc/LinksToDAG.tex
Added tikz-dependency to the paper. I want to be able to generate dependency graphs to see. The paper will read off exampleSentence.tikz and paste it into its example sentence category.

@@
 
+\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
+\usepackage{soul}
+
 \usepackage{algorithm}
@@
 
-\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
-\usepackage{soul}
+\usepackage{tikz-dependency}
+
 \newcommand{\Note}[1]{}
@@
 
+\begin{equation}
+  count_{L, left} = \sum_{a \in ARCS with label L} left_a
+\end{equation}
+\begin{equation}
+  count_{L, left} = M * allowed_{L,left} + slack_{L}
+\end{equation}
+Where $L \in LABELS$, and $M$ is the size of the corpus.
+The objective is:
+\begin{equation}
+  \min \sum_{L \in LABELS \atop DIR \in \{left, right\}} (allowed_{L,DIR}) \atop + COST(L) * \sum_L slack_L
+\end{equation}
+\begin{equation}
+  COST(L) = \frac{100}{(count_{L,left} + count_{L,right})}
+\end{equation}
+
 \begin{algorithm}
@@
 
+\begin{figure}[ht!]
+  \centering
+  \input{figure/exampleSentence.tikz}
+  \caption{Example}
+\end{figure}
+
 \section{Related Work}

### 2014-05-19 src/decoder.py
deleted 2 newlines from decoder.py

@@ -195,8 +195,7 @@ def dotOutput(sentence, wordTag, linkDep, linkLabel):
     ID = "".join(sentence.split(" "))
     ID = re.sub(r'\W+', '', ID)
     ID = ID[:30]
-    sentence = sentence.split()
-    
+    sentence = sentence.split()    
 
     directory = "/tmp/LinksToDAG_dot/"
     # If needed, make the directory to put all of our dot files. 

### 2014-05-20 doc/LinksToDAG.tex
removed LinksToDAG.sty. Redownloaded acl2014.sty and added it to the repo.
changed LinksToDAG.bst to acl2014.bst
moved exampleSentence.tikz to sentences.tikz
added smallcaps to everything in a math equation

--- 2014-05-18 doc/LinksToDAG.tex
@@
 \documentclass[11pt]{article}
-\usepackage{LinksToDAG}
+\usepackage{acl2014}
 \usepackage{times}
@@
 \usepackage{graphicx}
+\usepackage{caption}
+\usepackage{subcaption}
 
-\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
-\usepackage{soul}
+\usepackage{tikz-dependency}
+
 \newcommand{\Note}[1]{}
@@
 
+\begin{equation}
+  count_{L, \textsc{left}} = \sum_{a \in \textsc{ARCS}_L} left_a
+\end{equation}
+\begin{equation}
+  count_{L, \textsc{left}} = \textsc{M} \cdot \textsc{allowed}_{L,\textsc{left}} + \textsc{slack}_{L}
+\end{equation}
+Where $L \in \textsc{LABELS}$, and $\textsc{M}$ is the size of the corpus.
+The objective is:
+\begin{equation}
+  \min \sum_{L \in \textsc{LABELS} \atop \textsc{DIR} \in \{left, right\}} (\textsc{allowed}_{L,\textsc{DIR}}) \atop + \textsc{COST}(L) * \sum_L \textsc{slack}_L
+\end{equation}
+\begin{equation}
+  \textsc{COST}(L) = \frac{100}{(\textsc{count}_{L,\textsc{left}} + \textsc{count}_{L,\textsc{right}})}
+\end{equation}
+
 \begin{algorithm}
@@
 \begin{figure}[ht!]
-  \centering
   \includegraphics[width=\linewidth, keepaspectratio=true]{figure/precision_recall.png}
@@
 
+\section{Link Corpus}
+
+Compared to the original conll data, the link parses match about half the conll arcs in location and directionality. The mismatches in the other half can be accounted for in the following ways.
+
+\subsection{Blanks}
+Whenever the link parser failed to recognize a word from its dictionary, it would skip it and fail to attach any links. This accounts for about 10\% of the mismatches.
+
+\subsection{Directional Mismatches}
+In about 25\% of the cases, the conll arc's corresponding link would be assigned the opposite direction.
+
 \section{Discussion}
+
+\begin{figure*}[ht!]
+  \centering
+  \input{figure/sentences.tikz}
+  \caption{Example Sentences}
+\end{figure*}
 
@@
 
-\bibliographystyle{LinksToDAG}
+\bibliographystyle{acl2014}
 \bibliography{LinksToDAG}

### 2014-05-20 src/tools/experiments/conll_analysis.py
conll_analysis checks for blank words as well and counts them up.

@@ -4,18 +4,30 @@ import sys, os
 from pprint import pprint
 
 CONLL_DIR = "data/"
-SOL_DIR = "sol/"
-
 CONLL_LOC = CONLL_DIR+"english_bnews_train.conll"
-SOL_LOC = SOL_DIR+"links.conll"
-
 if len(sys.argv) > 2:
     CONLL_LOC = sys.argv[2]
+if not os.path.exists(CONLL_DIR):
+    os.makedirs(CONLL_DIR)
+
+SOL_DIR = "sol/"
+SOL_LOC = SOL_DIR+"links.conll"
+if not os.path.exists(SOL_DIR):
+    os.makedirs(SOL_DIR)
 
 ANALYSIS_DIR = "sol/conll_analysis/"
+ANALYSIS_FILE = ANALYSIS_DIR+"conll_analysis.txt"
 if not os.path.exists(ANALYSIS_DIR):
     os.makedirs(ANALYSIS_DIR)
-ANALYSIS_FILE = ANALYSIS_DIR+"conll_analysis.txt"
+
+TIKZ_DIR = "doc/figure/"
+TIKZ_LOC = TIKZ_DIR+"sentences.tikz"
+if not os.path.exists(TIKZ_DIR):
+    os.makedirs(TIKZ_DIR)
+TIKZ = open(TIKZ_LOC, 'w+')
+
+ALL_PARSES_LOC = TIKZ_DIR+"allparses.tikz"
+ALL_PARSES = open(ALL_PARSES_LOC, 'w+')
 
 
 
@@ -122,6 +134,7 @@ def analysis_match(conlls, links):
 # Analysis on the mismatches.
 def analysis_mismatch(conlls, links):
     mismatches = 0
+    blanks = 0
 
     # Directionality mismatch
     dir_miss = 0
@@ -142,6 +155,10 @@ def analysis_mismatch(conlls, links):
         link_labels = link[7].split(",")
 
         if conll_head not in link_heads:
+            # For all the times the link parser does [this] and fails to attach a word
+            if len(link_heads) == 1 and link_heads[0] == "-":
+                blanks += 1
+
             mismatches += 1
             head = int(conll[6])-1
 
@@ -149,12 +166,10 @@ def analysis_mismatch(conlls, links):
             candidate_heads = candidate[6].split(",")
             candidate_labels = candidate[7].split(",")
             
-
             if index not in candidate_heads:
                 continue
 
             dir_miss += 1
-
[…]

### 2014-05-20 src/tools/plotter.py
fixed plotter to get the legend out of the filename

@@ -19,9 +19,11 @@ args = parser.parse_args()
 labels = []
 file_lines = []
 for filename in args.strings:
-    label = filename.split(".")[0]
-    label = label.split("/")[-1]
+
+    label = filename.split("/")[-1]
+    label = label.split(".")[0]
     label = label.split("_")[0]
+
     labels.append(label)
     
     lines = []
@@ -31,6 +33,7 @@ for filename in args.strings:
         line = line.strip()
         lines.append(line)
     file_lines.append(lines)
+
 
 
 

### 2014-05-20 doc/LinksToDAG.tex
note for discussion

@@
 
+We should acknowledge that another option would be to post-process either CoNLL parses or our link parses to automatically add more multi-heading, e.g., to handle control phenomena ("John wants to skip" or "Jane ran for office in order to change the tax laws" should have John/Jane be the subject of two verbs).
+However, that requires per-language
+
 \section{Related Work}

### 2014-05-20 doc/LinksToDAG.tex
added notes from our meeting

@@
 %Despite this, the link parses offer additional links in addition to dependencies...
+
+\NoteJE{discuss multiheadedness.  cite Kromann 2003's danish
+  dependency treebank.  Prior work: McDonald and Pereira 2006, Sagae
+  and Tsujii 2008, Gomez-Rodriguez and Nivre 2013.}
+% http://aclweb.org/anthology/J/J13/J13-4002.pdf
+% http://citeseerx.ist.psu.edu/viewdoc/download?doi=10.1.1.156.3938&rep=rep1&type=pdf
 
@@
 
+\TODO{more in the appendix ...}
+
 We should acknowledge that another option would be to post-process either CoNLL parses or our link parses to automatically add more multi-heading, e.g., to handle control phenomena ("John wants to skip" or "Jane ran for office in order to change the tax laws" should have John/Jane be the subject of two verbs).
@@
 
+\appendix
+\section{English Link Types}
+
+\TODO{A table here that lists each link grammar type together with its
+  count, its allowed direction(s), its percentage of majority
+  directions by token, and its most common corresponding CoNLL tags
+  with percentages.  Readers could cut-and-paste this data to make a 
+  cheap post-processor for LG parses.}
+
+\section{Sample English Parses}
+
+\TODO{Show a bunch of English sentences with CoNLL vs.\@ link parses.
+  Explain how they were selected and how the results are colorized etc.}
+
 \end{document}

### 2014-05-24 src/tools/experiments/conll_analysis.py
changed and restructured the conll_analysis.py script. Broke it up into a few smaller entities. Introduced an adjacency list data structure to represent the parses. Displayed them using tikz_dependency.

@@ -2,6 +2,9 @@
 
 import sys, os
 from pprint import pprint
+from tikz_dependency import *
+from dependency_graph import *
+from conll_analysis_output import *
 
 CONLL_DIR = "data/"
 CONLL_LOC = CONLL_DIR+"english_bnews_train.conll"
@@ -30,11 +33,139 @@ ALL_PARSES_LOC = TIKZ_DIR+"allparses.tikz"
 ALL_PARSES = open(ALL_PARSES_LOC, 'w+')
 
 
+# Analysis
+def analysis(conlls, links):
+
+    word_data = analysis_words(conlls, links)
+    
+    graph_conlls = getGraph(conlls)
+    graph_links = getGraph(links)
+
+    matches = []
+    reverse_matches = []
+    extras = []
+
+    label_match = {}
+    label_match_count = {}
+
+    label_reverse = {}
+    label_reverse_count = {}
+
+    label_extra = {}
+    label_extra_count = {}
+    
+    mismatches = 0
+    blanks = 0
+    total = 0
+    
+    for head in graph_conlls.table:
+        children = graph_conlls.table[head]
+        total += len(children)
+
+        for child in children:
+            conll_label = graph_conlls.getEdge(head,child)
+
+            # If we have a match
+            if graph_links.existsEdge(head,child):
+                matches.append((head,child))
+                link_label = graph_links.getEdge(head,child)
+                
+                label_match[conll_label] = label_match.get(conll_label, set([])).union(set([link_label]))
+                label_match_count[(conll_label, link_label)] = label_match_count.get((conll_label, link_label),0) + 1
+
+            # If we have a reverse match
+            elif graph_links.existsEdge(child,head):
+                reverse_matches.append((child,head))
+                link_label = graph_links.getEdge(child,head)
+                
+                label_reverse[conll_label] = label_reverse.get(conll_label, set([])).union(set([link_label]))
+                label_reverse_count[(conll_label, link_label)] = label_reverse_count.get((conll_label, link_label), 0) + 1
+
+            # We have a mismatch
+            else:
+                mismatches += 1
+                if not graph_links.heads.get(child,[]):
+                    blanks +=1
+
+    # Delete the matches from the graphs, so we can work with the rest
+    for match in matches:
+        (head,child) = match
+        graph_conlls.deleteEdge(head,child)
+        graph_links.deleteEdge(head,child)
+
[…]

### 2014-05-24 src/tools/experiments/conll_analysis_output.py

@@ -1,0 +1,205 @@
+#!/usr/bin/python
+
+import sys, os
+from pprint import pprint
+from tikz_dependency import *
+from dependency_graph import *
+
+def result_numbers(final_total, 
+                   match_total, 
+                   extra_total, 
+                   blank_total, 
+                   mismatch_total, 
+                   reverse_match_total,
+                   dropped_word_total,
+                   multiheaded_word_count, 
+                   word_count_total, 
+                   dropped_sentence_count,
+                   multiheaded_sentence_count, 
+                   total_sentence_count):
+
+    match_percent = str(0)
+    mismatch_percent = str(0)
+    if final_total != 0:
+        match_percent = str(float(match_total)/final_total)
+        mismatch_percent = str(float(mismatch_total)/final_total)
+
+    blank_percent = str(0)
+    if blank_total != 0:
+        blank_percent = str(float(blank_total)/mismatch_total)
+
+    directional_percent = str(0)
+    if mismatch_total != 0:
+        directional_percent = str(float(reverse_match_total)/mismatch_total)
+
+    multiheaded_word_percent = str(0)
+    dropped_word_percent = str(0)
+    if word_count_total != 0:
+        multiheaded_word_percent = str(float(multiheaded_word_count)/word_count_total)
+        dropped_word_percent = str(float(dropped_word_total)/word_count_total)
+
+
+
+    return """------------------------------------------------------------
+ARC-LINK TOTALS
+------------------------------------------------------------
+
+How many conll arcs match a link. In both attachment and directionality.
+Match Total:\t\t\t\t"""+ str(match_total)+""" 
+Percent of all arcs:\t\t\t"""+ match_percent +"""
+
+How many conll arcs do not match a link in only directionality?
+Directional Mismatch Total:\t\t"""+str(reverse_match_total)+"""
+Percent of all mismatches:\t\t"""+ directional_percent +"""
+
+How many conll arcs do not match a link. In either attachment or directionality.
+Mismatch Total:\t\t\t\t"""+str(mismatch_total)+"""\t
+Percent of all arcs:\t\t\t"""+ mismatch_percent +"""
+
+How many conll arcs in total.
+Final Total:\t\t\t\t""" +str(final_total)+"""
+
+
+------------------------------------------------------------
+MISMATCHES
+------------------------------------------------------------
+
[…]

### 2014-05-24 src/tools/experiments/trial.sh
modified trial.sh

@@ -31,7 +31,6 @@ if [ $LANGUAGE ]; then
     mv sol/conll_analysis/conll_analysis.txt sol/conll_analysis/conll_analysis_$ITERATIONS\_$LANGUAGE.txt
 else
     cp sol/allowedLinks.txt /tmp/LinksToDAG_trial_sol/allowedLinks_$ITERATIONS.txt
-    
     python src/tools/experiments/conll_analysis.py $SENTENCES $CONLL
     mv sol/conll_analysis/conll_analysis.txt sol/conll_analysis/conll_analysis_$ITERATIONS.txt
 fi

### 2014-05-24 src/tools/experiments/conll_analysis.py
updated conll_analysis.py to output latex files to import into the paper

@@ -6,6 +6,7 @@ from tikz_dependency import *
 from dependency_graph import *
 from conll_analysis_output import *
 
+# The directory containing the original conll data file
 CONLL_DIR = "data/"
 CONLL_LOC = CONLL_DIR+"english_bnews_train.conll"
 if len(sys.argv) > 2:
@@ -13,23 +14,40 @@ if len(sys.argv) > 2:
 if not os.path.exists(CONLL_DIR):
     os.makedirs(CONLL_DIR)
 
+# The directory containing our link-conll result data file
 SOL_DIR = "sol/"
 SOL_LOC = SOL_DIR+"links.conll"
 if not os.path.exists(SOL_DIR):
     os.makedirs(SOL_DIR)
 
+# Analysis directory and files. Where we will store these results.
 ANALYSIS_DIR = "sol/conll_analysis/"
 ANALYSIS_FILE = ANALYSIS_DIR+"conll_analysis.txt"
 if not os.path.exists(ANALYSIS_DIR):
     os.makedirs(ANALYSIS_DIR)
 
+# The trial number. How many sentences we take in originally.
+TRIAL_NUM = 0
+if len(sys.argv) > 3:
+    TRIAL_NUM = int(sys.argv[3])
+
+# Latex directory and files. Where we will store these results in latex form.
+LATEX_DIR = "doc/figure/"
+LATEX_FILE_SENTENCE = LATEX_DIR+"conll_analysis_sentences.tex"
+LATEX_FILE_SENTENCE_DROPPED = LATEX_DIR+"conll_analysis_sentences_dropped.tex"
+LATEX_FILE_SENTENCE_MULTIHEADED = LATEX_DIR+"conll_analysis_sentences_multiheaded.tex"
+if not os.path.exists(LATEX_DIR):
+    os.makedirs(LATEX_DIR)
+
+
+# Tikz directory and files
 TIKZ_DIR = "doc/figure/"
 TIKZ_LOC = TIKZ_DIR+"sentences.tikz"
 if not os.path.exists(TIKZ_DIR):
     os.makedirs(TIKZ_DIR)
 TIKZ = open(TIKZ_LOC, 'w+')
 
-ALL_PARSES_LOC = TIKZ_DIR+"allparses.tikz"
+ALL_PARSES_LOC = TIKZ_DIR+"parses.tikz"
 ALL_PARSES = open(ALL_PARSES_LOC, 'w+')
 
 
[…]

### 2014-05-24 src/tools/experiments/conll_analysis_output.py

@@ -16,7 +16,10 @@ def result_numbers(final_total,
                    word_count_total, 
                    dropped_sentence_count,
                    multiheaded_sentence_count, 
+                   original_sentence_count,
+                   conll_analysis_sentence_count,
                    total_sentence_count):
+
 
     match_percent = str(0)
     mismatch_percent = str(0)
@@ -104,9 +107,18 @@ Total number of sentences with dropped words:
 Dropped Sentences:\t\t\t"""+str(dropped_sentence_count)+"""
 Percent of Sentences:\t\t\t"""+str(float(dropped_sentence_count) / total_sentence_count)+"""
 
-Total Sentence Count:\t\t\t"""+str(total_sentence_count)+"""
 
+Total Sentence Count used in this analysis:
+\t\t\t\t\t"""+str(conll_analysis_sentence_count)+"""
+\t\t\t\t\t("""+str(float(conll_analysis_sentence_count)/original_sentence_count)+""") of original sentences
+
+Total Sentence Count given after skipping the disconnected sentences in the link parser:
+\t\t\t\t\t"""+str(total_sentence_count)+"""
+\t\t\t\t\t("""+str(float(total_sentence_count)/original_sentence_count)+""") of original sentences
+
+Original Sentence Count:\t\t"""+str(original_sentence_count)+"""
 """
+
 
 
 
@@ -203,3 +215,27 @@ Of the cases where the links had different directionality, what are the "extra" 
 #print mismatch_extra_total
 
   
+
+
+
+
+
+def result_latex_corpus(original_sentence_count,
+                        total_sentence_count, 
+                        conll_analysis_sentence_count):
+
+    result = ""
+    result += "\t\\begin{tabular}{|l|l|}\n"
+    result += "\t\t\\hline\n"
+    result += "\t\tOriginal number of sentences in conll corpus & "+str(original_sentence_count)+"\\\\ \n"
+    result += "\t\t\\hline\n"
+    result += "\t\tSentences after discarding disconnected parses & "+str(total_sentence_count)+"\\\\ \n"
+    result += "\t\t\\hline\n"
+    result += "\t\tSentences used for experiment and analysis & "+str(conll_analysis_sentence_count)+"\\\\ \n"
+    result += "\t\t\\hline\n"
+    result += "\t\\end{tabular}\n"
+
+    return result
+
+
+

### 2014-05-24 doc/LinksToDAG.tex
updated the paper to include a table. Added some citations.

@@
 In this paper we explore whether these relationships include dependencies.
+
 To determine the directional dependencies within the link edge labels we will use integer linear programming, encoding the problem in the Zimpl little language \cite{Koch2004}. 
-It turns out that the link parses roughly only match half of the conll dependency corpus.
-However this is because \TODO{}.
+It turns out that the link parses roughly only match half of the conll dependency corpus. However this is because \TODO{}.
+
+Kromann 2003 Danish dependency treebank\cite{kromann2003danish}
+
+McDonald and Pereira 2006\cite{McDonald2006}
+
+Sagae and Tsujii 2008\cite{Sagae2008}
+
+Gomez-Rodriguez and Nivre 2013\cite{gomez-rodriguez_divisible_2013}
 
@@
 We ran the link parser on the english bnews corpus sentences \footnote{We ignored the link parses that the link parser could not find suitable attachments and returned a disconnected graph.
-This happened for roughly NNN\% of the corpus.}
+This happened for roughly 19\% of the sentences in the corpus.}
 
@@
 
-For example, setting the variables $$l_{LEFT} = TRUE, l_{RIGHT} = FALSE$$ would mean that the directional dependency of all links with label $l$ would be set to $LEFT$.
+For example, setting the variables $$l_{\textsc{LEFT}} = \textsc{TRUE}, l_{\textsc{RIGHT}} = \textsc{FALSE}$$ would mean that the directional dependency of all links with label $l$ would be set to $\textsc{LEFT}$.
 If both variables were set to true, then the link type is allowed to go in either direction.
@@
 
-\subsection{Slack Hierarchy}
+\subsection{Slack}
 
@@
 
-\TODO{} We also introduced slack on the link types such that link types with the same coarse grained label would try and align the same way as the majority in the group, where the preceding capital letters of the link type denote the coarse grained label, while the subscript letters denote further information.
-This slack places a prior on rare or never-before seen link types to be assigned in the same way as other similar link types.
-
-This slack hierarchy gave the ILP the flexibility to handle noise and novel link types while still trying to learn the overall link grammar.
+\begin{figure*}
+%\centering
 
 \begin{equation}
-  count_{L, \textsc{left}} = \sum_{a \in \textsc{ARCS}_L} left_a
+  count_{L, \textsc{left}} = \sum_{a \in \textsc{ARCS}_L} link(a,L)
 \end{equation}
@@
 \end{equation}
-Where $L \in \textsc{LABELS}$, and $\textsc{M}$ is the size of the corpus.
-The objective is:
+%Where $L \in \textsc{LABELS}$, and $\textsc{M}$ is the size of the corpus.
+OBJECTIVE:
 \begin{equation}
@@
 
-\begin{algorithm}
-\caption{ILP Encoding}\label{ILP}
-\begin{algorithmic}[1]
-%\Procedure{MyProcedure}{}
+\caption{The ILP formulation. L is a label and M is the size of the corpus.}
 
-\State $\text{min} \sum_{l \in labels} allowed_L(l) + allowed_R(l)$
-\State $\text{subject to}$
-\State $\text{Acyclicity}$
-\State $\text{All nodes reachable from the root.}$
-
-%\EndProcedure
-\end{algorithmic}
-\end{algorithm}
+\end{figure*}
 
@@
 \begin{figure}[ht!]
+  \small
   \includegraphics[width=\linewidth, keepaspectratio=true]{figure/precision_recall.png}
-  \caption{Precision and Recall of different runs on English}
+  \caption{\small Precision and recall of increasing runs of English sentences}
 \end{figure}
@@
 
-\section{Link Corpus}
+\section{Link Corpus Experiment and Results}
+
+\begin{figure}[ht!]
+  \centering
+  \small
+  \input{figure/conll_analysis_sentences.tex}
+  \caption{\small Of the sentences used for this experiment section about \input{figure/conll_analysis_sentences_dropped}had at least one dropped word, and about \input{figure/conll_analysis_sentences_multiheaded}had at least one multiheaded word}
+\end{figure}
 
@@
 
+\section{Future Work}
+
+\subsection{Slack Hierarchy}
+We would like to introduce slack on the link grammar types such that link types with the same coarse grained label would try and align the same way as the majority in the group, where the preceding capital letters of the link type denote the coarse grained label, while the subscript letters denote further information.
+This slack would place a prior on rare or never-before seen link types to be assigned in the same way as other similar link types.
+We think that this will give the ILP better flexibility to handle noise and novel link types while still trying to learn the overall link grammar.
+
 \section{Related Work}
@@
 
+\clearpage
 \appendix

### 2014-05-24 src/tools/experiments/trial.sh
modified trial.sh to take in the original number of sentences to compare with other numbers.

@@ -27,11 +27,11 @@ tail --lines 1 runtime >> /tmp/LinksToDAG_times
 if [ $LANGUAGE ]; then
 
     cp sol/allowedLinks.txt /tmp/LinksToDAG_trial_sol/allowedLinks_$ITERATIONS\_$LANGUAGE.txt
-    python src/tools/experiments/conll_analysis.py $SENTENCES $CONLL $LANGUAGE
+    python src/tools/experiments/conll_analysis.py $SENTENCES $CONLL $ITERATIONS $LANGUAGE
     mv sol/conll_analysis/conll_analysis.txt sol/conll_analysis/conll_analysis_$ITERATIONS\_$LANGUAGE.txt
 else
     cp sol/allowedLinks.txt /tmp/LinksToDAG_trial_sol/allowedLinks_$ITERATIONS.txt
-    python src/tools/experiments/conll_analysis.py $SENTENCES $CONLL
+    python src/tools/experiments/conll_analysis.py $SENTENCES $CONLL $ITERATIONS
     mv sol/conll_analysis/conll_analysis.txt sol/conll_analysis/conll_analysis_$ITERATIONS.txt
 fi
 

### 2014-05-26 doc/LinksToDAG.tex
expanded the introduction section.

@@
 \begin{abstract}
-
 % Copied from the google doc
-There has been recent theoretical work on multi-headed versions of dependency grammar (cite).
+There has been recent theoretical work on multi-headed versions of dependency grammar(cite).
 Such formalisms can be more syntactically and semantically expressive.
-However, there aren't any corpora at present (check this) except for Buch-Kromann's work on Danish (cite), so it is hard to investigate the benefit of such parsers or to work on making them faster or more accurate.
+However, there aren't any corpora at present (check this) except for Buch-Kromann's work on Danish \cite{kromann2003danish}, so it is hard to investigate the benefit of such parsers or to work on making them faster or more accurate.
 To fill this gap, we observe that link grammar produces parses that are similar to multi-headed planar dependency parses except that the links are undirected.
@@
 We also report ...
-
 \end{abstract}
@@
 
-Dependency parsing is the task of mapping a sentence to a projective (not always projective?) directed acyclic tree.
+\TODO{Intro}
+
+Projective dependency parsing is the task of mapping a sentence to a directed acyclic tree.
 Link parsing in contrast produces a multiheaded planar graph with undirected edges, where every edge has a label describing the relationship between two constituents in a parse.
-In this paper we explore whether these relationships include dependencies.
+In this paper we explore whether these relationships include dependencies, and whether the multiheadedness of the link grammar offers additional dependency relationships not found in other corpora.
 
-To determine the directional dependencies within the link edge labels we will use integer linear programming, encoding the problem in the Zimpl little language \cite{Koch2004}. 
-It turns out that the link parses roughly only match half of the conll dependency corpus. However this is because \TODO{}.
+We will accomplish this by generating link parses using the AbiWord/CMU link grammar parser, learning a direction to every encountered link label type, and then using these assignments for all link tokens in the link parse corpus.
+We formulate an integer linear programming problem that will find the smallest set of assignments that will still produce dependency parses that are acyclic with all nodes reachable from the root.
 
-Kromann 2003 Danish dependency treebank\cite{kromann2003danish}
+We publish our dictionary of label-to-direction assignments, that anyone can use to turn a link parse into a dependency parse.
 
-McDonald and Pereira 2006\cite{McDonald2006}
+Using the sentences of a dependency corpus as comparison, we find that the link parses do not wholly subsume dependency parses, and that the undirected links match roughly two thirds of the arcs in the conll dependency corpus.
+\TODO{This is because...}
+Of the links that match a corresponding dependency arc we measure whether the direction we chose is the same as the original dependency data. 
+Finally, we look to see that given a link label, whether the original conll label can be recovered.
 
-Sagae and Tsujii 2008\cite{Sagae2008}
+%about 20\%
 
-Gomez-Rodriguez and Nivre 2013\cite{gomez-rodriguez_divisible_2013}
-
-%Despite this, the link parses offer additional links in addition to dependencies...
+%Kromann 2003 Danish dependency treebank\cite{kromann2003danish}
+%McDonald and Pereira 2006\cite{McDonald2006}
+%Sagae and Tsujii 2008\cite{Sagae2008}
+%Gomez-Rodriguez and Nivre 2013\cite{gomez-rodriguez_divisible_2013}
 
@@
 
-We formulated the ILP to assign directional dependencies to a link grammar parse such that it is a connected DAG, with all of the nodes reachable from the root.
+We formulated an ILP to assign directional dependencies to all of the links in a set of link grammar parses, such that they are all connected DAGs, with the nodes reachable from the root.
 
@@
 Our ILP objective is to minimize the number of label\_direction variables set to true, while satisfying the other constraints.
+
+The encoding was written in the Zimpl little language \cite{Koch2004} and solved using the SCIP Optimization Suite (cite?).
 

### 2014-05-26 doc/LinksToDAG.tex
expanded the intro to the paper. filled out some more sections. Worked out the ILP formulation section in greater detail, and made it prettier.

@@
 \section{Introduction}
-
 \TODO{Intro}
-
 Projective dependency parsing is the task of mapping a sentence to a directed acyclic tree.
@@
 
-We publish our dictionary of label-to-direction assignments, that anyone can use to turn a link parse into a dependency parse.
+From the ILP, we produce a dictionary of label-to-direction assignments. We publish this so that anyone can easily turn a link parse into a dependency parse.
 
@@
 \section{Link Grammars}
+Link grammars are a grammatical system equivalent to context-free grammars that assign linking requirements to every given word\cite{SleatorTemperly91}.
+A link parser then tries to satisfy all of these requirements for every word of a sentence while still maintaining planarity.
+The resulting links describe the relationships between constituents in a parse.
 
-Link grammars describe the links, or relationships between constituents in a parse \cite{SleatorTemperly91}.
-
-We ran the link parser on the english bnews corpus sentences \footnote{We ignored the link parses that the link parser could not find suitable attachments and returned a disconnected graph.
-This happened for roughly 19\% of the sentences in the corpus.}
-
-The link grammar on which the link-parser makes its linking decisions based on a set of handwritten dictionaries.
-Instead of going through these dictionaries, we learned the link grammar using an ILP.
-This approach allows us to analyze any link grammar dictionary other than English.
+The link grammar is based on a set of handwritten dictionaries.
+Instead of going through these dictionaries, we learned the grammar using an ILP.
+This approach also potentially allows us to analyze link grammar dictionaries other than English.
+\TODO{explore other link grammar dictionaries}
 
 \section{Integer Linear Programming}
+We formulated an ILP to assign directionality to all of the links in a set of link grammar parses, while still mandating that the resulting dependency parses are connected DAGs, with the nodes reachable from the root.
 
-We formulated an ILP to assign directional dependencies to all of the links in a set of link grammar parses, such that they are all connected DAGs, with the nodes reachable from the root.
+In order to choose which direction the link labels are assigned, for every encountered label we generate two boolean label\_direction variables correspending to left and right.
+Setting the only one of these variables to \textsc{TRUE} chooses a consistent direction for a label throughout all parses, and because we assign a directionality to every link type at least one must be set to true.
+Setting both variables would allow us to go in either direction.
 
-The ILP chooses whether to set a link type to be assigned left, right, or both.
-For this, two label\_direction variables were made for each link type that appears in the parses, corresponding to whether all of the $l$ link edges were to be set to go left, right, or both.
-Because we assign a directionality to every link type, at least one of these variables must be set to true.
-
-For example, setting the variables $$l_{\textsc{LEFT}} = \textsc{TRUE}, l_{\textsc{RIGHT}} = \textsc{FALSE}$$ would mean that the directional dependency of all links with label $l$ would be set to $\textsc{LEFT}$.
-If both variables were set to true, then the link type is allowed to go in either direction.
-Our ILP objective is to minimize the number of label\_direction variables set to true, while satisfying the other constraints.
-
+Our ILP objective is to minimize the number of label\_direction variables set to true, while still satisfying the acyclicity and reachability constraints.
 The encoding was written in the Zimpl little language \cite{Koch2004} and solved using the SCIP Optimization Suite (cite?).
@@
 \subsection{Slack}
-
-We introduced slack on the directional dependency variable tokens such that link types were allowed to deviate from the majority up to (1\% ?) of the time before the ILP would assign both directional dependency variables to true.
-This addressed noise in the link parser's label assignments, while still allowing for the possibility that both directions could still be assigned.
-
-\begin{figure*}
+We introduced slack on the directional dependency variables such that the link types were allowed to deviate from the majority one percent of the time before the ILP would assign both directional dependency variables to true.
+This addressed noise in the link parser's label assignments, while still allowing for both directions to be assigned if needed.
+\begin{figure}
 %\centering
-
+  \small
+  \begin{align*}
+    &\textsc{dir} = \{\textsc{left}, \textsc{right}\}, \textsc{l} \in \textsc{labels} & \\
+    &\textsc{m} = \left|{\textsc{links}}\right|
+  \end{align*}
   \begin{equation}
-  count_{L, \textsc{left}} = \sum_{a \in \textsc{ARCS}_L} link(a,L)
+    \textsc{count}_{\textsc{l}, \textsc{dir}} = \sum \textsc{link}_{\textsc{l},\textsc{dir}}
   \end{equation}
   \begin{equation}
-  count_{L, \textsc{left}} = \textsc{M} \cdot \textsc{allowed}_{L,\textsc{left}} + \textsc{slack}_{L}
-\end{equation}
-%Where $L \in \textsc{LABELS}$, and $\textsc{M}$ is the size of the corpus.
-OBJECTIVE:
-\begin{equation}
-  \min \sum_{L \in \textsc{LABELS} \atop \textsc{DIR} \in \{left, right\}} (\textsc{allowed}_{L,\textsc{DIR}}) \atop + \textsc{COST}(L) * \sum_L \textsc{slack}_L
+    \textsc{count}_{\textsc{l}, \textsc{dir}} = \textsc{m} \cdot \textsc{allowed}_{\textsc{l},\textsc{dir}} + \textsc{slack}_{\textsc{l}} 
   \end{equation}
   \begin{equation}
-  \textsc{COST}(L) = \frac{100}{(\textsc{count}_{L,\textsc{left}} + \textsc{count}_{L,\textsc{right}})}
+    \textsc{cost}_{\textsc{l}} = \frac{100}{\textsc{count}_{\textsc{l},\textsc{left}} + \textsc{count}_{\textsc{l},\textsc{right}}}
   \end{equation}
+  \begin{equation}
+    \text{minimize} \sum_{(\textsc{l}, \textsc{dir}} \textsc{allowed}_{\textsc{l},\textsc{dir}}) + \textsc{cost}_{\textsc{l}} \cdot \sum_\textsc{l} \textsc{slack}_{\textsc{l}} 
+  \end{equation}
+  \begin{align*}
+    \text{subject to } &\forall_{i \in \textsc{sentence}}: \textsc{reachable}_{\textsc{root}, i} = \textsc{true} &\\
+                       &\forall_{\substack{i,j \in \textsc{sentence} \\ i\neq j }}: \textsc{cycle}_{i,j} = \textsc{false} &
+  \end{align*}
 
-\caption{The ILP formulation. L is a label and M is the size of the corpus.}
-
-\end{figure*}
+  \caption{\small The ILP formulation.}
+  %\\L is a label\\M is the size of the corpus.}
+\end{figure}
 
@@
 
-\section{Other languages}
+%\section{Other languages}
 
 \section{Link Corpus Experiment and Results}
+We ran the link parser on the english bnews corpus sentences.
+\footnote{We ignored the link parses that the link parser could not find suitable attachments and returned a disconnected graph.
+This happened for roughly 19\% of the sentences in the corpus.}
+From these parses we ran our ILP and generated directionalized multiheaded dependency data to compare with the original conll data.
 

### 2014-05-26 doc/LinksToDAG.tex
moved the ILP formulation equations around some more. Added a citation.

@@
 
-In order to choose which direction the link labels are assigned, for every encountered label we generate two boolean label\_direction variables correspending to left and right.
-Setting the only one of these variables to \textsc{TRUE} chooses a consistent direction for a label throughout all parses, and because we assign a directionality to every link type at least one must be set to true.
+In order to choose which direction the link labels are assigned, for every encountered label we generate two boolean label/direction variables correspending to left and right.
+Setting one of these variables to \textsc{true} allows only the specified direction for a label throughout all parses, and because we assign a directionality to every link type at least one of these variables must be set to true.
 Setting both variables would allow us to go in either direction.
 
-Our ILP objective is to minimize the number of label\_direction variables set to true, while still satisfying the acyclicity and reachability constraints.
-The encoding was written in the Zimpl little language \cite{Koch2004} and solved using the SCIP Optimization Suite (cite?).
+Our ILP objective is to minimize the number of label/direction variables set to \textsc{true}, while still satisfying the acyclicity and reachability constraints.
+The encoding was written in the Zimpl little language \cite{Koch2004} and solved using the SCIP Optimization Suite\cite{achterberg2009scip}.
 
@@
 \begin{figure}
-%\centering
   \small
-  \begin{align*}
-    &\textsc{dir} = \{\textsc{left}, \textsc{right}\}, \textsc{l} \in \textsc{labels} & \\
-    &\textsc{m} = \left|{\textsc{links}}\right|
-  \end{align*}
-  \begin{equation}
-    \textsc{count}_{\textsc{l}, \textsc{dir}} = \sum \textsc{link}_{\textsc{l},\textsc{dir}}
-  \end{equation}
-  \begin{equation}
-    \textsc{count}_{\textsc{l}, \textsc{dir}} = \textsc{m} \cdot \textsc{allowed}_{\textsc{l},\textsc{dir}} + \textsc{slack}_{\textsc{l}} 
-  \end{equation}
-  \begin{equation}
-    \textsc{cost}_{\textsc{l}} = \frac{100}{\textsc{count}_{\textsc{l},\textsc{left}} + \textsc{count}_{\textsc{l},\textsc{right}}}
-  \end{equation}
-  \begin{equation}
-    \text{minimize} \sum_{(\textsc{l}, \textsc{dir}} \textsc{allowed}_{\textsc{l},\textsc{dir}}) + \textsc{cost}_{\textsc{l}} \cdot \sum_\textsc{l} \textsc{slack}_{\textsc{l}} 
-  \end{equation}
-  \begin{align*}
+  \begin{align}
+    \textsc{dir} &= \{\textsc{left}, \textsc{right}\} &\\ 
+    \textsc{l} &\in \textsc{labels} & \\
+    \textsc{m} &= \left|{\textsc{links}}\right| &\\
+    \textsc{count}_{\textsc{l}, \textsc{dir}} &= \sum \textsc{link}_{\textsc{l},\textsc{dir}} &\\
+    \textsc{count}_{\textsc{l}, \textsc{dir}} &= \textsc{m} \cdot \textsc{allowed}_{\textsc{l},\textsc{dir}} + \textsc{slack}_{\textsc{l}} &\\
+    \textsc{cost}_{\textsc{l}} &= \frac{100}{\textsc{count}_{\textsc{l},\textsc{left}} + \textsc{count}_{\textsc{l},\textsc{right}}} &\\
+    \text{minimize} &\sum_{\textsc{l}, \textsc{dir}} (\textsc{allowed}_{\textsc{l},\textsc{dir}}) + \textsc{cost}_{\textsc{l}} \cdot \sum_\textsc{l} \textsc{slack}_{\textsc{l}} &\\
     \text{subject to } &\forall_{i \in \textsc{sentence}}: \textsc{reachable}_{\textsc{root}, i} = \textsc{true} &\\
-                       &\forall_{\substack{i,j \in \textsc{sentence} \\ i\neq j }}: \textsc{cycle}_{i,j} = \textsc{false} &
-  \end{align*}
-
+                       &\forall_{\substack{i,j \in \textsc{sentence} \\ i\neq j }}: \textsc{has\_cycle}_{i,j} = \textsc{false} &
+  \end{align}
   \caption{\small The ILP formulation.}
-  %\\L is a label\\M is the size of the corpus.}
 \end{figure}

### 2014-05-26 doc/LinksToDAG.tex
updated the precision_recall curve.

@@
 
-\subsection{Slack}
-We introduced slack on the directional dependency variables such that the link types were allowed to deviate from the majority one percent of the time before the ILP would assign both directional dependency variables to true.
-This addressed noise in the link parser's label assignments, while still allowing for both directions to be assigned if needed.
 \begin{figure}
@@
 
+\subsection{Slack}
+We introduced slack on the directional dependency variables such that the link types were allowed to deviate from the majority one percent of the time before the ILP would assign both directional dependency variables to true.
+This addressed noise in the link parser's label assignments, while still allowing for both directions to be assigned if needed.
+
 \subsection{Stability of Results}
-We tested to see how the solution changed over increasing numbers of processed sentences. 
+We measured how the solution changed over increasing numbers of processed sentences. 
 Taking the solution to the largest run as the answer key, we measured how much the previous runs deviated from it.
-We measured the precision of whether the assignments in the smaller runs matched the assignments of the largest run, and we measured the recall of whether the assignments in the largest run could be found in the smaller runs.
-
+We measured the precision of whether the assignments in the smaller runs matched the assignments of the largest run and we measured the recall of whether the assignments in the largest run could be found in the smaller runs.
 \begin{figure}[ht!]
@@
   \includegraphics[width=\linewidth, keepaspectratio=true]{figure/precision_recall.png}
-  \caption{\small Precision and recall of increasing runs of English sentences}
+  \caption{\small Precision/recall of results of English sentences}
 \end{figure}
-
-In the end, the precision values converged to 100\% while the recall values continued to grow throughout our data.
+We found that the precision values converged to 100\% while the recall values continued to grow throughout our data without converging.
 This indicates that the solutions of the previous runs were consistent with the largest run.

### 2014-05-27 doc/LinksToDAG.tex
started editing paper; added refs to bibfile

--- 2014-05-26 doc/LinksToDAG.tex
@@
 
-\newcommand{\Note}[1]{}
-\renewcommand{\Note}[1]{\hl{[#1]}}  % comment out this definition to suppress all Notes                                                                                
-\newcommand{\TODO}[1]{\Note{TODO: #1}}
-\newcommand{\NoteSigned}[3]{{\sethlcolor{#2}\Note{#1: #3}}}
-\newcommand{\NoteJE}[1]{\NoteSigned{JE}{LightBlue}{#1}}
-\newcommand{\NoteJH}[1]{\NoteSigned{JH}{YellowGreen}{#1}}
+\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}  % allows better color names
+\usepackage{todonotes}   % insert [disable] to disable all notes
+% Note that these macros accept optional arguments such as 
+% [size=\small,bordercolor=red].
+\newcommand{\Note}[4][]{\todo[author=#2,color=#3,fancyline,#1]{#4}}
+\newcommand{\noteJH}[2][]{\Note[#1]{MRG}{blue!40}{#2}}
+\newcommand{\noteJE}[2][]{\Note[#1]{JE}{green!40}{#2}}   
+\newcommand{\notewho}[3][]{\Note[#1]{#2}{orange!40}{#3}}  % extra arg with miscellaneous author
+\newcommand{\NoteJH}[2][]{\noteJH[inline,#1]{#2}}
+\newcommand{\NoteJE}[2][]{\noteJE[inline,#1]{#2}}
+\newcommand{\Notewho}[3][]{\notewho[inline,#1]{#2}{#3}}  % extra arg with miscellaneous author
 
-\title{Deriving Multi-headed Planar Dependency Parses from Link Grammar Parses}
+\title{Deriving Multi-Headed Projective Dependency Parses \\ from Link Grammar Parses}
+% Oriented link grammar: Creating a multi-headed dependency corpus.
 
-\author{Juneki Hong and Jason Eisner\\
-  Department of Computer Science \\
-  Johns Hopkins University \\
-  Baltimore, MD 21218, USA \\ 
-  {\tt \{juneki,jason\}@cs.jhu.edu} \\
-}
+% \author{Juneki Hong and Jason Eisner\\
+%   Department of Computer Science \\
+%   Johns Hopkins University \\
+%   Baltimore, MD 21218, USA \\ 
+%   {\tt \{juneki,jason\}@cs.jhu.edu} \\
+% }
 
@@
 \begin{abstract}
+
 % Copied from the google doc
-There has been recent theoretical work on multi-headed versions of dependency grammar(cite).
+Under multi-headed dependency grammar, a parse is a DAG rather than a tree.
 Such formalisms can be more syntactically and semantically expressive.
-However, there aren't any corpora at present (check this) except for Buch-Kromann's work on Danish \cite{kromann2003danish}, so it is hard to investigate the benefit of such parsers or to work on making them faster or more accurate.
-To fill this gap, we observe that link grammar produces parses that are similar to multi-headed planar dependency parses except that the links are undirected.
-We use Integer Linear Programming to assign consistent directions to the links in a corpus of NNN parses produced by the Link Grammar Parser, which has broad-coverage hand-written grammars of English, Russian, and more.
-(what statistics does it use?).
-We find that such directions can indeed be consistently assigned in a way that yields valid multi-headed dependency parses.
-The resulting parses in English appear linguistically plausible, although they are not in general consistent with CoNLL-style parses of the same sentences; we discuss the differences.
-We also report ...
+However, there are few multi-headed corpora at present, particularly for the projective case, so it is hard to investigate the benefit of such parsers or to work on making them faster or more accurate.
+To fill this gap, we observe that link grammar produces {\em undirected} projective graphs.
+% OLD
+% ... produces parses that are similar to multi-headed projective dependency parses, except that the links are undirected. 
+We use Integer Linear Programming to assign consistent directions to the labeled links in a corpus of !!!
+\noteJE{fill in the number} parses produced by the Link Grammar Parser, which has broad-coverage hand-written grammars of English, Russian, and more.
+\noteJE{what statistics does it use?} We find that such directions can indeed be consistently assigned in a way that yields valid multi-headed dependency parses.
+The resulting parses in English appear reasonably linguistically plausible, although they are not in general consistent with CoNLL-style parses of the same sentences; we discuss the differences.
+We also report on the convergence of our ILP method to a stable annotation scheme as the corpus size increases.
+
 \end{abstract}
@@
 \section{Introduction}
-\TODO{Intro}
-Projective dependency parsing is the task of mapping a sentence to a directed acyclic tree.
+
+Dependency parsing is the task of mapping a sentence to a projective (not always projective?) directed acyclic tree.
 Link parsing in contrast produces a multiheaded planar graph with undirected edges, where every edge has a label describing the relationship between two constituents in a parse.
-In this paper we explore whether these relationships include dependencies, and whether the multiheadedness of the link grammar offers additional dependency relationships not found in other corpora.
+In this paper we explore whether these relationships include dependencies.
+To determine the directional dependencies within the link edge labels we will use integer linear programming, encoding the problem in the Zimpl little language \cite{Koch2004}.
+It turns out that the link parses roughly only match half of the conll dependency corpus.
+However this is because \todo{}.
 
-We will accomplish this by generating link parses using the AbiWord/CMU link grammar parser, learning a direction to every encountered link label type, and then using these assignments for all link tokens in the link parse corpus.
-We formulate an integer linear programming problem that will find the smallest set of assignments that will still produce dependency parses that are acyclic with all nodes reachable from the root.
+%Despite this, the link parses offer additional links in addition to dependencies...
 
-From the ILP, we produce a dictionary of label-to-direction assignments. We publish this so that anyone can easily turn a link parse into a dependency parse.
+Some recent work has considered multi-headed ...
 
-Using the sentences of a dependency corpus as comparison, we find that the link parses do not wholly subsume dependency parses, and that the undirected links match roughly two thirds of the arcs in the conll dependency corpus.
-\TODO{This is because...}
-Of the links that match a corresponding dependency arc we measure whether the direction we chose is the same as the original dependency data. 
-Finally, we look to see that given a link label, whether the original conll label can be recovered.
+ by expressing the dependency of a word on multiple parents.
 
-%about 20\%
+\cite{mcdonald-pereira-2006,buchkromann-2006,sagae-tsujii-2008,gomezrodriguez-nivre-2013}
 
-%Kromann 2003 Danish dependency treebank\cite{kromann2003danish}
-%McDonald and Pereira 2006\cite{McDonald2006}
-%Sagae and Tsujii 2008\cite{Sagae2008}
-%Gomez-Rodriguez and Nivre 2013\cite{gomez-rodriguez_divisible_2013}
+ (check this) except for Buch-Kromann's work on Danish \cite{kromann-2003}
 
@@
 \section{Link Grammars}
-Link grammars are a grammatical system equivalent to context-free grammars that assign linking requirements to every given word\cite{SleatorTemperly91}.
-A link parser then tries to satisfy all of these requirements for every word of a sentence while still maintaining planarity.
-The resulting links describe the relationships between constituents in a parse.
 
-The link grammar is based on a set of handwritten dictionaries.
-Instead of going through these dictionaries, we learned the grammar using an ILP.
-This approach also potentially allows us to analyze link grammar dictionaries other than English.
-\TODO{explore other link grammar dictionaries}
+Link grammars describe the links, or relationships between constituents in a parse \cite{SleatorTemperly91}.
+
+We ran the link parser on the english bnews corpus sentences \footnote{We ignored the link parses that the link parser could not find suitable attachments and returned a disconnected graph.
+This happened for roughly NNN\% of the corpus.}
+
+The link grammar on which the link-parser makes its linking decisions based on a set of handwritten dictionaries.
+Instead of going through these dictionaries, we learned the link grammar using an ILP.
+This approach allows us to analyze any link grammar dictionary other than English.
 
 \section{Integer Linear Programming}
-We formulated an ILP to assign directionality to all of the links in a set of link grammar parses, while still mandating that the resulting dependency parses are connected DAGs, with the nodes reachable from the root.
 
-In order to choose which direction the link labels are assigned, for every encountered label we generate two boolean label\_direction variables correspending to left and right.
-Setting the only one of these variables to \textsc{TRUE} chooses a consistent direction for a label throughout all parses, and because we assign a directionality to every link type at least one must be set to true.
-Setting both variables would allow us to go in either direction.
+We formulated the ILP to assign directional dependencies to a link grammar parse such that it is a connected DAG, with all of the nodes reachable from the root.
 
-Our ILP objective is to minimize the number of label\_direction variables set to true, while still satisfying the acyclicity and reachability constraints.
-The encoding was written in the Zimpl little language \cite{Koch2004} and solved using the SCIP Optimization Suite (cite?).
+The ILP chooses whether to set a link type to be assigned left, right, or both.
+For this, two label\_direction variables were made for each link type that appears in the parses, corresponding to whether all of the $l$ link edges were to be set to go left, right, or both.
+Because we assign a directionality to every link type, at least one of these variables must be set to true.
 
-\TODO{cite model minimization papers}
+For example, setting the variables $$l_{LEFT} = TRUE, l_{RIGHT} = FALSE$$ would mean that the directional dependency of all links with label $l$ would be set to $LEFT$.
+If both variables were set to true, then the link type is allowed to go in either direction.
+Our ILP objective is to minimize the number of label\_direction variables set to true, while satisfying the other constraints.
 
-\subsection{Slack}
-We introduced slack on the directional dependency variables such that the link types were allowed to deviate from the majority one percent of the time before the ILP would assign both directional dependency variables to true.
-This addressed noise in the link parser's label assignments, while still allowing for both directions to be assigned if needed.
-\begin{figure}
-%\centering
-  \small
-  \begin{align*}
-    &\textsc{dir} = \{\textsc{left}, \textsc{right}\}, \textsc{l} \in \textsc{labels} & \\
-    &\textsc{m} = \left|{\textsc{links}}\right|
-  \end{align*}
+\todo{cite model minimization papers}
+
+\subsection{Slack Hierarchy}
+
+We introduced slack on the directional dependency variable tokens such that link types were allowed to deviate from the majority up to (1\% ?) of the time before the ILP would assign both directional dependency variables to true.
+This addressed noise in the link parser's label assignments, while still allowing for the possibility that both directions could still be assigned.
+
+\todo{} We also introduced slack on the link types such that link types with the same coarse grained label would try and align the same way as the majority in the group, where the preceding capital letters of the link type denote the coarse grained label, while the subscript letters denote further information.
+This slack places a prior on rare or never-before seen link types to be assigned in the same way as other similar link types.
+
+This slack hierarchy gave the ILP the flexibility to handle noise and novel link types while still trying to learn the overall link grammar.
+
 \begin{equation}
-    \textsc{count}_{\textsc{l}, \textsc{dir}} = \sum \textsc{link}_{\textsc{l},\textsc{dir}}
+  count_{L, \textsc{left}} = \sum_{a \in \textsc{ARCS}_L} left_a
 \end{equation}
 \begin{equation}
-    \textsc{count}_{\textsc{l}, \textsc{dir}} = \textsc{m} \cdot \textsc{allowed}_{\textsc{l},\textsc{dir}} + \textsc{slack}_{\textsc{l}} 
+  count_{L, \textsc{left}} = \textsc{M} \cdot \textsc{allowed}_{L,\textsc{left}} + \textsc{slack}_{L}
+\end{equation}
+Where $L \in \textsc{LABELS}$, and $\textsc{M}$ is the size of the corpus.
+The objective is:
+\begin{equation}
+  \min \sum_{L \in \textsc{LABELS} \atop \textsc{DIR} \in \{left, right\}} (\textsc{allowed}_{L,\textsc{DIR}}) \atop + \textsc{COST}(L) * \sum_L \textsc{slack}_L
 \end{equation}
 \begin{equation}
-    \textsc{cost}_{\textsc{l}} = \frac{100}{\textsc{count}_{\textsc{l},\textsc{left}} + \textsc{count}_{\textsc{l},\textsc{right}}}
+  \textsc{COST}(L) = \frac{100}{(\textsc{count}_{L,\textsc{left}} + \textsc{count}_{L,\textsc{right}})}
 \end{equation}
-  \begin{equation}
-    \text{minimize} \sum_{(\textsc{l}, \textsc{dir}} \textsc{allowed}_{\textsc{l},\textsc{dir}}) + \textsc{cost}_{\textsc{l}} \cdot \sum_\textsc{l} \textsc{slack}_{\textsc{l}} 
-  \end{equation}
-  \begin{align*}
-    \text{subject to } &\forall_{i \in \textsc{sentence}}: \textsc{reachable}_{\textsc{root}, i} = \textsc{true} &\\
-                       &\forall_{\substack{i,j \in \textsc{sentence} \\ i\neq j }}: \textsc{cycle}_{i,j} = \textsc{false} &
-  \end{align*}
 
-  \caption{\small The ILP formulation.}
-  %\\L is a label\\M is the size of the corpus.}
-\end{figure}
+\begin{algorithm}
+\caption{ILP Encoding}\label{ILP}
+\begin{algorithmic}[1]
+%\Procedure{MyProcedure}{}
+
+\State $\text{min} \sum_{l \in labels} allowed_L(l) + allowed_R(l)$
+\State $\text{subject to}$
+\State $\text{Acyclicity}$
+\State $\text{All nodes reachable from the root.}$
+
+%\EndProcedure
+\end{algorithmic}
+\end{algorithm}
 
@@
 \begin{figure}[ht!]
-  \small
   \includegraphics[width=\linewidth, keepaspectratio=true]{figure/precision_recall.png}
-  \caption{\small Precision and recall of increasing runs of English sentences}
+  \caption{Precision and Recall of different runs on English}
 \end{figure}
@@
 
-%\section{Other languages}
+\section{Other languages}
 
-\section{Link Corpus Experiment and Results}
-We ran the link parser on the english bnews corpus sentences.
-\footnote{We ignored the link parses that the link parser could not find suitable attachments and returned a disconnected graph.
-This happened for roughly 19\% of the sentences in the corpus.}
-From these parses we ran our ILP and generated directionalized multiheaded dependency data to compare with the original conll data.
-
-\begin{figure}[ht!]
-  \centering
-  \small
-  \input{figure/conll_analysis_sentences.tex}
-  \caption{\small Of the sentences used for this experiment section about \input{figure/conll_analysis_sentences_dropped}had at least one dropped word, and about \input{figure/conll_analysis_sentences_multiheaded}had at least one multiheaded word}
-\end{figure}
+\section{Link Corpus}
 
@@
 
-\TODO{more in the appendix ...}
+\todo{more in the appendix ...}
 
@@
 However, that requires per-language
-
-\section{Future Work}
-
-\subsection{Slack Hierarchy}
-We would like to introduce slack on the link grammar types such that link types with the same coarse grained label would try and align the same way as the majority in the group, where the preceding capital letters of the link type denote the coarse grained label, while the subscript letters denote further information.
-This slack would place a prior on rare or never-before seen link types to be assigned in the same way as other similar link types.
-We think that this will give the ILP better flexibility to handle noise and novel link types while still trying to learn the overall link grammar.
 
@@
 
-\clearpage
 \appendix
@@
 
-\TODO{A table here that lists each link grammar type together with its
+\todo{A table here that lists each link grammar type together with its
   count, its allowed direction(s), its percentage of majority
@@
 
-\TODO{Show a bunch of English sentences with CoNLL vs.\@ link parses.
+\todo{Show a bunch of English sentences with CoNLL vs.\@ link parses.
   Explain how they were selected and how the results are colorized etc.}

### 2014-05-27 doc/LinksToDAG.tex
added more to the paper.

--- 2014-05-26 doc/LinksToDAG.tex
@@
     \textsc{count}_{\textsc{l}, \textsc{dir}} &= \textsc{m} \cdot \textsc{allowed}_{\textsc{l},\textsc{dir}} + \textsc{slack}_{\textsc{l}} &\\
-    \textsc{cost}_{\textsc{l}} &= \frac{100}{\textsc{count}_{\textsc{l},\textsc{left}} + \textsc{count}_{\textsc{l},\textsc{right}}} &\\
-    \text{minimize} &\sum_{\textsc{l}, \textsc{dir}} (\textsc{allowed}_{\textsc{l},\textsc{dir}}) + \textsc{cost}_{\textsc{l}} \cdot \sum_\textsc{l} \textsc{slack}_{\textsc{l}} &\\
+    \textsc{cost}_{\textsc{l}} &= \frac{100}{\textsc{count}_{\textsc{l},\textsc{left}} + \textsc{count}_{\textsc{l},\textsc{right}}} &
+  \end{align}
+  \begin{equation} % I seperated out this line from the rest because it would disrupt the line number on the right side of the figure.
+    \text{minimize:} \sum_{\textsc{l}, \textsc{dir}} (\textsc{allowed}_{\textsc{l},\textsc{dir}}) + \textsc{cost}_{\textsc{l}} \cdot \sum_\textsc{l} \textsc{slack}_{\textsc{l}} 
+  \end{equation}
+  \begin{align}
     \text{subject to } &\forall_{i \in \textsc{sentence}}: \textsc{reachable}_{\textsc{root}, i} = \textsc{true} &\\
@@
 \end{figure}
-We found that the precision values converged to 100\% while the recall values continued to grow throughout our data without converging.
-This indicates that the solutions of the previous runs were consistent with the largest run.
-However as the data grew, we would continue to encounter novel link types.
+We found that the precision values converged to 100\% while the recall values continued to grow throughout our data.
+This indicates that the solutions of the previous runs were consistent with the largest run, and as the data grew we would continue to encounter novel link types and incorporate them.
 
@@
 \section{Link Corpus Experiment and Results}
-We ran the link parser on the english bnews corpus sentences.
+We ran the link parser and our ILP on the english bnews corpus sentences.
 \footnote{We ignored the link parses that the link parser could not find suitable attachments and returned a disconnected graph.
 This happened for roughly 19\% of the sentences in the corpus.}
-From these parses we ran our ILP and generated directionalized multiheaded dependency data to compare with the original conll data.
+From this, we generated directionalized multiheaded dependency data to compare with the original conll data.
 
@@
 \subsection{Blanks}
-Whenever the link parser failed to recognize a word from its dictionary, it would skip it and fail to attach any links. This accounts for about 10\% of the mismatches.
+The link parser will skip and attach no links to words that it failed to recognize from its dictionary. This accounts for about ten percent of the mismatches.
 
@@
   \caption{Example Sentences}
+  \label{fig:parses}
 \end{figure*}
@@
 
-We should acknowledge that another option would be to post-process either CoNLL parses or our link parses to automatically add more multi-heading, e.g., to handle control phenomena ("John wants to skip" or "Jane ran for office in order to change the tax laws" should have John/Jane be the subject of two verbs).
-However, that requires per-language
+\NoteJE{We should acknowledge that another option would be to post-process either CoNLL parses or our link parses to automatically add more multi-heading, e.g., to handle control phenomena ("John wants to skip" or "Jane ran for office in order to change the tax laws" should have John/Jane be the subject of two verbs).
+However, that requires per-language}
+
+We have a set of 100 parses similar to figure~\ref{fig:parses} in the appendix.
 

### 2014-05-27 src/tools/experiments/conll_analysis.py
modified the conll_analysis scripts to display the example sentences differently.

@@ -277,7 +277,7 @@ conll_analysis_sentence_count = 0
 
 sentenceCount = 0
 skip_sentenceCount = 2
-tikzLimit = 6
+tikzLimit = 3
 
 allParseCount = 0
 allParseLimit = 100
@@ -286,14 +286,20 @@ for linkSentence in link_results:
     if linkSentence in conll_results:
         conll_analysis_sentence_count += 1
 
-        if sentenceCount < tikzLimit and len(linkSentence.split()) > 4 and len(linkSentence.split()) < 7:
+        sentenceCheck = True
+        linkSentenceCheck = linkSentence.split()
+        sentenceCheck = "serwer" not in linkSentenceCheck and "salees" not in linkSentenceCheck
+        
+
+
+        if (sentenceCount < tikzLimit) and len(linkSentence.split()) > 4 and len(linkSentence.split()) < 6 and sentenceCheck:
             if skip_sentenceCount > 0:
                 skip_sentenceCount -= 1
             else:
-                tikz = tikz_dependency(conll_results[linkSentence], link_results[linkSentence], linkSentence, .9 / 2)            
+                tikz = tikz_dependency(conll_results[linkSentence], link_results[linkSentence], linkSentence, .97 / 3)            
                 TIKZ.write(tikz)
                 sentenceCount += 1
-                if sentenceCount % 2 == 0:
+                if sentenceCount % 3 == 0:
                     TIKZ.write("\n")
 
         if allParseCount < allParseLimit:

### 2014-05-27 doc/LinksToDAG.tex

--- 2014-05-26 doc/LinksToDAG.tex
@@
 \begin{abstract}
-% Copied from the google doc
-There has been recent theoretical work on multi-headed versions of dependency grammar(cite).
+Under multi-headed dependency grammar, a parse is a DAG rather than a tree.
 Such formalisms can be more syntactically and semantically expressive.
-However, there aren't any corpora at present (check this) except for Buch-Kromann's work on Danish \cite{kromann2003danish}, so it is hard to investigate the benefit of such parsers or to work on making them faster or more accurate.
-To fill this gap, we observe that link grammar produces parses that are similar to multi-headed planar dependency parses except that the links are undirected.
-We use Integer Linear Programming to assign consistent directions to the links in a corpus of NNN parses produced by the Link Grammar Parser, which has broad-coverage hand-written grammars of English, Russian, and more.
-(what statistics does it use?).
-We find that such directions can indeed be consistently assigned in a way that yields valid multi-headed dependency parses.
-The resulting parses in English appear linguistically plausible, although they are not in general consistent with CoNLL-style parses of the same sentences; we discuss the differences.
-We also report ...
+However, there are few multi-headed corpora at present, particularly for the projective case, so it is hard to investigate the benefit of such parsers or to work on making them faster or more accurate.
+To fill this gap, we observe that link grammar produces {\em undirected} projective graphs.
+% OLD
+% ... produces parses that are similar to multi-headed projective dependency parses, except that the links are undirected. 
+We use Integer Linear Programming to assign consistent directions to the labeled links in a corpus of !!!
+\noteJE{fill in the number} parses produced by the Link Grammar Parser, which has broad-coverage hand-written grammars of English, Russian, and more.
+\noteJE{what statistics does it use?} We find that such directions can indeed be consistently assigned in a way that yields valid multi-headed dependency parses.
+The resulting parses in English appear reasonably linguistically plausible, although they are not in general consistent with CoNLL-style parses of the same sentences; we discuss the differences.
+We also report on the convergence of our ILP method to a stable annotation scheme as the corpus size increases.
 \end{abstract}
@@
 \section{Introduction}
-\TODO{Intro}
+\todo{Intro}
 Projective dependency parsing is the task of mapping a sentence to a directed acyclic tree.
@@
 Using the sentences of a dependency corpus as comparison, we find that the link parses do not wholly subsume dependency parses, and that the undirected links match roughly two thirds of the arcs in the conll dependency corpus.
-\TODO{This is because...}
+\todo{This is because...}
 Of the links that match a corresponding dependency arc we measure whether the direction we chose is the same as the original dependency data. 
@@
 %Gomez-Rodriguez and Nivre 2013\cite{gomez-rodriguez_divisible_2013}
+
+Dependency parsing is the task of mapping a sentence to a projective (not always projective?) directed acyclic tree.
+Link parsing in contrast produces a multiheaded planar graph with undirected edges, where every edge has a label describing the relationship between two constituents in a parse.
+In this paper we explore whether these relationships include dependencies.
+To determine the directional dependencies within the link edge labels we will use integer linear programming, encoding the problem in the Zimpl little language \cite{Koch2004}.
+It turns out that the link parses roughly only match half of the conll dependency corpus.
+However this is because \todo{}.
+
+Some recent work has considered multi-headed ...
+
+ by expressing the dependency of a word on multiple parents.
+
+\cite{mcdonald-pereira-2006,buchkromann-2006,sagae-tsujii-2008,gomezrodriguez-nivre-2013}
+
+ (check this) except for Buch-Kromann's work on Danish \cite{kromann-2003}
 
@@
 This approach also potentially allows us to analyze link grammar dictionaries other than English.
-\TODO{explore other link grammar dictionaries}
+\todo{explore other link grammar dictionaries}
 
@@
 
-\TODO{cite model minimization papers}
+\todo{cite model minimization papers}
 
@@
 \end{figure}
+
+\subsection{Slack Hierarchy}
+
+We introduced slack on the directional dependency variable tokens such that link types were allowed to deviate from the majority up to (1\% ?) of the time before the ILP would assign both directional dependency variables to true.
+This addressed noise in the link parser's label assignments, while still allowing for the possibility that both directions could still be assigned.
+
+\todo{} We also introduced slack on the link types such that link types with the same coarse grained label would try and align the same way as the majority in the group, where the preceding capital letters of the link type denote the coarse grained label, while the subscript letters denote further information.
+This slack places a prior on rare or never-before seen link types to be assigned in the same way as other similar link types.
+
+This slack hierarchy gave the ILP the flexibility to handle noise and novel link types while still trying to learn the overall link grammar.
+
+\begin{equation}
+  count_{L, \textsc{left}} = \sum_{a \in \textsc{ARCS}_L} left_a
+\end{equation}
+\begin{equation}
+  count_{L, \textsc{left}} = \textsc{M} \cdot \textsc{allowed}_{L,\textsc{left}} + \textsc{slack}_{L}
+\end{equation}
+Where $L \in \textsc{LABELS}$, and $\textsc{M}$ is the size of the corpus.
+The objective is:
+\begin{equation}
+  \min \sum_{L \in \textsc{LABELS} \atop \textsc{DIR} \in \{left, right\}} (\textsc{allowed}_{L,\textsc{DIR}}) \atop + \textsc{COST}(L) * \sum_L \textsc{slack}_L
+\end{equation}
+\begin{equation}
+  \textsc{COST}(L) = \frac{100}{(\textsc{count}_{L,\textsc{left}} + \textsc{count}_{L,\textsc{right}})}
+\end{equation}
+
+\begin{algorithm}
+\caption{ILP Encoding}\label{ILP}
+\begin{algorithmic}[1]
+%\Procedure{MyProcedure}{}
+
+\State $\text{min} \sum_{l \in labels} allowed_L(l) + allowed_R(l)$
+\State $\text{subject to}$
+\State $\text{Acyclicity}$
+\State $\text{All nodes reachable from the root.}$
+\end{algorithmic}
+\end{algorithm}
 
@@
   \input{figure/conll_analysis_sentences.tex}
-  \caption{\small Of the sentences used for this experiment section about \input{figure/conll_analysis_sentences_dropped}had at least one dropped word, and about \input{figure/conll_analysis_sentences_multiheaded}had at least one multiheaded word}
+  \caption{\small Of the sentences used for this experiment section, about \input{figure/conll_analysis_sentences_dropped}
+    had at least one dropped word, and about \input{figure/conll_analysis_sentences_multiheaded}
+    had at least one multiheaded word}
 \end{figure}
@@
 
-\TODO{A table here that lists each link grammar type together with its
+\todo[inline]{A table here that lists each link grammar type together with its
   count, its allowed direction(s), its percentage of majority
@@
 
-\TODO{Show a bunch of English sentences with CoNLL vs.\@ link parses.
+\todo[inline]{Show a bunch of English sentences with CoNLL vs.\@ link parses.
   Explain how they were selected and how the results are colorized etc.}

### 2014-05-27 doc/LinksToDAG.tex

@@
 %\Procedure{MyProcedure}{}
-
 \State $\text{min} \sum_{l \in labels} allowed_L(l) + allowed_R(l)$

### 2014-05-27 doc/LinksToDAG.tex
Removed old versions of the ILP formulation. From Jason's merge.

@@
 
-\begin{equation}
-  count_{L, \textsc{left}} = \sum_{a \in \textsc{ARCS}_L} left_a
-\end{equation}
-\begin{equation}
-  count_{L, \textsc{left}} = \textsc{M} \cdot \textsc{allowed}_{L,\textsc{left}} + \textsc{slack}_{L}
-\end{equation}
-Where $L \in \textsc{LABELS}$, and $\textsc{M}$ is the size of the corpus.
-The objective is:
-\begin{equation}
-  \min \sum_{L \in \textsc{LABELS} \atop \textsc{DIR} \in \{left, right\}} (\textsc{allowed}_{L,\textsc{DIR}}) \atop + \textsc{COST}(L) * \sum_L \textsc{slack}_L
-\end{equation}
-\begin{equation}
-  \textsc{COST}(L) = \frac{100}{(\textsc{count}_{L,\textsc{left}} + \textsc{count}_{L,\textsc{right}})}
-\end{equation}
-
-\begin{algorithm}
-\caption{ILP Encoding}\label{ILP}
-\begin{algorithmic}[1]
-%\Procedure{MyProcedure}{}
-\State $\text{min} \sum_{l \in labels} allowed_L(l) + allowed_R(l)$
-\State $\text{subject to}$
-\State $\text{Acyclicity}$
-\State $\text{All nodes reachable from the root.}$
-\end{algorithmic}
-\end{algorithm}
-
 \subsection{Slack}

### 2014-05-27 doc/LinksToDAG.tex
small edits, and citation of link parser

--- 2014-05-27 doc/LinksToDAG.tex
@@
 Such formalisms can be more syntactically and semantically expressive.
-However, there are few multi-headed corpora at present, particularly for the projective case, so it is hard to investigate the benefit of such parsers or to work on making them faster or more accurate.
-To fill this gap, we observe that link grammar produces {\em undirected} projective graphs.
+However, few multi-headed corpora exist, particularly for the projective case, so it is hard to train, test, or improve such parsers.
+% OLD
+% investigate the benefit of such parsers or to work on making them faster or more accurate.  
+To help fill this gap, we observe that link grammar already produces {\em undirected} projective graphs.  
 % OLD
@@
 We use Integer Linear Programming to assign consistent directions to the labeled links in a corpus of !!!
-\noteJE{fill in the number} parses produced by the Link Grammar Parser, which has broad-coverage hand-written grammars of English, Russian, and more.
-\noteJE{what statistics does it use?} We find that such directions can indeed be consistently assigned in a way that yields valid multi-headed dependency parses.
+\noteJE{fill in the number} parses produced by the Link Grammar Parser, which has broad-coverage hand-written grammars of English, Russian, and other languages.
+We find that such directions can indeed be consistently assigned in a way that yields valid multi-headed dependency parses.
 The resulting parses in English appear reasonably linguistically plausible, although they are not in general consistent with CoNLL-style parses of the same sentences; we discuss the differences.
-We also report on the convergence of our ILP method to a stable annotation scheme as the corpus size increases.
 \end{abstract}
@@
 \todo{Intro}
-Projective dependency parsing is the task of mapping a sentence to a directed acyclic tree.
+Projective dependency parsing is the task of mapping a sentence to a directed acyclic graph.  
 Link parsing in contrast produces a multiheaded planar graph with undirected edges, where every edge has a label describing the relationship between two constituents in a parse.
@@
 
-We will accomplish this by generating link parses using the AbiWord/CMU link grammar parser, learning a direction to every encountered link label type, and then using these assignments for all link tokens in the link parse corpus.
+We will accomplish this by generating link parses using the AbiWord/CMU link grammar parser version 5.0.0 \noteJE{is this the version you used?} \cite{LINKPARSER-2014}, learning a direction to every encountered link label type, and then using these assignments for all link tokens in the link parse corpus.
 We formulate an integer linear programming problem that will find the smallest set of assignments that will still produce dependency parses that are acyclic with all nodes reachable from the root.
@@
 \todo{explore other link grammar dictionaries}
+
+\NoteJE{what statistics does it use?  what languages are available}
 
@@
 \subsection{Stability of Results}
+
+We report on the convergence of our ILP method to a stable annotation scheme as the corpus size increases.
+
 We measured how the solution changed over increasing numbers of processed sentences. 

### 2014-05-27 doc/LinksToDAG.tex
list gomez-rodriguez & nivre properties (need to check against their paper)

@@
 \usepackage{amsmath}
+\usepackage{xspace}
 
@@
 \newcommand{\Notewho}[3][]{\notewho[inline,#1]{#2}{#3}}  % extra arg with miscellaneous author
+
+% \newcommand{\root}{\texttt{\$}}
 
@@
 \todo{Intro}
-Projective dependency parsing is the task of mapping a sentence to a directed acyclic graph.  
+
+Dependency parsing maps a sentence to a directed graph whose vertices are the words $1, 2, \ldots, n$ of the sentence along with a distinguished ``root'' vertex 0.
+A labeled directed edge $u \stackrel{L}{\rightarrow} v$ or $v \stackrel{L}{\leftarrow} u$ indicates that the ``child'' $v$ is some kind of argument or modifier of its ``parent'' $u$.
+The edge label $L$ indicates the specific syntactic or semantic relationship between the two words.
+In the special case $u=0$, the edge designates $v$ as playing some special top-level role in the sentence, e.g., as the main verb.
+The case $v=0$ is not allowed.
+
+As recently discussed by \newcite{gomez-rodriguez_divisible_2013}, one might place several optional requirements on the graph:
+\begin{itemize}
+\item acyclic (there are no directed cycles)
+\item connected (for every pair of words $u,v$, there is an undirected path between them)
+\item accessible (for every word $v\neq 0$, there is a directed path from 0 to $v$)
+\item planar (no ``crossing edges'': if there is an edge between $u$ and $v$, then there is no edge between $i$ and $j$ where $u < i < v$ and not $u \leq j \leq v$)
+\item single-headed (each $v\neq 0$ has at most one parent)
+\end{itemize}
+It is common to impose all of these requirements at once, leading to a {\em projective dependency parser} that produces projective, rooted trees.
+However, parsing algorithms can be devised that relax any or all of the requirements \cite{gomez-rodriguez_divisible_2013}.
+
+In this paper, we are interested in relaxing the requirement of single-headedness.
+\NoteJE{discuss why.  Later, explain why our task isn't trivial.  We could just orient all edges from left to right, but we want to be consistent.}
+
 Link parsing in contrast produces a multiheaded planar graph with undirected edges, where every edge has a label describing the relationship between two constituents in a parse.

### 2014-05-27 doc/LinksToDAG.tex
rewrote sections 1-2

--- 2014-05-27 doc/LinksToDAG.tex
@@
 
-\section{Introduction}
-\todo{Intro}
-Projective dependency parsing is the task of mapping a sentence to a directed acyclic graph.  
-Link parsing in contrast produces a multiheaded planar graph with undirected edges, where every edge has a label describing the relationship between two constituents in a parse.
+\section{Multi-Headed Dependency Parsing}
+
+Dependency parsing maps a sentence to a directed graph whose vertices are the words $1, 2, \ldots, n$ of the sentence along with a distinguished ``root'' vertex 0.
+A labeled directed edge $u \stackrel{L}{\rightarrow} v$ or $v \stackrel{L}{\leftarrow} u$ indicates that the ``child'' $v$ is some kind of argument or modifier of its ``parent'' $u$.
+The edge label $L$ indicates the specific syntactic or semantic relationship between the two words.
+
+In the special case $u=0$, the edge designates $v$ as playing some special top-level role in the sentence, e.g., as the main verb.  We disallow $v=0$.
+
+As recently discussed by \newcite{gomezrodriguez-nivre-2013}, one might impose various requirements on the parse graph:
+\begin{itemize}
+\item {\sc Single-Head}: each word has at most one parent
+\item {\sc Acyclic}: there are no directed cycles
+\item {\sc Connected}: each pair of words has a undirected path between them
+\item {\sc Reachable}: each word can be reached from 0 by a directed path
+\item {\sc Planar}: edges may not ``cross'' \\ (if there are edges between $u,v$ and between $i,j$, where $u < i < v$, then $u \leq j \leq v$)
+\end{itemize}
+It is common to impose all of these requirements at once, leading to a {\em projective dependency parser} that produces projective trees rooted at 0.\footnote{{\sc Projective} basically means {\sc Planar} + {\sc Reachable}.}  However, parsing algorithms can be devised that relax any or all of the requirements \cite{gomezrodriguez-nivre-2013}.
+
+In this paper, we are interested in relaxing the {\sc Single-Head} requirement while preserving all the others.
+This means that the parse can have more than $n$ edges, allowing it to express more relationships between words.
+In English, for example, here are some constructions that seem to call for a multi-headed analysis.
+\begin{description}
+\item[control] In {\em ``Jill likes to skip,''} the word {\em Jill} is the subject of two verbs.
+In {\em ``Jill persuaded Jack to skip,''} {\em Jack} is the object of one verb and the subject of another.
+Without recognizing this, our parser would miss the syntactic invariant that {\em skip} always has a subject and {\em persuaded} always has an object.
+It would also be unable to exploit the selectional preferences of both verbs to help disambiguate the parse.
+This is why we prefer to make the parser aware of multi-headedness, rather than using a single-headed parser and then extracting the additional semantic roles from its output.
+\item[relativization] In {\em ``The boy that Jill skipped with fell down,''} the word {\em boy} is the object of {\em with} as well as the subject of {\em fell}.
+Without recognizing this, we would miss the syntactic invariant that {\em with} always has an object.
+\item[conjunction] In {\em ``Jack and Jill went up the hill,''} {\em Jack} and {\em Jill} serve as the two arguments to {\em and}, but they are also semantically subjects of {\em went}.
+Without recognizing this, we would have no (local) reason for expecting the arguments of {\em and} to be nouns.
+\end{description}
+
+In linguistics, it is common to analyze some of these structures using trees with ``empty categories.''  The subject of {\em skip} is taken to be a silent morpheme {\em PRO}:
+{\em ``Jill$_i$ likes PRO$_i$ to skip.''}  However, this is no longer a tree if one considers the implicit undirected edge between {\em Jill} and {\em PRO} (denoted by their shared index $i$).
+Our simpler representation contracts this coreference edge, eliminating {\em PRO} and creating a {\em Jill} $\leftarrow$ {\em skip} link.
+
+\section{Link Grammars}
+
+A few past NLP papers have explored multi-headed dependency parsing \cite{buchkromann-2006,mcdonald-pereira-2006,sagae-tsujii-2008,gomezrodriguez-nivre-2013}.
+Unfortunately, there seem to be no annotated corpora in this form other than the Danish Dependency Treebank \cite{kromann-2003}; researchers have sometimes converted corpora from other formats such as HPSG.
+
+All of these options result in non-projective parses, so the parsers must use non-projective or pseudo-projective algorithms.
+
+It seems at first that no one has worked out annotation conventions for {\em projective} multi-headed dependency parsing.
+However, this is not quite true.
+Link Grammar \cite{SleatorTemperly91} is a grammar-based formalism for projective dependency parsing with {\em undirected} links.
+It produces undirected connected planar graphs.
+The annotation conventions are implicit in the detailed Link Grammar lexicon,%
+\footnote{\url{http://www.abisource.com/projects/link-grammar/dict/introduction.html}.
+The link types in the English lexicon are documented at \url{http://www.abisource.com/projects/link-grammar/dict/summarize-links.html}.}
+which specifies for every word a constraint on the {\em sequence} of labeled leftward and rightward edges attached to it.
+As remarked by \newcite{eisner-2000-iwptbook}, this is analogous to dependency grammar's use of head automata to constrain a word's sequence of left and right children.
+For example, in {\em ``The boy that Jill skipped with fell down,''} the word {\em with} uses a lexical entry that requires it to link to a governing verb to the left, an extracted object farther to the left, and nothing to the right.
+
+Given a link grammar parse, it would be straightforward to convert it to an acyclic dependency parse by orienting all edges rightward.
+However, the result may violate the {\sc Reachability} constraint.
+Instead we could orient all edges by depth-first search from the root node, which yields a DAG satisfying all our constraints.
+However, this might result in inconsistent annotation conventions, with some \texttt{S} links pointing from subject to verb and others from verb to subject.
+
+We supposed that the link grammar lexicon designers actually had a consistent direction in mind for each edge type.
+We would expect verbs to point to their subject arguments in dependency grammar, and so we surmise that all \texttt{S} links are intended to point leftward (from verb to subject: {\em ``Jack $\stackrel{\texttt{S}}{\leftarrow}$ is falling''}).
+The link grammar designers take care to use a distinct \texttt{SI} label in cases of subject-verb inversion, and we surmise that \texttt{SI} links are intended to point rightward (again from verb to subject: {\em ``Jack $\stackrel{\texttt{SI}}{\rightarrow}$ is falling''}).
+
+Our goal in this paper is to recover these implicit directions by global optimization.
+We seek a fixed mapping from labels to directions such that we can interpret link grammar parses as directed dependency parses that satisfy all of our constraints.
+
+Our first thought was to seek a direction mapping such that no parsed word sequence allowed by the link grammar lexicon could possibly violate our constraints after directions were imposed.
+This is a well-defined constraint programming problem.
+For example, to prevent cyclicity, we would require (roughly speaking) that no word type could follow a sequence of directed rightward links through other word types and then a leftward link back to itself.
+
+However, working directly with the link grammar lexicon format is somewhat tricky.
+We also feared that there would not be a perfect solution---for example, because of errors in the lexicon, or linguistically unnatural word sequences not anticipated by the grammar designers.
+In this case it would not be clear how to relax our constraints.
+
+Thus, we chose to use a sample of {\em actual} sentences parsed by the link grammar, and to seek a direction mapping so that {\em these} parses would not violate our constraints after directions were imposed.
+If no such mapping exists, then we are willing to orient a few edge tokens in the wrong direction to ensure that the parses are still well-formed---but we minimize the number of such violations.
+In this way, the empirical distribution of sentences guides our assignment of directions.
+
+% Link grammars are a grammatical system equivalent to context-free grammars that assign linking requirements to every given word.
+A link parser then tries to satisfy all of these requirements for every word of a sentence while still maintaining planarity.
+The resulting links describe the relationships between constituents in a parse.
+
+% The link grammar is based on a set of handwritten dictionaries.
+Instead of going through these dictionaries, we learned the grammar using an ILP.
+This approach also potentially allows us to analyze link grammar dictionaries other than English.
+\todo{explore other link grammar dictionaries}
+
+% \NoteJE{what statistics does it use?  what languages are available}
+
+% Link parsing in contrast produces a multiheaded planar graph with undirected edges, where every edge has a label describing the relationship between two constituents in a parse.
 In this paper we explore whether these relationships include dependencies, and whether the multiheadedness of the link grammar offers additional dependency relationships not found in other corpora.
-
-We will accomplish this by generating link parses using the AbiWord/CMU link grammar parser version 5.0.0 \noteJE{is this the version you used?} \cite{LINKPARSER-2014}, learning a direction to every encountered link label type, and then using these assignments for all link tokens in the link parse corpus.
-We formulate an integer linear programming problem that will find the smallest set of assignments that will still produce dependency parses that are acyclic with all nodes reachable from the root.
-
-From the ILP, we produce a dictionary of label-to-direction assignments. We publish this so that anyone can easily turn a link parse into a dependency parse.
-
-Using the sentences of a dependency corpus as comparison, we find that the link parses do not wholly subsume dependency parses, and that the undirected links match roughly two thirds of the arcs in the conll dependency corpus.
-\todo{This is because...}
-Of the links that match a corresponding dependency arc we measure whether the direction we chose is the same as the original dependency data. 
-Finally, we look to see that given a link label, whether the original conll label can be recovered.
 
@@
 
-%Kromann 2003 Danish dependency treebank\cite{kromann2003danish}
-%McDonald and Pereira 2006\cite{McDonald2006}
-%Sagae and Tsujii 2008\cite{Sagae2008}
-%Gomez-Rodriguez and Nivre 2013\cite{gomez-rodriguez_divisible_2013}
-
-Dependency parsing is the task of mapping a sentence to a projective (not always projective?) directed acyclic tree.
+% Dependency parsing is the task of mapping a sentence to a projective (not always projective?) directed acyclic tree.
 Link parsing in contrast produces a multiheaded planar graph with undirected edges, where every edge has a label describing the relationship between two constituents in a parse.
@@
 However this is because \todo{}.
-
-Some recent work has considered multi-headed ...
-
- by expressing the dependency of a word on multiple parents.
-
-\cite{mcdonald-pereira-2006,buchkromann-2006,sagae-tsujii-2008,gomezrodriguez-nivre-2013}
-
- (check this) except for Buch-Kromann's work on Danish \cite{kromann-2003}
-
-\NoteJE{discuss multiheadedness.  cite Kromann 2003's danish
-  dependency treebank.  Prior work: McDonald and Pereira 2006, Sagae
-  and Tsujii 2008, Gomez-Rodriguez and Nivre 2013.}
-% http://aclweb.org/anthology/J/J13/J13-4002.pdf
-% http://citeseerx.ist.psu.edu/viewdoc/download?doi=10.1.1.156.3938&rep=rep1&type=pdf
-
-\section{Link Grammars}
-Link grammars are a grammatical system equivalent to context-free grammars that assign linking requirements to every given word\cite{SleatorTemperly91}.
-A link parser then tries to satisfy all of these requirements for every word of a sentence while still maintaining planarity.
-The resulting links describe the relationships between constituents in a parse.
-
-The link grammar is based on a set of handwritten dictionaries.
-Instead of going through these dictionaries, we learned the grammar using an ILP.
-This approach also potentially allows us to analyze link grammar dictionaries other than English.
-\todo{explore other link grammar dictionaries}
-
-\NoteJE{what statistics does it use?  what languages are available}
 
@@
 \section{Link Corpus Experiment and Results}
+
+We will accomplish this by generating link parses using the AbiWord/CMU link grammar parser version 5.0.0 \noteJE{is this the version you used?} \cite{LINKPARSER-2014}, learning a direction to every encountered link label type, and then using these assignments for all link tokens in the link parse corpus.
+We formulate an integer linear programming problem that will find the smallest set of assignments that will still produce dependency parses that are acyclic with all nodes reachable from the root.
+
+From the ILP, we produce a dictionary of label-to-direction assignments. We publish this so that anyone can easily turn a link parse into a dependency parse.
+
+Using the sentences of a dependency corpus as comparison, we find that the link parses do not wholly subsume dependency parses, and that the undirected links match roughly two thirds of the arcs in the conll dependency corpus.
+\todo{This is because...}
+Of the links that match a corresponding dependency arc we measure whether the direction we chose is the same as the original dependency data. 
+Finally, we look to see that given a link label, whether the original conll label can be recovered.
+
 We ran the link parser and our ILP on the english bnews corpus sentences.

### 2014-05-27 doc/LinksToDAG.tex
added note that there are 122 link types

@@
 \footnote{\url{http://www.abisource.com/projects/link-grammar/dict/introduction.html}.
-The link types in the English lexicon are documented at \url{http://www.abisource.com/projects/link-grammar/dict/summarize-links.html}.}
+The 122 link types in the English lexicon are documented at \url{http://www.abisource.com/projects/link-grammar/dict/summarize-links.html}.}
 which specifies for every word a constraint on the {\em sequence} of labeled leftward and rightward edges attached to it.
@@
 In this way, the empirical distribution of sentences guides our assignment of directions.
+The resulting directed corpus can be used for research on multi-headed dependency parsing.
 

### 2014-05-27 doc/LinksToDAG.tex
minor

@@
 \begin{abstract}
-Under multi-headed dependency grammar, a parse is a DAG rather than a tree.
-Such formalisms can be more syntactically and semantically expressive.
-However, few multi-headed corpora exist, particularly for the projective case, so it is hard to train, test, or improve such parsers.
+Under multi-headed dependency grammar, a parse is a DAG rather than a tree.  Such formalisms can be more syntactically and semantically expressive.  However, 
+it is hard to train, test, or improve multi-headed parsers because few multi-headed corpora exist, particularly for the projective case.
 % OLD
@@
 % ... produces parses that are similar to multi-headed projective dependency parses, except that the links are undirected. 
-We use Integer Linear Programming to assign consistent directions to the labeled links in a corpus of !!!
-\noteJE{fill in the number} parses produced by the Link Grammar Parser, which has broad-coverage hand-written grammars of English, Russian, and other languages.
+We use Integer Linear Programming to assign consistent directions to the labeled links in a corpus of 70,000 \noteJE{fill in the number} parses produced by the Link Grammar Parser, which has broad-coverage hand-written grammars of English as well as Russian and other languages.
 We find that such directions can indeed be consistently assigned in a way that yields valid multi-headed dependency parses.
@@
 In {\em ``Jill persuaded Jack to skip,''} {\em Jack} is the object of one verb and the subject of another.
-Without recognizing this, our parser would miss the syntactic invariant that {\em skip} always has a subject and {\em persuaded} always has an object.
+Without recognizing this, our parser would miss the syntactic invariants that {\em skip} always has a subject and {\em persuaded} always has an object.
 It would also be unable to exploit the selectional preferences of both verbs to help disambiguate the parse.

### 2014-05-29 src/encoder.py
minor white space in main.py and encoder.py

@@ -100,8 +100,6 @@ def getBatchDataFromLinkParses(lines):
     processedSentences = []
     linkData = []
 
- 
-
     i = 0
     temp = 0
     for j in xrange(len(lines)):

### 2014-05-29 src/main.py

@@ -37,8 +37,6 @@ if __name__=="__main__":
     args = parser.parse_args()
 
 
-
-    # TODO
     # already in the cache
     if args.ID != None and cache.check(args.ID):
         print "SCIP SOLUTION FOUND IN CACHE"
@@ -56,8 +54,7 @@ if __name__=="__main__":
         # Link Edge Encoder
         (processedSentences, links) = getBatchDataFromLinkParses(lines)
         (sentences,sizeOfCorpus) = getSentencesFromProcessedSentences(processedSentences)
-
-    
+            
         wordTags = []
         linkLabels = []
         for i in xrange(len(sentences)):
@@ -79,7 +76,6 @@ if __name__=="__main__":
         
         allowedLabels.sort()
         cache.store(args.ID, (sentences, linkLabels, wordTags, allowedLabels, linkDeps))
-
 
 
     # Save the allowed links solution file.

### 2014-05-29 src/tools/plotter.py
added functionality to plotter.py. It colors the graphs with colors, plots with marker dots, and adjusts the axis a little bit. This is in preparation for joint plots of english and russian data.

@@ -16,6 +16,12 @@ parser.add_argument("strings", nargs="*")
 args = parser.parse_args()
 
 
+PE = "Precision: English"
+RE = "Recall: English"
+PR = "Precision: Russian"
+RR = "Recall: Russian"
+
+
 labels = []
 file_lines = []
 for filename in args.strings:
@@ -23,6 +29,19 @@ for filename in args.strings:
     label = filename.split("/")[-1]
     label = label.split(".")[0]
     label = label.split("_")[0]
+
+    if len(label.split("#")) > 1:
+        label = label.split("#")
+        label = ": ".join(label)
+    else:
+        label += ": en"
+
+    label = {"precision: en": PE,
+             "recall: en": RE,
+             "precision: ru": PR,
+             "recall: ru": RR
+             }.get(label, label)
+
 
     labels.append(label)
     
@@ -82,13 +101,27 @@ ax.set_ylabel(args.yLabel)
 for l,coord in zip(labels,file_coord):
     Xs = [elem[0] for elem in coord]
     Ys = [elem[1] for elem in coord]
-    plt.plot(Xs,Ys, label=l)
+
+    c = {PE:"DarkOrange",
+         RE:"red",
+         PR:"DeepSkyBlue",
+         RR:"blue"
+         }.get(l,"blue")
+
+
+    m = {PE:"o",
+         RE:"o",
+         PR:"*",
+         RR:"*"}.get(l,".")
+
+
+    plt.plot(Xs,Ys, label=l, color=c, marker=m, linewidth=1.5)
 
 
 
 ax.legend(loc='lower right')
-plt.xlim(xmin, xmax)
-plt.ylim(ymin=0)
+plt.xlim(-0.015*xmax, xmax + 0.015*xmax)
+plt.ylim(0,1.04)
 
 
 

### 2014-05-29 src/tools/experiments/trial_ru.sh
russian trial script. Specifically for the russian data.

@@ -1,0 +1,32 @@
+#!/bin/sh
+mkdir -p /tmp/LinksToDAG_trial_sol/
+
+SENTENCES=$1
+ITERATIONS=$2
+LANGUAGE="ru"
+
+
+
+/usr/bin/time -f "%e" -o runtime ./LinksToDAG $SENTENCES $ITERATIONS $LANGUAGE
+
+
+
+
+echo -n $ITERATIONS " " >> /tmp/LinksToDAG_times
+tail --lines 1 runtime >> /tmp/LinksToDAG_times
+
+
+
+if [ $LANGUAGE ]; then
+    cp sol/allowedLinks.txt /tmp/LinksToDAG_trial_sol/allowedLinks_$ITERATIONS\_$LANGUAGE.txt
+    #python src/tools/experiments/conll_analysis.py $SENTENCES $CONLL $ITERATIONS $LANGUAGE
+    #mv sol/conll_analysis/conll_analysis.txt sol/conll_analysis/conll_analysis_$ITERATIONS\_$LANGUAGE.txt
+else
+    cp sol/allowedLinks.txt /tmp/LinksToDAG_trial_sol/allowedLinks_$ITERATIONS.txt
+    #python src/tools/experiments/conll_analysis.py $SENTENCES $CONLL $ITERATIONS
+    #mv sol/conll_analysis/conll_analysis.txt sol/conll_analysis/conll_analysis_$ITERATIONS.txt
+fi
+
+rm -f runtime
+
+

### 2014-05-29 src/link_parser.py
link_parser.py now times out the parse after 30 seconds. This was added because the link parser in russian doesn't seem to enter panic mode properly and takes way too long.

@@ -33,7 +33,6 @@ class linkParser:
             else:
                 parse = self.parse(sentence)
                 self.cache.store(ID,parse)
-
             output.write(parse)
         output.close()
 
@@ -43,11 +42,16 @@ class linkParser:
 
     def parse(self,sentence):
         echo_process = Popen(["echo",sentence], stdout=PIPE)
-        linkparser_process = Popen(["link-parser", self.language, "-!graphics=0", "-!links=0", "-!echo=1", "-!postscript=1", "-!panic=0"], stdin=echo_process.stdout, stdout=PIPE, stderr=subprocess.PIPE)
+
+        # Time out the link parser. Don't even need to bother with its internal timeout.
+        timeout = str(30)
+        linkparser_process = Popen(["timeout", timeout, "link-parser", self.language, "-!graphics=0", "-!links=0", "-!echo=1", "-!postscript=1", "-!panic=0"], stdin=echo_process.stdout, stdout=PIPE, stderr=subprocess.PIPE)
         echo_process.stdout.close()
 
         output, err = linkparser_process.communicate()        
-        linkparser_process.stdout.close()
+        linkparser_process.stdout.close()        
+        #output = output.strip()
+        
         return output
 
 

### 2014-05-29 trials.sh
updated trials.sh to run trial_ru.sh for the russian data.

@@ -4,8 +4,8 @@ typeset -i i END
 
 
 #END=10000
-END=60030
-#END=10
+#END=60030
+END=256
 #INCREMENT=1
 
 SENTENCES="data/english_bnews_train.sentences"
@@ -35,22 +35,30 @@ TYPE_AGREEMENT="sol/type_agreement/"
 TYPE_ANALYSIS=$TYPE_AGREEMENT"type_analysis.txt"
 
 
-
-
 rm -f $RUNTIMES $TXT_RUNTIME $SOL_RUNTIME $TXT_PRECISION $SOL_PRECISION $SOL_RECALL
 rm -rf /tmp/LinksToDAG_trial_sol/
+    
 
-CONLL_ANALYSIS="sol/conll_analysis/conll_analysis_"$END".txt"
+# Run the trial_ru.sh script if we are doing russian trials
+if [ $LANGUAGE -a $LANGUAGE = "ru" ]; then
+    for ((i=1;i<=$END;i=$((i*2))));
+    do
+        bash src/tools/experiments/trial_ru.sh $SENTENCES $i
+    done
+    
+    if [ $i -ne $END -a $((i/2)) -ne $END ]; then
+        bash src/tools/experiments/trial_ru.sh $SENTENCES $END
+    fi
+else     
+    for ((i=1;i<=$END;i=$((i*2))));
+    do
+        bash src/tools/experiments/trial.sh $SENTENCES $CONLL $i $LANGUAGE
+    done
 
-for ((i=1;i<=$END;i=$((i*2))));
-do
-    bash src/tools/experiments/trial.sh $SENTENCES $CONLL $i $LANGUAGE
-done
-
-if [ $i -ne $END -a $((i/2)) -ne $END ]; then
-    bash src/tools/experiments/trial.sh $SENTENCES $CONLL $END $LANGUAGE
+    if [ $i -ne $END -a $((i/2)) -ne $END ]; then
+        bash src/tools/experiments/trial.sh $SENTENCES $CONLL $END $LANGUAGE
+    fi
 fi
-
 
 
 
@@ -61,12 +69,23 @@ python src/tools/plotter.py -x "Sentences" -y "Runtime (seconds)" -o $SOL_RUNTIM
 
 
 python src/tools/experiments/type_analysis.py 
-cat $TXT_PRECISION | sort -V > temp; mv temp $TXT_PRECISION
-cat $TXT_RECALL | sort -V > temp; mv temp $TXT_RECALL
[…]

### 2014-05-29 src/tools/experiments/tikz_dependency.py
updated tikz_dependency.py to change the produced graphs a bit. There is still more improvements to be made.

@@ -61,7 +61,7 @@ def tikz_dependency(conlls, links, sentence, ratio = 0.3, subfigure = True):
                 graph_link.deleteEdge(child,head)
             
             else:
-                result += tikz_draw_edge(head, child, label, "above", "densely dotted")
+                result += tikz_draw_edge(head, child, label, "above", "dotted")
 
 
 
@@ -73,7 +73,7 @@ def tikz_dependency(conlls, links, sentence, ratio = 0.3, subfigure = True):
             if len(graph_link.heads[child]) > 1 :
                 result += tikz_draw_edge(head,child,label, "below", "ultra thick", "orange")
             else:
-                result += tikz_draw_edge(head,child,label, "below", "densely dotted")
+                result += tikz_draw_edge(head,child,label, "below", "dotted")
 
         
     result += "\t\\end{dependency}\n"

### 2014-05-29 src/tools/experiments/conll_analysis.py
changed, updated, and debugged the conll_analysis. We remove parses that have dropped words in our analysis. We have to update the paper with our data.

@@ -6,6 +6,8 @@ from tikz_dependency import *
 from dependency_graph import *
 from conll_analysis_output import *
 
+DEBUG = True
+
 # The directory containing the original conll data file
 CONLL_DIR = "data/"
 CONLL_LOC = CONLL_DIR+"english_bnews_train.conll"
@@ -39,21 +41,35 @@ LATEX_FILE_SENTENCE_MULTIHEADED = LATEX_DIR+"conll_analysis_sentences_multiheade
 if not os.path.exists(LATEX_DIR):
     os.makedirs(LATEX_DIR)
 
-
 # Tikz directory and files
 TIKZ_DIR = "doc/figure/"
-TIKZ_LOC = TIKZ_DIR+"sentences.tikz"
+EXAMPLE_SENTENCES_LOC = TIKZ_DIR+"sentences.tikz"
 if not os.path.exists(TIKZ_DIR):
     os.makedirs(TIKZ_DIR)
-TIKZ = open(TIKZ_LOC, 'w+')
+PAPER_TIKZ = open(EXAMPLE_SENTENCES_LOC, 'w+')
 
-ALL_PARSES_LOC = TIKZ_DIR+"parses.tikz"
-ALL_PARSES = open(ALL_PARSES_LOC, 'w+')
+EXAMPLE_PARSES_LOC = TIKZ_DIR+"parses.tikz"
+EXAMPLE_PARSES = open(EXAMPLE_PARSES_LOC, 'w+')
+
+
+
+# --------------------------------------------------------------------
+# Analysis while reading in data. Pertains mostly to sentence counts
+# --------------------------------------------------------------------
+# Number of sentences from the link data
+link_sentence_total = 0
+link_dropped_sentences = 0
+link_remaining_sentences = 0
+
+# These maps will map a sentence to its conll representation.
+# We will use this to query conll results from both our conll and link-conll files
+conll_results = {}
+link_results = {}
+
 
 
 # Analysis
 def analysis(conlls, links):
-
     word_data = analysis_words(conlls, links)
     
     graph_conlls = getGraph(conlls)
@@ -73,7 +89,7 @@ def analysis(conlls, links):
     label_extra_count = {}
     
     mismatches = 0
-    blanks = 0
+    #blanks = 0
     total = 0
     
     for head in graph_conlls.table:
@@ -141,7 +157,7 @@ def analysis(conlls, links):
     match_data = (matches, label_match, label_match_count)
     reverse_data = (reverse_matches, label_reverse, label_reverse_count)    
[…]

### 2014-05-29 src/tools/experiments/conll_analysis_output.py

@@ -8,17 +8,15 @@ from dependency_graph import *
 def result_numbers(final_total, 
                    match_total, 
                    extra_total, 
-                   blank_total, 
                    mismatch_total, 
                    reverse_match_total,
-                   dropped_word_total,
                    multiheaded_word_count, 
                    word_count_total, 
                    dropped_sentence_count,
                    multiheaded_sentence_count, 
                    original_sentence_count,
-                   conll_analysis_sentence_count,
-                   total_sentence_count):
+                   link_sentence_total,
+                   link_remaining_sentences):
 
 
     match_percent = str(0)
@@ -27,20 +25,13 @@ def result_numbers(final_total,
         match_percent = str(float(match_total)/final_total)
         mismatch_percent = str(float(mismatch_total)/final_total)
 
-    blank_percent = str(0)
-    if blank_total != 0:
-        blank_percent = str(float(blank_total)/mismatch_total)
-
     directional_percent = str(0)
     if mismatch_total != 0:
         directional_percent = str(float(reverse_match_total)/mismatch_total)
 
     multiheaded_word_percent = str(0)
-    dropped_word_percent = str(0)
     if word_count_total != 0:
         multiheaded_word_percent = str(float(multiheaded_word_count)/word_count_total)
-        dropped_word_percent = str(float(dropped_word_total)/word_count_total)
-
 
 
     return """------------------------------------------------------------
@@ -68,11 +59,6 @@ Final Total:\t\t\t\t""" +str(final_total)+"""
 MISMATCHES
 ------------------------------------------------------------
 
-How many conll arcs do not match a link because the link node was dropped and had no attachments.
-Blank Total:\t\t\t\t"""+ str(blank_total)+"""
-Percent of all mismatches:\t\t"""+ blank_percent +"""
-
-
 ------------------------------------------------------------
 EXTRA LINKS
 ------------------------------------------------------------
@@ -89,10 +75,6 @@ How many multiheaded words there are.
 Multiheaded Words:\t\t\t"""+str(multiheaded_word_count)+"""
 Percent of all words:\t\t\t"""+multiheaded_word_percent+"""
 
-How many dropped words there are.
[…]

### 2014-05-29 trials.sh
gave a language file extension to the precision_recall analysis files. This way, I can run different languages and have them save in different file names.

@@ -5,7 +5,7 @@ typeset -i i END
 
 #END=10000
 #END=60030
-END=256
+END=64
 #INCREMENT=1
 
 SENTENCES="data/english_bnews_train.sentences"
@@ -69,14 +69,19 @@ python src/tools/plotter.py -x "Sentences" -y "Runtime (seconds)" -o $SOL_RUNTIM
 
 
 python src/tools/experiments/type_analysis.py 
-cat $TXT_PRECISION | sort -V > temp_precision; 
-cat $TXT_RECALL | sort -V > temp_recall; 
+cat $TXT_PRECISION | sort -V > temp_precision; rm -f $TXT_PRECISION 
+cat $TXT_RECALL | sort -V > temp_recall; rm -f $TXT_RECALL
 
 if [ $LANGUAGE ]; then
-    mv temp_precision $PRECISION_RECALL"precision#"$LANGUAGE".txt"
-    mv temp_recall $PRECISION_RECALL"recall#"$LANGUAGE".txt"
-    python src/tools/plotter.py -x "Sentences" -y "Percent" -o $PRECISION_RECALL"precision_recall#"$LANGUAGE".png" $PRECISION_RECALL"precision#"$LANGUAGE".txt" $PRECISION_RECALL"recall#"$LANGUAGE".txt"
+    TXT_PRECISION=$PRECISION_RECALL"precision#"$LANGUAGE".txt"
+    TXT_RECALL=$PRECISION_RECALL"recall#"$LANGUAGE".txt"
+
+    mv temp_precision $TXT_PRECISION
+    mv temp_recall $TXT_RECALL
+    python src/tools/plotter.py -x "Sentences" -y "Percent" -o $PRECISION_RECALL"precision_recall#"$LANGUAGE".png" $TXT_PRECISION $TXT_RECALL
 else
+    TXT_PRECISION=$PRECISION_RECALL"precision#en.txt"
+    TXT_RECALL=$PRECISION_RECALL"recall#en.txt"
     mv temp_precision $TXT_PRECISION
     mv temp_recall $TXT_RECALL
     python src/tools/plotter.py -x "Sentences" -y "Percent" -o $SOL_PRECISION_RECALL $TXT_PRECISION $TXT_RECALL

### 2014-05-30 src/tools/experiments/conll_analysis.py
minor change to conll_analysis.py

@@ -6,7 +6,7 @@ from tikz_dependency import *
 from dependency_graph import *
 from conll_analysis_output import *
 
-DEBUG = True
+DEBUG = False
 
 # The directory containing the original conll data file
 CONLL_DIR = "data/"
@@ -425,7 +425,7 @@ f = open(LATEX_FILE_SENTENCE, "w+")
 f.write(result)
 f.close()
 
-dropped_sentence_percentage = float(link_dropped_sentences) / link_remaining_sentences * 100
+dropped_sentence_percentage = float(link_dropped_sentences) / link_sentence_total * 100
 dropped_sentence_percentage = "{:.2f}".format(dropped_sentence_percentage)+"\\%"
 
 multiheaded_sentence_percentage = float(multiheaded_sentence_count) / link_remaining_sentences * 100

### 2014-05-30 trials.sh
updated trials.sh. The plan is to run it on the clsp machines.

@@ -5,7 +5,7 @@ typeset -i i END
 
 #END=10000
 #END=60030
-END=64
+END=60030
 #INCREMENT=1
 
 SENTENCES="data/english_bnews_train.sentences"
@@ -92,5 +92,7 @@ fi
 
 
 
-#echo $RUNTIMES $END $'\n'The CLSP machines cant run plotter. Please download files and run: $'\n'python src/tools/plotter.py -x "Sentences" -y "Percent" -o $SOL_PRECISION_RECALL $TXT_PRECISION $TXT_RECALL | mutt -s $END -a $RUNTIMES -a $TXT_PRECISION -a $TXT_RECALL -a $TYPE_ANALYSIS -a $CONLL_ANALYSIS -- junekihong@gmail.com
+echo $RUNTIMES $END $'\n'The CLSP machines cant run plotter. Please download files and run: $'\n'python src/tools/plotter.py -x "Sentences" -y "Percent" -o $SOL_PRECISION_RECALL $TXT_PRECISION $TXT_RECALL | mutt -s $END -a $RUNTIMES -a $TXT_PRECISION -a $TXT_RECALL -- junekihong@gmail.com
+
+#-a $TYPE_ANALYSIS -a $CONLL_ANALYSIS -- junekihong@gmail.com
 

### 2014-05-30 doc/LinksToDAG.tex
updated the ILP formulation in the paper. Its really large now and does not fit in one column. But its much more complete of a description.

@@
 \newcommand{\Note}[4][]{\todo[author=#2,color=#3,fancyline,#1]{#4}}
-\newcommand{\noteJH}[2][]{\Note[#1]{MRG}{blue!40}{#2}}
+\newcommand{\noteJH}[2][]{\Note[#1]{JH}{blue!40}{#2}}
 \newcommand{\noteJE}[2][]{\Note[#1]{JE}{green!40}{#2}}   
@@
 \section{Integer Linear Programming}
-We formulated an ILP to assign directionality to all of the links in a set of link grammar parses, while still mandating that the resulting dependency parses are connected DAGs, with the nodes reachable from the root.
+We formulate an ILP to find the minimal set of link type to directionality assignments that would directionalize all the links in a set of link grammar parses, while still mandating that all of the resulting dependency parses would be connected DAGs, and with all the nodes reachable from the root.
 
-In order to choose which direction the link labels are assigned, for every encountered label we generate two boolean label/direction variables correspending to left and right.
-Setting one of these variables to \textsc{true} allows only the specified direction for a label throughout all parses, and because we assign a directionality to every link type at least one of these variables must be set to true.
-Setting both variables would allow us to go in either direction.
-
-Our ILP objective is to minimize the number of label/direction variables set to \textsc{true}, while still satisfying the acyclicity and reachability constraints.
+For every encountered label we generate two boolean label/direction variables (e.g.
+label/left, label/right).
+Setting one of these variables to \textsc{true} allows the label type to only go in the specified direction, while setting both would allow the label to go in either way.
+The objective is to minimize the number of label/direction variables set to \textsc{true}, while still satisfying acyclicity and reachability constraints.
 The encoding was written in the Zimpl little language \cite{Koch2004} and solved using the SCIP Optimization Suite\cite{achterberg2009scip}.
@@
 
-\begin{figure}
+\begin{figure*}
   \small
-  \begin{align}
-    \textsc{dir} &= \{\textsc{left}, \textsc{right}\} &\\ 
-    \textsc{l} &\in \textsc{labels} & \\
-    \textsc{m} &= \left|{\textsc{links}}\right| &\\
-    \textsc{count}_{\textsc{l}, \textsc{dir}} &= \sum \textsc{link}_{\textsc{l},\textsc{dir}} &\\
-    \textsc{count}_{\textsc{l}, \textsc{dir}} &= \textsc{m} \cdot \textsc{allowed}_{\textsc{l},\textsc{dir}} + \textsc{slack}_{\textsc{l}} &\\
-    \textsc{cost}_{\textsc{l}} &= \frac{100}{\textsc{count}_{\textsc{l},\textsc{left}} + \textsc{count}_{\textsc{l},\textsc{right}}} &
-  \end{align}
-  \begin{equation} % I seperated out this line from the rest because it would disrupt the line number on the right side of the figure.
-    \text{minimize:} \sum_{\textsc{l}, \textsc{dir}} (\textsc{allowed}_{\textsc{l},\textsc{dir}}) + \textsc{cost}_{\textsc{l}} \cdot \sum_\textsc{l} \textsc{slack}_{\textsc{l}} 
-  \end{equation}
-  \begin{align}
-    \text{subject to } &\forall_{i \in \textsc{sentence}}: \textsc{reachable}_{\textsc{root}, i} = \textsc{true} &\\
-                       &\forall_{\substack{i,j \in \textsc{sentence} \\ i\neq j }}: \textsc{has\_cycle}_{i,j} = \textsc{false} &
-  \end{align}
+  \begin{align*}
+    \text{Sets:}&&\\
+    &\textsc{link}_{i,j,\textsc{label},\textsc{sentence}} \in \textsc{links} &\\
+    &\textsc{label} \in \textsc{labels}&\\ 
+    &\textsc{dir} = \{\textsc{left}, \textsc{right}\} &\\
+    \text{Params:}&&\\
+    &\textsc{m}_\textsc{length} = \max (\forall{\textsc{link}_{i,j,\textsc{sentence}}} \max(i,j)) &\\
+    %&\textsc{count}_{\textsc{label}, \textsc{dir}} = \sum_{\textsc{link}_{\textsc{label},\textsc{dir}}}1 &\\
+    &\textsc{cost}_{\textsc{label}} = \frac{100}{ \sum_{\textsc{link}_{\textsc{label},\textsc{dir}}}1 } &\\
+    \text{Variables:}&&\\
+    &\textsc{slack}_{\textsc{link}} \geq 0 &\\ 
+    &\textsc{depth}_{i,\textsc{sentence}} \geq 0 &\\
+    &\textsc{allowed}_{\textsc{label},\textsc{dir}} = \{0,1\}&\\
+    &\textsc{arc}_{\textsc{link}} = \{0,1\} &\\
+    \text{minimize:}& &\\ 
+    &\sum_{\textsc{label}, \textsc{dir}} (\textsc{allowed}_{\textsc{label},\textsc{dir}}) + \textsc{cost}_{\textsc{label}} \cdot \sum_\textsc{link} \textsc{slack}_{\textsc{link}} &\\
+    &&\\
+    \text{subject to: }& &\\
+    &\text{Links go in the allowed directions, except for slack}&\\
+    &\forall{\textsc{link}_{i,j,\textsc{label}}}: 1-\textsc{arc}_{\textsc{link}} \leq \textsc{allowed}_{\textsc{label}, \textsc{left}} + \textsc{slack}_{\textsc{link}} &\\
+    &\forall{\textsc{link}_{i,j,\textsc{label}}}: \textsc{arc}_{\textsc{link}} \leq \textsc{allowed}_{\textsc{label}, \textsc{right}} + \textsc{slack}_{\textsc{link}} &\\
+    &\text{Depth constraints for acyclicity}&\\
+    &\forall{\textsc{link}_{i,\textsc{sentence}}}: \textsc{depth}_{\text{root},\textsc{sentence}} \leq \textsc{depth}_{i,\textsc{sentence}} &\\
+    &\forall{\textsc{link}_{i,j,\textsc{sentence}}}: \textsc{depth}_{i,\textsc{sentence}} + \textsc{m}_\textsc{length}\cdot\textsc{arc}_{i,j,\textsc{sentence}} \geq \textsc{depth}_{j,\textsc{sentence}} + 1 &\\
+    &\forall{\textsc{link}_{i,j,\textsc{sentence}}}: \textsc{depth}_{j,\textsc{sentence}} + \textsc{m}_\textsc{length}\cdot(1-\textsc{arc}_{i,j,\textsc{sentence}}) \geq \textsc{depth}_{i,\textsc{sentence}} + 1 &\\
+    &\text{Every token has a parent}&\\
+    &\forall{\textsc{link}_{i \neq \textsc{root},\textsc{sentence}}}: \sum_{\textsc{link}_{i,j,\textsc{sentence}}} (1-\textsc{arc}_{i,j,\textsc{sentence}}) + \sum_{\textsc{link}_{j,i,\textsc{sentence}}}(\textsc{arc}_{j,i,\textsc{sentence}}) \geq 1&\\
+  \end{align*}
   \caption{\small The ILP formulation.}
+\end{figure*}
+
+\subsection{Slack}
+We introduced slack such that the link types were allowed to deviate from the majority one percent of the time before the ILP would assign both label/direction variables to \textsc{true}.
+This allows for flexibility against noise in the link parser's label assignments, balanced with allowing for both directions to be assigned if necessary.
+
+\section{Experiments and Results}
+To evaluate ...
+
+We generated link parses using the AbiWord/CMU link grammar parser version 5.0.8 \cite{LINKPARSER-2014}.
+
+The experiments were conducted on Russian and English sentences.
+The Russian sentences were taken from a part of the Russian National Corpus (about 75,000 sentences)\noteJH{cite?}. 
+The English sentences were taken from the conll bnews training data (about 60,000 sentences)\noteJH{cite?}.
+We were able to compare the resulting English link dependency data with the original conll to analyze the differences.
+
+We publish the final link type to direction assignments so that anyone can easily turn a link parse into a dependency parse.
+
+We report on the convergence of the results of our ILP with increasing data.
+With increasing trials of sentences, we measure how similar the subsequent directional assignments would be with each other.
+Taking the largest run as the answer key, we compared how much the smaller runs deviated from it.
+We recorded the precision of whether the assignments in the smaller runs could be found in those of the largest run, and similarly the recall of whether the assignments in the largest run could be found in the smaller runs.
+
+\subsection{Stability of Results}
+\begin{figure}[ht!]
+  \small
+  \centering
+  \includegraphics[width=0.9\linewidth, keepaspectratio=true]{figure/precision_recall.png}
+  \caption{\small Precision/recall of increasing runs of English and Russian sentences}
 \end{figure}
 
-\subsection{Slack Hierarchy}
+We found that the precision values converged to 100\% while the recall values continued to grow throughout the data.
+This indicates that the solutions of the previous runs were consistent with the largest run, and as the data increased we would continue to encounter novel link types and incorporate them.
 
-We introduced slack on the directional dependency variable tokens such that link types were allowed to deviate from the majority up to (1\% ?) of the time before the ILP would assign both directional dependency variables to true.
-This addressed noise in the link parser's label assignments, while still allowing for the possibility that both directions could still be assigned.
+\subsection{Link Corpus Analysis}
 
-\todo{} We also introduced slack on the link types such that link types with the same coarse grained label would try and align the same way as the majority in the group, where the preceding capital letters of the link type denote the coarse grained label, while the subscript letters denote further information.
-This slack places a prior on rare or never-before seen link types to be assigned in the same way as other similar link types.
-
-This slack hierarchy gave the ILP the flexibility to handle noise and novel link types while still trying to learn the overall link grammar.
-
-\subsection{Slack}
-We introduced slack on the directional dependency variables such that the link types were allowed to deviate from the majority one percent of the time before the ILP would assign both directional dependency variables to true.
-This addressed noise in the link parser's label assignments, while still allowing for both directions to be assigned if needed.
-
-\subsection{Stability of Results}
-
-We report on the convergence of our ILP method to a stable annotation scheme as the corpus size increases.
-
-We measured how the solution changed over increasing numbers of processed sentences. 
-Taking the solution to the largest run as the answer key, we measured how much the previous runs deviated from it.
-We measured the precision of whether the assignments in the smaller runs matched the assignments of the largest run and we measured the recall of whether the assignments in the largest run could be found in the smaller runs.
-\begin{figure}[ht!]
-  \small
-  \includegraphics[width=\linewidth, keepaspectratio=true]{figure/precision_recall.png}
-  \caption{\small Precision/recall of results of English sentences}
-\end{figure}
-We found that the precision values converged to 100\% while the recall values continued to grow throughout our data.
-This indicates that the solutions of the previous runs were consistent with the largest run, and as the data grew we would continue to encounter novel link types and incorporate them.
-
-%\section{Other languages}
-
-\section{Link Corpus Experiment and Results}
-
-We will accomplish this by generating link parses using the AbiWord/CMU link grammar parser version 5.0.0 \noteJE{is this the version you used?} \cite{LINKPARSER-2014}, learning a direction to every encountered link label type, and then using these assignments for all link tokens in the link parse corpus.
-We formulate an integer linear programming problem that will find the smallest set of assignments that will still produce dependency parses that are acyclic with all nodes reachable from the root.
-
-From the ILP, we produce a dictionary of label-to-direction assignments. We publish this so that anyone can easily turn a link parse into a dependency parse.
+\begin{figure*}[ht!]
+  \centering
+  \input{figure/sentences.tikz}
+  \caption{Example Parses: The top blue arcs represent the conll dependencies, and the bottom red arcs are the directionalized link grammar parses}
+  \label{fig:parses}
+\end{figure*}
 
@@
   \input{figure/conll_analysis_sentences.tex}
-  \caption{\small Of the sentences used for this experiment section, about \input{figure/conll_analysis_sentences_dropped}
-    had at least one dropped word, and about \input{figure/conll_analysis_sentences_multiheaded}
-    had at least one multiheaded word}
+  \caption{\small From the original sentences, we discarded disconnected parses for the ILP.
+From these sentences, we only looked at the complete link parses and discarded sentences with dropped words for our analysis section.
+About \input{figure/conll_analysis_sentences_dropped} of the non-disconnected sentences contained dropped words.}
 \end{figure}
+
+\todo{Fix the caption in the conll analysis sentence chart}
+
+About \input{figure/conll_analysis_sentences_multiheaded} of the sentences with complete parses had at least one multiheaded word
 
@@
 \section{Discussion}
-
-\begin{figure*}[ht!]
-  \centering
-  \input{figure/sentences.tikz}
-  \caption{Example Sentences}
-  \label{fig:parses}
-\end{figure*}
 
@@
 
+Here we display a set of link parses as further examples
+
 \end{document}

### 2014-05-30 src/tools/experiments/conll_analysis.py
automatically generate the number of sentences used for the paper.

@@ -38,6 +38,7 @@ LATEX_DIR = "doc/figure/"
 LATEX_FILE_SENTENCE = LATEX_DIR+"conll_analysis_sentences.tex"
 LATEX_FILE_SENTENCE_DROPPED = LATEX_DIR+"conll_analysis_sentences_dropped.tex"
 LATEX_FILE_SENTENCE_MULTIHEADED = LATEX_DIR+"conll_analysis_sentences_multiheaded.tex"
+LATEX_FILE_SENTENCE_TOTAL = LATEX_DIR+"conll_analysis_sentences_total.tex"
 if not os.path.exists(LATEX_DIR):
     os.makedirs(LATEX_DIR)
 
@@ -425,9 +426,14 @@ f = open(LATEX_FILE_SENTENCE, "w+")
 f.write(result)
 f.close()
 
+
+
+#---------------------------------------------------------------------
+# Take our generated numbers and put them in latex files 
+# to be called by the paper
+#---------------------------------------------------------------------
 dropped_sentence_percentage = float(link_dropped_sentences) / link_sentence_total * 100
 dropped_sentence_percentage = "{:.2f}".format(dropped_sentence_percentage)+"\\%"
-
 multiheaded_sentence_percentage = float(multiheaded_sentence_count) / link_remaining_sentences * 100
 multiheaded_sentence_percentage = "{:.2f}".format(multiheaded_sentence_percentage)+"\\%"
 
@@ -439,6 +445,9 @@ f = open(LATEX_FILE_SENTENCE_MULTIHEADED, "w+")
 f.write(multiheaded_sentence_percentage)
 f.close()
 
+f = open(LATEX_FILE_SENTENCE_TOTAL, "w+")
+f.write("{:,d}".format(link_sentence_total))
+f.close()
 
 
 

### 2014-05-30 src/tools/plotter.py
changed the marker styles in the pyplot of plotter.py

@@ -109,10 +109,10 @@ for l,coord in zip(labels,file_coord):
          }.get(l,"blue")
 
 
-    m = {PE:"o",
-         RE:"o",
-         PR:"*",
-         RR:"*"}.get(l,".")
+    m = {PE:".",
+         RE:".",
+         PR:"+",
+         RR:"+"}.get(l,".")
 
 
     plt.plot(Xs,Ys, label=l, color=c, marker=m, linewidth=1.5)

### 2014-05-30 doc/LinksToDAG.tex
paper writing. Moved the ILP formulation to the appendix.

@@
 % ... produces parses that are similar to multi-headed projective dependency parses, except that the links are undirected. 
-We use Integer Linear Programming to assign consistent directions to the labeled links in a corpus of 70,000 \noteJE{fill in the number} parses produced by the Link Grammar Parser, which has broad-coverage hand-written grammars of English as well as Russian and other languages.
+We use Integer Linear Programming to assign consistent directions to the labeled links in a corpus of \input{figure/conll_analysis_sentences_total.tex}parses produced by the Link Grammar Parser, which has broad-coverage hand-written grammars of English as well as Russian and other languages.
 We find that such directions can indeed be consistently assigned in a way that yields valid multi-headed dependency parses.
@@
 
-%about 20\%
-
 % Dependency parsing is the task of mapping a sentence to a projective (not always projective?) directed acyclic tree.
@@
 We formulate an ILP to find the minimal set of link type to directionality assignments that would directionalize all the links in a set of link grammar parses, while still mandating that all of the resulting dependency parses would be connected DAGs, and with all the nodes reachable from the root.
+Our task of finding the smallest set of assignments that would explain the parsesis also called model minimization\cite{ravi2009}.
 
@@
 
-\todo{cite model minimization papers}
+\subsection{Slack}
+We introduced slack such that the link types were allowed to deviate from the majority one percent of the time before the ILP would assign both label/direction variables to \textsc{true}.
+This allows for flexibility against noise in the link parser's label assignments, balanced with allowing for both directions to be assigned if necessary.
+
+\section{Experiments and Results}
+
+We generated link parses using the AbiWord/CMU link grammar parser version 5.0.8 \cite{LINKPARSER-2014}.
+
+The experiments were conducted on Russian and English sentences.
+The Russian sentences were taken from a part of the Russian National Corpus (about 75,000 sentences)\noteJH{cite?}. 
+The English sentences were taken from the conll bnews training data (about 60,000 sentences)\noteJH{cite?}.
+We were also able to compare the resulting English link dependency data with the original conll to analyze the differences.
+
+In these languages, we also publish our link type to direction assignments so that anyone can use it to turn a link parse into a dependency parse.
+
+\subsection{Stability of Results}
+We report on the convergence of the results of our ILP with increasing data.
+With increasing trials of sentences, we measure how similar the subsequent directional assignments would be with each other.
+Taking the largest run as the answer key, we compared how much the smaller runs deviated from it.
+We recorded the precision of whether the assignments in the smaller runs could be found in those of the largest run, and similarly the recall of whether the assignments in the largest run could be found in the smaller runs.
+
+\begin{figure}[ht!]
+  \small
+  \centering
+  \includegraphics[width=\linewidth, keepaspectratio=true]{figure/precision_recall.png}
+  \caption{\small Precision/recall of increasing runs of English and Russian sentences. The precision values converge while the recall values continue to grow throughout the data.}
+\end{figure}
+Because the precision values were high, we found that the solutions of the previous runs were consistent with the largest run, and as the data increased so did the recall values, and so we would continue to encounter novel link types and then incorporate them.
+
+\subsection{English Link-Dependency Analysis}
+
+\begin{figure*}[ht!]
+  \centering
+  \input{figure/sentences.tikz}
+  \caption{Example Parses: The top blue arcs represent the conll dependencies, and the bottom red arcs are the directionalized link grammar parses}
+  \label{fig:parses}
+\end{figure*}
+
+In order to compare 
+Using the sentences of the dependency corpus as comparison, we find that the link parses do not wholly subsume dependency parses, and that the undirected links match roughly three fourths of the arcs in the conll dependency corpus.
+\todo{This is because...}
+
+Of the links that match a corresponding dependency arc
+
+we measure whether the direction we chose is the same as the original dependency data. 
+Finally, we look to see that given a link label, whether the original conll label can be recovered.
+
+From this, we generated directionalized multiheaded dependency data to compare with the original conll data.
+
+\begin{figure}[ht!]
+  \centering
+  \small
+  \input{figure/conll_analysis_sentences.tex}
+  \caption{\small The subsets of English sentences used for this work.}
+\end{figure}
+
+From the original sentences, we discarded disconnected parses for the ILP.
+From these sentences, we only looked at the complete link parses and discarded sentences with dropped words for our analysis section.
+We removed about \input{figure/conll_analysis_sentences_dropped} of the non-disconnected sentences this way.
+Of the sentences with complete parses, about \input{figure/conll_analysis_sentences_multiheaded}had at least one multiheaded word.
+
+Compared to the original conll data, the link parses match about half the conll arcs in location and directionality. The mismatches in the other half can be accounted for in the following ways.
+
+\section{Discussion}
+
+\todo{more in the appendix ...}
+
+\NoteJE{We should acknowledge that another option would be to post-process either CoNLL parses or our link parses to automatically add more multi-heading, e.g., to handle control phenomena ("John wants to skip" or "Jane ran for office in order to change the tax laws" should have John/Jane be the subject of two verbs).
+However, that requires per-language}
+
+We have a set of 100 parses similar to figure~\ref{fig:parses} in the appendix.
+
+\section{Future Work}
+
+\subsection{Slack Hierarchy}
+We would like to introduce slack on the link grammar types such that link types with the same coarse grained label would try and align the same way as the majority in the group, where the preceding capital letters of the link type denote the coarse grained label, while the subscript letters denote further information.
+This slack would place a prior on rare or never-before seen link types to be assigned in the same way as other similar link types.
+We think that this will give the ILP better flexibility to handle noise and novel link types while still trying to learn the overall link grammar.
+
+\section{Related Work}
+
+\bibliographystyle{acl2014}
+\bibliography{LinksToDAG}
+
+\clearpage
+\appendix
+\section{ILP Formulation}
+Here we will present our ILP formulation.
 
@@
 
-\subsection{Slack}
-We introduced slack such that the link types were allowed to deviate from the majority one percent of the time before the ILP would assign both label/direction variables to \textsc{true}.
-This allows for flexibility against noise in the link parser's label assignments, balanced with allowing for both directions to be assigned if necessary.
-
-\section{Experiments and Results}
-To evaluate ...
-
-We generated link parses using the AbiWord/CMU link grammar parser version 5.0.8 \cite{LINKPARSER-2014}.
-
-The experiments were conducted on Russian and English sentences.
-The Russian sentences were taken from a part of the Russian National Corpus (about 75,000 sentences)\noteJH{cite?}. 
-The English sentences were taken from the conll bnews training data (about 60,000 sentences)\noteJH{cite?}.
-We were able to compare the resulting English link dependency data with the original conll to analyze the differences.
-
-We publish the final link type to direction assignments so that anyone can easily turn a link parse into a dependency parse.
-
-We report on the convergence of the results of our ILP with increasing data.
-With increasing trials of sentences, we measure how similar the subsequent directional assignments would be with each other.
-Taking the largest run as the answer key, we compared how much the smaller runs deviated from it.
-We recorded the precision of whether the assignments in the smaller runs could be found in those of the largest run, and similarly the recall of whether the assignments in the largest run could be found in the smaller runs.
-
-\subsection{Stability of Results}
-\begin{figure}[ht!]
-  \small
-  \centering
-  \includegraphics[width=0.9\linewidth, keepaspectratio=true]{figure/precision_recall.png}
-  \caption{\small Precision/recall of increasing runs of English and Russian sentences}
-\end{figure}
-
-We found that the precision values converged to 100\% while the recall values continued to grow throughout the data.
-This indicates that the solutions of the previous runs were consistent with the largest run, and as the data increased we would continue to encounter novel link types and incorporate them.
-
-\subsection{Link Corpus Analysis}
-
-\begin{figure*}[ht!]
-  \centering
-  \input{figure/sentences.tikz}
-  \caption{Example Parses: The top blue arcs represent the conll dependencies, and the bottom red arcs are the directionalized link grammar parses}
-  \label{fig:parses}
-\end{figure*}
-
-Using the sentences of a dependency corpus as comparison, we find that the link parses do not wholly subsume dependency parses, and that the undirected links match roughly two thirds of the arcs in the conll dependency corpus.
-\todo{This is because...}
-Of the links that match a corresponding dependency arc we measure whether the direction we chose is the same as the original dependency data. 
-Finally, we look to see that given a link label, whether the original conll label can be recovered.
-
-We ran the link parser and our ILP on the english bnews corpus sentences.
-\footnote{We ignored the link parses that the link parser could not find suitable attachments and returned a disconnected graph.
-This happened for roughly 19\% of the sentences in the corpus.}
-From this, we generated directionalized multiheaded dependency data to compare with the original conll data.
-
-\begin{figure}[ht!]
-  \centering
-  \small
-  \input{figure/conll_analysis_sentences.tex}
-  \caption{\small From the original sentences, we discarded disconnected parses for the ILP.
-From these sentences, we only looked at the complete link parses and discarded sentences with dropped words for our analysis section.
-About \input{figure/conll_analysis_sentences_dropped} of the non-disconnected sentences contained dropped words.}
-\end{figure}
-
-\todo{Fix the caption in the conll analysis sentence chart}
-
-About \input{figure/conll_analysis_sentences_multiheaded} of the sentences with complete parses had at least one multiheaded word
-
-Compared to the original conll data, the link parses match about half the conll arcs in location and directionality. The mismatches in the other half can be accounted for in the following ways.
-
-\subsection{Blanks}
-The link parser will skip and attach no links to words that it failed to recognize from its dictionary. This accounts for about ten percent of the mismatches.
-
-\subsection{Directional Mismatches}
-In about 25\% of the cases, the conll arc's corresponding link would be assigned the opposite direction.
-
-\section{Discussion}
-
-\todo{more in the appendix ...}
-
-\NoteJE{We should acknowledge that another option would be to post-process either CoNLL parses or our link parses to automatically add more multi-heading, e.g., to handle control phenomena ("John wants to skip" or "Jane ran for office in order to change the tax laws" should have John/Jane be the subject of two verbs).
-However, that requires per-language}
-
-We have a set of 100 parses similar to figure~\ref{fig:parses} in the appendix.
-
-\section{Future Work}
-
-\subsection{Slack Hierarchy}
-We would like to introduce slack on the link grammar types such that link types with the same coarse grained label would try and align the same way as the majority in the group, where the preceding capital letters of the link type denote the coarse grained label, while the subscript letters denote further information.
-This slack would place a prior on rare or never-before seen link types to be assigned in the same way as other similar link types.
-We think that this will give the ILP better flexibility to handle noise and novel link types while still trying to learn the overall link grammar.
-
-\section{Related Work}
-
-\bibliographystyle{acl2014}
-\bibliography{LinksToDAG}
-
-\clearpage
-\appendix
 \section{English Link Types}

### 2014-05-31 doc/LinksToDAG.tex
updated the sample english parse section of the appendix in the paper.

@@
 \begin{abstract}
-Under multi-headed dependency grammar, a parse is a DAG rather than a tree.  Such formalisms can be more syntactically and semantically expressive.  However, 
+Under multi-headed dependency grammar, a parse is a directed acyclic graph rather than a tree.  Such formalisms can be more syntactically and semantically expressive.  However, 
 it is hard to train, test, or improve multi-headed parsers because few multi-headed corpora exist, particularly for the projective case.
@@
 The objective is to minimize the number of label/direction variables set to \textsc{true}, while still satisfying acyclicity and reachability constraints.
-The encoding was written in the Zimpl little language \cite{Koch2004} and solved using the SCIP Optimization Suite\cite{achterberg2009scip}.
+The encoding was written in the Zimpl little language \cite{Koch2004} and solved using the SCIP Optimization Suite \cite{achterberg2009scip}.
 
@@
 This allows for flexibility against noise in the link parser's label assignments, balanced with allowing for both directions to be assigned if necessary.
+
+\section{English and Russian Data}
+The data used for this work came from English and Russian corpuses.
+The English sentences were taken from the bnews conll training data set \noteJH{cite} and provided about 60,000 sentence examples in total.
+
+The Russian sentence data was taken from a part of the Russian National Corpus \noteJH{cite?}
 
@@
 \section{ILP Formulation}
-Here we will present our ILP formulation.
+We present our ILP formulation.
 
@@
 \section{Sample English Parses}
+In this section, we will display 100 sample English parses.
+The top half of each parse visualization represents the CoNLL dependency parse, while the bottom half shows our directionalized link grammar parse.
 
-\todo[inline]{Show a bunch of English sentences with CoNLL vs.\@ link parses.
-  Explain how they were selected and how the results are colorized etc.}
+\subsection{Sentence Selection}
+These sentences have been taken from the English bnews training data set.
+We chose to display sentences with complete link parses similar to the analysis section of this paper.
+We display the first 100 parses whose character lengths do not exceed the width of the page.
 
-Here we display a set of link parses as further examples
+\subsection{Parse Colorization}
+The top blue arcs represent the CoNLL dependency parses. For every arc:
+\begin{enumerate}
+\item If there is a corresponding link edge between the same nodes then a solid blue line is drawn.
+\item Or else a dotted blue line is drawn. 
+\end{enumerate}
+The bottom red arcs represent the directionalized link grammar parses. For every link edge:
+\begin{enumerate}
+\item If there is a corresponding CoNLL arc between the same nodes then a solid red line is drawn.
+\item If the matching arc is facing the opposite direction, then the link edge is drawn thicker.
+\item Of the remaining edges, if an edge is a multiheaded link, then it is drawn with thick orange.
+\item The rest of the remaining edges are drawn with dotted red lines.
+\end{enumerate}
+
+\subsection{Link Parser Annotations}
+The sentence used in the middle of the parse is taken from the output of the link parser.
+We removed sentences with dropped words, but they would have shown up in the parse as [bracketed].
+There are also words that the link parser did not recognize, but rather made a guess as to its part of speech.
+These words are tagged with a [!] symbol at the end.
+The dashes in the link grammar part of speech tag annotations mean that the link parser did not return one.
+
+% Uncomment this to add in all of the parses. Warning: compilation will take a long time.
+%\input{figure/parses.tikz}
 

### 2014-05-31 doc/LinksToDAG.tex
removed some cruft from the experiments section of the paper. Its still really draft-y and not polished though.

@@
 The data used for this work came from English and Russian corpuses.
-The English sentences were taken from the bnews conll training data set \noteJH{cite} and provided about 60,000 sentence examples in total.
+The English sentences were taken from the bnews CoNLL training data set \noteJH{cite}.
+We were able to compare the our link parses with the dependencies.
+The Russian sentence data was taken from a part of the Russian National Corpus \noteJH{cite?}. With only the sentence data, we did not have dependencies to compare with.
 
-The Russian sentence data was taken from a part of the Russian National Corpus \noteJH{cite?}
+We also publish our link type to direction assignments so that anyone can use it to turn a link parse into a dependency parse.
 
 \section{Experiments and Results}
-
 We generated link parses using the AbiWord/CMU link grammar parser version 5.0.8 \cite{LINKPARSER-2014}.
-
-The experiments were conducted on Russian and English sentences.
-The Russian sentences were taken from a part of the Russian National Corpus (about 75,000 sentences)\noteJH{cite?}. 
-The English sentences were taken from the conll bnews training data (about 60,000 sentences)\noteJH{cite?}.
-We were also able to compare the resulting English link dependency data with the original conll to analyze the differences.
-
-In these languages, we also publish our link type to direction assignments so that anyone can use it to turn a link parse into a dependency parse.
 

### 2014-06-01 doc/LinksToDAG.tex
fiddled with the ILP formulation slightly. Added another chart to the analysis section. Added the file too so that Jason can build the paper.

@@
 
-\section{English and Russian Data}
-The data used for this work came from English and Russian corpuses.
-The English sentences were taken from the bnews CoNLL training data set \noteJH{cite}.
-We were able to compare the our link parses with the dependencies.
-The Russian sentence data was taken from a part of the Russian National Corpus \noteJH{cite?}. With only the sentence data, we did not have dependencies to compare with.
+\section{Data Sets}
+The data used for this work came from English bnews CoNLL dependency training data \noteJH{cite?} and the Russian National Corpus\footnote{\url{http://ruscorpora.ru/en/}}.
+With the English sentences we compared our link parse results to the original dependency annotations, and with the Russian sentences we only produced link parses.
 
-We also publish our link type to direction assignments so that anyone can use it to turn a link parse into a dependency parse.
+\subsection{Discarding Incomplete Link Parses}
+Not all of the data was used for this work.
+From the given sentences, we discarded those that the link parser could not process and returned a parse with nodes disconnected from the root.
+\input{figure/conll_analysis_sentences_disconnected} of the English sentences were removed in this way.
+The ILP was run only on the connected parses to produce a corpus of directionalized link parses.
+
+For our analysis of these parses, we further discarded those that had dropped words, where the link parser could not attach links to every word in the sentence.
+About \input{figure/conll_analysis_sentences_dropped}of the non-disconnected sentences were removed, leaving us working only with complete link parses.
+
+\begin{figure}[ht!]
+  \centering
+  \small
+  \input{figure/conll_analysis_sentences.tex}
+  \caption{\small The English sentences used for this work.}
+\end{figure}
+
+\noteJH{The stability results come from the second row of this table. The analysis later in the paper uses the third row. Putting this section here might be confusing to readers?}
 
@@
 \subsection{Stability of Results}
-We report on the convergence of the results of our ILP with increasing data.
-With increasing trials of sentences, we measure how similar the subsequent directional assignments would be with each other.
-Taking the largest run as the answer key, we compared how much the smaller runs deviated from it.
-We recorded the precision of whether the assignments in the smaller runs could be found in those of the largest run, and similarly the recall of whether the assignments in the largest run could be found in the smaller runs.
-
 \begin{figure}[ht!]
@@
 \end{figure}
+We report on the convergence of the results of our ILP with increasing data.
+With increasing trials of sentences, we measure how similar the subsequent directional assignments would be with each other.
+Taking the largest run as the answer key, we compared how much the smaller runs deviated from it.
+We recorded the precision of whether the assignments in the smaller runs could be found in those of the largest run, and similarly the recall of whether the assignments in the largest run could be found in the smaller runs.
+
 Because the precision values were high, we found that the solutions of the previous runs were consistent with the largest run, and as the data increased so did the recall values, and so we would continue to encounter novel link types and then incorporate them.
@@
   \input{figure/sentences.tikz}
-  \caption{Example Parses: The top blue arcs represent the conll dependencies, and the bottom red arcs are the directionalized link grammar parses}
+  \caption{\small Example Parses: The top blue arcs represent conll dependencies; the bottom red arcs represent directionalized link grammar parses. See the Appendix for more details.}
   \label{fig:parses}
@@
 
-Of the links that match a corresponding dependency arc
+\begin{figure}[ht!]
+  \centering
+  \small
+  \input{figure/conll_analysis_arcs.tex}
+  \caption{Arcs that match links}
+\end{figure}
 
@@
 
-From this, we generated directionalized multiheaded dependency data to compare with the original conll data.
-
-\begin{figure}[ht!]
-  \centering
-  \small
-  \input{figure/conll_analysis_sentences.tex}
-  \caption{\small The subsets of English sentences used for this work.}
-\end{figure}
-
-From the original sentences, we discarded disconnected parses for the ILP.
-From these sentences, we only looked at the complete link parses and discarded sentences with dropped words for our analysis section.
-We removed about \input{figure/conll_analysis_sentences_dropped} of the non-disconnected sentences this way.
 Of the sentences with complete parses, about \input{figure/conll_analysis_sentences_multiheaded}had at least one multiheaded word.
-
-Compared to the original conll data, the link parses match about half the conll arcs in location and directionality. The mismatches in the other half can be accounted for in the following ways.
 
@@
 
-\begin{figure*}
+\begin{figure}[!htb]
   \small
@@
     \text{Sets:}&&\\
-    &\textsc{link}_{i,j,\textsc{label},\textsc{sentence}} \in \textsc{links} &\\
+    &\textsc{link}_{\textsc{sentence},\textsc{label}, i,j} \in \textsc{links} &\\
     &\textsc{label} \in \textsc{labels}&\\ 
@@
     \text{Params:}&&\\
-    &\textsc{m}_\textsc{length} = \max (\forall{\textsc{link}_{i,j,\textsc{sentence}}} \max(i,j)) &\\
+    &\textsc{m}_\textsc{length} = \max (\forall{\textsc{link}_{\textsc{sentence}, i,j}} \max(i,j)) &\\
     %&\textsc{count}_{\textsc{label}, \textsc{dir}} = \sum_{\textsc{link}_{\textsc{label},\textsc{dir}}}1 &\\
@@
     &\textsc{slack}_{\textsc{link}} \geq 0 &\\ 
-    &\textsc{depth}_{i,\textsc{sentence}} \geq 0 &\\
+    &\textsc{depth}_{\textsc{sentence},i} \geq 0 &\\
     &\textsc{allowed}_{\textsc{label},\textsc{dir}} = \{0,1\}&\\
@@
     &\text{Depth constraints for acyclicity}&\\
-    &\forall{\textsc{link}_{i,\textsc{sentence}}}: \textsc{depth}_{\text{root},\textsc{sentence}} \leq \textsc{depth}_{i,\textsc{sentence}} &\\
-    &\forall{\textsc{link}_{i,j,\textsc{sentence}}}: \textsc{depth}_{i,\textsc{sentence}} + \textsc{m}_\textsc{length}\cdot\textsc{arc}_{i,j,\textsc{sentence}} \geq \textsc{depth}_{j,\textsc{sentence}} + 1 &\\
-    &\forall{\textsc{link}_{i,j,\textsc{sentence}}}: \textsc{depth}_{j,\textsc{sentence}} + \textsc{m}_\textsc{length}\cdot(1-\textsc{arc}_{i,j,\textsc{sentence}}) \geq \textsc{depth}_{i,\textsc{sentence}} + 1 &\\
+    &\forall{\textsc{link}_{\textsc{sentence},i}}: \textsc{depth}_{\textsc{sentence},\text{root}} \leq \textsc{depth}_{\textsc{sentence},i} &\\
+    &\forall{\textsc{link}_{\textsc{sentence},i,j}}: \textsc{depth}_{\textsc{sentence},i} + \textsc{m}_\textsc{length}\cdot\textsc{arc}_{\textsc{sentence},i,j} \geq \textsc{depth}_{\textsc{sentence},j} + 1 &\\
+    &\forall{\textsc{link}_{\textsc{sentence},i,j}}: \textsc{depth}_{\textsc{sentence},j} + \textsc{m}_\textsc{length}\cdot(1-\textsc{arc}_{\textsc{sentence},i,j}) \geq \textsc{depth}_{\textsc{sentence},i} + 1 &\\
     &\text{Every token has a parent}&\\
-    &\forall{\textsc{link}_{i \neq \textsc{root},\textsc{sentence}}}: \sum_{\textsc{link}_{i,j,\textsc{sentence}}} (1-\textsc{arc}_{i,j,\textsc{sentence}}) + \sum_{\textsc{link}_{j,i,\textsc{sentence}}}(\textsc{arc}_{j,i,\textsc{sentence}}) \geq 1&\\
+    &\forall{\textsc{link}_{\textsc{sentence},i \neq \textsc{root}}}: \sum_{\textsc{link}_{\textsc{sentence},i,j}} (1-\textsc{arc}_{\textsc{sentence},i,j}) + \sum_{\textsc{link}_{\textsc{sentence},j,i}}(\textsc{arc}_{\textsc{sentence},j,i}) \geq 1&\\
   \end{align*}
   \caption{\small The ILP formulation.}
-\end{figure*}
+\end{figure}
+
+\clearpage
 

### 2014-06-01 src/tools/experiments/conll_analysis.py
adding the analysis script additions that make the new table in the paper.

@@ -36,9 +36,11 @@ if len(sys.argv) > 3:
 # Latex directory and files. Where we will store these results in latex form.
 LATEX_DIR = "doc/figure/"
 LATEX_FILE_SENTENCE = LATEX_DIR+"conll_analysis_sentences.tex"
+LATEX_FILE_SENTENCE_DISCONNECTED = LATEX_DIR+"conll_analysis_sentences_disconnected.tex"
 LATEX_FILE_SENTENCE_DROPPED = LATEX_DIR+"conll_analysis_sentences_dropped.tex"
 LATEX_FILE_SENTENCE_MULTIHEADED = LATEX_DIR+"conll_analysis_sentences_multiheaded.tex"
 LATEX_FILE_SENTENCE_TOTAL = LATEX_DIR+"conll_analysis_sentences_total.tex"
+LATEX_FILE_ARCS = LATEX_DIR+"conll_analysis_arcs.tex"
 if not os.path.exists(LATEX_DIR):
     os.makedirs(LATEX_DIR)
 
@@ -343,7 +345,7 @@ for linkSentence in link_results:
             if paper_sentence_count % 3 == 0:
                 PAPER_TIKZ.write("\n")
     # Filter out sentences to only up to length 16
-    elif example_parses_count < example_parses_limit and len(linkSentence.split()) <= 1500:
+    elif example_parses_count < example_parses_limit and len(linkSentence) <= 85:
         tikz = tikz_dependency(conll_results[linkSentence], link_results[linkSentence], linkSentence, 1.0, False)
         EXAMPLE_PARSES.write("\\begin{figure*}[ht!]\n")
         EXAMPLE_PARSES.write(tikz)
@@ -426,16 +428,23 @@ f = open(LATEX_FILE_SENTENCE, "w+")
 f.write(result)
 f.close()
 
+result = result_latex_arcs(match_total, reverse_match_total, mismatch_total, arc_total)
+f = open(LATEX_FILE_ARCS, "w+")
+f.write(result)
+f.close()
 
 
 #---------------------------------------------------------------------
 # Take our generated numbers and put them in latex files 
 # to be called by the paper
 #---------------------------------------------------------------------
-dropped_sentence_percentage = float(link_dropped_sentences) / link_sentence_total * 100
-dropped_sentence_percentage = "{:.2f}".format(dropped_sentence_percentage)+"\\%"
[…]

### 2014-06-01 src/tools/experiments/conll_analysis_output.py

@@ -206,19 +206,35 @@ Of the cases where the links had different directionality, what are the "extra" 
 def result_latex_corpus(original_sentence_count,
                         link_sentence_total, 
                         conll_analysis_sentence_count):
-
     result = ""
     result += "\t\\begin{tabular}{|l|l|}\n"
     result += "\t\t\\hline\n"
-    result += "\t\tOriginal number of sentences in conll corpus & "+str(original_sentence_count)+"\\\\ \n"
+    result += "\t\tNumber of English sentences used from the corpus & "+str(original_sentence_count)+"\\\\ \n"
     result += "\t\t\\hline\n"
     result += "\t\tSentences after discarding disconnected parses & "+str(link_sentence_total)+"\\\\ \n"
     result += "\t\t\\hline\n"
-    result += "\t\tSentences used for experiment and analysis & "+str(conll_analysis_sentence_count)+"\\\\ \n"
+    result += "\t\tSentences with complete parses used for analysis & "+str(conll_analysis_sentence_count)+"\\\\ \n"
     result += "\t\t\\hline\n"
     result += "\t\\end{tabular}\n"
 
     return result
 
 
-
+def result_latex_arcs(match_total, 
+                      reverse_match_total,
+                      mismatch_total,
+                      final_total):
+    result = "\t\\begin{tabular}{|l|l|l|}\n"
+    result += "\t\t\\hline\n"
+    result += "\t\t & CoNLL arcs & Percent of total \\\\ \n"
+    result += "\t\t\\hline\n"
+    result += "\t\tMatches: & "+str(match_total)+" & "+"{:.2f}\\%".format(float(match_total)/final_total*100)+"\\\\ \n"
+    result += "\t\t\\hline\n"
+    result += "\t\tFlipped direction: & "+str(reverse_match_total)+" & "+"{:.2f}\\%".format(float(reverse_match_total)/final_total*100)+"\\\\ \n"
+    result += "\t\t\\hline\n"
+    result += "\t\tMismatches: & "+str(mismatch_total)+" & "+"{:.2f}\\%".format(float(mismatch_total)/final_total*100)+"\\\\ \n"
+    result += "\t\t\\hline\n"
+    result += "\t\tTotal: & "+str(final_total)+" & "+str(int(float(final_total)/final_total)*100)+"\\% \\\\ \n"
+    result += "\t\t\\hline\n"
+    result += "\t\\end{tabular}\n"
+    return result

### 2014-06-02 doc/LinksToDAG.tex
explained ILP model; clarified data sets; other notes & small edits

@@
 
+\newcommand{\eqnref}[1]{\eqref{eqn:#1}}
+
 \usepackage[usenames,dvipsnames,svgnames,table]{xcolor}  % allows better color names
@@
 \date{}
+
+\setlength\titlebox{3cm}    % Expanding the titlebox
 
@@
 \begin{itemize}
-\item {\sc Single-Head}: each word has at most one parent
+\item {\sc Single-Head}: each word has $\leq 1$ parent
 \item {\sc Acyclic}: there are no directed cycles
@@
 \item {\sc Reachable}: each word can be reached from 0 by a directed path
+  \noteJE{note that this implies connected?
+or does connected not allow 0 words on the path, i.e., does it restrict 0 to one child, and if not, is there a principle for that?
+If no one, mention in footnote 1.}
 \item {\sc Planar}: edges may not ``cross'' \\ (if there are edges between $u,v$ and between $i,j$, where $u < i < v$, then $u \leq j \leq v$)
@@
 It produces undirected connected planar graphs.
-The annotation conventions are implicit in the detailed Link Grammar lexicon,%
+Annotation conventions are implicit in the detailed lexicon for the Link Grammar Parser,\noteJE{check caps}%
 \footnote{\url{http://www.abisource.com/projects/link-grammar/dict/introduction.html}.
@@
 Given a link grammar parse, it would be straightforward to convert it to an acyclic dependency parse by orienting all edges rightward.
-However, the result may violate the {\sc Reachability} constraint.
+However, the result may violate the {\sc Reachable} constraint.
 Instead we could orient all edges by depth-first search from the root node, which yields a DAG satisfying all our constraints.
-However, this might result in inconsistent annotation conventions, with some \texttt{S} links pointing from subject to verb and others from verb to subject.
+However, this might result in inconsistent annotation conventions, with some \texttt{S}-labeled links pointing from subject to verb and others from verb to subject.
 
@@
 We would expect verbs to point to their subject arguments in dependency grammar, and so we surmise that all \texttt{S} links are intended to point leftward (from verb to subject: {\em ``Jack $\stackrel{\texttt{S}}{\leftarrow}$ is falling''}).
-The link grammar designers take care to use a distinct \texttt{SI} label in cases of subject-verb inversion, and we surmise that \texttt{SI} links are intended to point rightward (again from verb to subject: {\em ``Jack $\stackrel{\texttt{SI}}{\rightarrow}$ is falling''}).
+The link grammar designers take care to use a distinct \texttt{SI} label in cases of subject-verb inversion, and we surmise that \texttt{SI} links are intended to point rightward (again from verb to subject: {\em ``Is $\stackrel{\texttt{SI}}{\rightarrow}$ Jack falling?''}).
 
@@
 This is a well-defined constraint programming problem.
-For example, to prevent cyclicity, we would require (roughly speaking) that no word type could follow a sequence of directed rightward links through other word types and then a leftward link back to itself.
+For example, to prevent cyclicity, we would require (roughly speaking) that no word type in the lexicon could follow a sequence of directed rightward links through other word types and then a leftward link back to itself.
 
@@
 
-\section{Integer Linear Programming}
-We formulate an ILP to find the minimal set of link type to directionality assignments that would directionalize all the links in a set of link grammar parses, while still mandating that all of the resulting dependency parses would be connected DAGs, and with all the nodes reachable from the root.
-Our task of finding the smallest set of assignments that would explain the parsesis also called model minimization\cite{ravi2009}.
+\section{Integer Linear Programming Model}
 
-For every encountered label we generate two boolean label/direction variables (e.g.
+For each undirected labeled edge $ij$ in the link corpus, where $i,j$ denote tokens of the same sentence with $i < j$, we introduce nonnegative integer variables $x_{ij}$ and $x_{ji}$ with a constraint $x_{ij}+x_{ji}=1$.
+We interpret $x_{ij}=1$ or $x_{ji}=1$ to mean that the link has direction $i \rightarrow j$ or $i \leftarrow j$, respectively.\footnote{In practice we halve the number of variables by replacing $x_{ji}$ with $1-x_{ij}$ for $j > i$, but that obscures the exposition.}
+
+% OLD VERSION
+% we introduce a variable $x_{ij}$ that is 0 or 1 according to whether $e$ points left or right.  We abbreviate $1-x_{ij}$ as $\bar{x}_{ij}$.
+
+For each non-0 token $v$, we ensure that it has at least one parent by constraining
+\begin{align}\label{eqn:oneparent}
+\sum_u x_{uv} & \geq 1
+\end{align}
+% OLD VERSION
+% $$\sum_{u < v} x_{uv} + \sum_{u > v} \bar{x}_{vu} \geq 1$$
+where $u$ ranges only over tokens such that the relevant variable exists.
+%
+To prevent cycles,\footnote{This also ensures {\sc reachable}, given \eqnref{oneparent}.} for each token $v$ we introduce a depth variable $d_v$ in the range $[0,n_v]$ (not constrained to be integer), where $n_v$ is the length of the sentence containing $v$.
+We require a child's depth to be at least 1 greater than each of its parents' depths---since this can be satisfied iff the sentence has no directed cycles:
+\begin{align}\label{eqn:nocycles}
+(\forall u)\; d_v + (1+n_v) \cdot (1-x_{uv}) & \geq 1+d_u
+% OLD VERSION
+% $$(\forall u < v) d_v + (1+n_v) \cdot \bar{x}_{uv} \geq 1+d_u$$
+% $$(\forall u > v) d_v + (1+n_v) \cdot x_{vu} \geq 1+d_u$$
+\end{align}
+The second summand ensures that \eqnref{nocycles} is trivially satisfied (hence has no effect) when $u$ is {\em not} the parent of $v$.
+As a speedup, we can require the 0 tokens to have depth 0.
+\noteJE{does this really still make a difference to speed?
+Possible better version: just add all the depths (perhaps multiplied by 0.001) to the objective function; this breaks ties in simplex.}
+
+\noteJE{what's the point of the root\_depth constraint in the .zpl file?  Almost trivially satisfied.}
+
+Finally, we encourage all links with the same label \noteJE{explain that we mean capital letters?} to have the same direction.
+For each label $L$, we introduce binary variables $r_v$ and $\ell_v$ which say that a link of type $L$ is ``allowed'' to point right or left, respectively.
+For each undirected edge $ij$ of label $L$, with $i < j$, we write
+\begin{align}
+x_{ij} &\leq r_L + s_{ij} & 
+x_{ji} &\leq \ell_L + s_{ij}
+\end{align}
+where $s_{ij} \geq 0$ is a slack variable that allows an edge token to point in a disallowed direction if needed to ensure \eqnref{oneparent}--\eqnref{nocycles}.
+
+Our objective tries to minimize the number of allowed directions (by link type---cf.\@ \newcite{ravi2009}) and the total slack (by link token):
+\begin{align}\label{eqn:obj}
+\min \left( \sum_{L} r_L + \ell_L \right) + \frac{100}{N_L} \cdot \sum_{ij} s_{ij}
+\end{align}
+where $N_L$ is the number of link tokens with label $L$.
+Objective \eqnref{obj} is willing to tolerate up to 1\% of those link tokens' using a disallowed direction before it prefers to allow {\em both} directions.
+% CUT FOR SPACE
+% (i.e., $r_L+\ell_L=2$).
+
+% OLD
+% We formulate an ILP to find the minimal set of link type to directionality assignments that would directionalize all the links in a set of link grammar parses, while still mandating that all of the resulting dependency parses would be connected DAGs, and with all the nodes reachable from the root.
+Our task of finding the smallest set of assignments that would explain the parses is also called model minimization\cite{ravi2009}.
+\noteJE{worth citing, but not quite the same thing}
+% For every encountered label we generate two boolean label/direction variables (e.g.
 label/left, label/right).
 Setting one of these variables to \textsc{true} allows the label type to only go in the specified direction, while setting both would allow the label to go in either way.
-The objective is to minimize the number of label/direction variables set to \textsc{true}, while still satisfying acyclicity and reachability constraints.
+% The objective is to minimize the number of label/direction variables set to \textsc{true}, while still satisfying acyclicity and reachability constraints.
 The encoding was written in the Zimpl little language \cite{Koch2004} and solved using the SCIP Optimization Suite \cite{achterberg2009scip}.
-
-\subsection{Slack}
-We introduced slack such that the link types were allowed to deviate from the majority one percent of the time before the ILP would assign both label/direction variables to \textsc{true}.
+% \subsection{Slack}
+% We introduced slack such that the link types were allowed to deviate from the majority one percent of the time before the ILP would assign both label/direction variables to \textsc{true}.
 This allows for flexibility against noise in the link parser's label assignments, balanced with allowing for both directions to be assigned if necessary.
@@
 \section{Data Sets}
-The data used for this work came from English bnews CoNLL dependency training data \noteJH{cite?} and the Russian National Corpus\footnote{\url{http://ruscorpora.ru/en/}}.
-With the English sentences we compared our link parse results to the original dependency annotations, and with the Russian sentences we only produced link parses.
+The data used for this work came from English bnews CoNLL dependency training data \noteJH{cite?} \noteJE{wait a minute, the bnews is not from CoNLL!
+It is merely in CoNLL format.
+It is from Ariya Rastrow, and is automatically parsed, I think.
+We should be comparing on hand-parsed data.}  and the Russian National Corpus\footnote{\url{http://ruscorpora.ru/en/}} \noteJE{give citation}.
+
+We generated link parses using the AbiWord/CMU link grammar parser version 5.0.8 \cite{LINKPARSER-2014}.
+The parser's coverage is not perfect: we obtained complete, connected parses for only !!!
+(of !!!) English sentences and !!!
+(of !!!) Russian sentences, discarding the other sentences.
+\noteJE{fill these !!!
+in automatically} These two languages have the most complete lexicons at present, although lexicons for !!!
+other languages are available.
+
+% With the English sentences we compared our link parse results to the original dependency annotations, and with the Russian sentences we only produced link parses.
 
 \subsection{Discarding Incomplete Link Parses}
+
+\NoteJE{ideally we could cut this section given what I wrote above.
+However, you seem to be saying that we did train on sentences with dropped words as long as the parses were connected.
+Didn't we decide to drop those sentences too?}
+
 Not all of the data was used for this work.
@@
 \section{Experiments and Results}
-We generated link parses using the AbiWord/CMU link grammar parser version 5.0.8 \cite{LINKPARSER-2014}.
+
+We solved our ILP problem using the SCIP Optimization Suite \cite{achterberg2009scip}, encoding it using the ZIMPL language \cite{Koch2004}.  
+\NoteJE{give runtime here}
 
@@
   \includegraphics[width=\linewidth, keepaspectratio=true]{figure/precision_recall.png}
-  \caption{\small Precision/recall of increasing runs of English and Russian sentences. The precision values converge while the recall values continue to grow throughout the data.}
+  \caption{\small Convergence to the results obtained on the largest run.  Precision relative to that run is very high even for small training sizes.  Recall grows as more link types are observed.}
 \end{figure}
@@
 We recorded the precision of whether the assignments in the smaller runs could be found in those of the largest run, and similarly the recall of whether the assignments in the largest run could be found in the smaller runs.
+\noteJE{this is not clear.
+are we comparing the direction table or the parses?
+If the parses, which parses?}
 
@@
 
-In order to compare 
+In the case of English, we had manually produced  single-headed 
+ CoNLL annotations \cite{!!!}.
+
 Using the sentences of the dependency corpus as comparison, we find that the link parses do not wholly subsume dependency parses, and that the undirected links match roughly three fourths of the arcs in the conll dependency corpus.
@@
 
-Of the sentences with complete parses, about \input{figure/conll_analysis_sentences_multiheaded}had at least one multiheaded word.
+Of the sentences with complete parses, about \input{figure/conll_analysis_sentences_multiheaded}
+\noteJE{why ``about''?}
+had at least one multiheaded word.
+
+\NoteJE{how often is the ``both'' option used?}
+
+\NoteJE{discuss differences!}
+
+\NoteJE{what do we say about Russian?}
 

### 2014-06-02 doc/LinksToDAG.tex
moved section on datasets
some clarifications etc.

@@
 
-\section{Integer Linear Programming Model}
-
-For each undirected labeled edge $ij$ in the link corpus, where $i,j$ denote tokens of the same sentence with $i < j$, we introduce nonnegative integer variables $x_{ij}$ and $x_{ji}$ with a constraint $x_{ij}+x_{ji}=1$.
-We interpret $x_{ij}=1$ or $x_{ji}=1$ to mean that the link has direction $i \rightarrow j$ or $i \leftarrow j$, respectively.\footnote{In practice we halve the number of variables by replacing $x_{ji}$ with $1-x_{ij}$ for $j > i$, but that obscures the exposition.}
-
-% OLD VERSION
-% we introduce a variable $x_{ij}$ that is 0 or 1 according to whether $e$ points left or right.  We abbreviate $1-x_{ij}$ as $\bar{x}_{ij}$.
-
-For each non-0 token $v$, we ensure that it has at least one parent by constraining
-\begin{align}\label{eqn:oneparent}
-\sum_u x_{uv} & \geq 1
-\end{align}
-% OLD VERSION
-% $$\sum_{u < v} x_{uv} + \sum_{u > v} \bar{x}_{vu} \geq 1$$
-where $u$ ranges only over tokens such that the relevant variable exists.
-%
-To prevent cycles,\footnote{This also ensures {\sc reachable}, given \eqnref{oneparent}.} for each token $v$ we introduce a depth variable $d_v$ in the range $[0,n_v]$ (not constrained to be integer), where $n_v$ is the length of the sentence containing $v$.
-We require a child's depth to be at least 1 greater than each of its parents' depths---since this can be satisfied iff the sentence has no directed cycles:
-\begin{align}\label{eqn:nocycles}
-(\forall u)\; d_v + (1+n_v) \cdot (1-x_{uv}) & \geq 1+d_u
-% OLD VERSION
-% $$(\forall u < v) d_v + (1+n_v) \cdot \bar{x}_{uv} \geq 1+d_u$$
-% $$(\forall u > v) d_v + (1+n_v) \cdot x_{vu} \geq 1+d_u$$
-\end{align}
-The second summand ensures that \eqnref{nocycles} is trivially satisfied (hence has no effect) when $u$ is {\em not} the parent of $v$.
-As a speedup, we can require the 0 tokens to have depth 0.
-\noteJE{does this really still make a difference to speed?
-Possible better version: just add all the depths (perhaps multiplied by 0.001) to the objective function; this breaks ties in simplex.}
-
-\noteJE{what's the point of the root\_depth constraint in the .zpl file?  Almost trivially satisfied.}
-
-Finally, we encourage all links with the same label \noteJE{explain that we mean capital letters?} to have the same direction.
-For each label $L$, we introduce binary variables $r_v$ and $\ell_v$ which say that a link of type $L$ is ``allowed'' to point right or left, respectively.
-For each undirected edge $ij$ of label $L$, with $i < j$, we write
-\begin{align}
-x_{ij} &\leq r_L + s_{ij} & 
-x_{ji} &\leq \ell_L + s_{ij}
-\end{align}
-where $s_{ij} \geq 0$ is a slack variable that allows an edge token to point in a disallowed direction if needed to ensure \eqnref{oneparent}--\eqnref{nocycles}.
-
-Our objective tries to minimize the number of allowed directions (by link type---cf.\@ \newcite{ravi2009}) and the total slack (by link token):
-\begin{align}\label{eqn:obj}
-\min \left( \sum_{L} r_L + \ell_L \right) + \frac{100}{N_L} \cdot \sum_{ij} s_{ij}
-\end{align}
-where $N_L$ is the number of link tokens with label $L$.
-Objective \eqnref{obj} is willing to tolerate up to 1\% of those link tokens' using a disallowed direction before it prefers to allow {\em both} directions.
-% CUT FOR SPACE
-% (i.e., $r_L+\ell_L=2$).
-
-% OLD
-% We formulate an ILP to find the minimal set of link type to directionality assignments that would directionalize all the links in a set of link grammar parses, while still mandating that all of the resulting dependency parses would be connected DAGs, and with all the nodes reachable from the root.
-Our task of finding the smallest set of assignments that would explain the parses is also called model minimization\cite{ravi2009}.
-\noteJE{worth citing, but not quite the same thing}
-% For every encountered label we generate two boolean label/direction variables (e.g.
-label/left, label/right).
-Setting one of these variables to \textsc{true} allows the label type to only go in the specified direction, while setting both would allow the label to go in either way.
-% The objective is to minimize the number of label/direction variables set to \textsc{true}, while still satisfying acyclicity and reachability constraints.
-The encoding was written in the Zimpl little language \cite{Koch2004} and solved using the SCIP Optimization Suite \cite{achterberg2009scip}.
-% \subsection{Slack}
-% We introduced slack such that the link types were allowed to deviate from the majority one percent of the time before the ILP would assign both label/direction variables to \textsc{true}.
-This allows for flexibility against noise in the link parser's label assignments, balanced with allowing for both directions to be assigned if necessary.
-
 \section{Data Sets}
@@
 
+\section{Integer Linear Programming Model}
+
+For each undirected labeled edge $ij$ in the link corpus, where $i,j$ denote tokens in the same sentence with $i < j$, we introduce nonnegative integer variables $x_{ij}$ and $x_{ji}$ with a constraint $x_{ij}+x_{ji}=1$.
+We interpret $x_{ij}=1$ or $x_{ji}=1$ to mean that the link has direction $i \rightarrow j$ or $i \leftarrow j$, respectively.\footnote{In practice we halve the number of variables by replacing $x_{ji}$ with $1-x_{ij}$ for $j > i$, but that obscures the exposition.}
+
+% OLD VERSION
+% we introduce a variable $x_{ij}$ that is 0 or 1 according to whether $e$ points left or right.  We abbreviate $1-x_{ij}$ as $\bar{x}_{ij}$.
+
+For each non-0 token $v$, we ensure that it has at least one parent by constraining\footnote{To denote two linked tokens, we use variables $i,j$ when $i$ is to the left of $j$, or variables $u,v$ when $u$ is the parent of $v$.}
+\begin{align}\label{eqn:oneparent}
+\sum_u x_{uv} & \geq 1
+\end{align}
+% OLD VERSION
+% $$\sum_{u < v} x_{uv} + \sum_{u > v} \bar{x}_{vu} \geq 1$$
+where $u$ ranges only over tokens such that the relevant variable exists.
+%
+To prevent cycles,\footnote{This also ensures {\sc reachable}, given \eqnref{oneparent}.} for each token $v$ we introduce a depth variable $d_v$ in the range $[0,n_v]$ (not constrained to be integer), where $n_v$ is the length of the sentence containing $v$.
+We require a child's depth to be at least 1 greater than each of its parents' depths---constraints that can be satisfied iff the sentence has no directed cycles:
+\begin{align}\label{eqn:nocycles}
+(\forall u)\; d_v + (1+n_v) \cdot (1-x_{uv}) & \geq 1+d_u
+% OLD VERSION
+% $$(\forall u < v) d_v + (1+n_v) \cdot \bar{x}_{uv} \geq 1+d_u$$
+% $$(\forall u > v) d_v + (1+n_v) \cdot x_{vu} \geq 1+d_u$$
+\end{align}
+The second summand ensures that \eqnref{nocycles} is trivially satisfied (hence has no effect) when $u$ is {\em not} the parent of $v$.
+As a speedup, we can require the 0 tokens to have depth 0.
+\noteJE{does this really still make a difference to speed?
+Possible better version: just add all the depths (perhaps multiplied by 0.001) to the objective function; this breaks ties in simplex.}
+
+\noteJE{what's the point of the root\_depth constraint in the .zpl file?  Almost trivially satisfied.}
+
+Finally, we encourage all links with the same label \noteJE{explain that we mean capital letters?} to have the same direction.
+For each label $L$, we introduce binary variables $r_L$ and $\ell_L$, which say whether a link of type $L$ is ``allowed'' to point right or left, respectively.
+For each undirected edge $ij$ of label $L$, with $i < j$, we write
+\begin{align}
+x_{ij} &\leq r_L + s_{ij} & 
+x_{ji} &\leq \ell_L + s_{ij}
+\end{align}
+where $s_{ij} \geq 0$ is a slack variable that allows an edge token to point in a disallowed direction if needed to ensure \eqnref{oneparent}--\eqnref{nocycles}.
+
+Our objective tries to minimize the number of allowed directions (by link type---cf.\@ \newcite{ravi2009}) and the total slack (by link token):
+\begin{align}\label{eqn:obj}
+\min \left( \sum_{L} r_L + \ell_L \right) + \frac{100}{N_L} \cdot \sum_{ij} s_{ij}
+\end{align}
+where $N_L$ is the number of link tokens with label $L$.
+Objective \eqnref{obj} is willing to tolerate up to 1\% of those link tokens' using a disallowed direction before it prefers to allow {\em both} directions.
+% CUT FOR SPACE
+% (i.e., $r_L+\ell_L=2$).
+
+% OLD
+% We formulate an ILP to find the minimal set of link type to directionality assignments that would directionalize all the links in a set of link grammar parses, while still mandating that all of the resulting dependency parses would be connected DAGs, and with all the nodes reachable from the root.
+Our task of finding the smallest set of assignments that would explain the parses is also called model minimization\cite{ravi2009}.
+\noteJE{worth citing, but not quite the same thing}
+% For every encountered label we generate two boolean label/direction variables (e.g.
+label/left, label/right).
+Setting one of these variables to \textsc{true} allows the label type to only go in the specified direction, while setting both would allow the label to go in either way.
+% The objective is to minimize the number of label/direction variables set to \textsc{true}, while still satisfying acyclicity and reachability constraints.
+The encoding was written in the Zimpl little language \cite{Koch2004} and solved using the SCIP Optimization Suite \cite{achterberg2009scip}.
+% \subsection{Slack}
+% We introduced slack such that the link types were allowed to deviate from the majority one percent of the time before the ILP would assign both label/direction variables to \textsc{true}.
+This allows for flexibility against noise in the link parser's label assignments, balanced with allowing for both directions to be assigned if necessary.
+
 \section{Experiments and Results}
@@
 Because the precision values were high, we found that the solutions of the previous runs were consistent with the largest run, and as the data increased so did the recall values, and so we would continue to encounter novel link types and then incorporate them.
+
+\NoteJE{how often is the ``both'' option used?  and how much slack do we have?}
 
@@
 
-we measure whether the direction we chose is the same as the original dependency data. 
-Finally, we look to see that given a link label, whether the original conll label can be recovered.
+we measure whether the direction we chose is the same as the original dependency data.  Finally, we look to see that given a link label, whether the original conll label can be recovered.
 
@@
 
-\NoteJE{how often is the ``both'' option used?}
-
-\NoteJE{discuss differences!}
+\NoteJE{discuss linguistic differences!}
 

### 2014-06-02 src/tools/experiments/conll_analysis.py
small merge with the paper. working on the conll_analysis.py scripts.

@@ -1,6 +1,6 @@
 #!/usr/bin/python
 
-import sys, os
+import sys, os, operator 
 from pprint import pprint
 from tikz_dependency import *
 from dependency_graph import *
@@ -193,6 +193,14 @@ def analysis_words(conlls,links):
     #dropped_data = (dropped_word_count, hasDropped)
     #return (dropped_data, multiheaded_data, word_count)
     return (multiheaded_data, word_count)
+
+
+
+
+# Analysis on the links. For the appendix of the paper
+def analysis_links(links):
+    print links
+
 
 
 
@@ -485,30 +493,20 @@ for (label_conll, label_link) in mismatch_directionality_counts:
     
     
 
-#pprint(link_label_prediction)
+pprint(link_label_prediction)
 
 predictions = []
 
-import operator 
 for label_link in link_label_prediction:
     conll_map = link_label_prediction[label_link]
     predictions.append((label_link, max(conll_map.iteritems(), key=operator.itemgetter(1))[0]))
 
 #predictions.sort()
-
-#print
-#print
 #pprint(predictions)
 
 
 
-#print "ALL EXTRA COUNTS"
-#pprint(all_extra_counts)
-#print
 
-#print "ALL BLANK COUNTS"
-#pprint(all_blank_counts)
-#print
 
 
 

### 2014-06-02 src/tools/experiments/conll_analysis_output.py

@@ -209,11 +209,13 @@ def result_latex_corpus(original_sentence_count,
     result = ""
     result += "\t\\begin{tabular}{|l|l|}\n"
     result += "\t\t\\hline\n"
-    result += "\t\tNumber of English sentences used from the corpus & "+str(original_sentence_count)+"\\\\ \n"
+    result += "\t & Sentences \\\\ \n"
     result += "\t\t\\hline\n"
-    result += "\t\tSentences after discarding disconnected parses & "+str(link_sentence_total)+"\\\\ \n"
+    result += "\t\tSentences used from the corpus & "+str(original_sentence_count)+"\\\\ \n"
     result += "\t\t\\hline\n"
-    result += "\t\tSentences with complete parses used for analysis & "+str(conll_analysis_sentence_count)+"\\\\ \n"
+    result += "\t\tAfter discarding disconnected parses & "+str(link_sentence_total)+"\\\\ \n"
+    result += "\t\t\\hline\n"
+    result += "\t\tComplete parses used for analysis & "+str(conll_analysis_sentence_count)+"\\\\ \n"
     result += "\t\t\\hline\n"
     result += "\t\\end{tabular}\n"
 

### 2014-06-02 trials.sh

@@ -3,9 +3,9 @@
 typeset -i i END
 
 
-#END=10000
+END=4096
 #END=60030
-END=60030
+#END=16384
 #INCREMENT=1
 
 SENTENCES="data/english_bnews_train.sentences"
@@ -92,7 +92,7 @@ fi
 
 
 
-echo $RUNTIMES $END $'\n'The CLSP machines cant run plotter. Please download files and run: $'\n'python src/tools/plotter.py -x "Sentences" -y "Percent" -o $SOL_PRECISION_RECALL $TXT_PRECISION $TXT_RECALL | mutt -s $END -a $RUNTIMES -a $TXT_PRECISION -a $TXT_RECALL -- junekihong@gmail.com
+#echo $RUNTIMES $END $'\n'The CLSP machines cant run plotter. Please download files and run: $'\n'python src/tools/plotter.py -x "Sentences" -y "Percent" -o $SOL_PRECISION_RECALL $TXT_PRECISION $TXT_RECALL | mutt -s $END -a $RUNTIMES -a $TXT_PRECISION -a $TXT_RECALL -- junekihong@gmail.com
 
 #-a $TYPE_ANALYSIS -a $CONLL_ANALYSIS -- junekihong@gmail.com
 

### 2014-06-02 src/tools/experiments/conll_analysis.py
Added a table to the appendix section. The English Link Types.

@@ -199,7 +199,24 @@ def analysis_words(conlls,links):
 
 # Analysis on the links. For the appendix of the paper
 def analysis_links(links):
-    print links
+    link_direction_counts = {}
+    
+    for link in links:
+        link = link.split()
+        index = int(link[0])
+        heads = link[6].split(",")
+        labels = link[7].split(",")
+
+        for (head,label) in zip(heads,labels):
+            head = int(head)
+
+            link_direction_counts[label] = link_direction_counts.get(label,{})
+            if head < index:
+                link_direction_counts[label]["left"] = link_direction_counts[label].get("left",0) + 1
+            elif head > index:
+                link_direction_counts[label]["right"] = link_direction_counts[label].get("right",0) + 1
+
+    return link_direction_counts
 
 
 
@@ -310,6 +327,8 @@ mismatch_directionality_counts = {}
 #mismatch_extra_total = 0
 #mismatch_extra_counts = {}
 
+link_direction_totals = {}
+
 # --------------------------------------------------------------------
 # Variables for word analysis
 # --------------------------------------------------------------------
@@ -361,6 +380,14 @@ for linkSentence in link_results:
         example_parses_count += 1
         if example_parses_count % 3 == 0:
             EXAMPLE_PARSES.write("\clearpage")
+
+
+    # analysis_links
+    link_direction_counts = analysis_links(link_results[linkSentence])
+    for label in link_direction_counts:
+        link_direction_totals[label] = link_direction_totals.get(label,{})
+        for direction in link_direction_counts[label]:
+            link_direction_totals[label][direction] = link_direction_totals[label].get(direction, 0) + link_direction_counts[label][direction]
 
 
     (match_data, reverse_data, extra_data, word_data, mismatch_data, total) = analysis(conll_results[linkSentence], link_results[linkSentence])
@@ -480,6 +507,14 @@ f.close()
 #print
 
 
+
+
+
+
+#---------------------------------------------------------------------
+# Setting up the matching link label counts and majority predictions
+# This will be used in a table of the appendix of the paper
+#---------------------------------------------------------------------
 link_label_prediction = {}
 for (label_conll, label_link) in all_match_counts:
     link_label_prediction[label_link] = link_label_prediction.get(label_link, {})
[…]

### 2014-06-02 src/link_parser.py
found an option in link_parser to turn off null links. I thought this would take a while, but only turned out to be about 30 seconds of work.

@@ -45,7 +45,7 @@ class linkParser:
 
         # Time out the link parser. Don't even need to bother with its internal timeout.
         timeout = str(30)
-        linkparser_process = Popen(["timeout", timeout, "link-parser", self.language, "-!graphics=0", "-!links=0", "-!echo=1", "-!postscript=1", "-!panic=0"], stdin=echo_process.stdout, stdout=PIPE, stderr=subprocess.PIPE)
+        linkparser_process = Popen(["timeout", timeout, "link-parser", self.language, "-!graphics=0", "-!links=0", "-!echo=1", "-!postscript=1", "-!panic=0", "-!null"], stdin=echo_process.stdout, stdout=PIPE, stderr=subprocess.PIPE)
         echo_process.stdout.close()
 
         output, err = linkparser_process.communicate()        

### 2014-06-02 src/tools/experiments/conll_analysis.py
small fix to conll_analysis.py

@@ -243,6 +243,8 @@ for line in f:
             ID = " ".join(sentence)
             ID = ID.lower()
             link_results[ID] = tuple(conll)
+
+
         
         sentence = []
         conll = []
@@ -260,6 +262,14 @@ for line in f:
         index = word.rfind("[!]")
         if index != -1:
             word = word[:(index)]
+
+        """index = -1
+        index = word.rfind("[?]")
+        if index != -1:
+            word = word[:index]+word[index+len("[?]"):]
+        print word"""
+
+
 
         # Get rid of the [] brackets around words that the link-parser could not attach.        
         #word = word.strip("[]")
@@ -282,6 +292,8 @@ f = f.readlines()
 sentence = []
 conll = []
 
+
+
 for line in f:
     line = line.strip()
 
@@ -289,12 +301,14 @@ for line in f:
         # The sentence is stored as lower case, because link-parser will sometimes lowercase the words that it does not know.
         ID = " ".join(sentence)
         ID = ID.lower()
+        
+
 
         # Check to see if the sentence is in the link data. 
         # This is a small optimization. We don't need to store all of the conll data. 
         if ID in link_results:
             conll_results[ID] = tuple(conll)
-        
+
         sentence = []
         conll = []
     else:
@@ -351,11 +365,16 @@ example_parses_limit = 100
 
 
 for linkSentence in link_results:
+    if linkSentence not in conll_results:
+        continue
+
+
     sentenceCheck = True
     linkSentenceCheck = linkSentence.split()
     # I am preventing certain sentences from appearing in the paper.
     # I just wanted to skip over some sentences because they didn't look cool enough
-    bannedWords = ["salees", "soldiers", "word", "serwer", "reason"]
+    #bannedWords = ["salees", "soldiers", "word", "serwer", "reason"]
+    bannedWords = []
     for word in bannedWords:
         if word in linkSentenceCheck:
             sentenceCheck = False
@@ -388,6 +407,7 @@ for linkSentence in link_results:
         link_direction_totals[label] = link_direction_totals.get(label,{})
         for direction in link_direction_counts[label]:
             link_direction_totals[label][direction] = link_direction_totals[label].get(direction, 0) + link_direction_counts[label][direction]
+
 
 
     (match_data, reverse_data, extra_data, word_data, mismatch_data, total) = analysis(conll_results[linkSentence], link_results[linkSentence])

### 2014-06-02 src/sentlen.sh
added sentlen.sh

@@ -1,0 +1,3 @@
+#!/bin/bash
+# Makes a file that maps every sentence ID to that sentence's length.
+cat LinksToDAG_links.txt | perl -ne 'chomp; @f = split(", *"); $len{$f[4]}=$f[0] if $len{$f[4]}<$f[0]; $len{$f[4]}=$f[1] if $len{$f[4]}<$f[1]; END { for $sent (keys %len) { print "$sent, $len{$sent}\n"; } }' > LinksToDAG_sentlen.txt

### 2014-06-02 src/tools/experiments/tikz_dependency.py
added Jasons sentlen script to the pipeline. Got it hooked up to the ILP solver.

@@ -28,6 +28,7 @@ def tikz_dependency(conlls, links, sentence, ratio = 0.3, subfigure = True):
     for POS in graph_conll.POS_sequence:
         POS_conll.append("{\\scriptsize "+POS+"}")
 
+    graph_link.sentence = (" ".join(graph_link.sentence)).replace("&", "\\&").replace("$", "\\$").split()
     link_sentence = " \& ".join(graph_link.sentence).replace("[", "\\lbrack ").replace("]", "\\rbrack")
     POS_sequence = " \& ".join(graph_link.POS_sequence).replace("$", "\\$")
 
@@ -46,7 +47,7 @@ def tikz_dependency(conlls, links, sentence, ratio = 0.3, subfigure = True):
  
             if head in graph_link.table and child in graph_link.table[head]:
                 result += tikz_draw_edge(head, child, graph_link.getEdge(head,child), "below", "thick")
-                result += tikz_draw_edge(head, child, label, "above", "thick")
+                result += tikz_draw_edge(head, child, "\\small "+label, "above", "thick")
                 graph_link.deleteEdge(head,child)
 
                 #multiheads = graph_link.heads.get(child, [])
@@ -57,11 +58,11 @@ def tikz_dependency(conlls, links, sentence, ratio = 0.3, subfigure = True):
             
             elif child in graph_link.table and head in graph_link.table[child]:
                 result += tikz_draw_edge(child, head, graph_link.getEdge(child,head), "below", "ultra thick")
-                result += tikz_draw_edge(head, child, label, "above", "thick")
+                result += tikz_draw_edge(head, child, "\\small "+label, "above", "thick")
                 graph_link.deleteEdge(child,head)
             
             else:
-                result += tikz_draw_edge(head, child, label, "above", "dotted")
+                result += tikz_draw_edge(head, child, "\\small "+label, "above", "dotted")
 
 
 

### 2014-06-02 src/tools/experiments/tikz_dependency.py
added extra handling of special characters when making the parse visualizations.

@@ -28,7 +28,7 @@ def tikz_dependency(conlls, links, sentence, ratio = 0.3, subfigure = True):
     for POS in graph_conll.POS_sequence:
         POS_conll.append("{\\scriptsize "+POS+"}")
 
-    graph_link.sentence = (" ".join(graph_link.sentence)).replace("&", "\\&").replace("$", "\\$").split()
+    graph_link.sentence = (" ".join(graph_link.sentence)).replace("&", "\\&").replace("$", "\\$").replace("%","\\%").split()
     link_sentence = " \& ".join(graph_link.sentence).replace("[", "\\lbrack ").replace("]", "\\rbrack")
     POS_sequence = " \& ".join(graph_link.POS_sequence).replace("$", "\\$")
 

### 2014-06-03 doc/LinksToDAG.tex
small merge with the paper. working on the conll_analysis.py scripts.
Added a table to the appendix section. The English Link Types.
Changed the table in the appendix to use coarse link labels. Made the table smaller.
updated the paper

@@
 \begin{itemize}
+
 \item {\sc Single-Head}: each word has $\leq 1$ parent
+\item {\sc Single-Head}: each word has at most one parent
 \item {\sc Acyclic}: there are no directed cycles
@@
 \section{Data Sets}
-The data used for this work came from English bnews CoNLL dependency training data \noteJH{cite?} \noteJE{wait a minute, the bnews is not from CoNLL!
-It is merely in CoNLL format.
-It is from Ariya Rastrow, and is automatically parsed, I think.
-We should be comparing on hand-parsed data.}  and the Russian National Corpus\footnote{\url{http://ruscorpora.ru/en/}} \noteJE{give citation}.
+The data used for this work came from English Penn Tree Bank CoNLL dependency training data \noteJH{cite?} 
+%\noteJE{wait a minute, the bnews is not from CoNLL!  It is merely in CoNLL format.  It is from Ariya Rastrow, and is automatically parsed, I think.  We should be comparing on hand-parsed data.}  
+and the Russian National Corpus\footnote{\url{http://ruscorpora.ru/en/}} \cite{russiancorpus} \noteJE{give citation}.
 
 We generated link parses using the AbiWord/CMU link grammar parser version 5.0.8 \cite{LINKPARSER-2014}.
-The parser's coverage is not perfect: we obtained complete, connected parses for only !!!
-(of !!!) English sentences and !!!
-(of !!!) Russian sentences, discarding the other sentences.
-\noteJE{fill these !!!
-in automatically} These two languages have the most complete lexicons at present, although lexicons for !!!
-other languages are available.
+The parser's coverage is not perfect: we obtained complete, connected parses for only \input{figure/conll_analysis_sentences_total.tex}of (\input{figure/conll_analysis_sentences_original.tex}) English sentences and only 206 (of 10,000) Russian sentences, discarding the other sentences.
+%\noteJE{fill these !!! in automatically} 
+These two languages have the most complete lexicons at present, although lexicons for seven other languages are available.
 
@@
 
-\subsection{Discarding Incomplete Link Parses}
+%\subsection{Discarding Incomplete Link Parses}
 
-\NoteJE{ideally we could cut this section given what I wrote above.
+%\NoteJE{ideally we could cut this section given what I wrote above.
 However, you seem to be saying that we did train on sentences with dropped words as long as the parses were connected.
@@
 
-Not all of the data was used for this work.
+%Not all of the data was used for this work.
 From the given sentences, we discarded those that the link parser could not process and returned a parse with nodes disconnected from the root.
@@
 
-For our analysis of these parses, we further discarded those that had dropped words, where the link parser could not attach links to every word in the sentence.
+%For our analysis of these parses, we further discarded those that had dropped words, where the link parser could not attach links to every word in the sentence.
 About \input{figure/conll_analysis_sentences_dropped}of the non-disconnected sentences were removed, leaving us working only with complete link parses.
 
-\begin{figure}[ht!]
-  \centering
-  \small
-  \input{figure/conll_analysis_sentences.tex}
-  \caption{\small The English sentences used for this work.}
-\end{figure}
+%\begin{figure}[ht!]
+%  \centering
+%  \small
+%  \input{figure/conll_analysis_sentences.tex}
+%  \caption{\small The English sentences used for this work.}
+%\end{figure}
 
-\noteJH{The stability results come from the second row of this table. The analysis later in the paper uses the third row. Putting this section here might be confusing to readers?}
+%\noteJH{The stability results come from the second row of this table. The analysis later in the paper uses the third row. Putting this section here might be confusing to readers?}
 
@@
 \section{Experiments and Results}
+\begin{figure*}[ht!]
+  \centering
+  \input{figure/sentences.tikz}
+  \caption{\small Example Parses: The top blue arcs represent conll dependencies; the bottom red arcs represent directionalized link grammar parses. See the Appendix for more details.}
+  \label{fig:parses}
+\end{figure*}
 
 We solved our ILP problem using the SCIP Optimization Suite \cite{achterberg2009scip}, encoding it using the ZIMPL language \cite{Koch2004}.  
-\NoteJE{give runtime here}
+\NoteJE{give runtime here}\NoteJH{A run of 8,192 sentences takes 1743 seconds}
 
@@
 With increasing trials of sentences, we measure how similar the subsequent directional assignments would be with each other.
-Taking the largest run as the answer key, we compared how much the smaller runs deviated from it.
+Taking the direction table of the largest run as the answer key, we compared how much the solutions of the smaller runs deviated from it.
 We recorded the precision of whether the assignments in the smaller runs could be found in those of the largest run, and similarly the recall of whether the assignments in the largest run could be found in the smaller runs.
-\noteJE{this is not clear.
-are we comparing the direction table or the parses?
-If the parses, which parses?}
+%\noteJE{this is not clear.  are we comparing the direction table or the parses?  If the parses, which parses?}
 
@@
 \subsection{English Link-Dependency Analysis}
-
-\begin{figure*}[ht!]
-  \centering
-  \input{figure/sentences.tikz}
-  \caption{\small Example Parses: The top blue arcs represent conll dependencies; the bottom red arcs represent directionalized link grammar parses. See the Appendix for more details.}
-  \label{fig:parses}
-\end{figure*}
 
@@
 
-Of the sentences with complete parses, about \input{figure/conll_analysis_sentences_multiheaded}
-\noteJE{why ``about''?}
-had at least one multiheaded word.
+Of the English sentences with complete parses, \input{figure/conll_analysis_sentences_multiheaded}had at least one multiheaded word.
 
@@
 
+\input{figure/link_analysis_coarse_table.tex}
+
 \section{Sample English Parses}

### 2014-06-03 doc/LinksToDAG.tex
minor edits

--- 2014-06-02 doc/LinksToDAG.tex
@@
 \usepackage{xspace}
+\usepackage{enumitem}
 
@@
 As recently discussed by \newcite{gomezrodriguez-nivre-2013}, one might impose various requirements on the parse graph:
-\begin{itemize}
+\begin{itemize}[noitemsep]
 \item {\sc Single-Head}: each word has $\leq 1$ parent
@@
 If no one, mention in footnote 1.}
-\item {\sc Planar}: edges may not ``cross'' \\ (if there are edges between $u,v$ and between $i,j$, where $u < i < v$, then $u \leq j \leq v$)
+\item {\sc Planar}: edges may not ``cross'' \\ (if there are edges between $i,j$ and between $u,v$, where $i < u < j$, then $i \leq v \leq j$)
 \end{itemize}
@@
 For example, in {\em ``The boy that Jill skipped with fell down,''} the word {\em with} uses a lexical entry that requires it to link to a governing verb to the left, an extracted object farther to the left, and nothing to the right.
+Each entry has a hand-assigned cost in \{0,1,2\} and the parser finds the parse of minimum total cost.
+% Information about cost is in this message from Linas Vepstas on 2014-02-01: 
+%   https://groups.google.com/d/msg/link-grammar/eeJw1Ofgc9U/diqPYSwuFfoJ
 
@@
 
+\input{figure/link_analysis_table.tex}
+
 \section{Sample English Parses}

### 2014-06-03 doc/LinksToDAG.tex
very minor

@@
 For our analysis of these parses, we further discarded those that had dropped words, where the link parser could not attach links to every word in the sentence.
-About \input{figure/conll_analysis_sentences_dropped}of the non-disconnected sentences were removed, leaving us working only with complete link parses.
+About \input{figure/conll_analysis_sentences_dropped} of the non-disconnected sentences were removed, leaving us working only with complete link parses.
 

### 2014-06-03 doc/LinksToDAG.tex

--- 2014-06-03 doc/LinksToDAG.tex
@@
 As recently discussed by \newcite{gomezrodriguez-nivre-2013}, one might impose various requirements on the parse graph:
-\begin{itemize}
-
+\begin{itemize}[noitemsep]
 \item {\sc Single-Head}: each word has $\leq 1$ parent
-\item {\sc Single-Head}: each word has at most one parent
 \item {\sc Acyclic}: there are no directed cycles
@@
 %For our analysis of these parses, we further discarded those that had dropped words, where the link parser could not attach links to every word in the sentence.
-About \input{figure/conll_analysis_sentences_dropped}of the non-disconnected sentences were removed, leaving us working only with complete link parses.
+About \input{figure/conll_analysis_sentences_dropped} of the non-disconnected sentences were removed, leaving us working only with complete link parses.
 

### 2014-06-03 doc/LinksToDAG.tex
clarified Russian data

@@
 \section{Data Sets}
-The data used for this work came from English Penn Tree Bank CoNLL dependency training data \noteJH{cite?} 
-%\noteJE{wait a minute, the bnews is not from CoNLL!  It is merely in CoNLL format.  It is from Ariya Rastrow, and is automatically parsed, I think.  We should be comparing on hand-parsed data.}  
-and the Russian National Corpus\footnote{\url{http://ruscorpora.ru/en/}} \cite{russiancorpus} \noteJE{give citation}.
+We used the English sentences from the CoNLL 2007 Shared Task \cite{CONLL-SHARED-2007}, a subset of the Penn Treebank for which {\em single}-headed reference parses are available.
+We also used a prefix of the Russian National Corpus.\footnote{That is, a prefix when considering the files in Unix sort order by filename.
+We dropped sentences with $\leq 2$ words, as well as file 22592319, which appeared to contain charts.}
 
 We generated link parses using the AbiWord/CMU link grammar parser version 5.0.8 \cite{LINKPARSER-2014}.
-The parser's coverage is not perfect: we obtained complete, connected parses for only \input{figure/conll_analysis_sentences_total.tex}of (\input{figure/conll_analysis_sentences_original.tex}) English sentences and only 206 (of 10,000) Russian sentences, discarding the other sentences.
-%\noteJE{fill these !!! in automatically} 
-These two languages have the most complete lexicons at present, although lexicons for seven other languages are available.
+The parser's coverage is not perfect: we obtained connected parses for only \input{figure/conll_analysis_sentences_total.tex}(of \input{figure/conll_analysis_sentences_original.tex}$\!\!$) English sentences and only 8,018 (of 10,000) Russian sentences, discarding the other sentences.\footnote{For Russian only, we permitted the parser to drop words if necessary.
+Otherwise only 208 sentences were parseable.}
+These two languages have the most complete lexicons at present, although lexicons for 7 other languages are available.
 

### 2014-06-03 src/tools/experiments/conll_analysis.py
updated conll_analysis.py. Removed words with [!] in the example sentences. Added rounding to the figures in the appendix table at the end.

@@ -371,8 +371,14 @@ for linkSentence in link_results:
         continue
 
 
+    # Remove sentences with [!] or [?] in them. I will have them later in the paper and explain what it means.
     sentenceCheck = True
-    linkSentenceCheck = linkSentence.split()
+    for line in link_results[linkSentence]:
+        line = line.split()
+        word = line[1]
+        if word.rfind("[!]") != -1 or word.rfind("[?]") != -1:
+            sentenceCheck = False
+            
     # I am preventing certain sentences from appearing in the paper.
     # I just wanted to skip over some sentences because they didn't look cool enough
     #bannedWords = ["salees", "soldiers", "word", "serwer", "reason"]
@@ -665,7 +671,7 @@ for label in labels:
         prediction_num = link_label_prediction[label][prediction]
         prediction_total = link_label_prediction_totals[label]
     if prediction_num and prediction_total:
-        prediction_percent = str(int(float(prediction_num) / prediction_total * 100)) + "\\%" + " (" + str(prediction_num) + "/" + str(prediction_total) + ")"
+        prediction_percent = str(int(float(prediction_num) / prediction_total * 100 + 0.5)) + "\\%" + " (" + str(prediction_num) + "/" + str(prediction_total) + ")"
 
 
     table += "\\hline\n"
@@ -726,7 +732,7 @@ for coarse_label in coarse_labels:
         prediction_num = link_label_coarse_prediction[coarse_label][prediction]
         prediction_total = link_label_coarse_prediction_totals[coarse_label]
     if prediction_num and prediction_total:
-        prediction_percent = str(int(float(prediction_num) / prediction_total * 100)) + "\\%" + " (" + str(prediction_num) + "/" + str(prediction_total) + ")"
+        prediction_percent = str(int(float(prediction_num) / prediction_total * 100 + 0.5)) + "\\%" + " (" + str(prediction_num) + "/" + str(prediction_total) + ")"
 
 
     table += "\\hline\n"

### 2014-06-03 src/tools/experiments/conll_analysis.py
updated the table in the appendix

@@ -201,12 +201,20 @@ def analysis_words(conlls,links):
 # Analysis on the links. For the appendix of the paper
 def analysis_links(links):
     link_direction_counts = {}
-    
+    link_label_multiheaded = {}
+
     for link in links:
         link = link.split()
         index = int(link[0])
         heads = link[6].split(",")
         labels = link[7].split(",")
+        
+        if len(heads) > 1:
+            for label in labels:
+                coarse_label = get_coarse_label(label)
+
+                link_label_multiheaded[coarse_label] = link_label_multiheaded.get(coarse_label, 0) + 1
+
 
         for (head,label) in zip(heads,labels):
             head = int(head)
@@ -217,7 +225,18 @@ def analysis_links(links):
             elif head > index:
                 link_direction_counts[label]["right"] = link_direction_counts[label].get("right",0) + 1
 
-    return link_direction_counts
+
+    return (link_direction_counts, link_label_multiheaded)
+
+
+
+# Gets the coarse label for any given specific label
+def get_coarse_label(label):
+    coarse_label = "".join(w for w in list(label) if w.isupper())    
+    if len(list(coarse_label)) >= 2 and list(coarse_label)[0] == "I" and list(coarse_label)[1] == "D":
+        coarse_label = "ID"
+    return coarse_label
+
 
 
 
@@ -344,6 +363,7 @@ mismatch_directionality_counts = {}
 
 link_direction_totals = {}
 link_direction_coarse_totals = {}
+link_label_isMultiheaded = {}
 
 # --------------------------------------------------------------------
 # Variables for word analysis
@@ -410,9 +430,9 @@ for linkSentence in link_results:
 
 
     # analysis_links
-    link_direction_counts = analysis_links(link_results[linkSentence])
+    (link_direction_counts, link_label_multiheaded) = analysis_links(link_results[linkSentence])
     for label in link_direction_counts:
-        coarse_label = "".join(w for w in list(label) if w.isupper())
+        coarse_label = get_coarse_label(label)
 
         link_direction_coarse_totals[coarse_label] = link_direction_coarse_totals.get(coarse_label,{})
         link_direction_totals[label] = link_direction_totals.get(label,{})
@@ -420,6 +440,11 @@ for linkSentence in link_results:
[…]

### 2014-06-03 doc/LinksToDAG.tex
a bunch of clarification, rearrangement, and deletion in the discussion of results

@@
 \usepackage{enumitem}
+\usepackage[small]{caption}
 
@@
 
+% ADDRESSED BY ABOVE?
+% \NoteJE{We should acknowledge that another option would be to post-process either CoNLL parses or our link parses to automatically add more multi-heading, e.g., to handle control phenomena ("John wants to skip" or "Jane ran for office in order to change the tax laws" should have John/Jane be the subject of two verbs).
+However, that requires per-language}
+
 All of these options result in non-projective parses, so the parsers must use non-projective or pseudo-projective algorithms.
@@
 \section{Data Sets}
-We used the English sentences from the CoNLL 2007 Shared Task \cite{CONLL-SHARED-2007}, a subset of the Penn Treebank for which {\em single}-headed reference parses are available.
-We also used a prefix of the Russian National Corpus.\footnote{That is, a prefix when considering the files in Unix sort order by filename.
-We dropped sentences with $\leq 2$ words, as well as file 22592319, which appeared to contain charts.}
+We used the English sentences from the CoNLL 2007 Shared Task \cite{CONLL-SHARED-2007}---a subset of the Penn Treebank for which {\em single}-headed reference parses are available.
+We also used a prefix of the Russian National Corpus,\footnote{That is, a prefix when considering the files in Unix sort order by filename.
+We dropped sentences with $\leq 2$ words, as well as file 22592319, which appeared to contain charts.} which is unparsed.
 
@@
 Otherwise only 208 sentences were parseable.}
-These two languages have the most complete lexicons at present, although lexicons for 7 other languages are available.
+These two languages have the most mature lexicons at present, although lexicons for 7 other languages are available.
+
+The link grammar parses do differ in style from the single-headed CoNLL parses of the same English sentences.
+They have !!!\% more edges overall.
+Only !!!\% of the links match CoNLL edges, and only !!!\% of the CoNLL edges match links.
+% DATA FOR THE ABOVE
+% \begin{figure}[ht!]
+%   \centering
+%   \small
+%   \input{figure/conll_analysis_arcs.tex}
+%   \caption{Arcs that match links}
+% \end{figure}
+% 
+% OLD
+% Using the sentences of the dependency corpus as comparison, we find that the link parses do not wholly subsume dependency parses, and that the undirected links match roughly three fourths of the arcs in the conll dependency corpus.
+\todo{This is because...}
 
@@
 The second summand ensures that \eqnref{nocycles} is trivially satisfied (hence has no effect) when $u$ is {\em not} the parent of $v$.  
-As a speedup, we can require the 0 tokens to have depth 0.
-\noteJE{does this really still make a difference to speed?
-Possible better version: just add all the depths (perhaps multiplied by 0.001) to the objective function; this breaks ties in simplex.}
-
-\noteJE{what's the point of the root\_depth constraint in the .zpl file?  Almost trivially satisfied.}
+\noteJE{in final version: figure out whether it is faster to require the 0 tokens to have depth 0, or add all the depths (perhaps multiplied by 0.001) to the objective function to break ties in simplex, or neither.
+See also other notes in tex file.}
+% As a speedup, we can require the 0 tokens to have depth 0. \noteJE{does this really still make a difference to speed?  
+% \noteJE{what's the point of the root\_depth constraint in the .zpl file?  Almost trivially satisfied.}
 
@@
   \input{figure/sentences.tikz}
-  \caption{\small Example Parses: The top blue arcs represent conll dependencies; the bottom red arcs represent directionalized link grammar parses. See the Appendix for more details.}
+  \caption{\small Example Parses: The top blue arcs show CoNLL dependency parses; the bottom red arcs show oriented link grammar parses.
+See Appendix~\ref{app:parses} for formatting details and more examples.}
   \label{fig:parses}
@@
 We solved our ILP problem using the SCIP Optimization Suite \cite{achterberg2009scip}, encoding it using the ZIMPL language \cite{Koch2004}.
-\NoteJE{give runtime here}\NoteJH{A run of 8,192 sentences takes 1743 seconds}
+Our largest run took under an hour.
+\noteJH{A run of 10,000 sentences takes 3142 seconds (update in final version!)}  Detailed results from this run are included as Appendix~\ref{app:linktypes}.
+On English, !!!
+of !!!
+link {\em types} allowed both directions, and only $\sum_{ij} s_{ij} = !!!$ of !!!
+link {\em tokens} required a disallowed direction via slack.
+\input{figure/conll_analysis_sentences_multiheaded}of the English sentences and !!! of the Russian sentences had at least one multiheaded word.
 
-\subsection{Stability of Results}
+\noteJE{fill these numbers in}
+
+\subsection{Stability of Results}\label{sec:stability}
 \begin{figure}[ht!]
@@
   \includegraphics[width=\linewidth, keepaspectratio=true]{figure/precision_recall.png}
-  \caption{\small Convergence to the results obtained on the largest run.  Precision relative to that run is very high even for small training sizes.  Recall grows as more link types are observed.}
-\end{figure}
-We report on the convergence of the results of our ILP with increasing data.
-With increasing trials of sentences, we measure how similar the subsequent directional assignments would be with each other.
-Taking the direction table of the largest run as the answer key, we compared how much the solutions of the smaller runs deviated from it.
-We recorded the precision of whether the assignments in the smaller runs could be found in those of the largest run, and similarly the recall of whether the assignments in the largest run could be found in the smaller runs.
-%\noteJE{this is not clear.  are we comparing the direction table or the parses?  If the parses, which parses?}
-
-Because the precision values were high, we found that the solutions of the previous runs were consistent with the largest run, and as the data increased so did the recall values, and so we would continue to encounter novel link types and then incorporate them.
-
-\NoteJE{how often is the ``both'' option used?  and how much slack do we have?}
-
-\subsection{English Link-Dependency Analysis}
-
-In the case of English, we had manually produced  single-headed 
- CoNLL annotations \cite{!!!}.
-
-Using the sentences of the dependency corpus as comparison, we find that the link parses do not wholly subsume dependency parses, and that the undirected links match roughly three fourths of the arcs in the conll dependency corpus.
-\todo{This is because...}
-
-\begin{figure}[ht!]
-  \centering
-  \small
-  \input{figure/conll_analysis_arcs.tex}
-  \caption{Arcs that match links}
+  \caption{Convergence to the direction mapping obtained on the largest run.
+Precision relative to that run is very high even for small datasets.
+Recall grows as more link types are seen.}\label{fig:convergence}
 \end{figure}
 
-we measure whether the direction we chose is the same as the original dependency data.  Finally, we look to see that given a link label, whether the original conll label can be recovered.
+We worried that the direction mapping might be unstable and sensitive to the input corpus.
+Figure~\ref{fig:convergence} shows otherwise (for both English and Russian).
+Using even a small prefix of the data got very high-precision results, in the sense that nearly all $r_L$ or $\ell_L$ variables that were 1 in this lightly trained mapping were also 1 in our largest run.
+The only disadvantage to using small data is low recall relative to the large run---many of the labels $L$ are not observed yet and so we do not yet allow either direction ($r_L=\ell_L=0$).
 
-Of the English sentences with complete parses, \input{figure/conll_analysis_sentences_multiheaded}had at least one multiheaded word.
+We used only coarse link tags as our labels, keeping only the capital letters of a tag (and merging all \texttt{ID} tags).
+This is because other characters in a tag indicate fine-grained features such as plurality that generally do not affect link direction.
+However, when we tried using fine tags as our labels instead, we found that all refinements of the same coarse tag would almost always spontaneously agree on their preferred direction.
+This indicates that there is indeed a ``natural'' direction for the coarse tag and that we can find it.
+
+% OLD
+% We report on the convergence of the results of our ILP with increasing data.
+With increasing trials of sentences, we measure how similar the subsequent direction mappings would be with each other.
+Taking the direction table of the largest run as the answer key, we compared how much the solutions of the smaller runs deviated from it.
+% We recorded the precision of whether the mappings in the smaller runs could be found in those of the largest run, and similarly the recall of whether the mappings in the largest run could be found in the smaller runs.
+
+% Because the precision values were high, we found that the solutions of the previous runs were consistent with the largest run, and as the data increased so did the recall values, and so we would continue to encounter novel link types and then incorporate them.
+
+\subsection{Linguistic Analysis}
+
+As noted earlier, the resulting English corpus uses an multi-headed annotation scheme that is somewhat different from the CoNLL annotations.
+Differences are tabulated in Appendix~\ref{app:linktypes}, and the actual parses are contrasted in Figure~\ref{fig:parses} and Appendix~\ref{app:parses}.
 
@@
 
-\NoteJE{what do we say about Russian?}
+\NoteJE{what do we say about Russian?  ask Ryan for final version?}
 
-\section{Discussion}
-
-\todo{more in the appendix ...}
-
-\NoteJE{We should acknowledge that another option would be to post-process either CoNLL parses or our link parses to automatically add more multi-heading, e.g., to handle control phenomena ("John wants to skip" or "Jane ran for office in order to change the tax laws" should have John/Jane be the subject of two verbs).
-However, that requires per-language}
-
-We have a set of 100 parses similar to figure~\ref{fig:parses} in the appendix.
-
-\section{Future Work}
-
-\subsection{Slack Hierarchy}
-We would like to introduce slack on the link grammar types such that link types with the same coarse grained label would try and align the same way as the majority in the group, where the preceding capital letters of the link type denote the coarse grained label, while the subscript letters denote further information.
+% DOESN'T SEEM NECESSARY FOR THE LANGUAGES WE TRIED
+% \section{Future Work}
+% \subsection{Slack Hierarchy}
+% We would like to introduce slack on the link grammar types such that link types with the same coarse grained label would try and align the same way as the majority in the group, where the preceding capital letters of the link type denote the coarse grained label, while the subscript letters denote further information.
 This slack would place a prior on rare or never-before seen link types to be assigned in the same way as other similar link types.
@@
 
-\section{Related Work}
-
+\newpage
 \bibliographystyle{acl2014}
@@
 \appendix
-\section{ILP Formulation}
-We present our ILP formulation.
 
-\begin{figure}[!htb]
-  \small
-  \begin{align*}
-    \text{Sets:}&&\\
-    &\textsc{link}_{\textsc{sentence},\textsc{label}, i,j} \in \textsc{links} &\\
-    &\textsc{label} \in \textsc{labels}&\\ 
-    &\textsc{dir} = \{\textsc{left}, \textsc{right}\} &\\
-    \text{Params:}&&\\
-    &\textsc{m}_\textsc{length} = \max (\forall{\textsc{link}_{\textsc{sentence}, i,j}} \max(i,j)) &\\
-    %&\textsc{count}_{\textsc{label}, \textsc{dir}} = \sum_{\textsc{link}_{\textsc{label},\textsc{dir}}}1 &\\
-    &\textsc{cost}_{\textsc{label}} = \frac{100}{ \sum_{\textsc{link}_{\textsc{label},\textsc{dir}}}1 } &\\
-    \text{Variables:}&&\\
-    &\textsc{slack}_{\textsc{link}} \geq 0 &\\ 
-    &\textsc{depth}_{\textsc{sentence},i} \geq 0 &\\
-    &\textsc{allowed}_{\textsc{label},\textsc{dir}} = \{0,1\}&\\
-    &\textsc{arc}_{\textsc{link}} = \{0,1\} &\\
-    \text{minimize:}& &\\ 
-    &\sum_{\textsc{label}, \textsc{dir}} (\textsc{allowed}_{\textsc{label},\textsc{dir}}) + \textsc{cost}_{\textsc{label}} \cdot \sum_\textsc{link} \textsc{slack}_{\textsc{link}} &\\
-    &&\\
-    \text{subject to: }& &\\
-    &\text{Links go in the allowed directions, except for slack}&\\
-    &\forall{\textsc{link}_{i,j,\textsc{label}}}: 1-\textsc{arc}_{\textsc{link}} \leq \textsc{allowed}_{\textsc{label}, \textsc{left}} + \textsc{slack}_{\textsc{link}} &\\
-    &\forall{\textsc{link}_{i,j,\textsc{label}}}: \textsc{arc}_{\textsc{link}} \leq \textsc{allowed}_{\textsc{label}, \textsc{right}} + \textsc{slack}_{\textsc{link}} &\\
-    &\text{Depth constraints for acyclicity}&\\
-    &\forall{\textsc{link}_{\textsc{sentence},i}}: \textsc{depth}_{\textsc{sentence},\text{root}} \leq \textsc{depth}_{\textsc{sentence},i} &\\
-    &\forall{\textsc{link}_{\textsc{sentence},i,j}}: \textsc{depth}_{\textsc{sentence},i} + \textsc{m}_\textsc{length}\cdot\textsc{arc}_{\textsc{sentence},i,j} \geq \textsc{depth}_{\textsc{sentence},j} + 1 &\\
-    &\forall{\textsc{link}_{\textsc{sentence},i,j}}: \textsc{depth}_{\textsc{sentence},j} + \textsc{m}_\textsc{length}\cdot(1-\textsc{arc}_{\textsc{sentence},i,j}) \geq \textsc{depth}_{\textsc{sentence},i} + 1 &\\
-    &\text{Every token has a parent}&\\
-    &\forall{\textsc{link}_{\textsc{sentence},i \neq \textsc{root}}}: \sum_{\textsc{link}_{\textsc{sentence},i,j}} (1-\textsc{arc}_{\textsc{sentence},i,j}) + \sum_{\textsc{link}_{\textsc{sentence},j,i}}(\textsc{arc}_{\textsc{sentence},j,i}) \geq 1&\\
-  \end{align*}
-  \caption{\small The ILP formulation.}
-\end{figure}
+% OLD
+% \section{ILP Formulation}
+% We present our ILP formulation. 
+% 
+% \begin{figure}[!htb]
+%   \small
+%   \begin{align*}
+%     \text{Sets:}&&\\
+%     &\textsc{link}_{\textsc{sentence},\textsc{label}, i,j} \in \textsc{links} &\\
+%     &\textsc{label} \in \textsc{labels}&\\ 
+%     &\textsc{dir} = \{\textsc{left}, \textsc{right}\} &\\
+%     \text{Params:}&&\\
+%     &\textsc{m}_\textsc{length} = \max (\forall{\textsc{link}_{\textsc{sentence}, i,j}} \max(i,j)) &\\
+%     %&\textsc{count}_{\textsc{label}, \textsc{dir}} = \sum_{\textsc{link}_{\textsc{label},\textsc{dir}}}1 &\\
+%     &\textsc{cost}_{\textsc{label}} = \frac{100}{ \sum_{\textsc{link}_{\textsc{label},\textsc{dir}}}1 } &\\
+%     \text{Variables:}&&\\
+%     &\textsc{slack}_{\textsc{link}} \geq 0 &\\ 
+%     &\textsc{depth}_{\textsc{sentence},i} \geq 0 &\\
+%     &\textsc{allowed}_{\textsc{label},\textsc{dir}} = \{0,1\}&\\
+%     &\textsc{arc}_{\textsc{link}} = \{0,1\} &\\
+%     \text{minimize:}& &\\ 
+%     &\sum_{\textsc{label}, \textsc{dir}} (\textsc{allowed}_{\textsc{label},\textsc{dir}}) + \textsc{cost}_{\textsc{label}} \cdot \sum_\textsc{link} \textsc{slack}_{\textsc{link}} &\\
+%     &&\\
+%     \text{subject to: }& &\\
+%     &\text{Links go in the allowed directions, except for slack}&\\
+%     &\forall{\textsc{link}_{i,j,\textsc{label}}}: 1-\textsc{arc}_{\textsc{link}} \leq \textsc{allowed}_{\textsc{label}, \textsc{left}} + \textsc{slack}_{\textsc{link}} &\\
+%     &\forall{\textsc{link}_{i,j,\textsc{label}}}: \textsc{arc}_{\textsc{link}} \leq \textsc{allowed}_{\textsc{label}, \textsc{right}} + \textsc{slack}_{\textsc{link}} &\\
+%     &\text{Depth constraints for acyclicity}&\\
+%     &\forall{\textsc{link}_{\textsc{sentence},i}}: \textsc{depth}_{\textsc{sentence},\text{root}} \leq \textsc{depth}_{\textsc{sentence},i} &\\
+%     &\forall{\textsc{link}_{\textsc{sentence},i,j}}: \textsc{depth}_{\textsc{sentence},i} + \textsc{m}_\textsc{length}\cdot\textsc{arc}_{\textsc{sentence},i,j} \geq \textsc{depth}_{\textsc{sentence},j} + 1 &\\
+%     &\forall{\textsc{link}_{\textsc{sentence},i,j}}: \textsc{depth}_{\textsc{sentence},j} + \textsc{m}_\textsc{length}\cdot(1-\textsc{arc}_{\textsc{sentence},i,j}) \geq \textsc{depth}_{\textsc{sentence},i} + 1 &\\
+%     &\text{Every token has a parent}&\\
+%     &\forall{\textsc{link}_{\textsc{sentence},i \neq \textsc{root}}}: \sum_{\textsc{link}_{\textsc{sentence},i,j}} (1-\textsc{arc}_{\textsc{sentence},i,j}) + \sum_{\textsc{link}_{\textsc{sentence},j,i}}(\textsc{arc}_{\textsc{sentence},j,i}) \geq 1&\\
+%   \end{align*}
+%   \caption{\small The ILP formulation.}
+% \end{figure}
+% 
+% 
+% \clearpage
 
-\clearpage
+\section{English Link Types}\label{app:linktypes}
 
-\section{English Link Types}
+Recall that \url{http://www.abisource.com/projects/link-grammar/dict/summarize-links.html} documents the 122 link types in the English link grammar lexicon.
+Below is the result of our largest English run, whose corpus uses !!!
+of these types.
 
-\todo[inline]{A table here that lists each link grammar type together with its
-  count, its allowed direction(s), its percentage of majority
-  directions by token, and its most common corresponding CoNLL tags
-  with percentages.  Readers could cut-and-paste this data to make a 
-  cheap post-processor for LG parses.}
+Other researchers may postprocess link grammar parses simply by using the first two columns, which indicate whether a given link label $L$ tended to be oriented rightward or leftward by our ILP.
+
+The labels that do not have a consistent direction are mainly those that connect a conjunction to its conjuncts.\footnote{See \url{http://www.abisource.com/projects/link-grammar/dict/coordination.html}.}  Our coarsening of the labels (see section~\ref{sec:stability}) discards the portion of the label that distinguishes left from right conjuncts.
+
+The third column indicates where multi-headedness happens, by showing how often the child of an $L$ link had additional parents as well.
+
+The remaining columns compare to the CoNLL parses.
+They show the fraction of $L$-labeled link tokens that were also present in the CoNLL reference parse---and on that fraction, whether the orientation matches CoNLL's and what the dominant CoNLL label was.
 
@@
 
-\section{Sample English Parses}
-In this section, we will display 100 sample English parses.
-The top half of each parse visualization represents the CoNLL dependency parse, while the bottom half shows our directionalized link grammar parse.
+\section{Sample English Parses}\label{app:parses}
+
+Below we display 100 sample English parses. The top half of each parse visualization represents the CoNLL dependency parse, while the bottom half shows our oriented link grammar parse.
 
@@
 \end{enumerate}
-The bottom red arcs represent the directionalized link grammar parses. For every link edge:
+The bottom red arcs represent the oriented link grammar parses. For every link edge:
 \begin{enumerate}

### 2014-06-03 src/tools/experiments/conll_analysis.py
added a comment in conll_analysis.py

@@ -736,7 +736,12 @@ f.close()
 #---------------------------------------------------------------------
 LATEX_FILE_COARSE_LINKS = LATEX_DIR+"link_analysis_coarse_table.tex"
 f = open(LATEX_FILE_COARSE_LINKS, "w+")
+
+
+# TODO use tabular
 latex_table = "\\begin{tabular}{|l|l|l|l|l|l|}\n"
+end_table = "\\end{tabular}\n"
+
 header = "Label & Rightward & Multiheaded & CoNLL Match & CoNLL Dir Match & CoNLL Label\\\\ \n"
 
 begin_figure = "\\begin{figure*}\n\\small\n\\centering\n"
@@ -788,9 +793,9 @@ for coarse_label in coarse_labels:
     line += 1
     
     # Break up the table into smaller tables
-    if line % 50 == 0:
+    if line % 55 == 0:
         table += "\\hline\n"
-        table += "\\end{tabular}\n"
+        table += end_table
         table += end_figure
         table += begin_figure
         table += latex_table
@@ -798,7 +803,7 @@ for coarse_label in coarse_labels:
         table += header
     
 table += "\\hline\n"
-table += "\\end{tabular}\n"
+table += end_table
 table += end_figure
 
 f.write(table)

### 2014-06-03 src/tools/experiments/conll_analysis.py
wrote a function, simplified code.

@@ -747,6 +747,9 @@ header = "Label & Rightward & Multiheaded & CoNLL Match & CoNLL Dir Match & CoNL
 begin_figure = "\\begin{figure*}\n\\small\n\\centering\n"
 end_figure = "\\end{figure*}\n"
 
+def make_percentage_figure(top,bottom):
+    return str(int(float(top) / float(bottom) * 100 + 0.5))+"\\%" + " ("+str(top)+"/"+str(bottom)+")"
+
 
 table = begin_figure
 table += latex_table
@@ -757,27 +760,26 @@ for coarse_label in coarse_labels:
     count = link_label_coarse_counts[coarse_label]
     
     right = link_direction_coarse_totals[coarse_label].get("right",0)
-    right = str(int(float(right) / count * 100))+"\\%" + " ("+str(right)+"/"+str(count)+")"
+    right = make_percentage_figure(right,count)
 
     prediction_num = ""
     prediction_total = ""
     match_percent = ""
     mismatch_percent = ""
-    multiheaded = str(int(float(link_label_isMultiheaded.get(coarse_label,0)) / float(count) * 100 + 0.5))+"\\% ("+str(link_label_isMultiheaded.get(coarse_label, 0)) + "/" + str(count) + ")"
-
+    multiheaded = make_percentage_figure(link_label_isMultiheaded.get(coarse_label,0), count)
 
 
     prediction = coarse_predictions.get(coarse_label, "-")
     if prediction != "-":
         prediction_num = link_label_coarse_prediction[coarse_label][prediction]
         prediction_total = link_label_coarse_prediction_totals[coarse_label]
-        prediction += " "+str(int(float(prediction_num) / float(prediction_total) * 100 + 0.5)) + "\\%"
-        prediction += " (" +str(prediction_num) + "/" + str(prediction_total) + ")"
+        prediction = make_percentage_figure(prediction_num, prediction_total)
+        
 
     if prediction_num and prediction_total:
-        match_percent = str(int(float(prediction_total) / count * 100 + 0.5)) + "\\%" + " (" + str(prediction_total) + "/" + str(count) + ")"
+        match_percent = make_percentage_figure(prediction_total, count)
 
[…]

### 2014-06-03 doc/LinksToDAG.tex
added longtable package
improved text in appendices

@@
 \usepackage[small]{caption}
+\usepackage{longtable}
 
@@
 % ... produces parses that are similar to multi-headed projective dependency parses, except that the links are undirected. 
-We use Integer Linear Programming to assign consistent directions to the labeled links in a corpus of \input{figure/conll_analysis_sentences_total.tex}parses produced by the Link Grammar Parser, which has broad-coverage hand-written grammars of English as well as Russian and other languages.
+We use Integer Linear Programming to assign consistent directions to the labeled links in a corpus of several thousand parses produced by the Link Grammar Parser, which has broad-coverage hand-written grammars of English as well as Russian and other languages.
 We find that such directions can indeed be consistently assigned in a way that yields valid multi-headed dependency parses.
-The resulting parses in English appear reasonably linguistically plausible, although they are not in general consistent with CoNLL-style parses of the same sentences; we discuss the differences.
+The resulting parses in English appear reasonably linguistically plausible, although their style does not match CoNLL-style parses of the same sentences; we discuss the differences.
 \end{abstract}
@@
 They have !!!\% more edges overall.
-Only !!!\% of the links match CoNLL edges, and only !!!\% of the CoNLL edges match links.
+Only !!!\% of the links match CoNLL arcs, and only !!!\% of the CoNLL arcs match links.
 % DATA FOR THE ABOVE
@@
 
-The labels that do not have a consistent direction are mainly those that connect a conjunction to its conjuncts.\footnote{See \url{http://www.abisource.com/projects/link-grammar/dict/coordination.html}.}  Our coarsening of the labels (see section~\ref{sec:stability}) discards the portion of the label that distinguishes left from right conjuncts.
+The labels that do not have a consistent direction are mainly those that connect a conjunction to its conjuncts.\footnote{See \url{http://www.abisource.com/projects/link-grammar/dict/coordination.html}.}
+Our coarsening of the labels (see section~\ref{sec:stability}) discards the characters that distinguish left from right conjuncts.
 
@@
 
+\renewcommand{\arraystretch}{1.2}    % more space between rows
 \input{figure/link_analysis_coarse_table.tex}
@@
 
-Below we display 100 sample English parses. The top half of each parse visualization represents the CoNLL dependency parse, while the bottom half shows our oriented link grammar parse.
+Below we display the first 100 English sentences whose widths do not exceed the width of this page.
+The blue top half shows the CoNLL tagging and dependency parse.
+The red bottom half shows the link grammar's tagging and our orientation of the link parse.
+% The dashes in the link grammar part of speech tag annotations mean that the link parser did not return one.
+Words that were unknown to the link parser are flagged with \texttt{[!]}.
 
-\subsection{Sentence Selection}
-These sentences have been taken from the English bnews training data set.
-We chose to display sentences with complete link parses similar to the analysis section of this paper.
-We display the first 100 parses whose character lengths do not exceed the width of the page.
+Edges are shown as dotted lines if they appear only in one parse and not in the other.
+The solid edges in the link parse are drawn thickly if the corresponding CoNLL arc points in the opposite direction.
+Multiple parents are shown using orange edges.
+\noteJE{not quite how orange is currently handled---fix}
 
-\subsection{Parse Colorization}
-The top blue arcs represent the CoNLL dependency parses. For every arc:
-\begin{enumerate}
-\item If there is a corresponding link edge between the same nodes then a solid blue line is drawn.
-\item Or else a dotted blue line is drawn. 
-\end{enumerate}
-The bottom red arcs represent the oriented link grammar parses. For every link edge:
-\begin{enumerate}
-\item If there is a corresponding CoNLL arc between the same nodes then a solid red line is drawn.
-\item If the matching arc is facing the opposite direction, then the link edge is drawn thicker.
-\item Of the remaining edges, if an edge is a multiheaded link, then it is drawn with thick orange.
-\item The rest of the remaining edges are drawn with dotted red lines.
-\end{enumerate}
+% OLD
+% For every arc:
+% \begin{enumerate}
+% \item If there is a corresponding link between the same nodes then a solid blue line is drawn.
+% \item Or else a dotted blue line is drawn. 
+% \end{enumerate}
+% The bottom red arcs represent the oriented link grammar parses. For every link edge:
+% \begin{enumerate}
+% \item If there is a corresponding CoNLL arc between the same nodes then a solid red line is drawn.
+% \item If the matching arc is facing the opposite direction, then the link edge is drawn thicker.
+% \item Of the remaining edges, if an edge is a multiheaded link, then it is drawn with thick orange.
+% \item The rest of the remaining edges are drawn with dotted red lines.
+% \end{enumerate}
 
-\subsection{Link Parser Annotations}
-The sentence used in the middle of the parse is taken from the output of the link parser.
-We removed sentences with dropped words, but they would have shown up in the parse as [bracketed].
-There are also words that the link parser did not recognize, but rather made a guess as to its part of speech.
-These words are tagged with a [!] symbol at the end.
-The dashes in the link grammar part of speech tag annotations mean that the link parser did not return one.
+% \subsection{Link Parser Annotations}
+% These sentences have been taken from the English bnews training data set. We chose to display sentences with complete link parses similar to the analysis section of this paper. 
+% 
+% \subsection{Parse Colorization}
+% The sentence used in the middle of the parse is taken from the output of the link parser. We removed sentences with dropped words, but they would have shown up in the parse as [bracketed].
 

### 2014-06-03 src/decoder.py
changed the project to work with the coarse grained labels instead of the fine grained labels. The changes are to the encoder.

@@ -18,6 +18,7 @@ def decodeSCIPsolution(links, solutionFile, getAllowedLabels=False):
     decodedSolutions = {}
 
 
+
     for line in lines:
         if line.find("rLink#") == 0:
             line = line.split()

### 2014-06-03 doc/LinksToDAG.tex

@@
 
-Other researchers may postprocess link grammar parses simply by using the first two columns, which indicate whether a given link label $L$ tended to be oriented rightward or leftward by our ILP.
+Other researchers could postprocess link grammar parses by using the first two columns.  These indicate whether a given link label $L$ tended to be oriented rightward or leftward by our ILP.
 
 The labels that do not have a consistent direction are mainly those that connect a conjunction to its conjuncts.\footnote{See \url{http://www.abisource.com/projects/link-grammar/dict/coordination.html}.}
-Our coarsening of the labels (see section~\ref{sec:stability}) discards the characters that distinguish left from right conjuncts.
+Our coarsening of the labels (see section~\ref{sec:stability}) discarded the characters that distinguish left from right conjuncts.
 

### 2014-06-03 doc/LinksToDAG.tex
clarified section 1

@@
 \item {\sc Connected}: each pair of words has a undirected path between them
-\item {\sc Reachable}: each word can be reached from 0 by a directed path
-  \noteJE{note that this implies connected?
-or does connected not allow 0 words on the path, i.e., does it restrict 0 to one child, and if not, is there a principle for that?
-If no one, mention in footnote 1.}
-\item {\sc Planar}: edges may not ``cross'' \\ (if there are edges between $i,j$ and between $u,v$, where $i < u < j$, then $i \leq v \leq j$)
+\item {\sc Reachable}: each word can be reached from 0 by a directed path ($\Rightarrow$ {\sc Connected})
+\item {\sc Planar}: edges may not ``cross''\footnote{That is, if there are edges between $i,j$ and between $u,v$, where $i < u < j$, then {\sc Planar} requires $i \leq v \leq j$.}
 \end{itemize}
-It is common to impose all of these requirements at once, leading to a {\em projective dependency parser} that produces projective trees rooted at 0.\footnote{{\sc Projective} basically means {\sc Planar} + {\sc Reachable}.}  However, parsing algorithms can be devised that relax any or all of the requirements \cite{gomezrodriguez-nivre-2013}.
+It is common to impose all of these requirements at once, leading to a {\em projective dependency parser} that produces projective trees rooted at 0.\footnote{{\sc Projective} basically means {\sc Planar} + {\sc Reachable}.
+Note: It does not prevent 0 from having multiple children.}  However, parsing algorithms can be devised that relax any or all of the requirements \cite{gomezrodriguez-nivre-2013}.
 
@@
 
+\NoteJE{improve table formatting}
+
 \renewcommand{\arraystretch}{1.2}    % more space between rows

### 2014-06-03 doc/LinksToDAG.tex
minor edits

@@
 We find that such directions can indeed be consistently assigned in a way that yields valid multi-headed dependency parses.
-The resulting parses in English appear reasonably linguistically plausible, although their style does not match CoNLL-style parses of the same sentences; we discuss the differences.
+The resulting parses in English appear reasonably linguistically plausible, though differing in style from CoNLL-style parses of the same sentences; we discuss the differences.
 \end{abstract}
@@
 Our goal in this paper is to recover these implicit directions by global optimization.
-We seek a fixed mapping from labels to directions such that we can interpret link grammar parses as directed dependency parses that satisfy all of our constraints.
+We seek a fixed mapping from labels to directions such that link grammar parses become directed dependency parses that satisfy all of our constraints.
 
@@
   \input{figure/sentences.tikz}
-  \caption{\small Example Parses: The top blue arcs show CoNLL dependency parses; the bottom red arcs show oriented link grammar parses.
+  \caption{\small Example Parses: The blue upper edges show CoNLL dependency parses; the red lower edges show oriented link grammar parses.
+Vertical edges are those with parent 0.
 See Appendix~\ref{app:parses} for formatting details and more examples.}
@@
 
-As noted earlier, the resulting English corpus uses an multi-headed annotation scheme that is somewhat different from the CoNLL annotations.
+The resulting English corpus uses an multi-headed annotation scheme that is somewhat different from the CoNLL annotations.
 Differences are tabulated in Appendix~\ref{app:linktypes}, and the actual parses are contrasted in Figure~\ref{fig:parses} and Appendix~\ref{app:parses}.
@@
 
-\NoteJE{improve table formatting}
+\NoteJE{improve table formatting.  Numeric justification; longtable; remove most hlines.}
 
@@
 Below we display the first 100 English sentences whose widths do not exceed the width of this page.
-The blue top half shows the CoNLL tagging and dependency parse.
-The red bottom half shows the link grammar's tagging and our orientation of the link parse.
+The blue upper half shows the CoNLL tagging and dependency parse.
+The red lower half shows the link grammar's tagging and our orientation of the link parse.
 % The dashes in the link grammar part of speech tag annotations mean that the link parser did not return one.
+A vertical edge to word token $v$ denotes an edge $0 \rightarrow v$.
 Words that were unknown to the link parser are flagged with \texttt{[!]}.

### 2014-06-03 src/tools/experiments/conll_analysis.py
updated the graph drawing to have thicker colors or dotted thicker colors.

@@ -419,7 +419,8 @@ for linkSentence in link_results:
             if paper_sentence_count % 3 == 0:
                 PAPER_TIKZ.write("\n")
     # Filter out sentences to only up to length 16
-    elif example_parses_count < example_parses_limit and len(linkSentence) <= 85:
+    elif example_parses_count < example_parses_limit and len(linkSentence) <= 100:
+    #elif example_parses_count < example_parses_limit:
         tikz = tikz_dependency(conll_results[linkSentence], link_results[linkSentence], linkSentence, 1.0, False)
         EXAMPLE_PARSES.write("\\begin{figure*}[ht!]\n")
         EXAMPLE_PARSES.write(tikz)

### 2014-06-03 src/tools/experiments/tikz_dependency.py

@@ -58,7 +58,7 @@ def tikz_dependency(conlls, links, sentence, ratio = 0.3, subfigure = True):
             
             elif child in graph_link.table and head in graph_link.table[child]:
                 result += tikz_draw_edge(child, head, graph_link.getEdge(child,head), "below", "ultra thick")
-                result += tikz_draw_edge(head, child, "\\small "+label, "above", "thick")
+                result += tikz_draw_edge(head, child, "\\small "+label, "above", "ultra thick")
                 graph_link.deleteEdge(child,head)
             
             else:
@@ -72,7 +72,7 @@ def tikz_dependency(conlls, links, sentence, ratio = 0.3, subfigure = True):
             label = graph_link.labels[(head,child)]
             
             if len(graph_link.heads[child]) > 1 :
-                result += tikz_draw_edge(head,child,label, "below", "ultra thick", "orange")
+                result += tikz_draw_edge(head,child,label, "below", "ultra thick, dotted", "orange")
             else:
                 result += tikz_draw_edge(head,child,label, "below", "dotted")
 

### 2014-06-03 src/tools/experiments/conll_analysis.py
found a bug. Reversed the polarity of the rightwards column of the appendix table.

@@ -221,9 +221,9 @@ def analysis_links(links):
 
             link_direction_counts[label] = link_direction_counts.get(label,{})
             if head < index:
-                link_direction_counts[label]["left"] = link_direction_counts[label].get("left",0) + 1
+                link_direction_counts[label]["right"] = link_direction_counts[label].get("left",0) + 1
             elif head > index:
-                link_direction_counts[label]["right"] = link_direction_counts[label].get("right",0) + 1
+                link_direction_counts[label]["left"] = link_direction_counts[label].get("right",0) + 1
 
 
     return (link_direction_counts, link_label_multiheaded)

### 2014-06-03 src/encoder.py
changed the encoder to only give the coarse grained solutions to the ILP.

@@ -88,7 +88,7 @@ def getDataFromLinkParse(lines):
         links[i][3] = links[i][3].strip("()")
 
         # Make the label coarse grained
-        links[i][3] = getCoarseLabel(links[i][3])
+        #links[i][3] = getCoarseLabel(links[i][3])
 
 
     return (processedSentence, links)

### 2014-06-03 src/tools/experiments/conll_analysis.py
new coarse table.

@@ -221,9 +221,9 @@ def analysis_links(links):
 
             link_direction_counts[label] = link_direction_counts.get(label,{})
             if head < index:
-                link_direction_counts[label]["right"] = link_direction_counts[label].get("left",0) + 1
+                link_direction_counts[label]["left"] = link_direction_counts[label].get("left",0) + 1
             elif head > index:
-                link_direction_counts[label]["left"] = link_direction_counts[label].get("right",0) + 1
+                link_direction_counts[label]["right"] = link_direction_counts[label].get("right",0) + 1
 
 
     return (link_direction_counts, link_label_multiheaded)
@@ -419,8 +419,7 @@ for linkSentence in link_results:
             if paper_sentence_count % 3 == 0:
                 PAPER_TIKZ.write("\n")
     # Filter out sentences to only up to length 16
-    elif example_parses_count < example_parses_limit and len(linkSentence) <= 100:
-    #elif example_parses_count < example_parses_limit:
+    elif example_parses_count < example_parses_limit and len(linkSentence) <= 85:
         tikz = tikz_dependency(conll_results[linkSentence], link_results[linkSentence], linkSentence, 1.0, False)
         EXAMPLE_PARSES.write("\\begin{figure*}[ht!]\n")
         EXAMPLE_PARSES.write(tikz)

### 2014-06-03 src/tools/experiments/conll_analysis.py
fixed a bug.

@@ -56,6 +56,37 @@ EXAMPLE_PARSES_LOC = TIKZ_DIR+"parses.tikz"
 EXAMPLE_PARSES = open(EXAMPLE_PARSES_LOC, 'w+')
 
 
+ALLOWED_LABELS = SOL_DIR + "allowedLinks.txt"
+ALLOWED_LABELS = open(ALLOWED_LABELS, "r")
+ALLOWED_LABELS_BOTH = TIKZ_DIR+"allowed_labels_both.tex"
+ALLOWED_LABELS_TOTAL = TIKZ_DIR+"allowed_labels_total.tex"
+
+allowedLabel = {}
+ALLOWED_LABELS = ALLOWED_LABELS.readlines()
+for line in ALLOWED_LABELS:
+    line = line.split()
+    if not line:
+        continue
+
+    label = line[0]
+    direction = line[1]
+    allowedLabel[label] = allowedLabel.get(label,{})
+    allowedLabel[label][direction] = True
+
+bothCount = 0
+for label in allowedLabel:
+    dirs = allowedLabel[label]
+    if len(dirs) > 1:
+        bothCount += 1
+
+f = open(ALLOWED_LABELS_BOTH,"w+")
+f.write(str(bothCount))
+f.close()
+
+f = open(ALLOWED_LABELS_TOTAL,"w+")
+f.write(str(len(allowedLabel)))
+f.close()
+
 
 # --------------------------------------------------------------------
 # Analysis while reading in data. Pertains mostly to sentence counts
@@ -221,9 +252,9 @@ def analysis_links(links):
 
             link_direction_counts[label] = link_direction_counts.get(label,{})
             if head < index:
-                link_direction_counts[label]["left"] = link_direction_counts[label].get("left",0) + 1
-            elif head > index:
                 link_direction_counts[label]["right"] = link_direction_counts[label].get("right",0) + 1
+            elif head > index:
+                link_direction_counts[label]["left"] = link_direction_counts[label].get("left",0) + 1
 
 
     return (link_direction_counts, link_label_multiheaded)
@@ -419,7 +450,8 @@ for linkSentence in link_results:
             if paper_sentence_count % 3 == 0:
                 PAPER_TIKZ.write("\n")
     # Filter out sentences to only up to length 16
-    elif example_parses_count < example_parses_limit and len(linkSentence) <= 85:
+    elif example_parses_count < example_parses_limit and len(linkSentence) <= 100:
+    #elif example_parses_count < example_parses_limit:
[…]

### 2014-06-03 src/encoder.py
put the encoder back

@@ -88,7 +88,7 @@ def getDataFromLinkParse(lines):
         links[i][3] = links[i][3].strip("()")
 
         # Make the label coarse grained
-        #links[i][3] = getCoarseLabel(links[i][3])
+        links[i][3] = getCoarseLabel(links[i][3])
 
 
     return (processedSentence, links)

### 2014-06-03 doc/LinksToDAG.tex
linguistic analysis

@@
 \begin{align}\label{eqn:obj}
-\min \left( \sum_{L} r_L + \ell_L \right) + \frac{100}{N_L} \cdot \sum_{ij} s_{ij}
+\min \left( \sum_{L} r_L + \ell_L \right) \cdot \frac{N_L}{4} + \sum_{ij} s_{ij}
 \end{align}
 where $N_L$ is the number of link tokens with label $L$.
-Objective \eqnref{obj} is willing to tolerate up to 1\% of those link tokens' using a disallowed direction before it prefers to allow {\em both} directions.
+Objective \eqnref{obj} is willing to tolerate up to 1/4 of those link tokens' using a disallowed direction before it prefers to allow {\em both} directions.
+\noteJE{check!}
 % CUT FOR SPACE
 % (i.e., $r_L+\ell_L=2$).
+
+\NoteJE{It might be a good idea to have the second allowed direction still carry some lesser slack penalty, so that we break ties in favor of the preferred direction.
+The two bits for each label would then say ``which direction is preferred?'' and ``is the other direction allowed?''}
 
@@
 We solved our ILP problem using the SCIP Optimization Suite \cite{achterberg2009scip}, encoding it using the ZIMPL language \cite{Koch2004}.
-Our largest run took under an hour.
-\noteJH{A run of 10,000 sentences takes 3142 seconds (update in final version!)}  Detailed results from this run are included as Appendix~\ref{app:linktypes}.
+Our largest run took 1.5 hours.
+\noteJE{update in final version}  Detailed results from this run are included as Appendix~\ref{app:linktypes}.
 On English, !!!
@@
 Figure~\ref{fig:convergence} shows otherwise (for both English and Russian).
-Using even a small prefix of the data got very high-precision results, in the sense that nearly all $r_L$ or $\ell_L$ variables that were 1 in this lightly trained mapping were also 1 in our largest run.
+\noteJE{IMPORTANT!
+current graph is with fine-grained tags; rerun with the coarse-grained tags as in the rest of the paper} Using even a small prefix of the data got very high-precision results, in the sense that nearly all $r_L$ or $\ell_L$ variables that were 1 in this lightly trained mapping were also 1 in our largest run.
 The only disadvantage to using small data is low recall relative to the large run---many of the labels $L$ are not observed yet and so we do not yet allow either direction ($r_L=\ell_L=0$).
@@
 
-The resulting English corpus uses an multi-headed annotation scheme that is somewhat different from the CoNLL annotations.
+The resulting English corpus uses a syntactic annotation scheme that is somewhat different from the CoNLL annotations.
 Differences are tabulated in Appendix~\ref{app:linktypes}, and the actual parses are contrasted in Figure~\ref{fig:parses} and Appendix~\ref{app:parses}.
 
-\NoteJE{discuss linguistic differences!}
+The link grammar results in multi-headed treatments of infinitivals, compound determiners, relative clauses, and embedding.
+The other annotation differences are generally reasonable, e.g., letting {\em 's} be the head of a possessive, and different handling of punctuation and lists.
+Of course, a few of the differences are due to parser attachment errors.
+
+The most vexing issue is that subject-verb links are handled inconsistently.
+Under the English link grammar, a verb that governs a clause (or 0 for the main clause) will link to both the clause's subject and its last (main) verb.
+This ought to give us the desired treatment of {\em ``Jill persuaded him to skip''}, in which {\em ``him''} has two parents.
+Unfortunately, the grammar is inconsistent when the subject is missing ({\em ``Jill wanted to skip''}): the governing verb here links instead to the clause's first (tensed) verb but {\em no longer} to the last as well.
+This favors a consistent solution where each auxiliary verb is a parent of the verb to its right ({\em to} $\rightarrow$ {\em skip}).
+As a result, when the subject is present, it must be the {\em parent} of the first verb, which otherwise would not have a parent ({\em him} $\rightarrow$ {\em to} $\rightarrow$ {\em skip}).
+Yet inconsistently, subjects must be {\em children} of verbs in a few  inconsistently handled constructions such as non-embedded reported speech ({\em ``It was impossible, they said''}), since the subject ({\em ``they''}) links only to the verb ({\em ``said''}) and so has no other possible parent.
+
+% Multiheaded:
+% IV (890/891) - multiheaded treatment of infinitivals
+% I (1783/2969) - multiheaded treatment of infinitivals
+% B (795/842) - leftward extraction from relative clauses
+% CV (1588/1799) - multiheaded treatment of subordination
+% AL (39/40) - compound determiners
+% PP (448/689) - subject attaches to aux as well as main verb
+% HA (3/3) -
+% JQ (2/2) - 
+% SF (85/111) - pleonastic subject
+% AF (6/12) - 
+% C (244/1813) - 
+% OF (155/291)
+% P (651/2060) - 
+% Q (12/30)- 
+% RS (300/305) - 
+% S (4256/7434) -
+% SX (4/6) - 
+% W (2321/4721) - 
+% WV (1800/4138) -
 
@@
 
+\appendix
 \clearpage
-\appendix
+\onecolumn
 
@@
 
-\NoteJE{improve table formatting.  Numeric justification; longtable; remove most hlines.}
+\NoteJE{improve table formatting.  Numeric justification; en dashes; longtable; remove most hlines.}
 
@@
 \section{Sample English Parses}\label{app:parses}
+
+\NoteJE{show the fine-grained link labels in the parses, even though the ILP deals with coarse-grained.  Juneki tried already to hack this in but was getting double links.}
 
@@
 
-Edges are shown as dotted lines if they appear only in one parse and not in the other.
-The solid edges in the link parse are drawn thickly if the corresponding CoNLL arc points in the opposite direction.
-Multiple parents are shown using orange edges.
-\noteJE{not quite how orange is currently handled---fix}
+Edges are shown as solid lines if they appear in both parses.  They are drawn more thickly if the directions do not match.
+
+Edges are shown as dotted lines if they appear only in one parse and not in the other.  Such edges are highlighted in orange if the child has multiple parents.
 

### 2014-06-03 doc/LinksToDAG.tex
conclusions

@@
 \usepackage[usenames,dvipsnames,svgnames,table]{xcolor}  % allows better color names
-\usepackage{todonotes}   % insert [disable] to disable all notes
+\usepackage[disable]{todonotes}   % insert [disable] to disable all notes
 % Note that these macros accept optional arguments such as 
@@
 \input{figure/conll_analysis_sentences_multiheaded}of the English sentences and !!! of the Russian sentences had at least one multiheaded word.
+\noteJE{2 significant digits is enough}
 
@@
 
-The most vexing issue is that subject-verb links are handled inconsistently.
-Under the English link grammar, a verb that governs a clause (or 0 for the main clause) will link to both the clause's subject and its last (main) verb.
+The main vexation is that subject-verb links are handled inconsistently.
+Under the English link grammar, the verb (or 0) that governs a clause will link to both the clause's subject and its last (main) verb.
 This ought to give us the desired treatment of {\em ``Jill persuaded him to skip''}, in which {\em ``him''} has two parents.
-Unfortunately, the grammar is inconsistent when the subject is missing ({\em ``Jill wanted to skip''}): the governing verb here links instead to the clause's first (tensed) verb but {\em no longer} to the last as well.
-This favors a consistent solution where each auxiliary verb is a parent of the verb to its right ({\em to} $\rightarrow$ {\em skip}).
+Unfortunately, it does not, because of inconsistencies in the link grammar itself.\footnote{When the subject is missing ({\em ``Jill wanted to skip''}), the governing verb now links instead to the clause's first (tensed) verb but {\em no longer} to the last as well.
+That favors a consistent solution where each auxiliary verb is a parent of the verb to its right ({\em to} $\rightarrow$ {\em skip}).
 As a result, when the subject is present, it must be the {\em parent} of the first verb, which otherwise would not have a parent ({\em him} $\rightarrow$ {\em to} $\rightarrow$ {\em skip}).
-Yet inconsistently, subjects must be {\em children} of verbs in a few  inconsistently handled constructions such as non-embedded reported speech ({\em ``It was impossible, they said''}), since the subject ({\em ``they''}) links only to the verb ({\em ``said''}) and so has no other possible parent.
+Yet also, subjects must be {\em children} of verbs in a few  inconsistently handled constructions such as non-embedded reported speech ({\em ``It was impossible, they said''}), since the subject ({\em ``they''}) links only to the verb ({\em ``said''}) and so has no other possible parent.}
 
@@
 We think that this will give the ILP better flexibility to handle noise and novel link types while still trying to learn the overall link grammar.
+
+\section{Conclusions}
+
+We have presented an automatic ILP-based method to ``orient'' link grammar parses in multiple languages, turning them into rooted connected DAGs.
+This improves their linguistic interpretability and provides new corpora for experimenting with multi-headed dependency parsers.
+
+More generally, ILP may be a valuable technology for enriching existing annotated corpora.
+For example, the Penn Treebank \cite{TREEBANK-1993} omitted types of annotations that plausibly could be reconstructed by rule.
 
@@
 
-% OLD
+% OLD NERDVIEW EXPLANATION
 % For every arc:
@@
 
-% \subsection{Link Parser Annotations}
-% These sentences have been taken from the English bnews training data set. We chose to display sentences with complete link parses similar to the analysis section of this paper. 
-% 
-% \subsection{Parse Colorization}
-% The sentence used in the middle of the parse is taken from the output of the link parser. We removed sentences with dropped words, but they would have shown up in the parse as [bracketed].
-
 % Uncomment this to add in all of the parses. Warning: compilation will take a long time.

### 2014-06-03 doc/LinksToDAG.tex
space hacking
my changes

--- 2014-06-03 doc/LinksToDAG.tex
@@
 The link grammar parses do differ in style from the single-headed CoNLL parses of the same English sentences.
-They have !!!\% more edges overall.
-Only !!!\% of the links match CoNLL arcs, and only !!!\% of the CoNLL arcs match links.
+They have \input{figure/conll_analysis_edge_percent.tex}more edges overall.
+Only \input{figure/conll_analysis_precision.tex}of the links match CoNLL arcs, and only \input{figure/conll_analysis_recall.tex}of the CoNLL arcs match links.
 % DATA FOR THE ABOVE
@@
 \noteJH{A run of 10,000 sentences takes 3142 seconds (update in final version!)}  Detailed results from this run are included as Appendix~\ref{app:linktypes}.
-On English, !!!
-of !!!
-link {\em types} allowed both directions, and only $\sum_{ij} s_{ij} = !!!$ of !!!
-link {\em tokens} required a disallowed direction via slack.
-\input{figure/conll_analysis_sentences_multiheaded}of the English sentences and !!! of the Russian sentences had at least one multiheaded word.
+On English, \input{figure/allowed_labels_both.tex}of \input{figure/allowed_labels_total.tex}link {\em types} allowed both directions, and only $\sum_{ij} s_{ij} = \input{figure/label_token_disallowed}$ of \input{figure/label_token_total}link {\em tokens} required a disallowed direction via slack.
+\input{figure/conll_analysis_sentences_multiheaded}of the English sentences and 0.15\% of the Russian sentences had at least one multiheaded word.
 

### 2014-06-03 doc/LinksToDAG.tex

--- 2014-06-03 doc/LinksToDAG.tex
@@
 \begin{figure*}[ht!]
+  \vspace{-6pt}
   \centering
@@
   \label{fig:parses}
+  \vspace{-6pt}
 \end{figure*}
@@
 Our largest run took 1.5 hours.
-\noteJE{update in final version}  Detailed results from this run are included as Appendix~\ref{app:linktypes}.
-On English, !!!
-of !!!
-link {\em types} allowed both directions, and only $\sum_{ij} s_{ij} = !!!$ of !!!
-link {\em tokens} required a disallowed direction via slack.
-\input{figure/conll_analysis_sentences_multiheaded}of the English sentences and !!! of the Russian sentences had at least one multiheaded word.
+\noteJH{A run of 10,000 sentences takes 3142 seconds (update in final version!)}  Detailed results from this run are included as Appendix~\ref{app:linktypes}.
+On English, \input{figure/allowed_labels_both.tex}of \input{figure/allowed_labels_total.tex}link {\em types} allowed both directions, and only $\sum_{ij} s_{ij} = \input{figure/label_token_disallowed}$ of \input{figure/label_token_total}link {\em tokens} required a disallowed direction via slack.
+\input{figure/conll_analysis_sentences_multiheaded}of the English sentences and 0.15\% of the Russian sentences had at least one multiheaded word.
 \noteJE{2 significant digits is enough}
-
-\noteJE{fill these numbers in}
+\noteJE{update these numbers}
 
@@
 Recall grows as more link types are seen.}\label{fig:convergence}
+\vspace{-6pt}
 \end{figure}

### 2014-06-03 src/tools/experiments/conll_analysis.py
analysis

@@ -42,6 +42,12 @@ LATEX_FILE_SENTENCE_MULTIHEADED = LATEX_DIR+"conll_analysis_sentences_multiheade
 LATEX_FILE_SENTENCE_TOTAL = LATEX_DIR+"conll_analysis_sentences_total.tex"
 LATEX_FILE_SENTENCE_ORIGINAL = LATEX_DIR+"conll_analysis_sentences_original.tex"
 LATEX_FILE_ARCS = LATEX_DIR+"conll_analysis_arcs.tex"
+LATEX_FILE_CONLL_RECALL = LATEX_DIR+"conll_analysis_recall.tex"
+LATEX_FILE_CONLL_PRECISION = LATEX_DIR+"conll_analysis_precision.tex"
+LATEX_FILE_EDGE_PERCENT = LATEX_DIR+"conll_analysis_edge_percent.tex"
+
+
+
 if not os.path.exists(LATEX_DIR):
     os.makedirs(LATEX_DIR)
 
@@ -60,6 +66,9 @@ ALLOWED_LABELS = SOL_DIR + "allowedLinks.txt"
 ALLOWED_LABELS = open(ALLOWED_LABELS, "r")
 ALLOWED_LABELS_BOTH = TIKZ_DIR+"allowed_labels_both.tex"
 ALLOWED_LABELS_TOTAL = TIKZ_DIR+"allowed_labels_total.tex"
+
+LABEL_TOKEN_DISALLOWED = TIKZ_DIR+"label_token_disallowed.tex"
+LABEL_TOKEN_TOTAL = TIKZ_DIR+"label_token_total.tex"
 
 allowedLabel = {}
 ALLOWED_LABELS = ALLOWED_LABELS.readlines()
@@ -558,6 +567,12 @@ f.write(result)
 f.close()
 
 
+
+
+
+
+
+
 #---------------------------------------------------------------------
 # Take our generated numbers and put them in latex files 
 # to be called by the paper
@@ -849,6 +864,52 @@ f.close()
 
 
 pprint(link_direction_coarse_totals)
-#pprint(
-#for coarse_label in link_direction_coarse_totals:
-#    direction
+#pprint(allowedLabel)
+
+errors = 0
+total = 0
+for coarse_label in link_direction_coarse_totals:
+    direction = link_direction_coarse_totals[coarse_label]
+
+    left = direction.get("left",0)
+    right = direction.get("right",0)
+
+    allowedDirection = allowedLabel[coarse_label]
[…]

### 2014-06-03 src/tools/experiments/conll_analysis.py
conll analysis

@@ -786,13 +786,13 @@ f = open(LATEX_FILE_COARSE_LINKS, "w+")
 
 
 # TODO use tabular
-latex_table = "\\begin{tabular}{|l|l|l|l|l|l|}\n"
-end_table = "\\end{tabular}\n"
+latex_table = "\\begin{longtable}{|l|l|l|l|l|l|}\n"
+end_table = "\\end{longtable}\n"
 
 header = "Label & Rightward & Multiheaded & CoNLL Match & CoNLL Dir Match & CoNLL Label\\\\ \n"
 
-begin_figure = "\\begin{figure*}\n\\small\n\\centering\n"
-end_figure = "\\end{figure*}\n"
+begin_figure = "\\begin{small}\n\\centering\n"
+end_figure = "\\end{small}\n"
 
 def make_percentage_figure(top,bottom):
     return str(int(float(top) / float(bottom) * 100 + 0.5))+"\\%" + " ("+str(top)+"/"+str(bottom)+")"

### 2014-06-03 doc/LinksToDAG.tex
everything
makefile now creates body.pdf and supplement.zip minor edits

@@
 \noteJH{A run of 10,000 sentences takes 3142 seconds (update in final version!)}  Detailed results from this run are included as Appendix~\ref{app:linktypes}.
-On English, \input{figure/allowed_labels_both.tex}of \input{figure/allowed_labels_total.tex}link {\em types} allowed both directions, and only $\sum_{ij} s_{ij} = \input{figure/label_token_disallowed}$ of \input{figure/label_token_total}link {\em tokens} required a disallowed direction via slack.
-\input{figure/conll_analysis_sentences_multiheaded}of the English sentences and 0.15\% of the Russian sentences had at least one multiheaded word.
+On English, only \input{figure/allowed_labels_both.tex}of \input{figure/allowed_labels_total.tex}link {\em types} allowed both directions, and only $\sum_{ij} s_{ij} = \input{figure/label_token_disallowed}$ of \input{figure/label_token_total}link {\em tokens} required a disallowed direction via slack.
+\input{figure/conll_analysis_sentences_multiheaded}of the English sentences but (disappointingly) only 0.15\% of the Russian sentences had at least one multiheaded word.
 \noteJE{2 significant digits is enough}
@@
 Recall that \url{http://www.abisource.com/projects/link-grammar/dict/summarize-links.html} documents the 122 link types in the English link grammar lexicon.
-Below is the result of our largest English run, whose corpus uses !!!
-of these types.
+Below is the result of our largest English run, whose corpus uses 107 of these types.
+\noteJH{automatically generate this number}
 
@@
 % Uncomment this to add in all of the parses. Warning: compilation will take a long time.
-%\input{figure/parses.tikz}
+\input{figure/parses.tikz}
 

### 2014-06-03 doc/LinksToDAG.tex
back to 4 pp

@@
 On English, only \input{figure/allowed_labels_both.tex}of \input{figure/allowed_labels_total.tex}link {\em types} allowed both directions, and only $\sum_{ij} s_{ij} = \input{figure/label_token_disallowed}$ of \input{figure/label_token_total}link {\em tokens} required a disallowed direction via slack.
-\input{figure/conll_analysis_sentences_multiheaded}of the English sentences but (disappointingly) only 0.15\% of the Russian sentences had at least one multiheaded word.
+\input{figure/conll_analysis_sentences_multiheaded}of the English sentences but only 0.15\% of the Russian sentences had at least one multiheaded word.
 \noteJE{2 significant digits is enough}

### 2014-06-03 doc/LinksToDAG.tex
back to 4 pp
FINAL SUBMISSION - BUT WARNING - THE APPENDICES USE TOO SMALL A RUN AND THE GRAPH IS OUT OF DATE AS WELL - TRY TO RESUBMIT BY EMAIL Edited and clarified business about subject-verb direction Munged table for Appendix A to use longtable Removed triple pdflatex call in makefil

@@
 On English, only \input{figure/allowed_labels_both.tex}of \input{figure/allowed_labels_total.tex}link {\em types} allowed both directions, and only $\sum_{ij} s_{ij} = \input{figure/label_token_disallowed}$ of \input{figure/label_token_total}link {\em tokens} required a disallowed direction via slack.
-\input{figure/conll_analysis_sentences_multiheaded}of the English sentences but only 0.15\% of the Russian sentences had at least one multiheaded word.
+\input{figure/conll_analysis_sentences_multiheaded}of the English sentences but only 0.15\% of the Russian ones had at least one multiheaded word.
 \noteJE{2 significant digits is enough}
@@
 
-The main vexation is that subject-verb links are handled inconsistently.
+The main vexation is the handling of subject-verb links.
 Under the English link grammar, the verb (or 0) that governs a clause will link to both the clause's subject and its last (main) verb.
 This ought to give us the desired treatment of {\em ``Jill persuaded him to skip''}, in which {\em ``him''} has two parents.
-Unfortunately, it does not, because of inconsistencies in the link grammar itself.\footnote{When the subject is missing ({\em ``Jill wanted to skip''}), the governing verb now links instead to the clause's first (tensed) verb but {\em no longer} to the last as well.
+Unfortunately, we get {\em him} $\rightarrow$ {\em to}, because of inconsistencies in the link grammar itself.\footnote{When the subject is missing ({\em ``Jill wanted to skip''}), the governing verb now links instead to the clause's first (tensed) verb but {\em no longer} to the last as well.
 That favors a consistent solution where each auxiliary verb is a parent of the verb to its right ({\em to} $\rightarrow$ {\em skip}).
 As a result, when the subject is present, it must be the {\em parent} of the first verb, which otherwise would not have a parent ({\em him} $\rightarrow$ {\em to} $\rightarrow$ {\em skip}).
-Yet also, subjects must be {\em children} of verbs in a few  inconsistently handled constructions such as non-embedded reported speech ({\em ``It was impossible, they said''}), since the subject ({\em ``they''}) links only to the verb ({\em ``said''}) and so has no other possible parent.}
+(Yet subjects are not {\em always} parents.
+The occasional construction {\em ``It was impossible, they said''} is handled inconsistently by the link grammar:
+the subject ({\em ``they''}) links only to the verb ({\em ``said''}) and so must be its child, no other parent being available.)}
 

### 2014-06-03 src/tools/experiments/conll_analysis.py
merged the table. for now.

@@ -863,7 +863,7 @@ f.close()
 
 
 
-pprint(link_direction_coarse_totals)
+#pprint(link_direction_coarse_totals)
 #pprint(allowedLabel)
 
 errors = 0

### 2014-06-04 doc/LinksToDAG.tex
merged the table. for now.
clarification in last paragraph before conclusions

@@
 Under the English link grammar, the verb (or 0) that governs a clause will link to both the clause's subject and its last (main) verb.
-This ought to give us the desired treatment of {\em ``Jill persuaded him to skip''}, in which {\em ``him''} has two parents.
-Unfortunately, we get {\em him} $\rightarrow$ {\em to}, because of inconsistencies in the link grammar itself.\footnote{When the subject is missing ({\em ``Jill wanted to skip''}), the governing verb now links instead to the clause's first (tensed) verb but {\em no longer} to the last as well.
-That favors a consistent solution where each auxiliary verb is a parent of the verb to its right ({\em to} $\rightarrow$ {\em skip}).
-As a result, when the subject is present, it must be the {\em parent} of the first verb, which otherwise would not have a parent ({\em him} $\rightarrow$ {\em to} $\rightarrow$ {\em skip}).
-(Yet subjects are not {\em always} parents.
-The occasional construction {\em ``It was impossible, they said''} is handled inconsistently by the link grammar:
+This would permit our desired treatment of {\em ``Jill persuaded him to skip''}, in which {\em ``him''} has two parents.
+But the ILP solution generally treats subjects as {\em parents} of verbs (thus we get {\em him} $\rightarrow$ {\em to}), owing to inconsistencies in the link grammar itself.\footnote{Why?
+When the subject is missing ({\em ``Jill wanted to skip''}), the governing verb now links instead to the clause's first (tensed) verb but {\em no longer} to the last as well.
+That means that a VP must be headed by its first verb ({\em to} $\rightarrow$ {\em skip}).
+As a result, when the subject is present, it must be the {\em parent} of the first verb ({\em him} $\rightarrow$ {\em to} $\rightarrow$ {\em skip}), since no other parent is available.
+({\em Note:} Subjects are not {\em always} parents.
+The construction {\em ``It was impossible, they said''} is handled inconsistently by the link grammar:
 the subject ({\em ``they''}) links only to the verb ({\em ``said''}) and so must be its child, no other parent being available.)}
@@
 % Uncomment this to add in all of the parses. Warning: compilation will take a long time.
-\input{figure/parses.tikz}
+%\input{figure/parses.tikz}
 

### 2014-06-10 src/tools/experiments/conll_analysis.py
fixed the production of the longtable at the end of the paper.

@@ -10,7 +10,7 @@ DEBUG = False
 
 # The directory containing the original conll data file
 CONLL_DIR = "data/"
-CONLL_LOC = CONLL_DIR+"english_bnews_train.conll"
+CONLL_LOC = CONLL_DIR+"english_ptb_train.conll"
 if len(sys.argv) > 2:
     CONLL_LOC = sys.argv[2]
 if not os.path.exists(CONLL_DIR):
@@ -786,10 +786,10 @@ f = open(LATEX_FILE_COARSE_LINKS, "w+")
 
 
 # TODO use tabular
-latex_table = "\\begin{longtable}{|l|l|l|l|l|l|}\n"
+latex_table = "\\begin{longtable}{|l|l|l|l|l|l|}\n\\hline\n"
 end_table = "\\end{longtable}\n"
 
-header = "Label & Rightward & Multiheaded & CoNLL Match & CoNLL Dir Match & CoNLL Label\\\\ \n"
+header = "Label & Rightward & Multiheaded & CoNLL Match & CoNLL Dir Match & CoNLL Label\\\\"
 
 begin_figure = "\\begin{small}\n\\centering\n"
 end_figure = "\\end{small}\n"
@@ -800,8 +800,13 @@ def make_percentage_figure(top,bottom):
 
 table = begin_figure
 table += latex_table
-table += "\\hline\n"
-table += header
+
+table += header + "\\hline\n"
+table += "\\endhead\n"
+table += "\n"
+
+table += "\\hline\n\\endfoot\n\n"
+
 line = 0
 for coarse_label in coarse_labels:
     count = link_label_coarse_counts[coarse_label]
@@ -833,7 +838,7 @@ for coarse_label in coarse_labels:
         match_percent = "-"
         mismatch_percent = "-"
 
-    table += "\\hline\n"
+    #table += "\\hline\n"
     table += coarse_label
     table += " & "+right
     table += " & "+multiheaded
@@ -844,7 +849,7 @@ for coarse_label in coarse_labels:
     line += 1
     
     # Break up the table into smaller tables
-    if line % 55 == 0:
+    """if line % 55 == 0:
         table += "\\hline\n"
         table += end_table
         table += end_figure
@@ -852,8 +857,8 @@ for coarse_label in coarse_labels:
         table += latex_table
         table += "\\hline\n"
         table += header
-    
-table += "\\hline\n"
+    """
+#table += "\\hline\n"
 table += end_table
 table += end_figure
 

### 2014-06-10 src/tools/plotter.py
added a few extra options with plotter.py. I moved the legend to the upper left corner for now. I think I'll add command line arguments that will allow us to specify where we want it to go.

@@ -1,6 +1,6 @@
 #!/usr/bin/env python
 
-import math, argparse, sys
+import math, numpy, argparse, sys
 import matplotlib.pyplot as plt
 from pprint import pprint
 
@@ -10,7 +10,11 @@ parser.add_argument("-o", "--output", dest="output", default="sol/runtimes.png",
 parser.add_argument("-x","--x", dest="xLabel", default="Sentences",
                     help="Specify the x axis.")
 parser.add_argument("-y","--y", dest="yLabel", default="Run Time (seconds)", 
-                    help="Speciyf the y axis.")
+                    help="Specify the y axis.")
+parser.add_argument("-loff","--legend_off", dest="legend", default=True, action="store_false",
+                    help="Turn the legend off")
+parser.add_argument("-ylog","--yaxis_log", dest="ylog", default=False, action="store_true",
+                    help="Use the log scale on the y axis")
 parser.add_argument("strings", nargs="*")
 
 args = parser.parse_args()
@@ -77,18 +81,32 @@ for lines in file_lines:
     file_coord.append(coord)
 
 
-# Get the minimum, maximum X values
+# Get the minimum, maximum X,Y values
 xmax = None
 xmin = 0
-for coord in file_coord:
-    temp_max = max(coord)
-    if xmax == None or xmax < temp_max:
-        xmax = temp_max
+ymax = None
+ymin = 0
 
-    temp_min = min(coord)
-    if xmin > temp_min:
-        xmin = temp_min
-xmax = xmax[0]
+for coord in file_coord:
+
+    temp_xmax = max([i[0] for i in coord])
+    if xmax == None or xmax < temp_xmax:
+        xmax = temp_xmax
+
+    temp_xmin = min([i[0] for i in coord])
+    if xmin > temp_xmin:
+        xmin = temp_xmin
+
+
+    temp_ymax = max([i[1] for i in coord])
+    if ymax == None or ymax < temp_ymax:
+        ymax = temp_ymax
+
+    temp_ymin = min([i[1] for i in coord])
+    if ymin > temp_ymin:
+        ymin = temp_ymin
+
+
 
 
 
@@ -106,7 +124,7 @@ for l,coord in zip(labels,file_coord):
          RE:"red",
          PR:"DeepSkyBlue",
          RR:"blue"
[…]

### 2014-07-07 src/tools/conll/multiheaded.py
multiheaded.py prints out the number of multiheaded sentences

@@ -1,0 +1,48 @@
+#!/usr/bin/env python
+
+import sys
+
+
+filename = "sol/links_ru.conll"
+if len(sys.argv) > 1:
+    filename = sys.argv[1]
+
+f = open(filename,"r")
+f = f.readlines()
+
+
+sentenceCount = 0
+multiheadedSentenceCount = 0
+sentence = []
+heads = []
+for line in f:
+    line = line.strip()
+
+    if not line:
+        
+        sentenceCount += 1
+
+        for head in heads:
+            if len(head) > 1:
+                multiheadedSentenceCount += 1
+                break
+
+        sentence = []
+        heads = []
+
+        continue
+
+    line = line.split()
+    head = line[6].split(",")
+    word = line[1]
+
+    sentence.append(word)
+    heads.append(tuple(head))
+    
+    if len(head) > 1:
+        print word, head
+
+
+print multiheadedSentenceCount
+#print sentenceCount
+#print float(multiheadedSentenceCount) / float(sentenceCount)

### 2014-07-07 src/tools/conll/multiheaded.py
added some basic conll analysis tools.

@@ -39,8 +39,8 @@ for line in f:
     sentence.append(word)
     heads.append(tuple(head))
     
-    if len(head) > 1:
-        print word, head
+    #if len(head) > 1:
+    #    print word, head
 
 
 print multiheadedSentenceCount

### 2014-07-07 src/tools/conll/printsentence.py

@@ -1,6 +1,24 @@
 #!/usr/bin/env python
 
-from utils import *
+import fileinput
+
+def readInput():
+    lines = []
+    for line in fileinput.input():
+        lines.append(line.strip())
+    return lines
+
+def getSentences(conll):
+    sentence = []
+    sentences = []
+    for line in conll:
+        if not line:
+            sentences.append(sentence)
+            sentence = []
+        else:
+            sentence.append(line)
+    return sentences
+
 
 def printSentences():
     lines = readInput()

### 2014-07-07 src/tools/experiments/conll_analysis.py
fiddled with the length parameters of tikz drawings. Also removed an example from the paper.

@@ -447,12 +447,12 @@ for linkSentence in link_results:
     # I am preventing certain sentences from appearing in the paper.
     # I just wanted to skip over some sentences because they didn't look cool enough
     #bannedWords = ["salees", "soldiers", "word", "serwer", "reason"]
-    """bannedWords = []
+    bannedWords = ["consensus"]
     for word in bannedWords:
         if word in linkSentenceCheck:
             sentenceCheck = False
             break
-    """
+    
 
 
     tempSentence = [] 
@@ -466,7 +466,7 @@ for linkSentence in link_results:
 
 
     # Link parses to put in the paper. Takes sentences of only length 5.
-    if (paper_sentence_count < paper_sentence_limit) and sentenceLength >= 25 and sentenceLength <= 30 and sentenceCheck:
+    if (paper_sentence_count < paper_sentence_limit) and sentenceLength >= 20 and sentenceLength <= 25 and sentenceCheck:
         if paper_sentence_skip > 0:
             paper_sentence_skip -= 1
         else:

### 2014-07-07 src/tools/experiments/trial_ru.sh
trial_ru.sh

@@ -19,6 +19,7 @@ tail --lines 1 runtime >> /tmp/LinksToDAG_times
 
 if [ $LANGUAGE ]; then
     cp sol/allowedLinks.txt /tmp/LinksToDAG_trial_sol/allowedLinks_$ITERATIONS\_$LANGUAGE.txt
+
     #python src/tools/experiments/conll_analysis.py $SENTENCES $CONLL $ITERATIONS $LANGUAGE
     #mv sol/conll_analysis/conll_analysis.txt sol/conll_analysis/conll_analysis_$ITERATIONS\_$LANGUAGE.txt
 else
@@ -26,6 +27,11 @@ else
     #python src/tools/experiments/conll_analysis.py $SENTENCES $CONLL $ITERATIONS
     #mv sol/conll_analysis/conll_analysis.txt sol/conll_analysis/conll_analysis_$ITERATIONS.txt
 fi
+
+
+
+
+
 
 rm -f runtime
 

### 2014-07-08 doc/LinksToDAG.tex
reenabled the TODO comments in the paper. Commented one out (making the longtable at the end in the appendix).
added reviews-initial.txt
updated the paper such that the russian numbers are automatically generated.

@@
 We used the English sentences from the CoNLL 2007 Shared Task \cite{CONLL-SHARED-2007}---a subset of the Penn Treebank for which {\em single}-headed reference parses are available. 
-We also used a prefix of the Russian National Corpus,\footnote{That is, a prefix when considering the files in Unix sort order by filename.
+%We also used a prefix of the Russian National Corpus,\footnote{That is, a prefix when considering the files in Unix sort order by filename.
 We dropped sentences with $\leq 2$ words, as well as file 22592319, which appeared to contain charts.} which is unparsed.
+We also used a prefix of the Russian News Commentary data from the ACL 2013 Shared Task of Machine Translation,\footnote{\url{http://www.statmt.org/wmt13/training-monolingual-nc-v8.tgz}} which is unparsed.
 
 We generated link parses using the AbiWord/CMU link grammar parser version 5.0.8 \cite{LINKPARSER-2014}.
-The parser's coverage is not perfect: we obtained connected parses for only \input{figure/conll_analysis_sentences_total.tex}(of \input{figure/conll_analysis_sentences_original.tex}$\!\!$) English sentences and only 8,018 (of 10,000) Russian sentences, discarding the other sentences.\footnote{For Russian only, we permitted the parser to drop words if necessary.
-Otherwise only 208 sentences were parseable.}
+The parser's coverage is not perfect: we obtained connected parses for only \input{figure/conll_analysis_sentences_total.tex}(of \input{figure/conll_analysis_sentences_original.tex}$\!\!$) English sentences and only \input{figure/russian_sentence_count.tex}(of \input{figure/russian_original_sentence_count.tex}$\!\!$) Russian sentences, discarding the other sentences.
 These two languages have the most mature lexicons at present, although lexicons for 7 other languages are available.
@@
 \noteJH{A run of 10,000 sentences takes 3142 seconds (update in final version!)}  Detailed results from this run are included as Appendix~\ref{app:linktypes}.
-On English, only \input{figure/allowed_labels_both.tex}of \input{figure/allowed_labels_total.tex}link {\em types} allowed both directions, and only $\sum_{ij} s_{ij} = \input{figure/label_token_disallowed}$ of \input{figure/label_token_total}link {\em tokens} required a disallowed direction via slack.
-\input{figure/conll_analysis_sentences_multiheaded}of the English sentences but only 0.15\% of the Russian ones had at least one multiheaded word.
-\noteJE{2 significant digits is enough}
-\noteJE{update these numbers}
+On English, only \input{figure/allowed_labels_both.tex}of \input{figure/allowed_labels_total.tex}link {\em types} allowed both directions, and only $\sum_{ij} s_{ij} = \input{figure/label_token_disallowed}$ of \input{figure/label_token_total.tex}link {\em tokens} required a disallowed direction via slack.
+\input{figure/conll_analysis_sentences_multiheaded.tex}of the English sentences but only \input{figure/russian_multiheaded.tex}of the Russian ones had at least one multiheaded word.
+%\noteJE{2 significant digits is enough}
+%\noteJE{update these numbers}
 
@@
 
-\newpage
+% \newpage
 \bibliographystyle{acl2014}
@@
 Recall that \url{http://www.abisource.com/projects/link-grammar/dict/summarize-links.html} documents the 122 link types in the English link grammar lexicon.
-Below is the result of our largest English run, whose corpus uses 107 of these types.
-\noteJH{automatically generate this number}
+Below is the result of our largest English run, whose corpus uses \input{figure/allowed_labels_total.tex}of these types.
+\noteJH{automatically generate this number correctly: 107}
 
@@
 
-\NoteJE{improve table formatting.  Numeric justification; en dashes; longtable; remove most hlines.}
+%\NoteJE{improve table formatting.  Numeric justification; en dashes; longtable; remove most hlines.}
 

### 2014-07-08 src/tools/conll/multiheaded.py
changed the output of multiheaded.py to a percentage.

@@ -43,6 +43,6 @@ for line in f:
     #    print word, head
 
 
-print multiheadedSentenceCount
+#print multiheadedSentenceCount
 #print sentenceCount
-#print float(multiheadedSentenceCount) / float(sentenceCount)
+print str(round(float(multiheadedSentenceCount) / float(sentenceCount), 2))+"\%"

### 2014-07-08 trials.sh
updated the russian trials to generate the necessary numbers for the paper.

@@ -3,7 +3,8 @@
 typeset -i i END
 
 
-END=10000
+END=100
+#END=10000
 #END=60030
 #END=16384
 #INCREMENT=1
@@ -49,6 +50,17 @@ if [ $LANGUAGE -a $LANGUAGE = "ru" ]; then
     if [ $i -ne $END -a $((i/2)) -ne $END ]; then
         bash src/tools/experiments/trial_ru.sh $SENTENCES $END
     fi
+
+    mv sol/links.conll sol/links_ru.conll
+    mv sol/allowedLinks.txt sol/allowedLinks_ru.txt
+
+    echo $END > doc/figure/russian_original_sentence_count.tex
+    python src/tools/conll/multiheaded.py sol/links_ru.conll > doc/figure/russian_multiheaded.tex
+    python src/tools/conll/printsentence.py sol/links_ru.conll > temp
+    wc -l temp | cut -d ' ' -f1 > doc/figure/russian_sentence_count.tex
+    rm -f temp
+
+
 else     
     for ((i=1;i<=$END;i=$((i*2))));
     do

### 2014-07-08 trials.sh
trials.sh

@@ -3,14 +3,15 @@
 typeset -i i END
 
 
-END=100
+
+#END=100
 #END=10000
 #END=60030
-#END=16384
+END=18577
 #INCREMENT=1
 
-SENTENCES="data/english_bnews_train.sentences"
-CONLL="data/english_bnews_train.conll"
+SENTENCES="data/english_ptb_train.sentences"
+CONLL="data/english_ptb_train.conll"
 LANGUAGE=""
 
 if [ $1 ]; then
@@ -104,7 +105,7 @@ fi
 
 
 
-echo $RUNTIMES $END $'\n'The CLSP machines cant run plotter. Please download files and run: $'\n'python src/tools/plotter.py -x "Sentences" -y "Percent" -o $SOL_PRECISION_RECALL $TXT_PRECISION $TXT_RECALL | mutt -s $END -a $RUNTIMES -a $TXT_PRECISION -a $TXT_RECALL -- junekihong@gmail.com
+#echo $RUNTIMES $END $'\n'The CLSP machines cant run plotter. Please download files and run: $'\n'python src/tools/plotter.py -x "Sentences" -y "Percent" -o $SOL_PRECISION_RECALL $TXT_PRECISION $TXT_RECALL | mutt -s $END -a $RUNTIMES -a $TXT_PRECISION -a $TXT_RECALL -- junekihong@gmail.com
 
 #-a $TYPE_ANALYSIS -a $CONLL_ANALYSIS -- junekihong@gmail.com
 

### 2014-07-08 src/tools/experiments/conll_analysis.py
quick fix for conll_analysis.py

@@ -449,9 +449,8 @@ for linkSentence in link_results:
     #bannedWords = ["salees", "soldiers", "word", "serwer", "reason"]
     bannedWords = ["consensus"]
     for word in bannedWords:
-        if word in linkSentenceCheck:
-            sentenceCheck = False
-            break
+        sentenceCheck = False
+        break
     
 
 

### 2014-07-10 doc/reviews-response-draft.txt
added author response

Happily, Reviewers #2 and #3's concerns about evaluation may be answered in the supplementary material. (You were not required to review the supplementary material, but we can point to it in our author response.)  We do not think that there is an evaluation gap in the paper, except for Reviewer #1's excellent suggestion that we could also compare with other semantic dependency representations (which may however be non-planar, disconnected, or not manually checked).

Appendix A in fact contains an evaluation like what is requested by reviewer #2.  Consider for example the first row.  It shows that translating each link grammar "A" edge to a leftward CoNLL "NMOD" edge would indeed be successful: this matches the CoNLL direction and label in 99% of the cases when CoNLL has an edge there at all (which is true for 86% of the 426 "A" links).

Scanning the last 3 columns of this table gives a clear picture of where the directed link parses line up with the CoNLL parses.  We could have artificially summarized the Appendix A table using one or two numbers.  However, we felt that the real information was in the line-by-line (type-by-type) pattern, in particular the fact that most link grammar label types mapped cleanly to a single direction, and that in most cases this agreed with the CoNLL direction and corresponded strongly to a single CoNLL label.  We also provided 100 actual parses in Appendix B so that researchers could see further whether our DAG parses would suit their research needs.

Note, however, that we are not holding out the converted link grammar parses as an ideal syntactic representation.  Indeed, we strongly believe that the "real" way to work on DAG parsers would be to produce an annotation scheme and a gold annotated resource.  Lacking one, the questions that we aimed to answer here were:

* Our experimental hypothesis was that the link grammar writers mentally associated a consistent direction with each link type, even though they did not annotate it.  We found that we could indeed recover consistent directions that satisfy the DAG constraints.  The first column of Appendix A can be directly applied to transform link parses into DAG parses without re-running the ILP.

* Are the converted link grammar parses good enough that people could use them to test PARSING ALGORITHMS?  This does not require that the parses agree with other annotation schemes (or even that they are linguistically defensible).  To experiment with the speed and accuracy of algorithms, one only needs a PLAUSIBLE distribution of data for train/test, which mimics the kind of data that one might have in a gold standard.  This is what we provide.  The point of the supplementary material is to demonstrate the plausibility.

* There has been a lot of ongoing development on link grammars for English and other languages, but it has been almost entirely outside the ACL community.  Is there a way to bring this work into the fold?  As far as we know, this is the first attempt.  These parses may serve as a source of extra features for other prediction or learning tasks (we actually had particular uses in mind that motivated the work).

* We pioneer the idea of using ILP to enhance incomplete annotations, using the ILP constraints to enforce internal consistency as well as formal requirements or linguistic preferences.

To Reviewer #3: There are no multi-planar link grammars at present.  If there were, yes, our ILP method would immediately extend.  In the meantime, one could do research on multi-planar DAG parsing by overlaying our DAG parses on the CoNLL trees (2 planes).

Thanks very much to Reviewer #1 for the additional citations and other suggestions.  (Regarding "converting corpora from other formats such as HPSG," we were referring to the work cited in the previous sentence.)

### 2014-07-10 doc/reviews-response.txt
added author response

Our SUPPLEMENTARY MATERIAL speaks to your evaluation concerns!

Reviewer #2: The last 3 columns of Appendix A show how the directed link parses line up with the CoNLL parses.  Most link grammar label types map cleanly to a single direction.  Usually this agrees with the CoNLL direction and corresponds strongly to a single CoNLL label.  We also provided 100 actual parses in Appendix B so that researchers could see further whether our DAG parses would suit their research needs.

(Row 1 shows that translating each link grammar "A" edge to a leftward CoNLL "NMOD" edge would match the CoNLL direction+label in 99% of the cases when CoNLL has a corresponding edge; that's 365 of 426 cases.)

But agreement with CoNLL wasn't our goal!  Our questions:

* Can we help people working on parsing algorithms?  Apparently yes.  We have produced a multi-headed projective DAG corpus.  The parses aren't gold, might not agree with other annotation schemes, or be linguistically defensible.  But they still give a GENERALLY PLAUSIBLE TRAINING/TEST SET for experimenting with the speed/accuracy of DAG parsing ALGORITHMS.  Someone should then build a gold corpus.

* Can ILP be used to enhance incomplete annotations?  Yes: we show how to enforce internal consistency along with well-formedness requirements.

* Did the link grammar writers have a link direction "in mind" even though it's not annotated?  Apparently yes.  In Appendix A we supply the solution.

* Can we ACLers make any use of the ignored labor of the link grammar writers?  Maybe.  These directed parses may serve as a source of extra features for other tasks.

Reviewer #3: No multi-planar link grammars exist at present.  If there were, our ILP method would immediately extend.  In the meantime, one can do research on 2-planar DAG parsing by overlaying our DAG parses on the CoNLL trees.

Reviewer #1: Thanks for the citations and suggestions.  We like the suggestions of other semantic dependency representations to compare (perhaps non-planar, disconnected, or not-manually-checked).

### 2014-07-10 src/tools/experiments/conll_analysis.py
small fix for conll_analysis.py that was preventing it from displaying any example sentences in the paper.

@@ -449,8 +449,9 @@ for linkSentence in link_results:
     #bannedWords = ["salees", "soldiers", "word", "serwer", "reason"]
     bannedWords = ["consensus"]
     for word in bannedWords:
-        sentenceCheck = False
-        break
+        if word in linkSentence:
+            sentenceCheck = False
+            break
     
 
 

### 2014-07-14 src/tools/addcommas.sh
found a script online that would add commas to a number. Used it to add commas in my russian sentence count figure.

@@ -1,0 +1,3 @@
+sed -e ': L
+s/\([0-9]\{1,19\}\)\([0-9]\{3\}\)/\1,\2/
+t L'

### 2014-07-14 trials.sh

@@ -4,10 +4,10 @@ typeset -i i END
 
 
 
-#END=100
+END=100
 #END=10000
 #END=60030
-END=18577
+#END=18577
 #INCREMENT=1
 
 SENTENCES="data/english_ptb_train.sentences"
@@ -57,8 +57,11 @@ if [ $LANGUAGE -a $LANGUAGE = "ru" ]; then
 
     echo $END > doc/figure/russian_original_sentence_count.tex
     python src/tools/conll/multiheaded.py sol/links_ru.conll > doc/figure/russian_multiheaded.tex
+
+
+
     python src/tools/conll/printsentence.py sol/links_ru.conll > temp
-    wc -l temp | cut -d ' ' -f1 > doc/figure/russian_sentence_count.tex
+    wc -l temp | cut -d ' ' -f1 | bash src/tools/addcommas.sh > doc/figure/russian_sentence_count.tex
     rm -f temp
 
 

### 2014-07-14 src/tools/experiments/conll_analysis.py
merged trials.sh

@@ -466,7 +466,7 @@ for linkSentence in link_results:
 
 
     # Link parses to put in the paper. Takes sentences of only length 5.
-    if (paper_sentence_count < paper_sentence_limit) and sentenceLength >= 20 and sentenceLength <= 25 and sentenceCheck:
+    if (paper_sentence_count < paper_sentence_limit) and sentenceLength >= 24 and sentenceLength <= 28 and sentenceCheck:
         if paper_sentence_skip > 0:
             paper_sentence_skip -= 1
         else:

### 2014-07-14 trials.sh

@@ -4,10 +4,11 @@ typeset -i i END
 
 
 
-END=100
+#END=100
+#END=64
 #END=10000
 #END=60030
-#END=18577
+END=18577
 #INCREMENT=1
 
 SENTENCES="data/english_ptb_train.sentences"
@@ -52,8 +53,12 @@ if [ $LANGUAGE -a $LANGUAGE = "ru" ]; then
         bash src/tools/experiments/trial_ru.sh $SENTENCES $END
     fi
 
-    mv sol/links.conll sol/links_ru.conll
-    mv sol/allowedLinks.txt sol/allowedLinks_ru.txt
+
+    #head sol/links.conll sol/allowedLinks.txt
+    
+
+    cp sol/links.conll sol/links_ru.conll
+    cp sol/allowedLinks.txt sol/allowedLinks_ru.txt
 
     echo $END > doc/figure/russian_original_sentence_count.tex
     python src/tools/conll/multiheaded.py sol/links_ru.conll > doc/figure/russian_multiheaded.tex

### 2014-07-14 src/tools/addcommas.sh
added a citation in addcommas.sh

@@ -1,3 +1,7 @@
+#!/bin/sh
+
+#taken from: http://stackoverflow.com/questions/14668094/add-commas-to-numeric-strings-in-unix
+
 sed -e ': L
 s/\([0-9]\{1,19\}\)\([0-9]\{3\}\)/\1,\2/
 t L'

### 2014-07-15 trials.sh
added commas for the figure of total original russian sentences.

@@ -60,7 +60,7 @@ if [ $LANGUAGE -a $LANGUAGE = "ru" ]; then
     cp sol/links.conll sol/links_ru.conll
     cp sol/allowedLinks.txt sol/allowedLinks_ru.txt
 
-    echo $END > doc/figure/russian_original_sentence_count.tex
+    echo $END | bash src/tools/addcommas.sh > doc/figure/russian_original_sentence_count.tex
     python src/tools/conll/multiheaded.py sol/links_ru.conll > doc/figure/russian_multiheaded.tex
 
 

### 2014-09-18 src/tools/experiments/conll_analysis.py
added some extra things for an additional 5th page of the paper.

@@ -52,14 +52,16 @@ if not os.path.exists(LATEX_DIR):
 
 # Tikz directory and files
 TIKZ_DIR = "doc/figure/"
-EXAMPLE_SENTENCES_LOC = TIKZ_DIR+"sentences.tikz"
 if not os.path.exists(TIKZ_DIR):
     os.makedirs(TIKZ_DIR)
+EXAMPLE_SENTENCES_LOC = TIKZ_DIR+"sentences.tikz"
 PAPER_TIKZ = open(EXAMPLE_SENTENCES_LOC, 'w+')
 
 EXAMPLE_PARSES_LOC = TIKZ_DIR+"parses.tikz"
 EXAMPLE_PARSES = open(EXAMPLE_PARSES_LOC, 'w+')
 
+SAMPLE_EXAMPLE_SENTENCES_LOC = TIKZ_DIR+"sample_parses.tikz"
+SAMPLE_EXAMPLE_PARSES = open(SAMPLE_EXAMPLE_SENTENCES_LOC, 'w+')
 
 ALLOWED_LABELS = SOL_DIR + "allowedLinks.txt"
 ALLOWED_LABELS = open(ALLOWED_LABELS, "r")
@@ -430,6 +432,9 @@ paper_sentence_limit = 3
 example_parses_count = 0
 example_parses_limit = 100
 
+sample_example_parses_count = 0
+sample_example_parses_limit = 4
+
 
 for linkSentence in link_results:
     if linkSentence not in conll_results:
@@ -475,6 +480,16 @@ for linkSentence in link_results:
             paper_sentence_count += 1
             if paper_sentence_count % 3 == 0:
                 PAPER_TIKZ.write("\n")
+    elif sample_example_parses_count < sample_example_parses_limit and sentenceLength >= 40 and sentenceLength <= 50:
+        tikz = tikz_dependency(conll_results[linkSentence], link_results[linkSentence], linkSentence, .97 / 2)
+        #SAMPLE_EXAMPLE_PARSES.write("\\begin{figure*}[ht!]\n")
+        SAMPLE_EXAMPLE_PARSES.write(tikz)
+        #SAMPLE_EXAMPLE_PARSES.write("\\end{figure*}\n\n")
+        sample_example_parses_count += 1
+        if sample_example_parses_count % 2 == 0:
+            SAMPLE_EXAMPLE_PARSES.write("\n")
+
+
     # Filter out sentences to only up to length 16
[…]

### 2014-09-18 src/tools/experiments/trial.sh

@@ -37,4 +37,3 @@ fi
 
 rm -f runtime
 
-

### 2014-09-18 trials.sh

@@ -11,6 +11,7 @@ typeset -i i END
 END=18577
 #INCREMENT=1
 
+
 SENTENCES="data/english_ptb_train.sentences"
 CONLL="data/english_ptb_train.conll"
 LANGUAGE=""

### 2014-09-21 doc/LinksToDAG.tex
Updated figures after rerunning the entire project.
commented out the call to the supplementary example parses in LinksToDAG.tex.
moved section 7 into figures with captions augmented conclusions some other minor edits

@@
 \section{Experiments and Results}
+
 \begin{figure*}[ht!]
@@
   \input{figure/sentences.tikz}
-  \caption{\small Example Parses: The blue upper edges show CoNLL dependency parses; the red lower edges show oriented link grammar parses.
-Vertical edges are those with parent 0.
-See Appendix~\ref{app:parses} for formatting details and more examples.}
+  \input{figure/sample_parses.tikz}
+  \caption{\small The blue upper edges show CoNLL dependency parses; the red lower edges show our oriented version of the link grammar parses.
+Edges are shown as dotted lines if they appear only in one parse; these are highlighted in orange if the child has multiple parents.
+Edges that appear in both parses are solid lines, drawn thicker if the directions do not match.
+Vertical edges have parent 0.
+For 100 example parses, see Appendix~\ref{app:parses} at \url{http://tidyurl.com/LinksToDAG_supplement}.}
   \label{fig:parses}
@@
 
-We solved our ILP problem using the SCIP Optimization Suite \cite{achterberg2009scip}, encoding it using the ZIMPL language \cite{Koch2004}.
-Our largest run took 1.5 hours.
-\noteJH{A run of 10,000 sentences takes 3142 seconds (update in final version!)}  Detailed results from this run are included as Appendix~\ref{app:linktypes}.
-On English, only \input{figure/allowed_labels_both.tex}of \input{figure/allowed_labels_total.tex}link {\em types} allowed both directions, and only $\sum_{ij} s_{ij} = \input{figure/label_token_disallowed}$ of \input{figure/label_token_total.tex}link {\em tokens} required a disallowed direction via slack.
-\input{figure/conll_analysis_sentences_multiheaded.tex}of the English sentences but only \input{figure/russian_multiheaded.tex}of the Russian ones had at least one multiheaded word.
-%\noteJE{2 significant digits is enough}
-%\noteJE{update these numbers}
+\begin{table*}[ht]
+% \renewcommand{\arraystretch}{1.2}    % more space between rows
+\input{figure/sample_link_analysis_coarse_table.tex}
+\caption{Our solution, i.e., our reconstruction of the ``intended'' direction for each link type in the English Link Grammar.
+We also indicate the extent to which each of these link types (1) has a single dominant direction, (2) participates in multi-headed constructions, and (3) corresponds to CoNLL links of a predictable direction and type.
+For space reasons, we show only the start of this table---the full table can be found in Appendix~\ref{app:linktypes} at \url{http://tidyurl.com/LinksToDAG_supplement}.}\label{tab:linktypes}
+\end{table*}
 
-\subsection{Stability of Results}\label{sec:stability}
-\begin{figure}[ht!]
+\begin{figure}[t!]
   \small
@@
 
+We solved our ILP problem using the SCIP Optimization Suite \cite{achterberg2009scip}, encoding it using the ZIMPL language \cite{Koch2004}.
+Our largest run took 1.5 hours.
+\noteJH{A run of 10,000 sentences takes 3142 seconds (update in final version!)}  On English, only \input{figure/allowed_labels_both.tex}of \input{figure/allowed_labels_total.tex}link {\em types} allowed both directions, and only $\sum_{ij} s_{ij} = \input{figure/label_token_disallowed}$ of \input{figure/label_token_total.tex}link {\em tokens} required a disallowed direction via slack.
+\input{figure/conll_analysis_sentences_multiheaded.tex}of the English sentences but (alas) only \input{figure/russian_multiheaded.tex}of the Russian ones had at least one multiheaded word.
+See Table~\ref{tab:linktypes} and
+Appendix~\ref{app:linktypes} for detailed results.
+%\noteJE{2 significant digits is enough}
+%\noteJE{update these numbers}
+
+\subsection{Stability of Results}\label{sec:stability}
+
 We worried that the direction mapping might be unstable and sensitive to the input corpus.
-Figure~\ref{fig:convergence} shows otherwise (for both English and Russian).
+Happily, Figure~\ref{fig:convergence} shows otherwise (for both English and Russian).
 \noteJE{IMPORTANT!
@@
 
-The resulting English corpus uses a syntactic annotation scheme that is somewhat different from the CoNLL annotations.
-Differences are tabulated in Appendix~\ref{app:linktypes}, and the actual parses are contrasted in Figure~\ref{fig:parses} and Appendix~\ref{app:parses}.
+The resulting English corpus uses a syntactic annotation scheme that is somewhat different from the CoNLL annotations.  Differences are tabulated in
+Appendix~\ref{app:linktypes}, while the actual parses are contrasted in Appendix~\ref{app:parses}.  By permission of the program chairs, these 
+appendices are available at the anonymous URL \url{http://tidyurl.com/LinksToDAG_supplement}.  Fragments are shown in 
+Table~\ref{tab:linktypes} and Figure~\ref{fig:parses}.
 
@@
 
-More generally, ILP may be a valuable technology for enriching existing annotated corpora.
-For example, the Penn Treebank \cite{TREEBANK-1993} omitted types of annotations that plausibly could be reconstructed by rule.
+ILP may {\em in general} be a valuable technology for enriching existing annotated corpora.
+For example, the Penn Treebank project \cite{TREEBANK-1993} deliberately omitted types of annotations that plausibly could be added automatically.
+ILP can help by leveraging unsupervised corpus-wide information \cite{ravi2009}, enforcing annotations that are simultaneously well-formed per sentence and consistent across sentences.
 
-% \newpage
+% \subsection{Sample of English Parses}
+% 
+% \begin{figure*}[ht!]
+%   \vspace{-6pt}
+%   \centering
+%   \input{figure/sample_parses.tikz}
+%   \vspace{-6pt}
+% \end{figure*}
+
+%\twocolumn
+
+\newpage
 \bibliographystyle{acl2014}

### 2014-09-21 doc/LinksToDAG.tex
edits from reviewer #1 of EMNLP short submission, including additions to bib

@@
 \begin{abstract}
-Under multi-headed dependency grammar, a parse is a directed acyclic graph rather than a tree.  Such formalisms can be more syntactically and semantically expressive.  However, 
+Under multi-headed dependency grammar, a parse is a connected DAG rather than a tree.  Such formalisms can be more syntactically and semantically expressive.  However, 
 it is hard to train, test, or improve multi-headed parsers because few multi-headed corpora exist, particularly for the projective case.
@@
 
-As recently discussed by \newcite{gomezrodriguez-nivre-2013}, one might impose various requirements on the parse graph:
+As discussed by \newcite{kuhlmann-nivre-2006,gomezrodriguez-nivre-2013}, one might impose various requirements on the parse graph:
 \begin{itemize}[noitemsep]
@@
 
+Graph representations of syntactic and semantic structure have been widely considered of late \cite{demarneffe-manning-2008,ivanova-et-al-2012,AMR-2013,oepen-et-al-2014}.
+
 A few past NLP papers have explored multi-headed dependency parsing \cite{buchkromann-2006,mcdonald-pereira-2006,sagae-tsujii-2008,gomezrodriguez-nivre-2013}.
-Unfortunately, there seem to be no annotated corpora in this form other than the Danish Dependency Treebank \cite{kromann-2003}; researchers have sometimes converted corpora from other formats such as HPSG.
+Unfortunately, there seem to be no annotated corpora in this form other than the Danish Dependency Treebank \cite{kromann-2003}; these papers  sometimes converted corpora from other formats such as HPSG.
 
@@
 It seems at first that no one has worked out annotation conventions for {\em projective} multi-headed dependency parsing.
-However, this is not quite true.
+However, this is only half-true.
 Link Grammar \cite{SleatorTemperly91} is a grammar-based formalism for projective dependency parsing with {\em undirected} links.
@@
 
-However, working directly with the link grammar lexicon format is somewhat tricky.
-We also feared that there would not be a perfect solution---for example, because of errors in the lexicon, or linguistically unnatural word sequences not anticipated by the grammar designers.
-In this case it would not be clear how to relax our constraints.
+% REMOVED AT REVIEWER SUGGESTION
+% However, working directly with the link grammar lexicon format is somewhat tricky. We also feared
+However, we feared that there would not be a feasible solution---because of errors in the lexicon or linguistically unnatural word sequences not anticipated by the grammar designers.
+In this case it would be unclear how to relax our constraints.
 

### 2014-09-21 src/tools/experiments/conll_analysis.py
changed the sample parses to be displayed. 2 short and 1 long for the paper.

@@ -60,8 +60,8 @@ PAPER_TIKZ = open(EXAMPLE_SENTENCES_LOC, 'w+')
 EXAMPLE_PARSES_LOC = TIKZ_DIR+"parses.tikz"
 EXAMPLE_PARSES = open(EXAMPLE_PARSES_LOC, 'w+')
 
-SAMPLE_EXAMPLE_SENTENCES_LOC = TIKZ_DIR+"sample_parses.tikz"
-SAMPLE_EXAMPLE_PARSES = open(SAMPLE_EXAMPLE_SENTENCES_LOC, 'w+')
+LONG_EXAMPLE_SENTENCES_LOC = TIKZ_DIR+"long_parses.tikz"
+LONG_EXAMPLE_PARSES = open(LONG_EXAMPLE_SENTENCES_LOC, 'w+')
 
 ALLOWED_LABELS = SOL_DIR + "allowedLinks.txt"
 ALLOWED_LABELS = open(ALLOWED_LABELS, "r")
@@ -427,13 +427,13 @@ multiheaded_sentence_count = 0
 # --------------------------------------------------------------------
 paper_sentence_count = 0
 paper_sentence_skip = 0
-paper_sentence_limit = 3
+paper_sentence_limit = 2
 
 example_parses_count = 0
 example_parses_limit = 100
 
-sample_example_parses_count = 0
-sample_example_parses_limit = 4
+long_example_parses_count = 0
+long_example_parses_limit = 1
 
 
 for linkSentence in link_results:
@@ -471,23 +471,23 @@ for linkSentence in link_results:
 
 
     # Link parses to put in the paper. Takes sentences of only length 5.
-    if (paper_sentence_count < paper_sentence_limit) and sentenceLength >= 24 and sentenceLength <= 28 and sentenceCheck:
+    if (paper_sentence_count < paper_sentence_limit) and sentenceLength >= 45 and sentenceLength <= 45 and sentenceCheck:
         if paper_sentence_skip > 0:
             paper_sentence_skip -= 1
         else:
-            tikz = tikz_dependency(conll_results[linkSentence], link_results[linkSentence], linkSentence, .97 / 3)            
+            tikz = tikz_dependency(conll_results[linkSentence], link_results[linkSentence], linkSentence, .97 / 2)      
             PAPER_TIKZ.write(tikz)
             paper_sentence_count += 1
-            if paper_sentence_count % 3 == 0:
+            if paper_sentence_count % 2 == 0:
[…]

### 2014-09-22 src/tools/experiments/conll_analysis.py
updated the conll_analysis.py script to output what we want on the paper.

@@ -63,6 +63,11 @@ EXAMPLE_PARSES = open(EXAMPLE_PARSES_LOC, 'w+')
 LONG_EXAMPLE_SENTENCES_LOC = TIKZ_DIR+"long_parses.tikz"
 LONG_EXAMPLE_PARSES = open(LONG_EXAMPLE_SENTENCES_LOC, 'w+')
 
+LONG_EXAMPLE_SENTENCES2_LOC = TIKZ_DIR+"long_parse2.tikz"
+LONG_EXAMPLE_PARSES2 = open(LONG_EXAMPLE_SENTENCES2_LOC, 'w+')
+
+
+
 ALLOWED_LABELS = SOL_DIR + "allowedLinks.txt"
 ALLOWED_LABELS = open(ALLOWED_LABELS, "r")
 ALLOWED_LABELS_BOTH = LATEX_DIR+"allowed_labels_both.tex"
@@ -430,11 +435,16 @@ paper_sentence_skip = 0
 paper_sentence_limit = 2
 
 example_parses_count = 0
+example_parses_skip = 0
 example_parses_limit = 100
 
 long_example_parses_count = 0
-long_example_parses_skip = 6
+long_example_parses_skip = 90
 long_example_parses_limit = 1
+
+long_example_parses2_count = 0
+long_example_parses2_skip = 0
+long_example_parses2_limit = 0
 
 
 for linkSentence in link_results:
@@ -481,11 +491,12 @@ for linkSentence in link_results:
             paper_sentence_count += 1
             if paper_sentence_count % 2 == 0:
                 PAPER_TIKZ.write("\n")
-    elif long_example_parses_count < long_example_parses_limit and sentenceLength >= 90 and sentenceLength <= 90 and sentenceCheck:
+    elif long_example_parses_count < long_example_parses_limit and sentenceLength >= 85 and sentenceLength <= 90 and sentenceCheck:
         if long_example_parses_skip > 0:
             long_example_parses_skip -= 1
         else:
             tikz = tikz_dependency(conll_results[linkSentence], link_results[linkSentence], linkSentence, .97)
+
             #LONG_EXAMPLE_PARSES.write("\\begin{figure*}[ht!]\n")
             LONG_EXAMPLE_PARSES.write(tikz)
             #LONG_EXAMPLE_PARSES.write("\\end{figure*}\n\n")
@@ -493,16 +504,37 @@ for linkSentence in link_results:
[…]

### 2014-09-23 src/tools/experiments/tikz_dependency.py
Changed the parse visualizations to make all multiheaded links to be colored orange.

@@ -50,7 +50,10 @@ def tikz_dependency(conlls, links, sentence, ratio = 0.3, subfigure = True):
             label = graph_conll.labels[(head,child)]
  
             if head in graph_link.table and child in graph_link.table[head]:
-                result += tikz_draw_edge(head, child, graph_link.getEdge(head,child), "below", "thick")
+                if len(graph_link.heads[child]) > 1:
+                    result += tikz_draw_edge(head, child, graph_link.getEdge(head,child), "below", "thick", "orange")
+                else:
+                    result += tikz_draw_edge(head, child, graph_link.getEdge(head,child), "below", "thick")
                 result += tikz_draw_edge(head, child, "\\small "+label, "above", "thick")
                 graph_link.deleteEdge(head,child)
 
@@ -61,7 +64,12 @@ def tikz_dependency(conlls, links, sentence, ratio = 0.3, subfigure = True):
                     
             
             elif child in graph_link.table and head in graph_link.table[child]:
-                result += tikz_draw_edge(child, head, graph_link.getEdge(child,head), "below", "ultra thick")
+
+
+                if len(graph_link.heads[head]) > 1:
+                    result += tikz_draw_edge(child, head, graph_link.getEdge(child,head), "below", "ultra thick", "orange")
+                else:
+                    result += tikz_draw_edge(child, head, graph_link.getEdge(child,head), "below", "ultra thick")
                 result += tikz_draw_edge(head, child, "\\small "+label, "above", "ultra thick")
                 graph_link.deleteEdge(child,head)
             

### 2014-09-23 doc/LinksToDAG.tex
changed the sample parses to be displayed. 2 short and 1 long for the paper.
increased the margin at the top, in order to accomodate the (hidden) author space
added a line adding in a second long parse. But commented it out.
mentioned that conll parses are gold - is that correct?
increased the size of the sample link analysis table to put in the paper.
updated the descriptions in figure1 and in the appendix about the parse visualizations making every multiheaded link orange.

@@
 
-% \author{Juneki Hong and Jason Eisner\\
+%\author{Juneki Hong and Jason Eisner\\
 %   Department of Computer Science \\
@@
 %   {\tt \{juneki,jason\}@cs.jhu.edu} \\
-% }
+%}
 
@@
 
-\setlength\titlebox{3cm}    % Expanding the titlebox
+\setlength\titlebox{6.5cm}    % Expanding the titlebox
 
@@
 
+% 
+% \begin{figure*}[ht!]
+%   \vspace{-6pt}
+%   \centering
+%   \input{figure/sample_parses.tikz}
+%   \vspace{-6pt}
+% \end{figure*}
+
 \section{Experiments and Results}
@@
   \input{figure/sentences.tikz}
-  \input{figure/sample_parses.tikz}
-  \caption{\small The blue upper edges show CoNLL dependency parses; the red lower edges show our oriented version of the link grammar parses.
-Edges are shown as dotted lines if they appear only in one parse; these are highlighted in orange if the child has multiple parents.
+  \input{figure/long_parses.tikz}
+  %\input{figure/long_parse2.tikz}
+  \caption{\small The blue upper edges show CoNLL gold dependency parses; the red lower edges show our oriented version of the link grammar parses.
+Edges are shown as dotted lines if they appear only in one parse.
+Edges are highlighted in orange if the child has multiple parents.
 Edges that appear in both parses are solid lines, drawn thicker if the directions do not match.
@@
 
-Edges are shown as dotted lines if they appear only in one parse and not in the other.  Such edges are highlighted in orange if the child has multiple parents.
+Edges are shown as dotted lines if they appear only in one parse and not in the other.
+
+Edges are highlighted in orange if the child has multiple parents.
 

### 2014-11-13 doc/LinksToDAG_final.tex
changed the style to TLT's specifications

--- 2014-09-23 doc/LinksToDAG.tex
@@
 
-\documentclass[11pt]{article}
-\usepackage{acl2014}
+\documentclass[a4paper,11pt]{article}
+\usepackage[latin1]{inputenc}
+\usepackage[T1]{fontenc}         
+\usepackage{pslatex}
+
+\usepackage{hyperref}
+\pagestyle{empty}
+
+\bibliographystyle{plain}
+
+%\documentclass[11pt]{article}
+%\usepackage{acl2014}
 \usepackage{times}
@@
 
-%\author{Juneki Hong and Jason Eisner\\
-%   Department of Computer Science \\
-%   Johns Hopkins University \\
-%   Baltimore, MD 21218, USA \\ 
-%   {\tt \{juneki,jason\}@cs.jhu.edu} \\
-%}
+\author{Juneki Hong and Jason Eisner\\
+   Department of Computer Science \\
+   Johns Hopkins University \\
+   Baltimore, MD 21218, USA \\ 
+   {\tt \{juneki,jason\}@cs.jhu.edu} \\
+}
 
@@
 
-\setlength\titlebox{6.5cm}    % Expanding the titlebox
+%\setlength\titlebox{6.5cm}    % Expanding the titlebox
 
@@
 \begin{abstract}
+\noindent
 Under multi-headed dependency grammar, a parse is a connected DAG rather than a tree.  Such formalisms can be more syntactically and semantically expressive.  However, 
@@
 
+\thispagestyle{empty}
+
 \section{Multi-Headed Dependency Parsing}
@@
 
-As discussed by \newcite{kuhlmann-nivre-2006,gomezrodriguez-nivre-2013}, one might impose various requirements on the parse graph:
+As discussed by \cite{kuhlmann-nivre-2006,gomezrodriguez-nivre-2013}, one might impose various requirements on the parse graph:
 \begin{itemize}[noitemsep]
@@
 which specifies for every word a constraint on the {\em sequence} of labeled leftward and rightward edges attached to it.
-As remarked by \newcite{eisner-2000-iwptbook}, this is analogous to dependency grammar's use of head automata to constrain a word's sequence of left and right children.
+As remarked by \cite{eisner-2000-iwptbook}, this is analogous to dependency grammar's use of head automata to constrain a word's sequence of left and right children.
 For example, in {\em ``The boy that Jill skipped with fell down,''} the word {\em with} uses a lexical entry that requires it to link to a governing verb to the left, an extracted object farther to the left, and nothing to the right.
@@
 
-Our objective tries to minimize the number of allowed directions (by link type---cf.\@ \newcite{ravi2009}) and the total slack (by link token):
+Our objective tries to minimize the number of allowed directions (by link type---cf.\@ \cite{ravi2009}) and the total slack (by link token):
 \begin{align}\label{eqn:obj}
@@
 
-\newpage
-\bibliographystyle{acl2014}
-\bibliography{LinksToDAG}
+%\newpage
+\begin{bibliography}{LinksToDAG_final}
+
+%\bibliographystyle{acl2014}
+%\bibliography{LinksToDAG}
+\end{bibliography}
 

### 2014-11-16 doc/LinksToDAG_final.tex
fixed the tables and charts of LinksToDAG_final.tex

@@
 \usepackage{times}
-\usepackage{url}
-\usepackage{latexsym}
+%\usepackage{url}
+%\usepackage{latexsym}
 \usepackage{amsmath}
-\usepackage{xspace}
+%\usepackage{xspace}
 \usepackage{enumitem}
@@
 \usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
-\usepackage{soul}
+%\usepackage{soul}
 
@@
 \usepackage{graphicx}
-\usepackage{caption}
+%\usepackage{caption}
 \usepackage{subcaption}
@@
 
+%\hspace*{-1.5in}
 \begin{figure*}[ht!]
+
   \vspace{-6pt}
   \centering
-  \input{figure/sentences.tikz}
+
+  %\input{figure/sentences.tikz}
+
+  \hspace*{-0.5in}
+  \begin{subfigure}[b]{0.49\textwidth}
+	\begin{dependency}
+	  \begin{deptext}
+		{\scriptsize DT} \& {\scriptsize NN} \& {\scriptsize MD} \& {\scriptsize RB} \& {\scriptsize RB} \& {\scriptsize VB} \& {\scriptsize VB} \& {\scriptsize IN} \& {\scriptsize NN} \& {\scriptsize .} \\
+		the \& matter \& may \& never \& even \& be \& tried \& in \& court \& . \\
+		- \& n-u \& v \& e \& e \& v \& v-d \& r \& n-u \& - \\
+	  \end{deptext}
+	  \deproot[edge above, edge style = {blue, dotted}]{3}{\small ROOT}
+	  \depedge[edge below, edge style = {red, ultra thick}]{2}{3}{S}
+	  \depedge[edge above, edge style = {blue, ultra thick}]{3}{2}{\small SBJ}
+	  \depedge[edge above, edge style = {blue, dotted}]{3}{4}{\small ADV}
+	  \depedge[edge above, edge style = {blue, dotted}]{3}{5}{\small ADV}
+	  \depedge[edge below, edge style = {red, thick}]{3}{6}{I}
+	  \depedge[edge above, edge style = {blue, thick}]{3}{6}{\small VC}
+	  \depedge[edge above, edge style = {blue, dotted}, edge unit distance =1.5ex]{3}{10}{\small P}
+	  \depedge[edge below, edge style = {red, thick}]{2}{1}{D}
+	  \depedge[edge above, edge style = {blue, thick}]{2}{1}{\small NMOD}
+	  \depedge[edge below, edge style = {red, thick}]{7}{8}{MV}
+	  \depedge[edge above, edge style = {blue, thick}]{7}{8}{\small ADV}
+	  \depedge[edge below, edge style = {orange, thick}]{6}{7}{P}
+	  \depedge[edge above, edge style = {blue, thick}]{6}{7}{\small VC}
+	  \depedge[edge below, edge style = {red, thick}]{8}{9}{J}
+	  \depedge[edge above, edge style = {blue, thick}]{8}{9}{\small PMOD}
+	  \deproot[edge below, edge style = {red, dotted}]{2}{W}
+	  \deproot[edge below, edge style = {orange, ultra thick, dotted}]{7}{WV}
+	  \deproot[edge below, edge style = {red, dotted}]{10}{X}
+	  \depedge[edge below, edge style = {red, dotted}]{6}{4}{E}
+	  \depedge[edge below, edge style = {red, dotted}]{6}{5}{E}
+	\end{dependency}
+  \end{subfigure}
+  \hspace*{0.5in}\begin{subfigure}[b]{0.49\textwidth}
+	\begin{dependency}
+	  \begin{deptext}
+		{\scriptsize DT} \& {\scriptsize NN} \& {\scriptsize NN} \& {\scriptsize VB} \& {\scriptsize RB} \& {\scriptsize VB} \& {\scriptsize DT} \& {\scriptsize NN} \& {\scriptsize .} \\
+		the \& Paris \& bank \& does \& n't \& publish \& that \& figure \& . \\
+		- \& b \& n \& v \& - \& v \& j-d \& n \& - \\
+	  \end{deptext}
+	  \deproot[edge above, edge style = {blue, dotted}]{4}{\small ROOT}
+	  \depedge[edge below, edge style = {red, thick}]{3}{1}{D}
+	  \depedge[edge above, edge style = {blue, thick}]{3}{1}{\small NMOD}
+	  \depedge[edge below, edge style = {red, thick}]{3}{2}{AN}
+	  \depedge[edge above, edge style = {blue, thick}]{3}{2}{\small NMOD}
+	  \depedge[edge below, edge style = {red, thick}]{6}{8}{O}
+	  \depedge[edge above, edge style = {blue, thick}]{6}{8}{\small OBJ}
+	  \depedge[edge below, edge style = {red, ultra thick}]{3}{4}{S}
+	  \depedge[edge above, edge style = {blue, ultra thick}]{4}{3}{\small SBJ}
+	  \depedge[edge below, edge style = {red, thick}]{4}{5}{N}
+	  \depedge[edge above, edge style = {blue, thick}]{4}{5}{\small VMOD}
+	  \depedge[edge below, edge style = {orange, thick}]{4}{6}{I}
+	  \depedge[edge above, edge style = {blue, thick}]{4}{6}{\small VC}
+	  \depedge[edge above, edge style = {blue, dotted}, edge unit distance = 1.7ex]{4}{9}{\small P}
+	  \depedge[edge below, edge style = {red, thick}]{8}{7}{D}
+	  \depedge[edge above, edge style = {blue, thick}]{8}{7}{\small NMOD}
+	  \deproot[edge below, edge style = {red, dotted}]{3}{W}
+	  \deproot[edge below, edge style = {orange, ultra thick, dotted}]{6}{WV}
+	  \deproot[edge below, edge style = {red, dotted}]{9}{X}
+	\end{dependency}
+  \end{subfigure}
+
+  \hspace*{-1in}
   \input{figure/long_parses.tikz}
+
   %\input{figure/long_parse2.tikz}
@@
   \vspace{-6pt}
+
 \end{figure*}
@@
 % \renewcommand{\arraystretch}{1.2}    % more space between rows
+\hspace*{-1in}
 \input{figure/sample_link_analysis_coarse_table.tex}
@@
 
-\renewcommand{\arraystretch}{1.2}    % more space between rows
+%\begin{table*}[ht]
+%\renewcommand{\arraystretch}{1.2}    % more space between rows
+%\hspace*{-1in}
 \input{figure/link_analysis_coarse_table.tex}
+%\end{table*}
 

### 2014-11-17 doc/LinksToDAG_final.tex
added in several of reviewer 2's comments. Tried to follow Jason's suggestion of changing the tone of the paper.

@@
 \bibliographystyle{plain}
+%\bibliographystyle{acl2014}
 
@@
 \usepackage{amsmath}
+
 %\usepackage{xspace}
@@
 
-% \newcommand{\root}{\texttt{\$}}
-
-\title{Deriving Multi-Headed Projective Dependency Parses \\ from Link Grammar Parses}
+\title{Deriving Multi-Headed Projective Dependency Parses from Link Grammar Parses}
 % Oriented link grammar: Creating a multi-headed dependency corpus.
 
-\author{Juneki Hong and Jason Eisner\\
-   Department of Computer Science \\
-   Johns Hopkins University \\
-   Baltimore, MD 21218, USA \\ 
-   {\tt \{juneki,jason\}@cs.jhu.edu} \\
+\author{Juneki Hong and Jason Eisner
+  \\[0.5cm] Department of Computer Science
+  \\Johns Hopkins University
+  \\Baltimore, MD 21218, USA 
+  \\{\tt \{juneki,jason\}@cs.jhu.edu}
 }
@@
 % investigate the benefit of such parsers or to work on making them faster or more accurate.  
-To help fill this gap, we observe that link grammar already produces {\em undirected} projective graphs.  
+To help fill this gap, we observe that link grammar already produces {\em undirected} projective graphs, and so we wanted to determine whether these could be converted into directionalized dependency parses.
 % OLD
@@
 We find that such directions can indeed be consistently assigned in a way that yields valid multi-headed dependency parses.
-The resulting parses in English appear reasonably linguistically plausible, though differing in style from CoNLL-style parses of the same sentences; we discuss the differences.
+%The resulting parses in English appear reasonably linguistically plausible, though differing in style from CoNLL-style parses of the same sentences; we discuss the differences.  
+However, the resulting parses in English differ in style from CoNLL-style parses of the same sentences; we discuss the differences.
+
 \end{abstract}
@@
 \thispagestyle{empty}
+
+%\newpage
 
@@
 
-Thus, we chose to use a sample of {\em actual} sentences parsed by the link grammar, and to seek a direction mapping so that {\em these} parses would not violate our constraints after directions were imposed.
+% OLD
+% Thus, we chose to use a sample of {\em actual} sentences parsed by the link grammar, and to seek a direction mapping so that {\em these} parses would not violate our constraints after directions were imposed.
+Thus, we chose to use a sample of sentences sucessfully parsed by the link grammar, and to seek a direction mapping that would not violate our constraints after directions were imposed.
 If no such mapping exists, then we are willing to orient a few edge tokens in the wrong direction to ensure that the parses are still well-formed---but we minimize the number of such violations.
 In this way, the empirical distribution of sentences guides our assignment of directions.
-The resulting directed corpus can be used for research on multi-headed dependency parsing.
+The resulting directed corpus could be used for research on multi-headed dependency parsing.
 
@@
 We generated link parses using the AbiWord/CMU link grammar parser version 5.0.8 \cite{LINKPARSER-2014}.
-The parser's coverage is not perfect: we obtained connected parses for only \input{figure/conll_analysis_sentences_total.tex}(of \input{figure/conll_analysis_sentences_original.tex}$\!\!$) English sentences and only \input{figure/russian_sentence_count.tex}(of \input{figure/russian_original_sentence_count.tex}$\!\!$) Russian sentences, discarding the other sentences.
+The parser's coverage is not perfect: we obtained connected parses for only \input{figure/conll_analysis_sentences_total.tex}(of \input{figure/conll_analysis_sentences_original.tex}$\!\!$) English sentences and only \input{figure/russian_sentence_count.tex}(of \input{figure/russian_original_sentence_count.tex}$\!\!$) Russian sentences, discarding the other sentences\footnote{When the link parser fails, it outputs a disconnected graph representing its best effort parse within a timelimit.
+We removed these sentences because our ILP encoding tries to satisfy the edges of an entire connected parse.
+Given a solution, it would be possible to return to these disconnected parses, directionalize the edges, and include them along with the rest of the corpus.
+It would also be possible to include these sentences for the ILP to solve by modifying the ILP to satisfy the edges of all disconnected components of a parse.}
 These two languages have the most mature lexicons at present, although lexicons for 7 other languages are available.
 
-The link grammar parses do differ in style from the single-headed CoNLL parses of the same English sentences.
-They have \input{figure/conll_analysis_edge_percent.tex}more edges overall.
+The link grammar parses differ in style from the single-headed CoNLL parses of the same English sentences, and the multiheaded arcs are used sparingly with only \input{figure/conll_analysis_edge_percent.tex}more edges overall.
 Only \input{figure/conll_analysis_precision.tex}of the links match CoNLL arcs, and only \input{figure/conll_analysis_recall.tex}of the CoNLL arcs match links.
@@
 
-\begin{table*}[ht]
-% \renewcommand{\arraystretch}{1.2}    % more space between rows
-\hspace*{-1in}
-\input{figure/sample_link_analysis_coarse_table.tex}
-\caption{Our solution, i.e., our reconstruction of the ``intended'' direction for each link type in the English Link Grammar.
-We also indicate the extent to which each of these link types (1) has a single dominant direction, (2) participates in multi-headed constructions, and (3) corresponds to CoNLL links of a predictable direction and type.
-For space reasons, we show only the start of this table---the full table can be found in Appendix~\ref{app:linktypes} at \url{http://tidyurl.com/LinksToDAG_supplement}.}\label{tab:linktypes}
-\end{table*}
-
 \begin{figure}[t!]
@@
 \end{figure}
+
+\begin{table*}[ht]
+% \renewcommand{\arraystretch}{1.2}    % more space between rows
+\hspace*{-1in}
+\input{figure/sample_link_analysis_coarse_table.tex}
+\caption{Our solution, i.e., our reconstruction of the ``intended'' direction for each link type in the English Link Grammar.
+We also indicate the extent to which each of these link types (1) has a single dominant direction, (2) participates in multi-headed constructions, and (3) corresponds to CoNLL links of a predictable direction and type.
+For space reasons, we show only the start of this table---the full table can be found in Appendix~\ref{app:linktypes} at \url{http://tidyurl.com/LinksToDAG_supplement}.}\label{tab:linktypes}
+\end{table*}
 
@@
 %\newpage
+
 \begin{bibliography}{LinksToDAG_final}
@@
 %\bibliographystyle{acl2014}
+
 %\bibliography{LinksToDAG}

### 2014-11-20 doc/LinksToDAG_final.tex
Shrunk the figure 2 graph. Changed the authorship section at the top

@@
 
-\author{Juneki Hong and Jason Eisner
-  \\[0.5cm] Department of Computer Science
-  \\Johns Hopkins University
-  \\Baltimore, MD 21218, USA 
-  \\{\tt \{juneki,jason\}@cs.jhu.edu}
+\author{
+  Juneki Hong \\
+  [0.5cm] Language Technologies Institute\\
+  Carnegie Mellon University\thanks{Work done while attending Johns Hopkins University}\\
+  Pittsburgh, PA 15213, USA\\
+  {\texttt{juneki@cs.cmu.edu}}
+  \and
+  Jason Eisner\\
+  [0.5cm] Department of Computer Science\\
+  Johns Hopkins University\\
+  Baltimore, MD 21218, USA \\
+  {\texttt{jason@cs.jhu.edu}}
 }
@@
 Vertical edges have parent 0.
-For 100 example parses, see Appendix~\ref{app:parses} at \url{http://tidyurl.com/LinksToDAG_supplement}.}
+For 100 example parses, see Appendix~\ref{app:parses} of the supplementary material.}
   \label{fig:parses}
@@
   \centering
-  \includegraphics[width=\linewidth, keepaspectratio=true]{figure/precision_recall.png}
+  \includegraphics[width=0.7\textwidth, keepaspectratio=true]{figure/precision_recall.png}
   \caption{Convergence to the direction mapping obtained on the largest run.
@@
 We also indicate the extent to which each of these link types (1) has a single dominant direction, (2) participates in multi-headed constructions, and (3) corresponds to CoNLL links of a predictable direction and type.
-For space reasons, we show only the start of this table---the full table can be found in Appendix~\ref{app:linktypes} at \url{http://tidyurl.com/LinksToDAG_supplement}.}\label{tab:linktypes}
+For space reasons, we show only the start of this table---the full table can be found in Appendix~\ref{app:linktypes} of the supplementary material.}\label{tab:linktypes}
 \end{table*}
@@
 The resulting English corpus uses a syntactic annotation scheme that is somewhat different from the CoNLL annotations.  Differences are tabulated in
-Appendix~\ref{app:linktypes}, while the actual parses are contrasted in Appendix~\ref{app:parses}.  By permission of the program chairs, these 
-appendices are available at the anonymous URL \url{http://tidyurl.com/LinksToDAG_supplement}.  Fragments are shown in 
-Table~\ref{tab:linktypes} and Figure~\ref{fig:parses}.
+Appendix~\ref{app:linktypes}, while the actual parses are contrasted in Appendix~\ref{app:parses}.  
+By permission of the program chairs, these appendices are available at the anonymous URL \url{http://tidyurl.com/LinksToDAG_supplement}.  
+Fragments are shown in Table~\ref{tab:linktypes} and Figure~\ref{fig:parses}.
 

### 2014-11-20 doc/LinksToDAG_final.tex
removed the last reference to the supplementary material url

@@
 The resulting English corpus uses a syntactic annotation scheme that is somewhat different from the CoNLL annotations.  Differences are tabulated in
-Appendix~\ref{app:linktypes}, while the actual parses are contrasted in Appendix~\ref{app:parses}.  
-By permission of the program chairs, these appendices are available at the anonymous URL \url{http://tidyurl.com/LinksToDAG_supplement}.  
-Fragments are shown in Table~\ref{tab:linktypes} and Figure~\ref{fig:parses}.
+Appendix~\ref{app:linktypes}, while the actual parses are contrasted in Appendix~\ref{app:parses}, and fragments of these are shown in Table~\ref{tab:linktypes} and Figure~\ref{fig:parses}.
+%By permission of the program chairs, these appendices are available at the anonymous URL \url{http://tidyurl.com/LinksToDAG_supplement}.
 

### 2014-11-20 doc/LinksToDAG_final.tex
Expanded the caption of figure 2

@@
   \caption{Convergence to the direction mapping obtained on the largest run.
-Precision relative to that run is very high even for small datasets.
-Recall grows as more link types are seen.}\label{fig:convergence}
+Precision of the directional assignments in smaller runs relative to the largest run is high even for small datasets.
+Recall grows as more link types are seen and directionalized.}\label{fig:convergence}
   \vspace{-6pt}

### 2014-11-20 doc/LinksToDAG_final.tex
clarified or reverted some changes from Juneki; added ack to NSF; restored todonotes; introduced macro to map \newcite -> \cite

--- 2014-11-20 doc/LinksToDAG_final.tex
@@
 
+%\bibliographystyle{acl2014}
 \bibliographystyle{plain}
-%\bibliographystyle{acl2014}
+\newcommand{\newcite}[1]{\cite{#1}}   % ACL's newcite is not available in plain bib style, so just use cite
 
@@
 
-\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}  % allows better color names
-\usepackage[disable]{todonotes}   % insert [disable] to disable all notes
+\usepackage{todonotes}   % insert [disable] to disable all notes
 % Note that these macros accept optional arguments such as 
@@
 \newcommand{\Note}[4][]{\todo[author=#2,color=#3,fancyline,#1]{#4}}
-\newcommand{\noteJH}[2][]{\Note[#1]{JH}{blue!40}{#2}}
-\newcommand{\noteJE}[2][]{\Note[#1]{JE}{green!40}{#2}}   
-\newcommand{\notewho}[3][]{\Note[#1]{#2}{orange!40}{#3}}  % extra arg with miscellaneous author
+\newcommand{\noteJH}[2][]{\Note[size=\small,#1]{JH}{blue!40}{#2}}
+\newcommand{\noteJE}[2][]{\Note[size=\small,#1]{JE}{green!40}{#2}}   
+\newcommand{\notewho}[3][]{\Note[size=\small,#1]{#2}{orange!40}{#3}}  % extra arg with miscellaneous author
 \newcommand{\NoteJH}[2][]{\noteJH[inline,#1]{#2}}
@@
 
-\title{Deriving Multi-Headed Projective Dependency Parses from Link Grammar Parses}
+\title{Deriving Multi-Headed Projective Dependency Parses from Link
+  Grammar Parses\thanks{This material is based upon work supported by the National
+    Science Foundation under Grant No. 1423276.  The work was mainly
+    conducted while the first author was at Johns Hopkins
+    University.\noteJE{also thank Russian link grammar people for
+      giving us data?  anyone else?}}}
 % Oriented link grammar: Creating a multi-headed dependency corpus.
@@
   [0.5cm] Language Technologies Institute\\
-  Carnegie Mellon University\thanks{Work done while attending Johns Hopkins University}\\
+  Carnegie Mellon University\\
   Pittsburgh, PA 15213, USA\\
@@
 We find that such directions can indeed be consistently assigned in a way that yields valid multi-headed dependency parses.
-%The resulting parses in English appear reasonably linguistically plausible, though differing in style from CoNLL-style parses of the same sentences; we discuss the differences.  
-However, the resulting parses in English differ in style from CoNLL-style parses of the same sentences; we discuss the differences.
+The resulting parses in English appear reasonably linguistically plausible, though differing in style from CoNLL-style parses of the same sentences; we discuss the differences.
 
@@
 
-As discussed by \cite{kuhlmann-nivre-2006,gomezrodriguez-nivre-2013}, one might impose various requirements on the parse graph:
+As discussed by \newcite{kuhlmann-nivre-2006,gomezrodriguez-nivre-2013}, one might impose various requirements on the parse graph:
 \begin{itemize}[noitemsep]
@@
 which specifies for every word a constraint on the {\em sequence} of labeled leftward and rightward edges attached to it.
-As remarked by \cite{eisner-2000-iwptbook}, this is analogous to dependency grammar's use of head automata to constrain a word's sequence of left and right children.
+As remarked by \newcite{eisner-2000-iwptbook}, this is analogous to dependency grammar's use of head automata to constrain a word's sequence of left and right children.
 For example, in {\em ``The boy that Jill skipped with fell down,''} the word {\em with} uses a lexical entry that requires it to link to a governing verb to the left, an extracted object farther to the left, and nothing to the right.
@@
 
-% OLD
-% Thus, we chose to use a sample of {\em actual} sentences parsed by the link grammar, and to seek a direction mapping so that {\em these} parses would not violate our constraints after directions were imposed.
-Thus, we chose to use a sample of sentences sucessfully parsed by the link grammar, and to seek a direction mapping that would not violate our constraints after directions were imposed.
+Thus, we chose to use a sample of {\em actual} sentences parsed by the link grammar, and to seek a direction mapping so that {\em these} parses would not violate our constraints after directions were imposed.
+\noteJE{clarify per reviewer comment}
 If no such mapping exists, then we are willing to orient a few edge tokens in the wrong direction to ensure that the parses are still well-formed---but we minimize the number of such violations.
@@
 
-% Link parsing in contrast produces a multiheaded planar graph with undirected edges, where every edge has a label describing the relationship between two constituents in a parse.
-In this paper we explore whether these relationships include dependencies, and whether the multiheadedness of the link grammar offers additional dependency relationships not found in other corpora.
+% Link parsing in contrast produces a multi-headed planar graph with undirected edges, where every edge has a label describing the relationship between two constituents in a parse.
+In this paper we explore whether these relationships include dependencies, and whether the multi-headedness of the link grammar offers additional dependency relationships not found in other corpora.
 
 % Dependency parsing is the task of mapping a sentence to a projective (not always projective?) directed acyclic tree.
-Link parsing in contrast produces a multiheaded planar graph with undirected edges, where every edge has a label describing the relationship between two constituents in a parse.
+Link parsing in contrast produces a multi-headed planar graph with undirected edges, where every edge has a label describing the relationship between two constituents in a parse.
 In this paper we explore whether these relationships include dependencies.
@@
 We generated link parses using the AbiWord/CMU link grammar parser version 5.0.8 \cite{LINKPARSER-2014}.
-The parser's coverage is not perfect: we obtained connected parses for only \input{figure/conll_analysis_sentences_total.tex}(of \input{figure/conll_analysis_sentences_original.tex}$\!\!$) English sentences and only \input{figure/russian_sentence_count.tex}(of \input{figure/russian_original_sentence_count.tex}$\!\!$) Russian sentences, discarding the other sentences\footnote{When the link parser fails, it outputs a disconnected graph representing its best effort parse within a timelimit.
+The parser's coverage is not perfect: we obtained connected parses for only \input{figure/conll_analysis_sentences_total.tex}(of \input{figure/conll_analysis_sentences_original.tex}$\!\!$) English sentences and only \input{figure/russian_sentence_count.tex}(of \input{figure/russian_original_sentence_count.tex}$\!\!$) Russian sentences, discarding the other sentences\footnote{When the link parser fails, it outputs a disconnected graph representing its best effort parse within a time limit.
 We removed these sentences because our ILP encoding tries to satisfy the edges of an entire connected parse.
-Given a solution, it would be possible to return to these disconnected parses, directionalize the edges, and include them along with the rest of the corpus.
+\NoteJE{no, we removed them because we have no confidence that the link grammar is doing anything reasonable in this case} Given a solution, it would be possible to return to these disconnected parses, directionalize the edges, and include them along with the rest of the corpus.
 It would also be possible to include these sentences for the ILP to solve by modifying the ILP to satisfy the edges of all disconnected components of a parse.}
@@
 
-The link grammar parses differ in style from the single-headed CoNLL parses of the same English sentences, and the multiheaded arcs are used sparingly with only \input{figure/conll_analysis_edge_percent.tex}more edges overall.
+The link grammar parses have \input{figure/conll_analysis_edge_percent.tex}more edges overall, indicating that their directed versions will have a few multi-headed constructions per sentence.  They
+do differ in style from the single-headed CoNLL parses of the same English sentences.
 Only \input{figure/conll_analysis_precision.tex}of the links match CoNLL arcs, and only \input{figure/conll_analysis_recall.tex}of the CoNLL arcs match links.
@@
 The second summand ensures that \eqnref{nocycles} is trivially satisfied (hence has no effect) when $u$ is {\em not} the parent of $v$.  
-\noteJE{in final version: figure out whether it is faster to require the 0 tokens to have depth 0, or add all the depths (perhaps multiplied by 0.001) to the objective function to break ties in simplex, or neither.
+\NoteJE{in final version: figure out whether it is faster to require the 0 tokens to have depth 0, or add all the depths (perhaps multiplied by 0.001) to the objective function to break ties in simplex, or neither.
 See also other notes in tex file.}
@@
 
-Our objective tries to minimize the number of allowed directions (by link type---cf.\@ \cite{ravi2009}) and the total slack (by link token):
+Our objective tries to minimize the number of allowed directions (by link type---cf.\@ \newcite{ravi2009}) and the total slack (by link token):
 \begin{align}\label{eqn:obj}
@@
   \includegraphics[width=0.7\textwidth, keepaspectratio=true]{figure/precision_recall.png}
-  \caption{Convergence to the direction mapping obtained on the largest run.
-Precision relative to that run is very high even for small datasets.
-Recall grows as more link types are seen.}\label{fig:convergence}
+  \caption{Rapid convergence to the direction mapping obtained on the largest dataset.
+The direction mappings obtained on small datasets have high {\em precision} relative to the one obtained on the largest dataset.
+{\em Recall} grows as more link types are seen and directionalized.}\label{fig:convergence}
   \vspace{-6pt}
@@
 \noteJH{A run of 10,000 sentences takes 3142 seconds (update in final version!)}  On English, only \input{figure/allowed_labels_both.tex}of \input{figure/allowed_labels_total.tex}link {\em types} allowed both directions, and only $\sum_{ij} s_{ij} = \input{figure/label_token_disallowed}$ of \input{figure/label_token_total.tex}link {\em tokens} required a disallowed direction via slack.
-\input{figure/conll_analysis_sentences_multiheaded.tex}of the English sentences but (alas) only \input{figure/russian_multiheaded.tex}of the Russian ones had at least one multiheaded word.
+\input{figure/conll_analysis_sentences_multiheaded.tex}of the English sentences but (alas) only \input{figure/russian_multiheaded.tex}of the Russian ones had at least one multi-headed word.
 See Table~\ref{tab:linktypes} and
@@
 The resulting English corpus uses a syntactic annotation scheme that is somewhat different from the CoNLL annotations.  Differences are tabulated in
-Appendix~\ref{app:linktypes}, while the actual parses are contrasted in Appendix~\ref{app:parses}, and fragments of these are shown in Table~\ref{tab:linktypes} and Figure~\ref{fig:parses}.
+Appendix~\ref{app:linktypes}.  The actual parses are contrasted in Appendix~\ref{app:parses}, and fragments of these are shown in Table~\ref{tab:linktypes} and Figure~\ref{fig:parses}.
 %By permission of the program chairs, these appendices are available at the anonymous URL \url{http://tidyurl.com/LinksToDAG_supplement}.

### 2014-11-21 doc/LinksToDAG_final.tex
added new intro

@@
 
-\title{Deriving Multi-Headed Projective Dependency Parses from Link
+\title{Deriving Multi-Headed Planar Dependency Parses from Link
   Grammar Parses\thanks{This material is based upon work supported by the National
@@
 Under multi-headed dependency grammar, a parse is a connected DAG rather than a tree.  Such formalisms can be more syntactically and semantically expressive.  However, 
-it is hard to train, test, or improve multi-headed parsers because few multi-headed corpora exist, particularly for the projective case.
+it is hard to train, test, or improve multi-headed parsers because few multi-headed corpora exist, particularly for the projective or planar case.
 % OLD
 % investigate the benefit of such parsers or to work on making them faster or more accurate.  
-To help fill this gap, we observe that link grammar already produces {\em undirected} projective graphs, and so we wanted to determine whether these could be converted into directionalized dependency parses.
+To help fill this gap, we observe that link grammar already produces {\em undirected} planar graphs, and so we wanted to determine whether these could be converted into directionalized dependency parses.
 % OLD
@@
 %\newpage
+
+\section{Motivation}
+
+Link Grammar \cite{SleatorTemperly91} is a syntactic formalism in which a parse of a sentence is an undirected, edge-labeled, planar graph.
+The labeled edges of the graph represent syntactic relationships among the words.
+The vertices of the graph are simply the words $1, 2, \ldots, n$ of the sentence, along with a distinguished ``root'' vertex 0.
+
+The small Link Grammar community has invested effort in creating link grammars for several languages.
+In this short paper, we consider whether their {\em undirected} parses can be converted automatically to {\em directed} ones.
+We have four motivations:
+
+\begin{enumerate}
+
+\item We were curious about the relation between link grammar annotation and dependency grammar annotation.
+We suspected that the link grammar parses could be actually interpreted as multi-headed dependency grammar parses.
+Although the link grammar authors did not bother to specify directions for the different edge types, we suspected that they essentially had such directions in mind.
+
+\item Our problem provides a good case study for how to automatically enrich a corpus.
+Hand-constructed grammars or corpora sometimes provide a lighter level of annotation than desired.
+In our setting, the edges lack directions; in other settings, the syntactic categories may be coarser than desired, or some relations may be omitted.
+One may want to automatically enrich the annotation in such cases, whether by doing some kind of learning \cite[et seq.]{matsuzaki-et-al-2005}, or by exploiting implicit constraints.
+In this paper, we exploit implicit constraints of consistency and acyclicity.
+
+\item The resulting parses may be useful data for evaluating new parsing {\em algorithms}.
+There has been a good deal of recent research on projective dependency parsing, variously using global optimization or sequential classification.
+\noteJE{add citations} Some of these algorithms could be extended to the {\em multi-headed} case, which is of linguistic interest for reasons discussed below.
+However, it is not clear how to evaluate such algorithms.
+For training and testing, one needs a plausible sample of multi-headed planar dependency parses of real sentences.
+Our method cheaply manufactures such a sample, to compensate for the lack of gold-standard data of this form.\footnote{A reasonable objection to this line of algorithmic work is that the projectivity or planarity restriction is linguistically questionable even for single-headed parsing, and even more so for multi-headed parsing.
+So why not go straight to {\em non}-projective multi-headed parsing?
+Mainly, because efficient algorithms often exist in the projective case, and these projective algorithms can be useful as an intermediate step in producing non-planar parses.
+Since natural-language parses tend to be ``almost projective'' in practice (\cite{pitler}), one can use a fast projective parser as the first step in a coarse-to-fine or stacking architecture, in order to obtain preliminary edge scores that are supplied to a non-projective parser.
+Another approach is
+``pseudo-projectivization,'' where we reversibly transform non-projective parses into a projective annotation style so that projective parsers can be used.
+\noteJE{cite Pitler}}
+
+\item Our automatic method for reconstructing the latent directions had an unexpected benefit.
+It uncovered an inconsistency in the English link grammar, regarding the handling of embedded sentences with missing (PRO) subjects.
+
+\end{enumerate}
 

### 2014-11-21 doc/LinksToDAG_final.tex
added introduction, edited bib, updated todonotes.sty to later version

@@
 For training and testing, one needs a plausible sample of multi-headed planar dependency parses of real sentences.
-Our method cheaply manufactures such a sample, to compensate for the lack of gold-standard data of this form.\footnote{A reasonable objection to this line of algorithmic work is that the projectivity or planarity restriction is linguistically questionable even for single-headed parsing, and even more so for multi-headed parsing.
+Our method cheaply manufactures such a sample, to compensate for the current lack of gold-standard data of this form.\footnote{A reasonable objection to this line of algorithmic work is that the projectivity or planarity restriction is linguistically questionable even for single-headed parsing, and even more so for multi-headed parsing.
 So why not go straight to {\em non}-projective multi-headed parsing?
@@
 The direction mappings obtained on small datasets have high {\em precision} relative to the one obtained on the largest dataset.
-{\em Recall} grows as more link types are seen and directionalized.}\label{fig:convergence}
+Their {\em recall} grows as more link types are seen and directionalized.}\label{fig:convergence}
   \vspace{-6pt}

### 2014-11-21 doc/LinksToDAG_final.tex
many improvements to paper and bibliography ...

@@
     conducted while the first author was at Johns Hopkins
-    University.\noteJE{also thank Russian link grammar people for
-      giving us data?  anyone else?}}}
+    University.\noteJE{anyone else to thank?}}}
 % Oriented link grammar: Creating a multi-headed dependency corpus.
@@
 
-\section{Motivation}
+\section{Motivation}\label{sec:motivation}
 
-Link Grammar \cite{SleatorTemperly91} is a syntactic formalism in which a parse of a sentence is an undirected, edge-labeled, planar graph.
+Link Grammar \cite{sleator-temperley-1991} is a syntactic formalism in which a parse of a sentence is an undirected, edge-labeled, planar graph.
 The labeled edges of the graph represent syntactic relationships among the words.
@@
 In this short paper, we consider whether their {\em undirected} parses can be converted automatically to {\em directed} ones.
-We have four motivations:
+We have three motivations:
 
@@
 \item We were curious about the relation between link grammar annotation and dependency grammar annotation.
-We suspected that the link grammar parses could be actually interpreted as multi-headed dependency grammar parses.
+We suspected that the link grammar parses could be interpreted as multi-headed dependency grammar parses.
 Although the link grammar authors did not bother to specify directions for the different edge types, we suspected that they essentially had such directions in mind.
@@
 One may want to automatically enrich the annotation in such cases, whether by doing some kind of learning \cite[et seq.]{matsuzaki-et-al-2005}, or by exploiting implicit constraints.
-In this paper, we exploit implicit constraints of consistency and acyclicity.
+In this paper, we use Integer Linear Programming to exploit implicit constraints of consistency and acyclicity.
 
-\item The resulting parses may be useful data for evaluating new parsing {\em algorithms}.
-There has been a good deal of recent research on projective dependency parsing, variously using global optimization or sequential classification.
-\noteJE{add citations} Some of these algorithms could be extended to the {\em multi-headed} case, which is of linguistic interest for reasons discussed below.
-However, it is not clear how to evaluate such algorithms.
-For training and testing, one needs a plausible sample of multi-headed planar dependency parses of real sentences.
-Our method cheaply manufactures such a sample, to compensate for the current lack of gold-standard data of this form.\footnote{A reasonable objection to this line of algorithmic work is that the projectivity or planarity restriction is linguistically questionable even for single-headed parsing, and even more so for multi-headed parsing.
-So why not go straight to {\em non}-projective multi-headed parsing?
-Mainly, because efficient algorithms often exist in the projective case, and these projective algorithms can be useful as an intermediate step in producing non-planar parses.
-Since natural-language parses tend to be ``almost projective'' in practice (\cite{pitler}), one can use a fast projective parser as the first step in a coarse-to-fine or stacking architecture, in order to obtain preliminary edge scores that are supplied to a non-projective parser.
-Another approach is
-``pseudo-projectivization,'' where we reversibly transform non-projective parses into a projective annotation style so that projective parsers can be used.
-\noteJE{cite Pitler}}
-
-\item Our automatic method for reconstructing the latent directions had an unexpected benefit.
-It uncovered an inconsistency in the English link grammar, regarding the handling of embedded sentences with missing (PRO) subjects.
+\item The resulting parses may be useful data for experimenting with new parsing {\em algorithms}.
+There has been a good deal of recent research on projective dependency parsing, variously using global optimization or sequential classification (see \cite{martins-et-al-2008,bohnet-2011,chen-li-zhang-2014} for surveys).
+Some of these algorithms could be extended to the {\em multi-headed} case, which is of linguistic and computational interest for reasons discussed below.
+However, for training and testing such algorithms, one would need a plausible sample of multi-headed projective dependency parses of real sentences.
+Our method cheaply manufactures such a sample, to compensate for the current lack of gold-standard data of this form.
 
 \end{enumerate}
+
+\noindent Our automatic method for reconstructing the latent directions also had an unexpected benefit.
+It revealed an inconsistency in the hand-written English link grammar, regarding the handling of embedded sentences with missing (PRO) subjects.
 
@@
 \begin{itemize}[noitemsep]
-\item {\sc Single-Head}: each word has $\leq 1$ parent
-\item {\sc Acyclic}: there are no directed cycles
-\item {\sc Connected}: each pair of words has a undirected path between them
-\item {\sc Reachable}: each word can be reached from 0 by a directed path ($\Rightarrow$ {\sc Connected})
-\item {\sc Planar}: edges may not ``cross''\footnote{That is, if there are edges between $i,j$ and between $u,v$, where $i < u < j$, then {\sc Planar} requires $i \leq v \leq j$.}
+\item {\sc Single-Head}: Each word has $\leq 1$ parent.
+\item {\sc Acyclic}: There are no directed cycles.
+\item {\sc Connected}: Each pair of words has a undirected path between them.
+\item {\sc Reachable}: Each word can be reached from 0 by a directed path (which implies {\sc Connected}).  Note that 0 may have multiple children.
+\item {\sc Planar}: edges may not ``cross.''  That is, if there are edges between $i,j$ and between $u,v$, where $i < u < j$, then {\sc Planar} requires $i \leq v \leq j$.
 \end{itemize}
-It is common to impose all of these requirements at once, leading to a {\em projective dependency parser} that produces projective trees rooted at 0.\footnote{{\sc Projective} basically means {\sc Planar} + {\sc Reachable}.
-Note: It does not prevent 0 from having multiple children.}  However, parsing algorithms can be devised that relax any or all of the requirements \cite{gomezrodriguez-nivre-2013}.
+It is common to impose all of these requirements at once, leading to a {\em projective dependency parser} that produces projective trees rooted at 0.
+However, parsing algorithms can be devised that relax any or all of the requirements \cite{gomezrodriguez-nivre-2013}.
 
 In this paper, we are interested in relaxing the {\sc Single-Head} requirement while preserving all the others.
-This means that the parse can have more than $n$ edges, allowing it to express more relationships between words.
-In English, for example, here are some constructions that seem to call for a multi-headed analysis.
+This is the setting of {\em multi-headed projective dependency parsing}.
+Just as in the single-headed case, the other requirements ensure that all edges are projective.
+(A projective edge is one where the parent is an ancestor of all words between the parent and the child \cite{nivre-slides}.)
+
+Relaxing {\sc Single-Head} means that the parse can have more than $n$ edges, allowing it to express more relationships between words.  
+In English, for example, here are some constructions that seem to call for a multi-headed analysis:
 \begin{description}
@@
 
-In linguistics, it is common to analyze some of these structures using trees with ``empty categories.''  The subject of {\em skip} is taken to be a silent morpheme {\em PRO}:
+\noindent In linguistics, it is common to analyze some of these structures using trees with ``empty categories.''  The subject of {\em skip} is taken to be a silent morpheme {\em PRO}:
 {\em ``Jill$_i$ likes PRO$_i$ to skip.''}  However, this is no longer a tree if one considers the implicit undirected edge between {\em Jill} and {\em PRO} (denoted by their shared index $i$).
@@
 
+\medskip
+
+An anonymous reviewer objects to research on projective parsing algorithms, since the {\sc Planar} restriction is linguistically questionable even for single-headed parsing, and even more so for multi-headed parsing.
+However, efficient algorithms often exist in the projective case, and these projective algorithms---which are typically first developed on a projective corpus of the sort that we will construct---can be useful even when the true parses are not quite projective.
+Natural-language parses have a low rate of non-projective edges \cite{nivre-slides} and the non-projective parses tend to be ``almost projective'' \cite{pitler-et-al-2013}.
+Thus, one can apply a fast projective parser as an approximate method, or as one ingredient in a more complex model \cite{martins-et-al-2013}, or as the first step in a coarse-to-fine or stacking architecture \cite{rush-petrov-2012,martins-et-al-2008} in order to obtain preliminary edge scores that are then supplied to a non-projective parser.
+Another approach is to transform non-projective parses into a projective annotation style so that projective parsers can be used \cite{nivre-nilsson-2005,nivre-2009}.
+
 \section{Link Grammars}
 
-Graph representations of syntactic and semantic structure have been widely considered of late \cite{demarneffe-manning-2008,ivanova-et-al-2012,AMR-2013,oepen-et-al-2014}.
-
+Graph representations of syntactic and semantic structure have been widely considered of late \cite{demarneffe-manning-2008,demarneffe-et-al-2013,ivanova-et-al-2012,AMR-2013,oepen-et-al-2014}.
 A few past NLP papers have explored multi-headed dependency parsing \cite{buchkromann-2006,mcdonald-pereira-2006,sagae-tsujii-2008,gomezrodriguez-nivre-2013}.
-Unfortunately, there seem to be no annotated corpora in this form other than the Danish Dependency Treebank \cite{kromann-2003}; these papers  sometimes converted corpora from other formats such as HPSG.
+They constructed their multi-headed dependency corpora by automatically converting from other formats such as HPSG.
+Currently there seem to be no corpora that were directly annotated in this form, other than the Danish Dependency Treebank \cite{kromann-2003}.
 
@@
 
-All of these options result in non-projective parses, so the parsers must use non-projective or pseudo-projective algorithms.
-
+The above work considered non-projective parses.
 It seems at first that no one has worked out annotation conventions for {\em projective} multi-headed dependency parsing.
 However, this is only half-true.
-Link Grammar \cite{SleatorTemperly91} is a grammar-based formalism for projective dependency parsing with {\em undirected} links.
+Link Grammar \cite{sleator-temperley-1991} is a grammar-based formalism for projective dependency parsing with {\em undirected} links.
 It produces undirected connected planar graphs.
-Annotation conventions are implicit in the detailed lexicon for the Link Grammar Parser,\noteJE{check caps}%
-\footnote{\url{http://www.abisource.com/projects/link-grammar/dict/introduction.html}.
-The 122 link types in the English lexicon are documented at \url{http://www.abisource.com/projects/link-grammar/dict/summarize-links.html}.}
+Annotation conventions are implicit in the detailed lexicon for the Link Grammar Parser \cite{LINK-1999}.
+The 122 link types in the English lexicon are documented at \url{http://www.abisource.com/projects/link-grammar/dict/summarize-links.html},
 which specifies for every word a constraint on the {\em sequence} of labeled leftward and rightward edges attached to it.
@@
 For example, in {\em ``The boy that Jill skipped with fell down,''} the word {\em with} uses a lexical entry that requires it to link to a governing verb to the left, an extracted object farther to the left, and nothing to the right.
-Each entry has a hand-assigned cost in \{0,1,2\} and the parser finds the parse of minimum total cost.
-% Information about cost is in this message from Linas Vepstas on 2014-02-01: 
-%   https://groups.google.com/d/msg/link-grammar/eeJw1Ofgc9U/diqPYSwuFfoJ
+Each entry has a hand-assigned cost in \{0,1,2\}, and the parser finds the parse of minimum total cost \cite{LINK-1999,vepstas-2014}.
 
@@
 
-We supposed that the link grammar lexicon designers actually had a consistent direction in mind for each edge type.
-We would expect verbs to point to their subject arguments in dependency grammar, and so we surmise that all \texttt{S} links are intended to point leftward (from verb to subject: {\em ``Jack $\stackrel{\texttt{S}}{\leftarrow}$ is falling''}).
-The link grammar designers take care to use a distinct \texttt{SI} label in cases of subject-verb inversion, and we surmise that \texttt{SI} links are intended to point rightward (again from verb to subject: {\em ``Is $\stackrel{\texttt{SI}}{\rightarrow}$ Jack falling?''}).
+In the English link grammar, an \texttt{S} edge encodes a ``subject-verb'' relation {\em whose left word serves as the subject}.
+We would expect verbs to point to their subject arguments in dependency grammar, and so we surmise that all \texttt{S} links should be interpreted as pointing leftward (from verb to subject: {\em ``\mbox{Jack $\stackrel{\texttt{S}}{\leftarrow}$ is} falling''}).
+
+In general, we supposed that the link grammar lexicon designers actually had a {\em consistent} direction in mind for each {\em edge label}.
+This does not imply that English subjects must always appear to the left of the verb!
+The link grammar designers took care to use a distinct \texttt{SI} label in cases of subject-verb inversion, and we surmise that \texttt{SI} links are intended to point rightward (again from verb to subject: {\em ``\mbox{Is $\stackrel{\texttt{SI}}{\rightarrow}$ Jack} falling?''}).
+Similarly, different edge labels are used for English ``object-verb'' relations according to whether it is the left or the right word that serves as the object.
+These labels are presumably intended to encode different edge directions.
+
+% PREVIOUS ATTEMPT
+% This does not rule out variation in the word order used to express a semantic relation.
+For example, the English link grammar allows both SVO and VSO sentences, but with different edge labels.
+An \texttt{S} edge encodes a ``subject-verb'' relation whose the left word serves as the subject, while an \texttt{SI} edge encodes the same relation when the the right word serves as the subject.
 
@@
 
-Thus, we chose to use a sample of {\em actual} sentences parsed by the link grammar, and to seek a direction mapping so that {\em these} parses would not violate our constraints after directions were imposed.
-\noteJE{clarify per reviewer comment}
+Thus, rather than considering all theoretically possible word sequences, we chose to use a sample of {\em naturally occurring} sentences parsed by the link grammar, and to seek a direction mapping so that {\em these} parses would not violate our constraints after directions were imposed.
 If no such mapping exists, then we are willing to orient a few edge tokens in the wrong direction to ensure that the parses are still well-formed---but we minimize the number of such violations.
 In this way, the empirical distribution of sentences guides our assignment of directions.
-The resulting directed corpus could be used for research on multi-headed dependency parsing.
+We are releasing the resulting multi-headed directed corpus via our personal websites.
+\noteJE{is this the best way?}
 
@@
 We generated link parses using the AbiWord/CMU link grammar parser version 5.0.8 \cite{LINKPARSER-2014}.
-The parser's coverage is not perfect: we obtained connected parses for only \input{figure/conll_analysis_sentences_total.tex}(of \input{figure/conll_analysis_sentences_original.tex}$\!\!$) English sentences and only \input{figure/russian_sentence_count.tex}(of \input{figure/russian_original_sentence_count.tex}$\!\!$) Russian sentences, discarding the other sentences\footnote{When the link parser fails, it outputs a disconnected graph representing its best effort parse within a time limit.
-We removed these sentences because our ILP encoding tries to satisfy the edges of an entire connected parse.
-\NoteJE{no, we removed them because we have no confidence that the link grammar is doing anything reasonable in this case} Given a solution, it would be possible to return to these disconnected parses, directionalize the edges, and include them along with the rest of the corpus.
-It would also be possible to include these sentences for the ILP to solve by modifying the ILP to satisfy the edges of all disconnected components of a parse.}
+The parser's coverage is less than ideal: we obtained connected parses for only \input{figure/conll_analysis_sentences_total.tex}(of \input{figure/conll_analysis_sentences_original.tex}$\!\!$) English sentences and only \input{figure/russian_sentence_count.tex}(of \mbox{\input{figure/russian_original_sentence_count.tex}$\!\!$)} Russian sentences, discarding the other sentences.\footnote{When the link parser fails, it outputs a disconnected graph representing its best effort parse within a time limit.
+We removed these sentences for fear that the parses would be unreliable.
+% REMARK BY JUNEKI, PROBABLY TOO DISTRACTING
+% our ILP encoding tries to satisfy the edges of an entire connected parse.
+Given a solution, it would be possible to return to these disconnected parses, directionalize the edges, and include them along with the rest of the corpus.
+It would also be possible to include these sentences for the ILP to solve by modifying the ILP to satisfy the edges of all disconnected components of a parse.
+}
 These two languages have the most mature lexicons at present, although lexicons for 7 other languages are available.
 
-The link grammar parses have \input{figure/conll_analysis_edge_percent.tex}more edges overall, indicating that their directed versions will have a few multi-headed constructions per sentence.  They
+On English, the link grammar parses have \input{figure/conll_analysis_edge_percent.tex}more edges overall, indicating that their directed versions will have a few multi-headed constructions per sentence.
+They
 do differ in style from the single-headed CoNLL parses of the same English sentences.
@@
 The resulting English corpus uses a syntactic annotation scheme that is somewhat different from the CoNLL annotations.  Differences are tabulated in
-Appendix~\ref{app:linktypes}.  The actual parses are contrasted in Appendix~\ref{app:parses}, and fragments of these are shown in Table~\ref{tab:linktypes} and Figure~\ref{fig:parses}.
-%By permission of the program chairs, these appendices are available at the anonymous URL \url{http://tidyurl.com/LinksToDAG_supplement}.
+Appendix~\ref{app:linktypes} of the supplementary material, while the actual parses are contrasted in Appendix~\ref{app:parses}.
+Fragments of these appendices are shown in Table~\ref{tab:linktypes} and Figure~\ref{fig:parses}.
+% By permission of the program chairs, these appendices are available at the anonymous URL \url{http://tidyurl.com/LinksToDAG_supplement}.
 
@@
 This would permit our desired treatment of {\em ``Jill persuaded him to skip''}, in which {\em ``him''} has two parents.
-But the ILP solution generally treats subjects as {\em parents} of verbs (thus we get {\em him} $\rightarrow$ {\em to}), owing to inconsistencies in the link grammar itself.\footnote{Why?
-When the subject is missing ({\em ``Jill wanted to skip''}), the governing verb now links instead to the clause's first (tensed) verb but {\em no longer} to the last as well.
-That means that a VP must be headed by its first verb ({\em to} $\rightarrow$ {\em skip}).
-As a result, when the subject is present, it must be the {\em parent} of the first verb ({\em him} $\rightarrow$ {\em to} $\rightarrow$ {\em skip}), since no other parent is available.
-({\em Note:} Subjects are not {\em always} parents.
-The construction {\em ``It was impossible, they said''} is handled inconsistently by the link grammar:
-the subject ({\em ``they''}) links only to the verb ({\em ``said''}) and so must be its child, no other parent being available.)}
+But the ILP solution generally treats subjects as {\em parents} of verbs (thus we get {\em him} $\rightarrow$ {\em to}).
+The reason for this is an inconsistency in the link grammar itself.\footnote{Specifically, the link grammar is inconsistent about how it handles clauses with missing subjects ({\em ``Jill wanted to skip''}).
+In this case, the governing verb links to the clause's first (tensed) verb in lieu of its subject.
+The problem is that it no longer also links to the clause's last (main) verb as it would if the subject were present.
+Hence our method concludes that a VP is always headed by its first verb ({\em to} $\rightarrow$ {\em skip}).
+That is a respectable convention on its own, but recall that unfortunately, the governing verb does {\em not} link to this first verb when the subject is present.
+As a result, the only remaining possible parent for the first verb is the subject ({\em him} $\rightarrow$ {\em to} $\rightarrow$ {\em skip}).
+This leads to the odd solution where subjects are treated as parents of verbs in general.
+
+Note that subjects are not {\em always} parents, due to another inconsistency in the link grammar.
+In the construction {\em ``It was impossible, they said''}, the subject ({\em ``they''}) is inconsistently not linked to 0 but only to the verb ({\em ``said''}) and so must be the verb's child, no other parent being available.}
+Fixing the link grammar would presumably correct the link direction.
+As noted in section~\ref{sec:motivation}, it is arguably a positive result that our method was able to detect a problem in the grammar engineering.
 
@@
 
+\clearpage
 \appendix
-\clearpage
-\onecolumn
 

### 2014-11-21 doc/LinksToDAG_final.tex
answered a couple more reviewer issues

@@
 \noteJE{check!}
+One could experiment with adjusting the constant 1/4; we selected that value simply because it seemed like a reasonable threshold {\em a priori}.
 % CUT FOR SPACE
@@
 The other annotation differences are generally reasonable, e.g., letting {\em 's} be the head of a possessive, and different handling of punctuation and lists.
+One could easily modify the ILP to explicitly encourage agreement with the CoNLL link directions (for word pairs that are linked in CoNLL).
 Of course, a few of the differences are due to parser attachment errors.

### 2014-11-22 doc/LinksToDAG_final.tex
resolved some TODOs

@@
 \noteJE{IMPORTANT!
-current graph is with fine-grained tags; rerun with the coarse-grained tags as in the rest of the paper} Using even a small prefix of the data got very high-precision results, in the sense that nearly all $r_L$ or $\ell_L$ variables that were 1 in this lightly trained mapping were also 1 in our largest run.
+current graph is with fine-grained tags; rerun with the coarse-grained tags as in the rest of the paper} \noteJH{done} Using even a small prefix of the data got very high-precision results, in the sense that nearly all $r_L$ or $\ell_L$ variables that were 1 in this lightly trained mapping were also 1 in our largest run.
 The only disadvantage to using small data is low recall relative to the large run---many of the labels $L$ are not observed yet and so we do not yet allow either direction ($r_L=\ell_L=0$).
@@
 Below is the result of our largest English run, whose corpus uses \input{figure/allowed_labels_total.tex}of these types.
-\noteJH{automatically generate this number correctly: 107}
 

### 2014-12-06 doc/slides/hong+eisner.TLT13.slides.tex
initial outline of paper

--- 2014-11-21 doc/LinksToDAG_final.tex
@@
-%
-% File acl2014.tex
-%
-% Contact: koller@ling.uni-potsdam.de, yusuke@nii.ac.jp
-%%
-%% Based on the style files for ACL-2013, which were, in turn,
-%% Based on the style files for ACL-2012, which were, in turn,
-%% based on the style files for ACL-2011, which were, in turn, 
-%% based on the style files for ACL-2010, which were, in turn, 
-%% based on the style files for ACL-IJCNLP-2009, which were, in turn,
-%% based on the style files for EACL-2009 and IJCNLP-2008...
-%% Based on the style files for EACL 2006 by 
-%%e.agirre@ehu.es or Sergi.Balari@uab.es
-%% and that of ACL 08 by Joakim Nivre and Noah Smith
+\documentclass{beamer}
 
-\documentclass[a4paper,11pt]{article}
-\usepackage[latin1]{inputenc}
-\usepackage[T1]{fontenc}         
-\usepackage{pslatex}
+% For more theme options:
+%/usr/share/texmf/tex/latex/beamer/base/themes/theme/compatibility
+\usepackage{beamerthemelined}
 
-\usepackage{hyperref}
-\pagestyle{empty}
-
-%\bibliographystyle{acl2014}
-\bibliographystyle{plain}
-\newcommand{\newcite}[1]{\cite{#1}}   % ACL's newcite is not available in plain bib style, so just use cite
-
-%\documentclass[11pt]{article}
-%\usepackage{acl2014}
-\usepackage{times}
-%\usepackage{url}
-%\usepackage{latexsym}
-\usepackage{amsmath}
-
-%\usepackage{xspace}
-\usepackage{enumitem}
-\usepackage[small]{caption}
-\usepackage{longtable}
-
-\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
-%\usepackage{soul}
+% this package seems to throw an error for me. -Juneki 12/6/14
+%\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
+\usepackage{soul}
 
@@
 \usepackage{graphicx}
-%\usepackage{caption}
+\usepackage{caption}
 \usepackage{subcaption}
@@
 \newcommand{\eqnref}[1]{\eqref{eqn:#1}}
-
+%\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}  % allows better color names
 \usepackage{todonotes}   % insert [disable] to disable all notes
-% Note that these macros accept optional arguments such as 
-% [size=\small,bordercolor=red].
 \newcommand{\Note}[4][]{\todo[author=#2,color=#3,fancyline,#1]{#4}}
-\newcommand{\noteJH}[2][]{\Note[size=\small,#1]{JH}{blue!40}{#2}}
-\newcommand{\noteJE}[2][]{\Note[size=\small,#1]{JE}{green!40}{#2}}   
-\newcommand{\notewho}[3][]{\Note[size=\small,#1]{#2}{orange!40}{#3}}  % extra arg with miscellaneous author
+\newcommand{\noteJH}[2][]{\Note[#1]{JH}{blue!40}{#2}}
+\newcommand{\noteJE}[2][]{\Note[#1]{JE}{green!40}{#2}}
+\newcommand{\notewho}[3][]{\Note[#1]{#2}{orange!40}{#3}}  % extra arg with miscellaneous author
 \newcommand{\NoteJH}[2][]{\noteJH[inline,#1]{#2}}
@@
 
-\title{Deriving Multi-Headed Planar Dependency Parses from Link
-  Grammar Parses\thanks{This material is based upon work supported by the National
-    Science Foundation under Grant No. 1423276.  The work was mainly
-    conducted while the first author was at Johns Hopkins
-    University.\noteJE{anyone else to thank?}}}
-% Oriented link grammar: Creating a multi-headed dependency corpus.
+\begin{document}
+\title{Deriving Multi-Headed Planar Dependency Parses from Link Grammar Parses}
+\author{Juneki Hong, Jason Eisner}
+\date{\today}
+\frame{\titlepage}
 
-\author{
-  Juneki Hong \\
-  [0.5cm] Language Technologies Institute\\
-  Carnegie Mellon University\\
-  Pittsburgh, PA 15213, USA\\
-  {\texttt{juneki@cs.cmu.edu}}
-  \and
-  Jason Eisner\\
-  [0.5cm] Department of Computer Science\\
-  Johns Hopkins University\\
-  Baltimore, MD 21218, USA \\
-  {\texttt{jason@cs.jhu.edu}}
+\section{Motivation}
+\frame{\frametitle{Motivation}
+
 }
 
-\date{}
+\section{Multi-headed}
+\frame{\frametitle{Multi-headed Dependency Parsing}
 
-%\setlength\titlebox{6.5cm}    % Expanding the titlebox
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{multiheaded}
+  \caption{English is not completely multiheaded}
+  \NoteJH{image taken from Chris Dyer's slide: http://demo.clab.cs.cmu.edu/fa2014-11711/images/0/0e/Depparsing.pdf}
+\end{figure}
 
-\begin{document}
-\maketitle
-
-\begin{abstract}
-\noindent
-Under multi-headed dependency grammar, a parse is a connected DAG rather than a tree.  Such formalisms can be more syntactically and semantically expressive.  However, 
-it is hard to train, test, or improve multi-headed parsers because few multi-headed corpora exist, particularly for the projective or planar case.
-% OLD
-% investigate the benefit of such parsers or to work on making them faster or more accurate.  
-To help fill this gap, we observe that link grammar already produces {\em undirected} planar graphs, and so we wanted to determine whether these could be converted into directionalized dependency parses.
-% OLD
-% ... produces parses that are similar to multi-headed projective dependency parses, except that the links are undirected. 
-We use Integer Linear Programming to assign consistent directions to the labeled links in a corpus of several thousand parses produced by the Link Grammar Parser, which has broad-coverage hand-written grammars of English as well as Russian and other languages.
-We find that such directions can indeed be consistently assigned in a way that yields valid multi-headed dependency parses.
-The resulting parses in English appear reasonably linguistically plausible, though differing in style from CoNLL-style parses of the same sentences; we discuss the differences.
-
-\end{abstract}
-
-\thispagestyle{empty}
-
-%\newpage
-
-\section{Motivation}\label{sec:motivation}
-
-Link Grammar \cite{sleator-temperley-1991} is a syntactic formalism in which a parse of a sentence is an undirected, edge-labeled, planar graph.
-The labeled edges of the graph represent syntactic relationships among the words.
-The vertices of the graph are simply the words $1, 2, \ldots, n$ of the sentence, along with a distinguished ``root'' vertex 0.
-
-The small Link Grammar community has invested effort in creating link grammars for several languages.
-In this short paper, we consider whether their {\em undirected} parses can be converted automatically to {\em directed} ones.
-We have three motivations:
-
-\begin{enumerate}
-
-\item We were curious about the relation between link grammar annotation and dependency grammar annotation.
-We suspected that the link grammar parses could be interpreted as multi-headed dependency grammar parses.
-Although the link grammar authors did not bother to specify directions for the different edge types, we suspected that they essentially had such directions in mind.
-
-\item Our problem provides a good case study for how to automatically enrich a corpus.
-Hand-constructed grammars or corpora sometimes provide a lighter level of annotation than desired.
-In our setting, the edges lack directions; in other settings, the syntactic categories may be coarser than desired, or some relations may be omitted.
-One may want to automatically enrich the annotation in such cases, whether by doing some kind of learning \cite[et seq.]{matsuzaki-et-al-2005}, or by exploiting implicit constraints.
-In this paper, we use Integer Linear Programming to exploit implicit constraints of consistency and acyclicity.
-
-\item The resulting parses may be useful data for experimenting with new parsing {\em algorithms}.
-There has been a good deal of recent research on projective dependency parsing, variously using global optimization or sequential classification (see \cite{martins-et-al-2008,bohnet-2011,chen-li-zhang-2014} for surveys).
-Some of these algorithms could be extended to the {\em multi-headed} case, which is of linguistic and computational interest for reasons discussed below.
-However, for training and testing such algorithms, one would need a plausible sample of multi-headed projective dependency parses of real sentences.
-Our method cheaply manufactures such a sample, to compensate for the current lack of gold-standard data of this form.
-
-\end{enumerate}
-
-\noindent Our automatic method for reconstructing the latent directions also had an unexpected benefit.
-It revealed an inconsistency in the hand-written English link grammar, regarding the handling of embedded sentences with missing (PRO) subjects.
-
-\section{Multi-Headed Dependency Parsing}
-
-Dependency parsing maps a sentence to a directed graph whose vertices are the words $1, 2, \ldots, n$ of the sentence along with a distinguished ``root'' vertex 0.
-A labeled directed edge $u \stackrel{L}{\rightarrow} v$ or $v \stackrel{L}{\leftarrow} u$ indicates that the ``child'' $v$ is some kind of argument or modifier of its ``parent'' $u$.
-The edge label $L$ indicates the specific syntactic or semantic relationship between the two words.
-
-In the special case $u=0$, the edge designates $v$ as playing some special top-level role in the sentence, e.g., as the main verb.  We disallow $v=0$.
-
-As discussed by \newcite{kuhlmann-nivre-2006,gomezrodriguez-nivre-2013}, one might impose various requirements on the parse graph:
-\begin{itemize}[noitemsep]
-\item {\sc Single-Head}: Each word has $\leq 1$ parent.
-\item {\sc Acyclic}: There are no directed cycles.
-\item {\sc Connected}: Each pair of words has a undirected path between them.
-\item {\sc Reachable}: Each word can be reached from 0 by a directed path (which implies {\sc Connected}).  Note that 0 may have multiple children.
-\item {\sc Planar}: edges may not ``cross.''  That is, if there are edges between $i,j$ and between $u,v$, where $i < u < j$, then {\sc Planar} requires $i \leq v \leq j$.
-\end{itemize}
-It is common to impose all of these requirements at once, leading to a {\em projective dependency parser} that produces projective trees rooted at 0.
-However, parsing algorithms can be devised that relax any or all of the requirements \cite{gomezrodriguez-nivre-2013}.
-
-In this paper, we are interested in relaxing the {\sc Single-Head} requirement while preserving all the others.
-This is the setting of {\em multi-headed projective dependency parsing}.
-Just as in the single-headed case, the other requirements ensure that all edges are projective.
-(A projective edge is one where the parent is an ancestor of all words between the parent and the child \cite{nivre-slides}.)
-
-Relaxing {\sc Single-Head} means that the parse can have more than $n$ edges, allowing it to express more relationships between words.  
-In English, for example, here are some constructions that seem to call for a multi-headed analysis:
-\begin{description}
-\item[control] In {\em ``Jill likes to skip,''} the word {\em Jill} is the subject of two verbs.
-In {\em ``Jill persuaded Jack to skip,''} {\em Jack} is the object of one verb and the subject of another.
-Without recognizing this, our parser would miss the syntactic invariants that {\em skip} always has a subject and {\em persuaded} always has an object.
-It would also be unable to exploit the selectional preferences of both verbs to help disambiguate the parse.
-This is why we prefer to make the parser aware of multi-headedness, rather than using a single-headed parser and then extracting the additional semantic roles from its output.
-\item[relativization] In {\em ``The boy that Jill skipped with fell down,''} the word {\em boy} is the object of {\em with} as well as the subject of {\em fell}.
-Without recognizing this, we would miss the syntactic invariant that {\em with} always has an object.
-\item[conjunction] In {\em ``Jack and Jill went up the hill,''} {\em Jack} and {\em Jill} serve as the two arguments to {\em and}, but they are also semantically subjects of {\em went}.
-Without recognizing this, we would have no (local) reason for expecting the arguments of {\em and} to be nouns.
-\end{description}
-
-\noindent In linguistics, it is common to analyze some of these structures using trees with ``empty categories.''  The subject of {\em skip} is taken to be a silent morpheme {\em PRO}:
-{\em ``Jill$_i$ likes PRO$_i$ to skip.''}  However, this is no longer a tree if one considers the implicit undirected edge between {\em Jill} and {\em PRO} (denoted by their shared index $i$).
-Our simpler representation contracts this coreference edge, eliminating {\em PRO} and creating a {\em Jill} $\leftarrow$ {\em skip} link.
-
-\medskip
-
-An anonymous reviewer objects to research on projective parsing algorithms, since the {\sc Planar} restriction is linguistically questionable even for single-headed parsing, and even more so for multi-headed parsing.
-However, efficient algorithms often exist in the projective case, and these projective algorithms---which are typically first developed on a projective corpus of the sort that we will construct---can be useful even when the true parses are not quite projective.
-Natural-language parses have a low rate of non-projective edges \cite{nivre-slides} and the non-projective parses tend to be ``almost projective'' \cite{pitler-et-al-2013}.
-Thus, one can apply a fast projective parser as an approximate method, or as one ingredient in a more complex model \cite{martins-et-al-2013}, or as the first step in a coarse-to-fine or stacking architecture \cite{rush-petrov-2012,martins-et-al-2008} in order to obtain preliminary edge scores that are then supplied to a non-projective parser.
-Another approach is to transform non-projective parses into a projective annotation style so that projective parsers can be used \cite{nivre-nilsson-2005,nivre-2009}.
+}
 
 \section{Link Grammars}
+\frame{\frametitle{Link Grammars}
 
-Graph representations of syntactic and semantic structure have been widely considered of late \cite{demarneffe-manning-2008,demarneffe-et-al-2013,ivanova-et-al-2012,AMR-2013,oepen-et-al-2014}.
-A few past NLP papers have explored multi-headed dependency parsing \cite{buchkromann-2006,mcdonald-pereira-2006,sagae-tsujii-2008,gomezrodriguez-nivre-2013}.
-They constructed their multi-headed dependency corpora by automatically converting from other formats such as HPSG.
-Currently there seem to be no corpora that were directly annotated in this form, other than the Danish Dependency Treebank \cite{kromann-2003}.
+\begin{figure}
+\begin{subfigure}[b]{0.4850\textwidth}
+	\begin{dependency}
+		\begin{deptext}
+			the \& matter \& may \& never \& even \& be \& tried \& in \& court \& . \\
+			- \& n-u \& v \& e \& e \& v \& v-d \& r \& n-u \& - \\
+		\end{deptext}
+		\depedge[edge above, edge style = {red}]{2}{3}{S}
+		\depedge[edge above, edge style = {red}]{3}{6}{I}
+		\depedge[edge above, edge style = {red}]{2}{1}{D}
+		\depedge[edge above, edge style = {red}]{7}{8}{MV}
+		\depedge[edge above, edge style = {red}]{6}{7}{P}
+		\depedge[edge above, edge style = {red}]{8}{9}{J}
+		\deproot[edge above, edge style = {red}]{2}{W}
+		\deproot[edge above, edge style = {red}]{7}{WV}
+		\deproot[edge above, edge style = {red}]{10}{X}
+		\depedge[edge above, edge style = {red}]{6}{4}{E}
+		\depedge[edge above, edge style = {red}]{6}{5}{E}
+	\end{dependency}
+\end{subfigure}
 
-% ADDRESSED BY ABOVE?
-% \NoteJE{We should acknowledge that another option would be to post-process either CoNLL parses or our link parses to automatically add more multi-heading, e.g., to handle control phenomena ("John wants to skip" or "Jane ran for office in order to change the tax laws" should have John/Jane be the subject of two verbs).
-However, that requires per-language}
+\end{figure}
+\NoteJH{figure not centered}
+\NoteJH{Make the edges undirected}
 
-The above work considered non-projective parses.
-It seems at first that no one has worked out annotation conventions for {\em projective} multi-headed dependency parsing.
-However, this is only half-true.
-Link Grammar \cite{sleator-temperley-1991} is a grammar-based formalism for projective dependency parsing with {\em undirected} links.
-It produces undirected connected planar graphs.
-Annotation conventions are implicit in the detailed lexicon for the Link Grammar Parser \cite{LINK-1999}.
-The 122 link types in the English lexicon are documented at \url{http://www.abisource.com/projects/link-grammar/dict/summarize-links.html},
-which specifies for every word a constraint on the {\em sequence} of labeled leftward and rightward edges attached to it.
-As remarked by \newcite{eisner-2000-iwptbook}, this is analogous to dependency grammar's use of head automata to constrain a word's sequence of left and right children.
-For example, in {\em ``The boy that Jill skipped with fell down,''} the word {\em with} uses a lexical entry that requires it to link to a governing verb to the left, an extracted object farther to the left, and nothing to the right.
-Each entry has a hand-assigned cost in \{0,1,2\}, and the parser finds the parse of minimum total cost \cite{LINK-1999,vepstas-2014}.
+}
 
-Given a link grammar parse, it would be straightforward to convert it to an acyclic dependency parse by orienting all edges rightward.
-However, the result may violate the {\sc Reachable} constraint.
-Instead we could orient all edges by depth-first search from the root node, which yields a DAG satisfying all our constraints.
-However, this might result in inconsistent annotation conventions, with some \texttt{S}-labeled links pointing from subject to verb and others from verb to subject.
-
-In the English link grammar, an \texttt{S} edge encodes a ``subject-verb'' relation {\em whose left word serves as the subject}.
-We would expect verbs to point to their subject arguments in dependency grammar, and so we surmise that all \texttt{S} links should be interpreted as pointing leftward (from verb to subject: {\em ``\mbox{Jack $\stackrel{\texttt{S}}{\leftarrow}$ is} falling''}).
-
-In general, we supposed that the link grammar lexicon designers actually had a {\em consistent} direction in mind for each {\em edge label}.
-This does not imply that English subjects must always appear to the left of the verb!
-The link grammar designers took care to use a distinct \texttt{SI} label in cases of subject-verb inversion, and we surmise that \texttt{SI} links are intended to point rightward (again from verb to subject: {\em ``\mbox{Is $\stackrel{\texttt{SI}}{\rightarrow}$ Jack} falling?''}).
-Similarly, different edge labels are used for English ``object-verb'' relations according to whether it is the left or the right word that serves as the object.
-These labels are presumably intended to encode different edge directions.
-
-% PREVIOUS ATTEMPT
-% This does not rule out variation in the word order used to express a semantic relation.
-For example, the English link grammar allows both SVO and VSO sentences, but with different edge labels.
-An \texttt{S} edge encodes a ``subject-verb'' relation whose the left word serves as the subject, while an \texttt{SI} edge encodes the same relation when the the right word serves as the subject.
-
-Our goal in this paper is to recover these implicit directions by global optimization.
-We seek a fixed mapping from labels to directions such that link grammar parses become directed dependency parses that satisfy all of our constraints.
-
-Our first thought was to seek a direction mapping such that no parsed word sequence allowed by the link grammar lexicon could possibly violate our constraints after directions were imposed.
-This is a well-defined constraint programming problem.
-For example, to prevent cyclicity, we would require (roughly speaking) that no word type in the lexicon could follow a sequence of directed rightward links through other word types and then a leftward link back to itself.
-
-% REMOVED AT REVIEWER SUGGESTION
-% However, working directly with the link grammar lexicon format is somewhat tricky. We also feared
-However, we feared that there would not be a feasible solution---because of errors in the lexicon or linguistically unnatural word sequences not anticipated by the grammar designers.
-In this case it would be unclear how to relax our constraints.
-
-Thus, rather than considering all theoretically possible word sequences, we chose to use a sample of {\em naturally occurring} sentences parsed by the link grammar, and to seek a direction mapping so that {\em these} parses would not violate our constraints after directions were imposed.
-If no such mapping exists, then we are willing to orient a few edge tokens in the wrong direction to ensure that the parses are still well-formed---but we minimize the number of such violations.
-In this way, the empirical distribution of sentences guides our assignment of directions.
-We are releasing the resulting multi-headed directed corpus via our personal websites.
-\noteJE{is this the best way?}
-
-% Link grammars are a grammatical system equivalent to context-free grammars that assign linking requirements to every given word.
-A link parser then tries to satisfy all of these requirements for every word of a sentence while still maintaining planarity.
-The resulting links describe the relationships between constituents in a parse.
-
-% The link grammar is based on a set of handwritten dictionaries.
-Instead of going through these dictionaries, we learned the grammar using an ILP.
-This approach also potentially allows us to analyze link grammar dictionaries other than English.
-\todo{explore other link grammar dictionaries}
-
-% \NoteJE{what statistics does it use?  what languages are available}
-
-% Link parsing in contrast produces a multi-headed planar graph with undirected edges, where every edge has a label describing the relationship between two constituents in a parse.
-In this paper we explore whether these relationships include dependencies, and whether the multi-headedness of the link grammar offers additional dependency relationships not found in other corpora.
-
-% Dependency parsing is the task of mapping a sentence to a projective (not always projective?) directed acyclic tree.
-Link parsing in contrast produces a multi-headed planar graph with undirected edges, where every edge has a label describing the relationship between two constituents in a parse.
-In this paper we explore whether these relationships include dependencies.
-To determine the directional dependencies within the link edge labels we will use integer linear programming, encoding the problem in the Zimpl little language \cite{Koch2004}.
-It turns out that the link parses roughly only match half of the conll dependency corpus.
-However this is because \todo{}.
-
-\section{Data Sets}
-We used the English sentences from the CoNLL 2007 Shared Task \cite{CONLL-SHARED-2007}---a subset of the Penn Treebank for which {\em single}-headed reference parses are available. 
-%We also used a prefix of the Russian National Corpus,\footnote{That is, a prefix when considering the files in Unix sort order by filename.
-We dropped sentences with $\leq 2$ words, as well as file 22592319, which appeared to contain charts.} which is unparsed.
-We also used a prefix of the Russian News Commentary data from the ACL 2013 Shared Task of Machine Translation,\footnote{\url{http://www.statmt.org/wmt13/training-monolingual-nc-v8.tgz}} which is unparsed.
-
-We generated link parses using the AbiWord/CMU link grammar parser version 5.0.8 \cite{LINKPARSER-2014}.
-The parser's coverage is less than ideal: we obtained connected parses for only \input{figure/conll_analysis_sentences_total.tex}(of \input{figure/conll_analysis_sentences_original.tex}$\!\!$) English sentences and only \input{figure/russian_sentence_count.tex}(of \mbox{\input{figure/russian_original_sentence_count.tex}$\!\!$)} Russian sentences, discarding the other sentences.\footnote{When the link parser fails, it outputs a disconnected graph representing its best effort parse within a time limit.
-We removed these sentences for fear that the parses would be unreliable.
-% REMARK BY JUNEKI, PROBABLY TOO DISTRACTING
-% our ILP encoding tries to satisfy the edges of an entire connected parse.
-Given a solution, it would be possible to return to these disconnected parses, directionalize the edges, and include them along with the rest of the corpus.
-It would also be possible to include these sentences for the ILP to solve by modifying the ILP to satisfy the edges of all disconnected components of a parse.
-}
-These two languages have the most mature lexicons at present, although lexicons for 7 other languages are available.
-
-On English, the link grammar parses have \input{figure/conll_analysis_edge_percent.tex}more edges overall, indicating that their directed versions will have a few multi-headed constructions per sentence.
-They
-do differ in style from the single-headed CoNLL parses of the same English sentences.
-Only \input{figure/conll_analysis_precision.tex}of the links match CoNLL arcs, and only \input{figure/conll_analysis_recall.tex}of the CoNLL arcs match links.
-% DATA FOR THE ABOVE
-% \begin{figure}[ht!]
-%   \centering
-%   \small
-%   \input{figure/conll_analysis_arcs.tex}
-%   \caption{Arcs that match links}
-% \end{figure}
-% 
-% OLD
-% Using the sentences of the dependency corpus as comparison, we find that the link parses do not wholly subsume dependency parses, and that the undirected links match roughly three fourths of the arcs in the conll dependency corpus.
-\todo{This is because...}
-
-% With the English sentences we compared our link parse results to the original dependency annotations, and with the Russian sentences we only produced link parses.
-
-%\subsection{Discarding Incomplete Link Parses}
-
-%\NoteJE{ideally we could cut this section given what I wrote above.
-However, you seem to be saying that we did train on sentences with dropped words as long as the parses were connected.
-Didn't we decide to drop those sentences too?}
-
-%Not all of the data was used for this work.
-From the given sentences, we discarded those that the link parser could not process and returned a parse with nodes disconnected from the root.
-\input{figure/conll_analysis_sentences_disconnected} of the English sentences were removed in this way.
-The ILP was run only on the connected parses to produce a corpus of directionalized link parses.
-
-%For our analysis of these parses, we further discarded those that had dropped words, where the link parser could not attach links to every word in the sentence.
-About \input{figure/conll_analysis_sentences_dropped} of the non-disconnected sentences were removed, leaving us working only with complete link parses.
-
-%\begin{figure}[ht!]
-%  \centering
-%  \small
-%  \input{figure/conll_analysis_sentences.tex}
-%  \caption{\small The English sentences used for this work.}
-%\end{figure}
-
-%\noteJH{The stability results come from the second row of this table. The analysis later in the paper uses the third row. Putting this section here might be confusing to readers?}
-
-\section{Integer Linear Programming Model}
-
-For each undirected labeled edge $ij$ in the link corpus, where $i,j$ denote tokens in the same sentence with $i < j$, we introduce nonnegative integer variables $x_{ij}$ and $x_{ji}$ with a constraint $x_{ij}+x_{ji}=1$.
-We interpret $x_{ij}=1$ or $x_{ji}=1$ to mean that the link has direction $i \rightarrow j$ or $i \leftarrow j$, respectively.\footnote{In practice we halve the number of variables by replacing $x_{ji}$ with $1-x_{ij}$ for $j > i$, but that obscures the exposition.}
-
-% OLD VERSION
-% we introduce a variable $x_{ij}$ that is 0 or 1 according to whether $e$ points left or right.  We abbreviate $1-x_{ij}$ as $\bar{x}_{ij}$.
-
-For each non-0 token $v$, we ensure that it has at least one parent by constraining\footnote{To denote two linked tokens, we use variables $i,j$ when $i$ is to the left of $j$, or variables $u,v$ when $u$ is the parent of $v$.}
-\begin{align}\label{eqn:oneparent}
-\sum_u x_{uv} & \geq 1
-\end{align}
-% OLD VERSION
-% $$\sum_{u < v} x_{uv} + \sum_{u > v} \bar{x}_{vu} \geq 1$$
-where $u$ ranges only over tokens such that the relevant variable exists.
-%
-To prevent cycles,\footnote{This also ensures {\sc reachable}, given \eqnref{oneparent}.} for each token $v$ we introduce a depth variable $d_v$ in the range $[0,n_v]$ (not constrained to be integer), where $n_v$ is the length of the sentence containing $v$.
-We require a child's depth to be at least 1 greater than each of its parents' depths---constraints that can be satisfied iff the sentence has no directed cycles:
-\begin{align}\label{eqn:nocycles}
-(\forall u)\; d_v + (1+n_v) \cdot (1-x_{uv}) & \geq 1+d_u
-% OLD VERSION
-% $$(\forall u < v) d_v + (1+n_v) \cdot \bar{x}_{uv} \geq 1+d_u$$
-% $$(\forall u > v) d_v + (1+n_v) \cdot x_{vu} \geq 1+d_u$$
-\end{align}
-The second summand ensures that \eqnref{nocycles} is trivially satisfied (hence has no effect) when $u$ is {\em not} the parent of $v$.  
-\NoteJE{in final version: figure out whether it is faster to require the 0 tokens to have depth 0, or add all the depths (perhaps multiplied by 0.001) to the objective function to break ties in simplex, or neither.
-See also other notes in tex file.}
-% As a speedup, we can require the 0 tokens to have depth 0. \noteJE{does this really still make a difference to speed?  
-% \noteJE{what's the point of the root\_depth constraint in the .zpl file?  Almost trivially satisfied.}
-
-Finally, we encourage all links with the same label \noteJE{explain that we mean capital letters?} to have the same direction.
-For each label $L$, we introduce binary variables $r_L$ and $\ell_L$, which say whether a link of type $L$ is ``allowed'' to point right or left, respectively.
-For each undirected edge $ij$ of label $L$, with $i < j$, we write
-\begin{align}
-x_{ij} &\leq r_L + s_{ij} & 
-x_{ji} &\leq \ell_L + s_{ij}
-\end{align}
-where $s_{ij} \geq 0$ is a slack variable that allows an edge token to point in a disallowed direction if needed to ensure \eqnref{oneparent}--\eqnref{nocycles}.
-
-Our objective tries to minimize the number of allowed directions (by link type---cf.\@ \newcite{ravi2009}) and the total slack (by link token):
-\begin{align}\label{eqn:obj}
-\min \left( \sum_{L} r_L + \ell_L \right) \cdot \frac{N_L}{4} + \sum_{ij} s_{ij}
-\end{align}
-where $N_L$ is the number of link tokens with label $L$.
-Objective \eqnref{obj} is willing to tolerate up to 1/4 of those link tokens' using a disallowed direction before it prefers to allow {\em both} directions.
-\noteJE{check!}
-% CUT FOR SPACE
-% (i.e., $r_L+\ell_L=2$).
-
-\NoteJE{It might be a good idea to have the second allowed direction still carry some lesser slack penalty, so that we break ties in favor of the preferred direction.
-The two bits for each label would then say ``which direction is preferred?'' and ``is the other direction allowed?''}
-
-% OLD
-% We formulate an ILP to find the minimal set of link type to directionality assignments that would directionalize all the links in a set of link grammar parses, while still mandating that all of the resulting dependency parses would be connected DAGs, and with all the nodes reachable from the root.
-Our task of finding the smallest set of assignments that would explain the parses is also called model minimization\cite{ravi2009}.
-\noteJE{worth citing, but not quite the same thing}
-% For every encountered label we generate two boolean label/direction variables (e.g.
-label/left, label/right).
-Setting one of these variables to \textsc{true} allows the label type to only go in the specified direction, while setting both would allow the label to go in either way.
-% The objective is to minimize the number of label/direction variables set to \textsc{true}, while still satisfying acyclicity and reachability constraints.
-The encoding was written in the Zimpl little language \cite{Koch2004} and solved using the SCIP Optimization Suite \cite{achterberg2009scip}.
-% \subsection{Slack}
-% We introduced slack such that the link types were allowed to deviate from the majority one percent of the time before the ILP would assign both label/direction variables to \textsc{true}.
-This allows for flexibility against noise in the link parser's label assignments, balanced with allowing for both directions to be assigned if necessary.
-
-% 
-% \begin{figure*}[ht!]
-%   \vspace{-6pt}
-%   \centering
-%   \input{figure/sample_parses.tikz}
-%   \vspace{-6pt}
-% \end{figure*}
-
-\section{Experiments and Results}
-
-%\hspace*{-1.5in}
-\begin{figure*}[ht!]
-
-  \vspace{-6pt}
-  \centering
-
-  %\input{figure/sentences.tikz}
-
-  \hspace*{-0.5in}
-  \begin{subfigure}[b]{0.49\textwidth}
+\frame{\frametitle{Link Grammars}
+\begin{figure}
+\begin{subfigure}[b]{0.4850\textwidth}
 	\begin{dependency}
@@
 \end{subfigure}
-  \hspace*{0.5in}\begin{subfigure}[b]{0.49\textwidth}
-	\begin{dependency}
-	  \begin{deptext}
-		{\scriptsize DT} \& {\scriptsize NN} \& {\scriptsize NN} \& {\scriptsize VB} \& {\scriptsize RB} \& {\scriptsize VB} \& {\scriptsize DT} \& {\scriptsize NN} \& {\scriptsize .} \\
-		the \& Paris \& bank \& does \& n't \& publish \& that \& figure \& . \\
-		- \& b \& n \& v \& - \& v \& j-d \& n \& - \\
-	  \end{deptext}
-	  \deproot[edge above, edge style = {blue, dotted}]{4}{\small ROOT}
-	  \depedge[edge below, edge style = {red, thick}]{3}{1}{D}
-	  \depedge[edge above, edge style = {blue, thick}]{3}{1}{\small NMOD}
-	  \depedge[edge below, edge style = {red, thick}]{3}{2}{AN}
-	  \depedge[edge above, edge style = {blue, thick}]{3}{2}{\small NMOD}
-	  \depedge[edge below, edge style = {red, thick}]{6}{8}{O}
-	  \depedge[edge above, edge style = {blue, thick}]{6}{8}{\small OBJ}
-	  \depedge[edge below, edge style = {red, ultra thick}]{3}{4}{S}
-	  \depedge[edge above, edge style = {blue, ultra thick}]{4}{3}{\small SBJ}
-	  \depedge[edge below, edge style = {red, thick}]{4}{5}{N}
-	  \depedge[edge above, edge style = {blue, thick}]{4}{5}{\small VMOD}
-	  \depedge[edge below, edge style = {orange, thick}]{4}{6}{I}
-	  \depedge[edge above, edge style = {blue, thick}]{4}{6}{\small VC}
-	  \depedge[edge above, edge style = {blue, dotted}, edge unit distance = 1.7ex]{4}{9}{\small P}
-	  \depedge[edge below, edge style = {red, thick}]{8}{7}{D}
-	  \depedge[edge above, edge style = {blue, thick}]{8}{7}{\small NMOD}
-	  \deproot[edge below, edge style = {red, dotted}]{3}{W}
-	  \deproot[edge below, edge style = {orange, ultra thick, dotted}]{6}{WV}
-	  \deproot[edge below, edge style = {red, dotted}]{9}{X}
-	\end{dependency}
-  \end{subfigure}
 
-  \hspace*{-1in}
-  \input{figure/long_parses.tikz}
-
-  %\input{figure/long_parse2.tikz}
-  \caption{\small The blue upper edges show CoNLL gold dependency parses; the red lower edges show our oriented version of the link grammar parses.
-Edges are shown as dotted lines if they appear only in one parse.
-Edges are highlighted in orange if the child has multiple parents.
-Edges that appear in both parses are solid lines, drawn thicker if the directions do not match.
-Vertical edges have parent 0.
-For 100 example parses, see Appendix~\ref{app:parses} of the supplementary material.}
-  \label{fig:parses}
-  \vspace{-6pt}
-
-\end{figure*}
-
-\begin{figure}[t!]
-  \small
-  \centering
-  \includegraphics[width=0.7\textwidth, keepaspectratio=true]{figure/precision_recall.png}
-  \caption{Rapid convergence to the direction mapping obtained on the largest dataset.
-The direction mappings obtained on small datasets have high {\em precision} relative to the one obtained on the largest dataset.
-Their {\em recall} grows as more link types are seen and directionalized.}\label{fig:convergence}
-  \vspace{-6pt}
 \end{figure}
 
-\begin{table*}[ht]
-% \renewcommand{\arraystretch}{1.2}    % more space between rows
-  \hspace*{-1in}
-  \input{figure/sample_link_analysis_coarse_table.tex}
-  \caption{Our solution, i.e., our reconstruction of the ``intended'' direction for each link type in the English Link Grammar.
-We also indicate the extent to which each of these link types (1) has a single dominant direction, (2) participates in multi-headed constructions, and (3) corresponds to CoNLL links of a predictable direction and type.
-For space reasons, we show only the start of this table---the full table can be found in Appendix~\ref{app:linktypes} of the supplementary material.}\label{tab:linktypes}
-\end{table*}
+}
 
-We solved our ILP problem using the SCIP Optimization Suite \cite{achterberg2009scip}, encoding it using the ZIMPL language \cite{Koch2004}.
-Our largest run took 1.5 hours.
-\noteJH{A run of 10,000 sentences takes 3142 seconds (update in final version!)}  On English, only \input{figure/allowed_labels_both.tex}of \input{figure/allowed_labels_total.tex}link {\em types} allowed both directions, and only $\sum_{ij} s_{ij} = \input{figure/label_token_disallowed}$ of \input{figure/label_token_total.tex}link {\em tokens} required a disallowed direction via slack.
-\input{figure/conll_analysis_sentences_multiheaded.tex}of the English sentences but (alas) only \input{figure/russian_multiheaded.tex}of the Russian ones had at least one multi-headed word.
-See Table~\ref{tab:linktypes} and
-Appendix~\ref{app:linktypes} for detailed results.
-%\noteJE{2 significant digits is enough}
-%\noteJE{update these numbers}
+\section{ILP}
+\frame{\frametitle{Integer Linear Programming}
 
-\subsection{Stability of Results}\label{sec:stability}
+\NoteJH{Will the audience know about ILP?}
 
-We worried that the direction mapping might be unstable and sensitive to the input corpus.
-Happily, Figure~\ref{fig:convergence} shows otherwise (for both English and Russian).
-\noteJE{IMPORTANT!
-current graph is with fine-grained tags; rerun with the coarse-grained tags as in the rest of the paper} Using even a small prefix of the data got very high-precision results, in the sense that nearly all $r_L$ or $\ell_L$ variables that were 1 in this lightly trained mapping were also 1 in our largest run.
-The only disadvantage to using small data is low recall relative to the large run---many of the labels $L$ are not observed yet and so we do not yet allow either direction ($r_L=\ell_L=0$).
+}
 
-We used only coarse link tags as our labels, keeping only the capital letters of a tag (and merging all \texttt{ID} tags).
-This is because other characters in a tag indicate fine-grained features such as plurality that generally do not affect link direction.
-However, when we tried using fine tags as our labels instead, we found that all refinements of the same coarse tag would almost always spontaneously agree on their preferred direction.
-This indicates that there is indeed a ``natural'' direction for the coarse tag and that we can find it.
+\section{ILP Model}
+\frame{\frametitle{Integer Linear Programming Model}
 
-% OLD
-% We report on the convergence of the results of our ILP with increasing data.
-With increasing trials of sentences, we measure how similar the subsequent direction mappings would be with each other.
-Taking the direction table of the largest run as the answer key, we compared how much the solutions of the smaller runs deviated from it.
-% We recorded the precision of whether the mappings in the smaller runs could be found in those of the largest run, and similarly the recall of whether the mappings in the largest run could be found in the smaller runs.
+\NoteJH{Take from the paper.}
 
-% Because the precision values were high, we found that the solutions of the previous runs were consistent with the largest run, and as the data increased so did the recall values, and so we would continue to encounter novel link types and then incorporate them.
+}
 
-\subsection{Linguistic Analysis}
+\section{Experiments and Results}
+\frame{\frametitle{Experiments and Results}
 
-The resulting English corpus uses a syntactic annotation scheme that is somewhat different from the CoNLL annotations.  Differences are tabulated in
-Appendix~\ref{app:linktypes} of the supplementary material, while the actual parses are contrasted in Appendix~\ref{app:parses}.
-Fragments of these appendices are shown in Table~\ref{tab:linktypes} and Figure~\ref{fig:parses}.
-% By permission of the program chairs, these appendices are available at the anonymous URL \url{http://tidyurl.com/LinksToDAG_supplement}.
-
-The link grammar results in multi-headed treatments of infinitivals, compound determiners, relative clauses, and embedding.
-The other annotation differences are generally reasonable, e.g., letting {\em 's} be the head of a possessive, and different handling of punctuation and lists.
-Of course, a few of the differences are due to parser attachment errors.
-
-The main vexation is the handling of subject-verb links.
-Under the English link grammar, the verb (or 0) that governs a clause will link to both the clause's subject and its last (main) verb.
-This would permit our desired treatment of {\em ``Jill persuaded him to skip''}, in which {\em ``him''} has two parents.
-But the ILP solution generally treats subjects as {\em parents} of verbs (thus we get {\em him} $\rightarrow$ {\em to}).
-The reason for this is an inconsistency in the link grammar itself.\footnote{Specifically, the link grammar is inconsistent about how it handles clauses with missing subjects ({\em ``Jill wanted to skip''}).
-In this case, the governing verb links to the clause's first (tensed) verb in lieu of its subject.
-The problem is that it no longer also links to the clause's last (main) verb as it would if the subject were present.
-Hence our method concludes that a VP is always headed by its first verb ({\em to} $\rightarrow$ {\em skip}).
-That is a respectable convention on its own, but recall that unfortunately, the governing verb does {\em not} link to this first verb when the subject is present.
-As a result, the only remaining possible parent for the first verb is the subject ({\em him} $\rightarrow$ {\em to} $\rightarrow$ {\em skip}).
-This leads to the odd solution where subjects are treated as parents of verbs in general.
-
-Note that subjects are not {\em always} parents, due to another inconsistency in the link grammar.
-In the construction {\em ``It was impossible, they said''}, the subject ({\em ``they''}) is inconsistently not linked to 0 but only to the verb ({\em ``said''}) and so must be the verb's child, no other parent being available.}
-Fixing the link grammar would presumably correct the link direction.
-As noted in section~\ref{sec:motivation}, it is arguably a positive result that our method was able to detect a problem in the grammar engineering.
-
-% Multiheaded:
-% IV (890/891) - multiheaded treatment of infinitivals
-% I (1783/2969) - multiheaded treatment of infinitivals
-% B (795/842) - leftward extraction from relative clauses
-% CV (1588/1799) - multiheaded treatment of subordination
-% AL (39/40) - compound determiners
-% PP (448/689) - subject attaches to aux as well as main verb
-% HA (3/3) -
-% JQ (2/2) - 
-% SF (85/111) - pleonastic subject
-% AF (6/12) - 
-% C (244/1813) - 
-% OF (155/291)
-% P (651/2060) - 
-% Q (12/30)- 
-% RS (300/305) - 
-% S (4256/7434) -
-% SX (4/6) - 
-% W (2321/4721) - 
-% WV (1800/4138) -
-
-\NoteJE{what do we say about Russian?  ask Ryan for final version?}
-
-% DOESN'T SEEM NECESSARY FOR THE LANGUAGES WE TRIED
-% \section{Future Work}
-% \subsection{Slack Hierarchy}
-% We would like to introduce slack on the link grammar types such that link types with the same coarse grained label would try and align the same way as the majority in the group, where the preceding capital letters of the link type denote the coarse grained label, while the subscript letters denote further information.
-This slack would place a prior on rare or never-before seen link types to be assigned in the same way as other similar link types.
-We think that this will give the ILP better flexibility to handle noise and novel link types while still trying to learn the overall link grammar.
+}
 
 \section{Conclusions}
+\frame{\frametitle{Conclusions}
 
-We have presented an automatic ILP-based method to ``orient'' link grammar parses in multiple languages, turning them into rooted connected DAGs.
-This improves their linguistic interpretability and provides new corpora for experimenting with multi-headed dependency parsers.
-
-ILP may {\em in general} be a valuable technology for enriching existing annotated corpora.
-For example, the Penn Treebank project \cite{TREEBANK-1993} deliberately omitted types of annotations that plausibly could be added automatically.
-ILP can help by leveraging unsupervised corpus-wide information \cite{ravi2009}, enforcing annotations that are simultaneously well-formed per sentence and consistent across sentences.
-
-% \subsection{Sample of English Parses}
-% 
-% \begin{figure*}[ht!]
-%   \vspace{-6pt}
-%   \centering
-%   \input{figure/sample_parses.tikz}
-%   \vspace{-6pt}
-% \end{figure*}
-
-%\twocolumn
-
-%\newpage
-
-\begin{bibliography}{LinksToDAG_final}
-
-%\bibliographystyle{acl2014}
-
-%\bibliography{LinksToDAG}
-\end{bibliography}
-
-\clearpage
-\appendix
-
-% OLD
-% \section{ILP Formulation}
-% We present our ILP formulation. 
-% 
-% \begin{figure}[!htb]
-%   \small
-%   \begin{align*}
-%     \text{Sets:}&&\\
-%     &\textsc{link}_{\textsc{sentence},\textsc{label}, i,j} \in \textsc{links} &\\
-%     &\textsc{label} \in \textsc{labels}&\\ 
-%     &\textsc{dir} = \{\textsc{left}, \textsc{right}\} &\\
-%     \text{Params:}&&\\
-%     &\textsc{m}_\textsc{length} = \max (\forall{\textsc{link}_{\textsc{sentence}, i,j}} \max(i,j)) &\\
-%     %&\textsc{count}_{\textsc{label}, \textsc{dir}} = \sum_{\textsc{link}_{\textsc{label},\textsc{dir}}}1 &\\
-%     &\textsc{cost}_{\textsc{label}} = \frac{100}{ \sum_{\textsc{link}_{\textsc{label},\textsc{dir}}}1 } &\\
-%     \text{Variables:}&&\\
-%     &\textsc{slack}_{\textsc{link}} \geq 0 &\\ 
-%     &\textsc{depth}_{\textsc{sentence},i} \geq 0 &\\
-%     &\textsc{allowed}_{\textsc{label},\textsc{dir}} = \{0,1\}&\\
-%     &\textsc{arc}_{\textsc{link}} = \{0,1\} &\\
-%     \text{minimize:}& &\\ 
-%     &\sum_{\textsc{label}, \textsc{dir}} (\textsc{allowed}_{\textsc{label},\textsc{dir}}) + \textsc{cost}_{\textsc{label}} \cdot \sum_\textsc{link} \textsc{slack}_{\textsc{link}} &\\
-%     &&\\
-%     \text{subject to: }& &\\
-%     &\text{Links go in the allowed directions, except for slack}&\\
-%     &\forall{\textsc{link}_{i,j,\textsc{label}}}: 1-\textsc{arc}_{\textsc{link}} \leq \textsc{allowed}_{\textsc{label}, \textsc{left}} + \textsc{slack}_{\textsc{link}} &\\
-%     &\forall{\textsc{link}_{i,j,\textsc{label}}}: \textsc{arc}_{\textsc{link}} \leq \textsc{allowed}_{\textsc{label}, \textsc{right}} + \textsc{slack}_{\textsc{link}} &\\
-%     &\text{Depth constraints for acyclicity}&\\
-%     &\forall{\textsc{link}_{\textsc{sentence},i}}: \textsc{depth}_{\textsc{sentence},\text{root}} \leq \textsc{depth}_{\textsc{sentence},i} &\\
-%     &\forall{\textsc{link}_{\textsc{sentence},i,j}}: \textsc{depth}_{\textsc{sentence},i} + \textsc{m}_\textsc{length}\cdot\textsc{arc}_{\textsc{sentence},i,j} \geq \textsc{depth}_{\textsc{sentence},j} + 1 &\\
-%     &\forall{\textsc{link}_{\textsc{sentence},i,j}}: \textsc{depth}_{\textsc{sentence},j} + \textsc{m}_\textsc{length}\cdot(1-\textsc{arc}_{\textsc{sentence},i,j}) \geq \textsc{depth}_{\textsc{sentence},i} + 1 &\\
-%     &\text{Every token has a parent}&\\
-%     &\forall{\textsc{link}_{\textsc{sentence},i \neq \textsc{root}}}: \sum_{\textsc{link}_{\textsc{sentence},i,j}} (1-\textsc{arc}_{\textsc{sentence},i,j}) + \sum_{\textsc{link}_{\textsc{sentence},j,i}}(\textsc{arc}_{\textsc{sentence},j,i}) \geq 1&\\
-%   \end{align*}
-%   \caption{\small The ILP formulation.}
-% \end{figure}
-% 
-% 
-% \clearpage
-
-\section{English Link Types}\label{app:linktypes}
-
-Recall that \url{http://www.abisource.com/projects/link-grammar/dict/summarize-links.html} documents the 122 link types in the English link grammar lexicon.
-Below is the result of our largest English run, whose corpus uses \input{figure/allowed_labels_total.tex}of these types.
-\noteJH{automatically generate this number correctly: 107}
-
-Other researchers could postprocess link grammar parses by using the first two columns.  These indicate whether a given link label $L$ tended to be oriented rightward or leftward by our ILP.
-
-The labels that do not have a consistent direction are mainly those that connect a conjunction to its conjuncts.\footnote{See \url{http://www.abisource.com/projects/link-grammar/dict/coordination.html}.}
-Our coarsening of the labels (see section~\ref{sec:stability}) discarded the characters that distinguish left from right conjuncts.
-
-The third column indicates where multi-headedness happens, by showing how often the child of an $L$ link had additional parents as well.
-
-The remaining columns compare to the CoNLL parses.
-They show the fraction of $L$-labeled link tokens that were also present in the CoNLL reference parse---and on that fraction, whether the orientation matches CoNLL's and what the dominant CoNLL label was.
-
-%\NoteJE{improve table formatting.  Numeric justification; en dashes; longtable; remove most hlines.}
-
-%\begin{table*}[ht]
-%\renewcommand{\arraystretch}{1.2}    % more space between rows
-%\hspace*{-1in}
-\input{figure/link_analysis_coarse_table.tex}
-%\end{table*}
-
-\section{Sample English Parses}\label{app:parses}
-
-\NoteJE{show the fine-grained link labels in the parses, even though the ILP deals with coarse-grained.  Juneki tried already to hack this in but was getting double links.}
-
-Below we display the first 100 English sentences whose widths do not exceed the width of this page.
-The blue upper half shows the CoNLL tagging and dependency parse.
-The red lower half shows the link grammar's tagging and our orientation of the link parse.
-% The dashes in the link grammar part of speech tag annotations mean that the link parser did not return one.
-A vertical edge to word token $v$ denotes an edge $0 \rightarrow v$.
-Words that were unknown to the link parser are flagged with \texttt{[!]}.
-
-Edges are shown as solid lines if they appear in both parses.  They are drawn more thickly if the directions do not match.
-
-Edges are shown as dotted lines if they appear only in one parse and not in the other.
-
-Edges are highlighted in orange if the child has multiple parents.
-
-% OLD NERDVIEW EXPLANATION
-% For every arc:
-% \begin{enumerate}
-% \item If there is a corresponding link between the same nodes then a solid blue line is drawn.
-% \item Or else a dotted blue line is drawn. 
-% \end{enumerate}
-% The bottom red arcs represent the oriented link grammar parses. For every link edge:
-% \begin{enumerate}
-% \item If there is a corresponding CoNLL arc between the same nodes then a solid red line is drawn.
-% \item If the matching arc is facing the opposite direction, then the link edge is drawn thicker.
-% \item Of the remaining edges, if an edge is a multiheaded link, then it is drawn with thick orange.
-% \item The rest of the remaining edges are drawn with dotted red lines.
-% \end{enumerate}
-
-% Uncomment this to add in all of the parses. Warning: compilation will take a long time.
-%\input{figure/parses.tikz}
+}
 

### 2014-12-06 doc/slides/hong+eisner.TLT13.slides.tex
updated parse visualization explanations

@@
 \begin{figure}
-\begin{subfigure}[b]{0.4850\textwidth}
+  \begin{dependency}[edge style={-}]
+	\begin{deptext}
+	  the \& matter \& may \& never \& even \& be \& tried \& in \& court \& . \\
+	  - \& n-u \& v \& e \& e \& v \& v-d \& r \& n-u \& - \\
+	\end{deptext}
+	\depedge[edge above, edge style = {red}]{2}{3}{S}
+	\depedge[edge above, edge style = {red}]{3}{6}{I}
+	\depedge[edge above, edge style = {red}]{2}{1}{D}
+	\depedge[edge above, edge style = {red}]{7}{8}{MV}
+	\depedge[edge above, edge style = {red}]{6}{7}{P}
+	\depedge[edge above, edge style = {red}]{8}{9}{J}
+	\deproot[edge above, edge style = {red}]{2}{W}
+	\deproot[edge above, edge style = {red}]{7}{WV}
+	\deproot[edge above, edge style = {red}]{10}{X}
+	\depedge[edge above, edge style = {red}]{6}{4}{E}
+	\depedge[edge above, edge style = {red}]{6}{5}{E}
+  \end{dependency}
+  \caption{Link Parse of a sentence from Penn Tree Bank}
+\end{figure}
+
+}
+
+\frame{\frametitle{Link Grammars}
+\begin{figure}
   \begin{dependency}
@@
   \end{dependency}
-\end{subfigure}
 
+  \caption{Directionalize the edges}
 \end{figure}
-\NoteJH{figure not centered}
-\NoteJH{Make the edges undirected}
+}
+
+\frame{\frametitle{Link Grammars}
+\begin{figure}
+  \begin{dependency}
+	\begin{deptext}
+	  {\scriptsize DT} \& {\scriptsize NN} \& {\scriptsize MD} \& {\scriptsize RB} \& {\scriptsize RB} \& {\scriptsize VB} \& {\scriptsize VB} \& {\scriptsize IN} \& {\scriptsize NN} \& {\scriptsize .} \\
+	  the \& matter \& may \& never \& even \& be \& tried \& in \& court \& . \\
+	  - \& n-u \& v \& e \& e \& v \& v-d \& r \& n-u \& - \\
+	\end{deptext}
+	\deproot[edge above, edge style = {blue}]{3}{\small ROOT}
+	\depedge[edge below, edge style = {red}]{2}{3}{S}
+	\depedge[edge above, edge style = {blue}]{3}{2}{\small SBJ}
+	\depedge[edge above, edge style = {blue}]{3}{4}{\small ADV}
+	\depedge[edge above, edge style = {blue}]{3}{5}{\small ADV}
+	\depedge[edge below, edge style = {red}]{3}{6}{I}
+	\depedge[edge above, edge style = {blue}]{3}{6}{\small VC}
+	\depedge[edge above, edge style = {blue}, edge unit distance =1.5ex]{3}{10}{\small P}
+	\depedge[edge below, edge style = {red}]{2}{1}{D}
+	\depedge[edge above, edge style = {blue}]{2}{1}{\small NMOD}
+	\depedge[edge below, edge style = {red}]{7}{8}{MV}
+	\depedge[edge above, edge style = {blue}]{7}{8}{\small ADV}
+	\depedge[edge below, edge style = {red}]{6}{7}{P}
+	\depedge[edge above, edge style = {blue}]{6}{7}{\small VC}
+	\depedge[edge below, edge style = {red}]{8}{9}{J}
+	\depedge[edge above, edge style = {blue}]{8}{9}{\small PMOD}
+	\deproot[edge below, edge style = {red}]{2}{W}
+	\deproot[edge below, edge style = {red}]{7}{WV}
+	\deproot[edge below, edge style = {red}]{10}{X}
+	\depedge[edge below, edge style = {red}]{6}{4}{E}
+	\depedge[edge below, edge style = {red}]{6}{5}{E}
+  \end{dependency}
+
+  \caption{Compare resulting dependency parse with Penn Treebank annotations}
+\end{figure}
 
@@
 \begin{figure}
-\begin{subfigure}[b]{0.4850\textwidth}
   \begin{dependency}
@@
   \end{dependency}
-\end{subfigure}
 
+  \caption{Compare resulting dependency parse with Penn Treebank annotations}
 \end{figure}

### 2014-12-07 doc/slides/hong+eisner.TLT13.slides.tex
added multiheaded slides. Still need to do Motivation, ILP, Experiments, Conclusions.

--- 2014-12-06 doc/slides/hong+eisner.TLT13.slides.tex
@@
 \frame{\frametitle{Motivation}
-
+  \begin{itemize}
+    \item We need a multiheaded dependency corpus for parsing
+  \end{itemize}
 }
@@
 \section{Multi-headed}
-\frame{\frametitle{Multi-headed Dependency Parsing}
+
+\frame{\frametitle{Multi-headedness}
+  \begin{itemize}
+    \item Control
+    \item Relativization
+    \item Conjunction
+  \end{itemize}
+}
+
+\subsection{Control}
+\frame{\frametitle{Control}
+  \begin{figure}
+    \begin{dependency}
+      \begin{deptext}
+        Jill \& likes \& to \& skip \\
+      \end{deptext}
+      \depedge[edge above, thick, hide label]{2}{1}{}
+      \depedge[edge below, ultra thick, hide label, edge style = {purple}, edge unit distance = 1.5ex]{4}{1}{}
+      \depedge[edge above, thick, hide label]{2}{3}{}
+      \depedge[edge above, thick, hide label]{3}{4}{}
+    \end{dependency}
+    \caption*{Jill is the subject of two verbs}
+  \end{figure}
 
   \begin{figure}
-  \centering
-  \includegraphics[width=\linewidth]{multiheaded}
-  \caption{English is not completely multiheaded}
+    \begin{dependency}
+      \begin{deptext}
+        Jill \& persuaded \& Jack \& to \& skip \\
+      \end{deptext}
+      \depedge[edge above, thick, hide label]{2}{1}{}
+      \depedge[edge above, thick, hide label]{2}{3}{}
+      \depedge[edge above, thick, hide label]{3}{4}{}
+      \depedge[edge above, thick, hide label]{4}{5}{}
+      \depedge[edge below, ultra thick, hide label, edge style = {purple}, edge unit distance = 1.5ex]{3}{5}{}
+    \end{dependency}
+    \caption*{Jack is the object of one verb and the subject of another}
+  \end{figure}
+}
+
+\subsection{Relativization}
+\frame{\frametitle{Relativization}
+  \begin{figure}
+    \begin{dependency}
+      \begin{deptext}
+        The \& boy \& that \& Jill \& skipped \& with \& fell \& down \\
+      \end{deptext}
+      \depedge[edge above, thick, hide label]{2}{1}{}
+      \depedge[edge above, thick, hide label]{2}{3}{}
+      \depedge[edge above, thick, hide label]{3}{4}{}
+      \depedge[edge above, thick, hide label]{4}{5}{}
+      \depedge[edge above, thick, hide label]{5}{6}{}
+      \depedge[edge above, thick, hide label]{2}{7}{}
+      \depedge[edge above, thick, hide label]{7}{8}{}
+      \depedge[edge below, ultra thick, hide label, edge style = {purple}, edge unit distance = 1.5ex]{6}{2}{}
+    \end{dependency}
+    \caption*{The boy is the object of ``with'' as well as the subject of ``fell''.}
+  \end{figure}
+}
+
+\subsection{Conjunction}
+\frame{\frametitle{Conjunction}
+  \begin{figure}
+    \begin{dependency}
+      \begin{deptext}
+        Jack \& and \& Jill \& went \& up \& the \& hill \\
+      \end{deptext}
+      \depedge[edge above, thick, hide label]{2}{1}{}
+      \depedge[edge above, thick, hide label]{2}{3}{}
+      \depedge[edge above, thick, hide label]{4}{2}{}
+      \depedge[edge above, thick, hide label]{4}{5}{}
+      \depedge[edge above, thick, hide label]{5}{7}{}
+      \depedge[edge above, thick, hide label]{7}{6}{}
+
+      \depedge[edge below, ultra thick, hide label, edge style = {purple}, edge unit distance = 1.5ex]{4}{1}{}
+      \depedge[edge below, ultra thick, hide label, edge style = {purple}, edge unit distance = 1.5ex]{4}{3}{}
+
+    \end{dependency}
+    \caption*{Jack and Jill serve as the two arguments to and, but they are also semantically subjects of went.}
+  \end{figure}
+}
+
+%\frame{\frametitle{Multi-headedness}
+%  \begin{figure}
+%    \begin{dependency}
+%      \begin{deptext}
+%      I \& think \& the \& man \& who \& you \& saw \& wants \& to \& eat \& sushi \\
+%      \end{deptext}
+%      \deproot[edge above, thick]{2}{ROOT}
+%      \depedge[edge above, thick, hide label]{2}{1}{} % think -> I
+%      \depedge[edge above, thick, hide label]{4}{3}{} % man -> the
+%      \depedge[edge above, thick, hide label]{4}{5}{} % man -> who
+%      \depedge[edge above, thick, hide label]{7}{6}{} % saw -> you
+%      \depedge[edge above, thick, hide label]{8}{4}{} % wants -> man
+%      \depedge[edge above, thick, hide label]{4}{7}{} % man -> saw
+%      \depedge[edge above, thick, hide label]{2}{8}{} % think -> wants
+%      \depedge[edge above, thick, hide label]{8}{10}{} % wants -> eat
+%      \depedge[edge above, thick, hide label]{10}{9}{} % eat -> to
+%      \depedge[edge above, thick, hide label]{10}{11}{} % eat -> sushi
+%      \depedge[edge below, edge style = {purple}, hide label, edge unit distance = 1.5ex, ultra thick]{10}{4}{} % eat -> man
+%    \end{dependency}
+%    \caption*{Example of a Multiheaded Parse}
+%  \end{figure}
+%}
+
+% Shouldn't mention non-projectivity. 
+%\frame{\frametitle{Non-projectivity}
+%\begin{figure}
+%  \includegraphics[width=\linewidth]{multiheaded}
+%  \caption*{English is not a completely projective language}
+%  \NoteJH{image taken from Chris Dyer's slide: http://demo.clab.cs.cmu.edu/fa2014-11711/images/0/0e/Depparsing.pdf}
+%  \NoteJH{I should also ask Chris for permission to use this}
+%\end{figure}
+%}
+
+\section{Non Projectivity}
+\frame{\frametitle{We do not address: Non-Projectivity}
+\begin{itemize}
+  \item Projectivity assumptions allow us to have efficient parsing algorithms that utilize dynamic programming
+  \item English is mostly projective
+\end{itemize}
+}
+
+\frame{\frametitle{We do not address: Non-Projectivity}
+\begin{itemize}
+  \item Projectivity assumptions allow us to have efficient parsing algorithms that utilize dynamic programming
+  \item English is mostly projective
+\end{itemize}
+
+\begin{figure}
+  \includegraphics[width=0.7\linewidth]{multiheaded}
+  \caption*{But not always}
   \NoteJH{image taken from Chris Dyer's slide: http://demo.clab.cs.cmu.edu/fa2014-11711/images/0/0e/Depparsing.pdf}
+  \NoteJH{I should also ask Chris for permission to use this}
 \end{figure}
@@
 \section{Link Grammars}
+
 \frame{\frametitle{Link Grammars}
+  \begin{enumerate}
+    \item Grammar-based formalism for projective dependency parsing with undirected links
+  \end{enumerate}
+}
 
+\frame{\frametitle{Link Grammars}
 \begin{figure}
-\begin{subfigure}[b]{0.4850\textwidth}
+  \begin{dependency}[edge style={-}]
+	\begin{deptext}
+	  the \& matter \& may \& never \& even \& be \& tried \& in \& court \& . \\
+	  - \& n-u \& v \& e \& e \& v \& v-d \& r \& n-u \& - \\
+	\end{deptext}
+	\depedge[edge above, thick, edge style = {red}]{2}{3}{S}
+	\depedge[edge above, thick, edge style = {red}]{3}{6}{I}
+	\depedge[edge above, thick, edge style = {red}]{2}{1}{D}
+	\depedge[edge above, thick, edge style = {red}]{7}{8}{MV}
+	\depedge[edge above, thick, edge style = {red}]{6}{7}{P}
+	\depedge[edge above, thick, edge style = {red}]{8}{9}{J}
+	\deproot[edge above, thick, edge style = {red}]{2}{W}
+	\deproot[edge above, thick, edge style = {red}]{7}{WV}
+	\deproot[edge above, thick, edge style = {red}]{10}{X}
+	\depedge[edge above, thick, edge style = {red}]{6}{4}{E}
+	\depedge[edge above, thick, edge style = {red}]{6}{5}{E}
+  \end{dependency}
+  \caption*{Link Parse of a sentence from Penn Tree Bank}
+\end{figure}
+
+}
+
+\frame{\frametitle{Link Grammars}
+\begin{figure}
   \begin{dependency}
@@
 	\end{deptext}
-		\depedge[edge above, edge style = {red}]{2}{3}{S}
-		\depedge[edge above, edge style = {red}]{3}{6}{I}
-		\depedge[edge above, edge style = {red}]{2}{1}{D}
-		\depedge[edge above, edge style = {red}]{7}{8}{MV}
-		\depedge[edge above, edge style = {red}]{6}{7}{P}
-		\depedge[edge above, edge style = {red}]{8}{9}{J}
-		\deproot[edge above, edge style = {red}]{2}{W}
-		\deproot[edge above, edge style = {red}]{7}{WV}
-		\deproot[edge above, edge style = {red}]{10}{X}
-		\depedge[edge above, edge style = {red}]{6}{4}{E}
-		\depedge[edge above, edge style = {red}]{6}{5}{E}
+	\depedge[edge above, thick, edge style = {red}]{2}{3}{S}
+	\depedge[edge above, thick, edge style = {red}]{3}{6}{I}
+	\depedge[edge above, thick, edge style = {red}]{2}{1}{D}
+	\depedge[edge above, thick, edge style = {red}]{7}{8}{MV}
+	\depedge[edge above, thick, edge style = {red}]{6}{7}{P}
+	\depedge[edge above, thick, edge style = {red}]{8}{9}{J}
+	\deproot[edge above, thick, edge style = {red}]{2}{W}
+	\deproot[edge above, thick, edge style = {red}]{7}{WV}
+	\deproot[edge above, thick, edge style = {red}]{10}{X}
+	\depedge[edge above, thick, edge style = {red}]{6}{4}{E}
+	\depedge[edge above, thick, edge style = {red}]{6}{5}{E}
   \end{dependency}
-\end{subfigure}
 
+  \caption*{Directionalize the edges}
 \end{figure}
-\NoteJH{figure not centered}
-\NoteJH{Make the edges undirected}
+}
+
+\frame{\frametitle{Link Grammars}
+\begin{figure}
+  \begin{dependency}
+	\begin{deptext}
+	  {\scriptsize DT} \& {\scriptsize NN} \& {\scriptsize MD} \& {\scriptsize RB} \& {\scriptsize RB} \& {\scriptsize VB} \& {\scriptsize VB} \& {\scriptsize IN} \& {\scriptsize NN} \& {\scriptsize .} \\
+	  the \& matter \& may \& never \& even \& be \& tried \& in \& court \& . \\
+	  - \& n-u \& v \& e \& e \& v \& v-d \& r \& n-u \& - \\
+	\end{deptext}
+	\deproot[edge above, thick, edge style = {blue}]{3}{\small ROOT}
+	\depedge[edge below, thick, edge style = {red}]{2}{3}{S}
+	\depedge[edge above, thick, edge style = {blue}]{3}{2}{\small SBJ}
+	\depedge[edge above, thick, edge style = {blue}]{3}{4}{\small ADV}
+	\depedge[edge above, thick, edge style = {blue}]{3}{5}{\small ADV}
+	\depedge[edge below, thick, edge style = {red}]{3}{6}{I}
+	\depedge[edge above, thick, edge style = {blue}]{3}{6}{\small VC}
+	\depedge[edge above, thick, edge style = {blue}, edge unit distance =1.5ex]{3}{10}{\small P}
+	\depedge[edge below, thick, edge style = {red}]{2}{1}{D}
+	\depedge[edge above, thick, edge style = {blue}]{2}{1}{\small NMOD}
+	\depedge[edge below, thick, edge style = {red}]{7}{8}{MV}
+	\depedge[edge above, thick, edge style = {blue}]{7}{8}{\small ADV}
+	\depedge[edge below, thick, edge style = {red}]{6}{7}{P}
+	\depedge[edge above, thick, edge style = {blue}]{6}{7}{\small VC}
+	\depedge[edge below, thick, edge style = {red}]{8}{9}{J}
+	\depedge[edge above, thick, edge style = {blue}]{8}{9}{\small PMOD}
+	\deproot[edge below, thick, edge style = {red}]{2}{W}
+	\deproot[edge below, thick, edge style = {red}]{7}{WV}
+	\deproot[edge below, thick, edge style = {red}]{10}{X}
+	\depedge[edge below, thick, edge style = {red}]{6}{4}{E}
+	\depedge[edge below, thick, edge style = {red}]{6}{5}{E}
+  \end{dependency}
+
+  \caption*{Compare resulting dependency parse with Penn Treebank annotations}
+\end{figure}
 
@@
   \end{dependency}
-\end{subfigure}
 
+  \caption*{Compare resulting dependency parse with Penn Treebank annotations}
 \end{figure}
@@
 
+%\nocite{*}
+%\bibliographystyle{plain}
+%\bibliography{hong+eisner.TLT13.slides.bib}
+%\appendix
+
 \end{document}

### 2014-12-07 doc/slides/hong+eisner.TLT13.slides.tex
comment
added more slides regarding the ILP, Experiments, and Conclusions section. ILP section is incomplete. Motivation section also needs work.

@@
 \end{itemize}
+
+\NoteJH{Should I mention non-projectivity? This slide is mostly here because of a reviewer comment.}
 }
@@
 \frame{\frametitle{Link Grammars}
-  \begin{enumerate}
+  \begin{itemize}
     \item Grammar-based formalism for projective dependency parsing with undirected links
-  \end{enumerate}
+    \item We set out to discover if the undirected links are actually directed relationships.
+  \end{itemize}
 }
 
-\frame{\frametitle{Link Grammars}
+\frame{\frametitle{Link Grammars: Example Parse}
 \begin{figure}
@@
 
-\frame{\frametitle{Link Grammars}
+\frame{\frametitle{Link Grammars: Converting into a Directed Acyclic Graph}
 \begin{figure}
@@
 
-  \caption*{Compare resulting dependency parse with Penn Treebank annotations}
+  \caption*{Compare resulting dependency parse with CoNLL 2007 shared task}
 \end{figure}
@@
 
-\section{ILP}
-\frame{\frametitle{Integer Linear Programming}
+%\section{ILP}
+%\frame{\frametitle{Integer Linear Programming}
 
-\NoteJH{Will the audience know about ILP?}
+%\NoteJH{Will the audience know about ILP?}
+
+%}
+
+\section{ILP Model}
+\frame{\frametitle{Integer Linear Programming Model: Variables}
+
+For each sentence, for each edge $i,j$ where $i < j$, we introduce variables $x_{ij}$ and $x_{ji}$.
+
+\NoteJH{TODO: Incomplete slide. Need to describe things like depth of the variable, label}
+}
+
+\frame{\frametitle{Integer Linear Programming Model: Encoded Constraints}
+
+\begin{itemize}
+  \item Connectedness
+  \item Acyclicity
+  \item Consistency of Directionalized Links
+\end{itemize}
+
+\NoteJH{Take from the paper.}
+}
+
+\subsection{Connectedness, Acyclicity}
+\frame{\frametitle{Connectedness, Acyclicity}
+\begin{itemize}
+  \item[] Connectedness
+    \begin{align}
+      \sum_u x_{uv} & \geq 1
+    \end{align}
+
+  \item[] Acyclic
+    \begin{align}
+      (\forall_u)\; d_v + (1 + n_v) \cdot (1 - x_{uv}) & \geq 1+d_u
+    \end{align}
+\end{itemize}
 
@@
 
-\section{ILP Model}
-\frame{\frametitle{Integer Linear Programming Model}
+\frame{\frametitle{Consistence of Directionalized Links}
+\begin{itemize}
+  \item[] Consistency of Directionalized Links
+    \begin{align}\label{direction+slack}
+      x_{ij} &\leq r_L + s_{ij} &
+      x_{ji} &\leq \ell_L + s_{ij}
+    \end{align}
 
-\NoteJH{Take from the paper.}
+    \begin{align}\label{eqn:obj}
+      \min \left( \sum_L r_L + \ell_L \right)
+    \end{align}
+
+    %$$x_{ij} \leq r_L + s_{ij} $$
+    %$$x_{ji} \leq \ell_L + s_{ij}$$
+    %$$\min \left( \sum_L r_L + \ell_L \right)$$
+
+\end{itemize}
+}
+
+\frame{\frametitle{Consistence of Directionalized Links with Slack}
+\begin{itemize}
+  \item[] Consistency of Directionalized Links
+    \begin{align}\label{direction+slack}
+      x_{ij} &\leq r_L + s_{ij} &
+      x_{ji} &\leq \ell_L + s_{ij}
+    \end{align}
+
+    \begin{align}\label{eqn:obj}
+      \min \left( \sum_L r_L + \ell_L \right) \cdot \frac{N_L}{4} + \sum_{ij}s_{ij}
+    \end{align}
+
+    %$$x_{ij} \leq r_L + s_{ij} $$
+    %$$x_{ji} \leq \ell_L + s_{ij}$$
+    %$$\min \left( \sum_L r_L + \ell_L \right) \cdot \frac{N_L}{4} + \sum_{ij}s_{ij}$$
+
+\end{itemize}
+}
+
+\section{Experiments and Results}
+%\frame{\frametitle{Experiments and Results}
+%\begin{itemize}
+%  \item 18,577 English sentences with gold CoNLL. 
+%  \item 18,577 unlabeled Russian sentences.
+%\end{itemize}
+%}
+
+\frame{\frametitle{Data Sets}
+\begin{itemize}
+  \item[] 18,577 English sentences with 10,960 connected parses
+  \item[] 18,577 unlabeled Russian sentences with 4,913 connected parses
+\end{itemize}
 
@@
 
-\section{Experiments and Results}
-\frame{\frametitle{Experiments and Results}
+\frame{\frametitle{On the English data set}
+\begin{itemize}
+\item[] Link Data has 8\% additional edges over the CoNLL.
+\item[] 52\% of links match CoNLL arcs
+\item[] 57\% of CoNLL arcs match links
+\end{itemize}
 
+\NoteJH{I skipped the ``Stability of Results'' section}
+}
+
+\frame{\frametitle{ILP Results}
+  \begin{itemize}
+    \item[] Link types that allowed both directions:
+      \begin{itemize}
+        \item[] $\frac{7}{113} = 6.19\%$ 
+      \end{itemize}
+
+    \item[] Link tokens that required disallowed direction via slack:
+      \begin{itemize}
+        \item[] $\frac{4043}{195000} = 2.07\%$ 
+      \end{itemize}
+    \item[]
+  \end{itemize}
+}
+
+\frame{\frametitle{ILP Results}
+  \begin{itemize}
+    \item[] Link types that allowed both directions:
+      \begin{itemize}
+        \item[] $\frac{7}{113} = 6.19\%$ 
+      \end{itemize}
+
+    \item[] Link tokens that required disallowed direction via slack:
+      \begin{itemize}
+        \item[] $\frac{4043}{195000} = 2.07\%$ 
+      \end{itemize}
+    \item[] The link labels mostly have a consistent direction.
+  \end{itemize}
+}
+
+\frame{\frametitle{Link Results}
+  \begin{itemize}
+    \item Subject-Verb links are backwards.
+  \end{itemize}
+
+\begin{figure}
+  \begin{dependency}
+	\begin{deptext}
+	  {\scriptsize DT} \& {\scriptsize NN} \& {\scriptsize MD} \& {\scriptsize RB} \& {\scriptsize RB} \& {\scriptsize VB} \& {\scriptsize VB} \& {\scriptsize IN} \& {\scriptsize NN} \& {\scriptsize .} \\
+	  the \& matter \& may \& never \& even \& be \& tried \& in \& court \& . \\
+	  - \& n-u \& v \& e \& e \& v \& v-d \& r \& n-u \& - \\
+	\end{deptext}
+	\deproot[edge above, edge style = {blue, dotted}]{3}{\small ROOT}
+	\depedge[edge below, edge style = {red, ultra thick}]{2}{3}{S}
+	\depedge[edge above, edge style = {blue, ultra thick}]{3}{2}{\small SBJ}
+	\depedge[edge above, edge style = {blue, dotted}]{3}{4}{\small ADV}
+	\depedge[edge above, edge style = {blue, dotted}]{3}{5}{\small ADV}
+	\depedge[edge below, edge style = {red, thick}]{3}{6}{I}
+	\depedge[edge above, edge style = {blue, thick}]{3}{6}{\small VC}
+	\depedge[edge above, edge style = {blue, dotted}, edge unit distance =1.5ex]{3}{10}{\small P}
+	\depedge[edge below, edge style = {red, thick}]{2}{1}{D}
+	\depedge[edge above, edge style = {blue, thick}]{2}{1}{\small NMOD}
+	\depedge[edge below, edge style = {red, thick}]{7}{8}{MV}
+	\depedge[edge above, edge style = {blue, thick}]{7}{8}{\small ADV}
+	\depedge[edge below, edge style = {orange, thick}]{6}{7}{P}
+	\depedge[edge above, edge style = {blue, thick}]{6}{7}{\small VC}
+	\depedge[edge below, edge style = {red, thick}]{8}{9}{J}
+	\depedge[edge above, edge style = {blue, thick}]{8}{9}{\small PMOD}
+	\deproot[edge below, edge style = {red, dotted}]{2}{W}
+	\deproot[edge below, edge style = {orange, ultra thick, dotted}]{7}{WV}
+	\deproot[edge below, edge style = {red, dotted}]{10}{X}
+	\depedge[edge below, edge style = {red, dotted}]{6}{4}{E}
+	\depedge[edge below, edge style = {red, dotted}]{6}{5}{E}
+  \end{dependency}
+\end{figure}
+
+\begin{itemize}
+\item[] %This is probably due to an inconsistency of the Link Grammar, discovered by our method.
+\end{itemize}
+
+}
+
+\frame{\frametitle{Link Results}
+  \begin{itemize}
+    \item Subject-Verb links are backwards.
+  \end{itemize}
+
+\begin{figure}
+  \begin{dependency}
+	\begin{deptext}
+	  {\scriptsize DT} \& {\scriptsize NN} \& {\scriptsize MD} \& {\scriptsize RB} \& {\scriptsize RB} \& {\scriptsize VB} \& {\scriptsize VB} \& {\scriptsize IN} \& {\scriptsize NN} \& {\scriptsize .} \\
+	  the \& matter \& may \& never \& even \& be \& tried \& in \& court \& . \\
+	  - \& n-u \& v \& e \& e \& v \& v-d \& r \& n-u \& - \\
+	\end{deptext}
+	%\deproot[edge above, edge style = {blue, dotted}]{3}{\small ROOT}
+	\depedge[edge below, edge style = {red, ultra thick}]{2}{3}{S}
+	\depedge[edge above, edge style = {blue, ultra thick}]{3}{2}{\small SBJ}
+	%\depedge[edge above, edge style = {blue, dotted}]{3}{4}{\small ADV}
+	%\depedge[edge above, edge style = {blue, dotted}]{3}{5}{\small ADV}
+	%\depedge[edge below, edge style = {red, thick}]{3}{6}{I}
+	%\depedge[edge above, edge style = {blue, thick}]{3}{6}{\small VC}
+	%\depedge[edge above, edge style = {blue, dotted}, edge unit distance =1.5ex]{3}{10}{\small P}
+	%\depedge[edge below, edge style = {red, thick}]{2}{1}{D}
+	%\depedge[edge above, edge style = {blue, thick}]{2}{1}{\small NMOD}
+	%\depedge[edge below, edge style = {red, thick}]{7}{8}{MV}
+	%\depedge[edge above, edge style = {blue, thick}]{7}{8}{\small ADV}
+	%\depedge[edge below, edge style = {orange, thick}]{6}{7}{P}
+	%\depedge[edge above, edge style = {blue, thick}]{6}{7}{\small VC}
+	%\depedge[edge below, edge style = {red, thick}]{8}{9}{J}
+	%\depedge[edge above, edge style = {blue, thick}]{8}{9}{\small PMOD}
+	%\deproot[edge below, edge style = {red, dotted}]{2}{W}
+	%\deproot[edge below, edge style = {orange, ultra thick, dotted}]{7}{WV}
+	%\deproot[edge below, edge style = {red, dotted}]{10}{X}
+	%\depedge[edge below, edge style = {red, dotted}]{6}{4}{E}
+	%\depedge[edge below, edge style = {red, dotted}]{6}{5}{E}
+  \end{dependency}
+\end{figure}
+
+\begin{itemize}
+\item This is due to an inconsistency of the Link Grammar, discovered by our method. \NoteJH{The reason why is in footnote 6 of page 10 in our paper. Reread after sleeping.}
+\end{itemize}
+
+\NoteJH{Should I include a section of the appendix table? Which rows should I include?}
 }
@@
 \frame{\frametitle{Conclusions}
+  \begin{itemize}
+  \item Link Grammar parses can be oriented into connected DAGs
+  \item New corpora for building multi-headed dependency parsers
+  \item ILP can be used to build or annotate corpora
+  \end{itemize}
+}
 
+\frame{\frametitle{}
+  \begin{itemize}
+  \item[] Questions?
+  \end{itemize}
 }

### 2014-12-07 doc/slides/hong+eisner.TLT13.slides.tex
updated the motivation. There is an ILP section still left TODO. But the rough draft of the presentation is mostly covered.

@@
   \begin{itemize}
-    \item We need a multiheaded dependency corpus for parsing
+    \item A multiheaded dependency corpus would be useful for building new parsers
+  \end{itemize}
+}
+
+\section{Motivation}
+\frame{\frametitle{Motivation}
+  \begin{itemize}
+    \item A multiheaded dependency corpus would be useful for testing new parsing algorithms
+    \item Such a corpus could be automatically annotated using Integer Linear Programming
+  \end{itemize}
+}
+
+\section{Motivation}
+\frame{\frametitle{Motivation}
+  \begin{itemize}
+    \item A multiheaded dependency corpus would be useful for testing new parsing algorithms
+    \item Such a corpus could be automatically annotated using Integer Linear Programming
+    \item We explored whether the Link Grammar could be adapted for this purpose.
+  \end{itemize}
+}
+
+\section{Motivation}
+\frame{\frametitle{Motivation}
+  \begin{itemize}
+    \item A multiheaded dependency corpus would be useful for testing new parsing algorithms
+    \item Such a corpus could be automatically annotated using Integer Linear Programming
+    \item We explored whether the Link Grammar could be adapted for this purpose.
+    \item The results of this are mixed, but provides a good case study.
   \end{itemize}

### 2014-12-07 doc/slides/hong+eisner.TLT13.slides.tex
fixed the slack in the ILP slides. Added roots to example parses.

@@
 
-\section{Motivation}
 \frame{\frametitle{Motivation}
@@
       \end{deptext}
+      \deproot[edge above, thick, hide label, edge unit distance = 1.5ex]{2}{}
       \depedge[edge above, thick, hide label]{2}{1}{}
-      \depedge[edge below, ultra thick, hide label, edge style = {purple}, edge unit distance = 1.5ex]{4}{1}{}
+      \depedge[edge below, ultra thick, hide label, edge style = {purple}, edge unit distance = 1.2ex]{4}{1}{}
       \depedge[edge above, thick, hide label]{2}{3}{}
@@
       \end{deptext}
+      \deproot[edge above, thick, hide label, edge unit distance = 1.5ex]{2}{}
       \depedge[edge above, thick, hide label]{2}{1}{}
@@
       \end{deptext}
+      \deproot[edge above, thick, hide label, edge unit distance = 1.5ex]{7}{}
       \depedge[edge above, thick, hide label]{2}{1}{}
       \depedge[edge above, thick, hide label]{2}{3}{}
-      \depedge[edge above, thick, hide label]{3}{4}{}
-      \depedge[edge above, thick, hide label]{4}{5}{}
+      \depedge[edge above, thick, hide label]{3}{5}{}
+      \depedge[edge above, thick, hide label]{5}{4}{}
       \depedge[edge above, thick, hide label]{5}{6}{}
-      \depedge[edge above, thick, hide label]{2}{7}{}
+      \depedge[edge above, thick, hide label, edge unit distance = 1.8ex]{7}{2}{}
       \depedge[edge above, thick, hide label]{7}{8}{}
@@
       \end{deptext}
+      \deproot[edge above, thick, hide label, edge unit distance = 1.5ex]{4}{}
       \depedge[edge above, thick, hide label]{2}{1}{}
@@
     \begin{align}\label{direction+slack}
-      x_{ij} &\leq r_L + s_{ij} &
-      x_{ji} &\leq \ell_L + s_{ij}
+      x_{ij} &\leq r_L &
+      x_{ji} &\leq \ell_L 
     \end{align}

### 2014-12-07 doc/slides/hong+eisner.TLT13.slides.tex
fixed typos

@@
   \begin{itemize}
-    \item A multiheaded dependency corpus would be useful for building new parsers
+    \item A multiheaded dependency corpus would be useful for testing new parsing algorithms
   \end{itemize}
@@
     \end{dependency}
-    \caption*{The boy is the object of ``with'' as well as the subject of ``fell''.}
+    \caption*{The boy is the object of \textit{with} as well as the subject of \textit{fell}.}
   \end{figure}
@@
     \end{dependency}
-    \caption*{Jack and Jill serve as the two arguments to and, but they are also semantically subjects of went.}
+    \caption*{Jack and Jill serve as the two arguments to \textit{and}, but are also subjects of \textit{went}.}
   \end{figure}

### 2014-12-07 doc/slides/hong+eisner.TLT13.slides.tex
typos. captions

@@
   \begin{itemize}
-    \item Grammar-based formalism for projective dependency parsing with undirected links
-    \item We set out to discover if the undirected links are actually directed relationships.
+    \item[] Grammar-based formalism for projective dependency parsing with undirected links
+    %\item We set out to discover if the undirected links are actually directed relationships.
   \end{itemize}
@@
 \frame{\frametitle{Link Grammars}
+Compare resulting dependency parse with CoNLL 2007 shared task. 
 \begin{figure}
@@
 
-  \caption*{Compare resulting dependency parse with CoNLL 2007 shared task}
+  \caption*{Top half is CoNLL. Bottom half is the directionalized link parse.}
 \end{figure}

### 2014-12-07 doc/slides/hong+eisner.TLT13.slides.tex
removed some slides. Moved some slides. Working on the ILP slides

@@
       \end{deptext}
-      \deproot[edge above, thick, hide label, edge unit distance = 1.5ex]{7}{}
+      \deproot[edge above, thick, hide label, edge unit distance = 2ex]{7}{}
       \depedge[edge above, thick, hide label]{2}{1}{}
@@
       \end{deptext}
-      \deproot[edge above, thick, hide label, edge unit distance = 1.5ex]{4}{}
+      \deproot[edge above, thick, hide label, edge unit distance = 2ex]{4}{}
       \depedge[edge above, thick, hide label]{2}{1}{}
@@
 
-\section{Non Projectivity}
-\frame{\frametitle{We do not address: Non-Projectivity}
-\begin{itemize}
-  \item Projectivity assumptions allow us to have efficient parsing algorithms that utilize dynamic programming
-  \item English is mostly projective
-\end{itemize}
+%\section{Non Projectivity}
+%\frame{\frametitle{We do not address: Non-Projectivity}
+%\begin{itemize}
+%  \item Projectivity assumptions allow us to have efficient parsing algorithms that utilize dynamic programming
+%  \item English is mostly projective
+%\end{itemize}
 
-\NoteJH{Should I mention non-projectivity? This slide is mostly here because of a reviewer comment.}
-}
+%\NoteJH{Should I mention non-projectivity? This slide is mostly here because of a reviewer comment.}
+%}
 
-\frame{\frametitle{We do not address: Non-Projectivity}
-\begin{itemize}
-  \item Projectivity assumptions allow us to have efficient parsing algorithms that utilize dynamic programming
-  \item English is mostly projective
-\end{itemize}
+%\frame{\frametitle{We do not address: Non-Projectivity}
+%\begin{itemize}
+%  \item Projectivity assumptions allow us to have efficient parsing algorithms that utilize dynamic programming
+%  \item English is mostly projective
+%\end{itemize}
 
-\begin{figure}
-  \includegraphics[width=0.7\linewidth]{multiheaded}
-  \caption*{But not always}
-  \NoteJH{image taken from Chris Dyer's slide: http://demo.clab.cs.cmu.edu/fa2014-11711/images/0/0e/Depparsing.pdf}
-  \NoteJH{I should also ask Chris for permission to use this}
-\end{figure}
+%\begin{figure}
+%  \includegraphics[width=0.7\linewidth]{multiheaded}
+%  \caption*{But not always}
+%  \NoteJH{image taken from Chris Dyer's slide: http://demo.clab.cs.cmu.edu/fa2014-11711/images/0/0e/Depparsing.pdf}
+%  \NoteJH{I should also ask Chris for permission to use this}
+%\end{figure}
 
-}
+%}
 
@@
 \section{ILP Model}
-\frame{\frametitle{Integer Linear Programming Model: Variables}
 
-For each sentence, for each edge $i,j$ where $i < j$, we introduce variables $x_{ij}$ and $x_{ji}$.
+\frame{\frametitle{Integer Linear Programming Model}
 
-\NoteJH{TODO: Incomplete slide. Need to describe things like depth of the variable, label}
-}
-
-\frame{\frametitle{Integer Linear Programming Model: Encoded Constraints}
-
+Encoded Constraints:
 \begin{itemize}
@@
 \NoteJH{Take from the paper.}
+}
+
+\frame{\frametitle{Integer Linear Programming Model}
+
+For each sentence, for each edge $i,j$, where $i < j$
+
+\begin{figure}
+  \begin{dependency}[edge style={-}]
+    \begin{deptext}
+      . \& . \& . \& $i$ \& . \& . \& . \& $j$ \& . \& . \& . \\
+    \end{deptext}
+    \depedge[edge above, thick, edge unit distance = 1.1ex]{4}{8}{L}
+  \end{dependency}
+\end{figure}
+
+Variables:
+
+\begin{itemize}
+\item $x_{ij}, x_{ji} \in \mathbb{Z} \geq 0$ representing the orientation of each link
+
+$x_{ij} + x_{ji} = 1$
+
+%\item $x_{ij} + x_{ji} = 1$ (A link can only be oriented left or right)
+
+\end{itemize}
+
+%\NoteJH{TODO: Incomplete slide. Need to describe things like depth of the variable, label}
 }
@@
 
-  \item[] Acyclic
+  \item[] Acyclicity
+    \begin{itemize}
+      \item a
+    \end{itemize}
+
     \begin{align}
@@
       \end{itemize}
+  \end{itemize}
+
+  \begin{itemize}
   \item[]
@@
       \end{itemize}
+  \end{itemize}
+
+  \begin{itemize}
   \item[] The link labels mostly have a consistent direction.

### 2014-12-07 doc/slides/hong+eisner.TLT13.slides.tex
updated ILP slides

@@
 Variables: 
-
 \begin{itemize}
-\item $x_{ij}, x_{ji} \in \mathbb{Z} \geq 0$ representing the orientation of each link
-
-$x_{ij} + x_{ji} = 1$
+\item[] $x_{ij}, x_{ji} \in \mathbb{Z} \geq 0$: orientation of each link
+\item[] $x_{ij} + x_{ji} = 1$
 
@@
 
-\subsection{Connectedness, Acyclicity}
 \frame{\frametitle{Connectedness, Acyclicity}
@@
     \begin{itemize}
-      \item a
+      \item[] Given that node $u$ is the parent of $v$
+      \item[] $n_v$: length of the sentence containing node $v$
+      \item[] $d_v \in [0, n_v]$: depth of the node from the root of the sentence
     \end{itemize}
@@
 
-\frame{\frametitle{Consistence of Directionalized Links}
+\frame{\frametitle{Consistency of Directionalized Links}
 \begin{itemize}
   \item[] Consistency of Directionalized Links
+    \begin{itemize}
+      \item[] $r_L, \ell_L \in \{0,1\}$: whether links with label $L$ allowed left/right
+    \end{itemize}
     \begin{align}\label{direction+slack}
@@
     \end{align}
-
     \begin{align}\label{eqn:obj}
@@
     \end{align}
+    \begin{itemize}
+      \item[]  
+      \item[]  
+    \end{itemize}
 
-    %$$x_{ij} \leq r_L + s_{ij} $$
-    %$$x_{ji} \leq \ell_L + s_{ij}$$
-    %$$\min \left( \sum_L r_L + \ell_L \right)$$
+    \begin{itemize}
+      \item[]
+    \end{itemize}
 
@@
 
-\frame{\frametitle{Consistence of Directionalized Links with Slack}
+\frame{\frametitle{Consistency of Directionalized Links with Slack}
 \begin{itemize}
   \item[] Consistency of Directionalized Links
+    \begin{itemize}
+      \item[] $r_L, \ell_L \in \{0,1\}$: whether links with label $L$ allowed left/right
+    \end{itemize}
     \begin{align}\label{direction+slack}
@@
     \end{align}
+    \begin{itemize}
+      \item[] $s_{ij} \in \mathbb{R} \geq 0$: slack variable 
+      \item[] $N_L$: Number of link tokens with label $L$
+    \end{itemize}
 
-    %$$x_{ij} \leq r_L + s_{ij} $$
-    %$$x_{ji} \leq \ell_L + s_{ij}$$
-    %$$\min \left( \sum_L r_L + \ell_L \right) \cdot \frac{N_L}{4} + \sum_{ij}s_{ij}$$
+    \begin{itemize}
+      \item[] Slack allows a few links with label $L$ in disallowed directions
+    \end{itemize}
 
@@
       \end{itemize}
-
     \item[] Link tokens that required disallowed direction via slack:
@@
   \end{itemize}
-
   \begin{itemize}
@@
 
-\frame{\frametitle{Link Results}
-  \begin{itemize}
-    \item Subject-Verb links are backwards.
-  \end{itemize}
-
+\frame{\frametitle{Link Results: Subject-Verb links are backwards}
 \begin{figure}
@@
 	\end{deptext}
-	%\deproot[edge above, edge style = {blue, dotted}]{3}{\small ROOT}
+	\deproot[edge above]{3}{\small ROOT}
 	\depedge[edge below, edge style = {red, ultra thick}]{2}{3}{S}
 	\depedge[edge above, edge style = {blue, ultra thick}]{3}{2}{\small SBJ}
-	%\depedge[edge above, edge style = {blue, dotted}]{3}{4}{\small ADV}
-	%\depedge[edge above, edge style = {blue, dotted}]{3}{5}{\small ADV}
-	%\depedge[edge below, edge style = {red, thick}]{3}{6}{I}
-	%\depedge[edge above, edge style = {blue, thick}]{3}{6}{\small VC}
-	%\depedge[edge above, edge style = {blue, dotted}, edge unit distance =1.5ex]{3}{10}{\small P}
-	%\depedge[edge below, edge style = {red, thick}]{2}{1}{D}
-	%\depedge[edge above, edge style = {blue, thick}]{2}{1}{\small NMOD}
-	%\depedge[edge below, edge style = {red, thick}]{7}{8}{MV}
-	%\depedge[edge above, edge style = {blue, thick}]{7}{8}{\small ADV}
-	%\depedge[edge below, edge style = {orange, thick}]{6}{7}{P}
-	%\depedge[edge above, edge style = {blue, thick}]{6}{7}{\small VC}
-	%\depedge[edge below, edge style = {red, thick}]{8}{9}{J}
-	%\depedge[edge above, edge style = {blue, thick}]{8}{9}{\small PMOD}
-	%\deproot[edge below, edge style = {red, dotted}]{2}{W}
-	%\deproot[edge below, edge style = {orange, ultra thick, dotted}]{7}{WV}
-	%\deproot[edge below, edge style = {red, dotted}]{10}{X}
-	%\depedge[edge below, edge style = {red, dotted}]{6}{4}{E}
-	%\depedge[edge below, edge style = {red, dotted}]{6}{5}{E}
+	\depedge[edge above]{3}{4}{\small ADV}
+	\depedge[edge above]{3}{5}{\small ADV}
+	\depedge[edge below]{3}{6}{I}
+	\depedge[edge above]{3}{6}{\small VC}
+	\depedge[edge above, edge unit distance =1.5ex]{3}{10}{\small P}
+	\depedge[edge below]{2}{1}{D}
+	\depedge[edge above]{2}{1}{\small NMOD}
+	\depedge[edge below]{7}{8}{MV}
+    \depedge[edge above]{7}{8}{\small ADV}
+	\depedge[edge below]{6}{7}{P}
+	\depedge[edge above]{6}{7}{\small VC}
+	\depedge[edge below]{8}{9}{J}
+	\depedge[edge above]{8}{9}{\small PMOD}
+	\deproot[edge below]{2}{W}
+    \deproot[edge below]{7}{WV}
+	\deproot[edge below]{10}{X}
+	\depedge[edge below]{6}{4}{E}
+	\depedge[edge below]{6}{5}{E}
   \end{dependency}

### 2014-12-08 doc/slides/hong+eisner.TLT13.slides.tex
added a note
put in a table
small changes
slides so far. example sentence updated (and sent to jason)

@@
 \date{\today}
-\frame{\titlepage}
+\frame{\titlepage \NoteJH{Include pictures of us?}}
 
@@
     \item A multiheaded dependency corpus would be useful for testing new parsing algorithms
+    \item[]  
+    \item[]  
+    \item[]  
   \end{itemize}
@@
     \item Such a corpus could be automatically annotated using Integer Linear Programming
+    \item[]  
+    \item[]  
   \end{itemize}
@@
     \item We explored whether the Link Grammar could be adapted for this purpose.
+    \item[]  
   \end{itemize}
@@
 \end{itemize}
-
-\NoteJH{Take from the paper.}
 }
@@
 \frame{\frametitle{Data Sets}
-\begin{itemize}
-  \item[] 18,577 English sentences with 10,960 connected parses
-  \item[] 18,577 unlabeled Russian sentences with 4,913 connected parses
-\end{itemize}
+%\begin{itemize}
+%  \item[] 18,577 English sentences with 10,960 connected parses
+%  \item[] 18,577 unlabeled Russian sentences with 4,913 connected parses
+%\end{itemize}
+  \begin{table}[h]
+    \begin{tabular}{|l|l|l|}
+      \hline
+      & \# Sentences & \# Connected Parses \\ \hline
+      English             & 18,577        & 10,960               \\ \hline
+      Russian (unlabeled) & 18,577        & 4,913                \\ \hline
+    \end{tabular}
+  \end{table}
 
-}
+  On the English data set:
 
-\frame{\frametitle{On the English data set}
   \begin{itemize}
@@
 
+%\frame{\frametitle{On the English data set}
+%\begin{itemize}
+%\item[] Link Data has 8\% additional edges over the CoNLL.
+%\item[] 52\% of links match CoNLL arcs
+%\item[] 57\% of CoNLL arcs match links
+%\end{itemize}
+
+%
+%}
+
 \frame{\frametitle{ILP Results}
+%  \begin{table}[h]
+%    \begin{tabular}{|l|l||l|}
+%      \hline
+%      & & out of total \\ \hline
+%      Link types & & \\allowing both directions & 7 & 113 \\ \hline
+%      Link tokens & & \\requiring disallowed direction & & \\via slack & 4043 & 195,000 \\ \hline
+%    \end{tabular}
+%  \end{table}
   \begin{itemize}
@@
     \begin{itemize}
-        \item[] $\frac{7}{113} = 6.19\%$ 
+    \item[] 7 / 113 = 6.19\%
     \end{itemize}
@@
     \begin{itemize}
-        \item[] $\frac{4043}{195000} = 2.07\%$ 
+    \item[] 4043 / 195,000 = 2.07\%
     \end{itemize}
@@
       \begin{itemize}
-        \item[] $\frac{7}{113} = 6.19\%$ 
+      \item[] 7 / 113 = 6.19\%
+      %\item[] $\frac{7}{113} = 6.19\%$ 
       \end{itemize}
@@
       \begin{itemize}
-        \item[] $\frac{4043}{195000} = 2.07\%$ 
+        %\item[] $\frac{4043}{195000} = 2.07\%$ 
+        \item[] 4043 / 195,000 = 2.07\%
       \end{itemize}
@@
 	\end{deptext}
-	\deproot[edge above]{3}{\small ROOT}
+	\deproot[edge above, edge style = {gray}, hide label]{3}{\small ROOT}
 	\depedge[edge below, edge style = {red, ultra thick}]{2}{3}{S}
 	\depedge[edge above, edge style = {blue, ultra thick}]{3}{2}{\small SBJ}
-	\depedge[edge above]{3}{4}{\small ADV}
-	\depedge[edge above]{3}{5}{\small ADV}
-	\depedge[edge below]{3}{6}{I}
-	\depedge[edge above]{3}{6}{\small VC}
-	\depedge[edge above, edge unit distance =1.5ex]{3}{10}{\small P}
-	\depedge[edge below]{2}{1}{D}
-	\depedge[edge above]{2}{1}{\small NMOD}
-	\depedge[edge below]{7}{8}{MV}
-    \depedge[edge above]{7}{8}{\small ADV}
-	\depedge[edge below]{6}{7}{P}
-	\depedge[edge above]{6}{7}{\small VC}
-	\depedge[edge below]{8}{9}{J}
-	\depedge[edge above]{8}{9}{\small PMOD}
-	\deproot[edge below]{2}{W}
-    \deproot[edge below]{7}{WV}
-	\deproot[edge below]{10}{X}
-	\depedge[edge below]{6}{4}{E}
-	\depedge[edge below]{6}{5}{E}
+	\depedge[edge above, edge style = {gray}, hide label]{3}{4}{\small ADV}
+	\depedge[edge above, edge style = {gray}, hide label]{3}{5}{\small ADV}
+	\depedge[edge below, edge style = {gray}, hide label]{3}{6}{I}
+	\depedge[edge above, edge style = {gray}, hide label]{3}{6}{\small VC}
+	\depedge[edge above, , edge style = {gray}, hide label, edge unit distance =1.5ex]{3}{10}{\small P}
+	\depedge[edge below, edge style = {gray}, hide label]{2}{1}{D}
+	\depedge[edge above, edge style = {gray}, hide label]{2}{1}{\small NMOD}
+	\depedge[edge below, edge style = {gray}, hide label]{7}{8}{MV}
+    \depedge[edge above, edge style = {gray}, hide label]{7}{8}{\small ADV}
+	\depedge[edge below, edge style = {gray}, hide label]{6}{7}{P}
+	\depedge[edge above, edge style = {gray}, hide label]{6}{7}{\small VC}
+	\depedge[edge below, edge style = {gray}, hide label]{8}{9}{J}
+	\depedge[edge above, edge style = {gray}, hide label]{8}{9}{\small PMOD}
+	\deproot[edge below, edge style = {gray}, hide label]{2}{W}
+    \deproot[edge below, edge style = {gray}, hide label]{7}{WV}
+	\deproot[edge below, edge style = {gray}, hide label]{10}{X}
+	\depedge[edge below, edge style = {gray}, hide label]{6}{4}{E}
+	\depedge[edge below, edge style = {gray}, hide label]{6}{5}{E}
   \end{dependency}
@@
 \begin{itemize}
-\item This is due to an inconsistency of the Link Grammar, discovered by our method. \NoteJH{The reason why is in footnote 6 of page 10 in our paper. Reread after sleeping.}
+\item This is due to an inconsistency of the Link Grammar, discovered by our method. 
 \end{itemize}

### 2014-12-08 doc/abstract/LinksToDAG_abstract.tex
updated abstract. For submission to MASC

--- 2014-11-21 doc/LinksToDAG_final.tex
@@
 
-\documentclass[a4paper,11pt]{article}
-\usepackage[latin1]{inputenc}
-\usepackage[T1]{fontenc}         
-\usepackage{pslatex}
-
-\usepackage{hyperref}
-\pagestyle{empty}
-
-%\bibliographystyle{acl2014}
-\bibliographystyle{plain}
-\newcommand{\newcite}[1]{\cite{#1}}   % ACL's newcite is not available in plain bib style, so just use cite
-
-%\documentclass[11pt]{article}
-%\usepackage{acl2014}
+\documentclass[11pt]{article}
+\usepackage{acl2014}
 \usepackage{times}
-%\usepackage{url}
-%\usepackage{latexsym}
+\usepackage{url}
+\usepackage{latexsym}
 \usepackage{amsmath}
-
-%\usepackage{xspace}
+\usepackage{xspace}
 \usepackage{enumitem}
@@
 \usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
-%\usepackage{soul}
+\usepackage{soul}
 
@@
 
-\usepackage{tikz-dependency}
+%\usepackage{tikz-dependency}
 
 \newcommand{\eqnref}[1]{\eqref{eqn:#1}}
+\bibliographystyle{acl2014}
 
-\usepackage{todonotes}   % insert [disable] to disable all notes
+\usepackage[disable]{todonotes}   % insert [disable] to disable all notes
 % Note that these macros accept optional arguments such as 
@@
 
+%\author{Juneki Hong and Jason Eisner\\
+%   Department of Computer Science \\
+%   Johns Hopkins University \\
+%   Baltimore, MD 21218, USA \\ 
+%   {\tt \{juneki,jason\}@cs.jhu.edu} \\
+%}
+
 \author{
   Juneki Hong \\
-  [0.5cm] Language Technologies Institute\\
+  Language Technologies Institute\\
   Carnegie Mellon University\\
@@
   {\texttt{juneki@cs.cmu.edu}}
-  \and
+  \And
   Jason Eisner\\
-  [0.5cm] Department of Computer Science\\
+  Department of Computer Science\\
   Johns Hopkins University\\
@@
 
-\begin{abstract}
-\noindent
-Under multi-headed dependency grammar, a parse is a connected DAG rather than a tree.  Such formalisms can be more syntactically and semantically expressive.  However, 
-it is hard to train, test, or improve multi-headed parsers because few multi-headed corpora exist, particularly for the projective or planar case.
-% OLD
-% investigate the benefit of such parsers or to work on making them faster or more accurate.  
+%\begin{abstract}
+%\noindent
+Under multi-headed dependency grammar, a parse is a connected DAG rather than a tree.
+Such formalisms can be more syntactically and semantically expressive.
+However, it is hard to train, test, or improve multi-headed parsers because few multi-headed corpora exist, particularly for the projective or planar case.
 To help fill this gap, we observe that link grammar already produces {\em undirected} planar graphs, and so we wanted to determine whether these could be converted into directionalized dependency parses.
-% OLD
-% ... produces parses that are similar to multi-headed projective dependency parses, except that the links are undirected. 
-We use Integer Linear Programming to assign consistent directions to the labeled links in a corpus of several thousand parses produced by the Link Grammar Parser, which has broad-coverage hand-written grammars of English as well as Russian and other languages.
-We find that such directions can indeed be consistently assigned in a way that yields valid multi-headed dependency parses.
-The resulting parses in English appear reasonably linguistically plausible, though differing in style from CoNLL-style parses of the same sentences; we discuss the differences.
-
-\end{abstract}
-
-\thispagestyle{empty}
-
-%\newpage
-
-\section{Motivation}\label{sec:motivation}
 
@@
 The labeled edges of the graph represent syntactic relationships among the words.
-The vertices of the graph are simply the words $1, 2, \ldots, n$ of the sentence, along with a distinguished ``root'' vertex 0.
-
+%The vertices of the graph are simply the words $1, 2, \ldots, n$ of the sentence, along with a distinguished ``root'' vertex 0.  
 The small Link Grammar community has invested effort in creating link grammars for several languages.
@@
 One may want to automatically enrich the annotation in such cases, whether by doing some kind of learning \cite[et seq.]{matsuzaki-et-al-2005}, or by exploiting implicit constraints.
-In this paper, we use Integer Linear Programming to exploit implicit constraints of consistency and acyclicity.
 
@@
 There has been a good deal of recent research on projective dependency parsing, variously using global optimization or sequential classification (see \cite{martins-et-al-2008,bohnet-2011,chen-li-zhang-2014} for surveys).
-Some of these algorithms could be extended to the {\em multi-headed} case, which is of linguistic and computational interest for reasons discussed below.
+Some of these algorithms could be extended to the {\em multi-headed} case, which is of linguistic and computational interest.
 However, for training and testing such algorithms, one would need a plausible sample of multi-headed projective dependency parses of real sentences.
@@
 
-\noindent Our automatic method for reconstructing the latent directions also had an unexpected benefit.
-It revealed an inconsistency in the hand-written English link grammar, regarding the handling of embedded sentences with missing (PRO) subjects.
+We use Integer Linear Programming to exploit implicit constraints of consistency and acyclicity, and assign consistent directions to the labeled links in a corpus of several thousand parses produced by the Link Grammar Parser, which has broad-coverage hand-written grammars of English as well as Russian and other languages.
 
-\section{Multi-Headed Dependency Parsing}
+We find that such directions can indeed be consistently assigned in a way that yields valid multi-headed dependency parses.
+The resulting parses in English appear reasonably linguistically plausible, though differing in style from CoNLL-style parses of the same sentences; we discuss the differences.
 
-Dependency parsing maps a sentence to a directed graph whose vertices are the words $1, 2, \ldots, n$ of the sentence along with a distinguished ``root'' vertex 0.
-A labeled directed edge $u \stackrel{L}{\rightarrow} v$ or $v \stackrel{L}{\leftarrow} u$ indicates that the ``child'' $v$ is some kind of argument or modifier of its ``parent'' $u$.
-The edge label $L$ indicates the specific syntactic or semantic relationship between the two words.
+\newpage
 
-In the special case $u=0$, the edge designates $v$ as playing some special top-level role in the sentence, e.g., as the main verb.  We disallow $v=0$.
-
-As discussed by \newcite{kuhlmann-nivre-2006,gomezrodriguez-nivre-2013}, one might impose various requirements on the parse graph:
-\begin{itemize}[noitemsep]
-\item {\sc Single-Head}: Each word has $\leq 1$ parent.
-\item {\sc Acyclic}: There are no directed cycles.
-\item {\sc Connected}: Each pair of words has a undirected path between them.
-\item {\sc Reachable}: Each word can be reached from 0 by a directed path (which implies {\sc Connected}).  Note that 0 may have multiple children.
-\item {\sc Planar}: edges may not ``cross.''  That is, if there are edges between $i,j$ and between $u,v$, where $i < u < j$, then {\sc Planar} requires $i \leq v \leq j$.
-\end{itemize}
-It is common to impose all of these requirements at once, leading to a {\em projective dependency parser} that produces projective trees rooted at 0.
-However, parsing algorithms can be devised that relax any or all of the requirements \cite{gomezrodriguez-nivre-2013}.
-
-In this paper, we are interested in relaxing the {\sc Single-Head} requirement while preserving all the others.
-This is the setting of {\em multi-headed projective dependency parsing}.
-Just as in the single-headed case, the other requirements ensure that all edges are projective.
-(A projective edge is one where the parent is an ancestor of all words between the parent and the child \cite{nivre-slides}.)
-
-Relaxing {\sc Single-Head} means that the parse can have more than $n$ edges, allowing it to express more relationships between words.  
-In English, for example, here are some constructions that seem to call for a multi-headed analysis:
-\begin{description}
-\item[control] In {\em ``Jill likes to skip,''} the word {\em Jill} is the subject of two verbs.
-In {\em ``Jill persuaded Jack to skip,''} {\em Jack} is the object of one verb and the subject of another.
-Without recognizing this, our parser would miss the syntactic invariants that {\em skip} always has a subject and {\em persuaded} always has an object.
-It would also be unable to exploit the selectional preferences of both verbs to help disambiguate the parse.
-This is why we prefer to make the parser aware of multi-headedness, rather than using a single-headed parser and then extracting the additional semantic roles from its output.
-\item[relativization] In {\em ``The boy that Jill skipped with fell down,''} the word {\em boy} is the object of {\em with} as well as the subject of {\em fell}.
-Without recognizing this, we would miss the syntactic invariant that {\em with} always has an object.
-\item[conjunction] In {\em ``Jack and Jill went up the hill,''} {\em Jack} and {\em Jill} serve as the two arguments to {\em and}, but they are also semantically subjects of {\em went}.
-Without recognizing this, we would have no (local) reason for expecting the arguments of {\em and} to be nouns.
-\end{description}
-
-\noindent In linguistics, it is common to analyze some of these structures using trees with ``empty categories.''  The subject of {\em skip} is taken to be a silent morpheme {\em PRO}:
-{\em ``Jill$_i$ likes PRO$_i$ to skip.''}  However, this is no longer a tree if one considers the implicit undirected edge between {\em Jill} and {\em PRO} (denoted by their shared index $i$).
-Our simpler representation contracts this coreference edge, eliminating {\em PRO} and creating a {\em Jill} $\leftarrow$ {\em skip} link.
-
-\medskip
-
-An anonymous reviewer objects to research on projective parsing algorithms, since the {\sc Planar} restriction is linguistically questionable even for single-headed parsing, and even more so for multi-headed parsing.
-However, efficient algorithms often exist in the projective case, and these projective algorithms---which are typically first developed on a projective corpus of the sort that we will construct---can be useful even when the true parses are not quite projective.
-Natural-language parses have a low rate of non-projective edges \cite{nivre-slides} and the non-projective parses tend to be ``almost projective'' \cite{pitler-et-al-2013}.
-Thus, one can apply a fast projective parser as an approximate method, or as one ingredient in a more complex model \cite{martins-et-al-2013}, or as the first step in a coarse-to-fine or stacking architecture \cite{rush-petrov-2012,martins-et-al-2008} in order to obtain preliminary edge scores that are then supplied to a non-projective parser.
-Another approach is to transform non-projective parses into a projective annotation style so that projective parsers can be used \cite{nivre-nilsson-2005,nivre-2009}.
-
-\section{Link Grammars}
-
-Graph representations of syntactic and semantic structure have been widely considered of late \cite{demarneffe-manning-2008,demarneffe-et-al-2013,ivanova-et-al-2012,AMR-2013,oepen-et-al-2014}.
-A few past NLP papers have explored multi-headed dependency parsing \cite{buchkromann-2006,mcdonald-pereira-2006,sagae-tsujii-2008,gomezrodriguez-nivre-2013}.
-They constructed their multi-headed dependency corpora by automatically converting from other formats such as HPSG.
-Currently there seem to be no corpora that were directly annotated in this form, other than the Danish Dependency Treebank \cite{kromann-2003}.
-
-% ADDRESSED BY ABOVE?
-% \NoteJE{We should acknowledge that another option would be to post-process either CoNLL parses or our link parses to automatically add more multi-heading, e.g., to handle control phenomena ("John wants to skip" or "Jane ran for office in order to change the tax laws" should have John/Jane be the subject of two verbs).
-However, that requires per-language}
-
-The above work considered non-projective parses.
-It seems at first that no one has worked out annotation conventions for {\em projective} multi-headed dependency parsing.
-However, this is only half-true.
-Link Grammar \cite{sleator-temperley-1991} is a grammar-based formalism for projective dependency parsing with {\em undirected} links.
-It produces undirected connected planar graphs.
-Annotation conventions are implicit in the detailed lexicon for the Link Grammar Parser \cite{LINK-1999}.
-The 122 link types in the English lexicon are documented at \url{http://www.abisource.com/projects/link-grammar/dict/summarize-links.html},
-which specifies for every word a constraint on the {\em sequence} of labeled leftward and rightward edges attached to it.
-As remarked by \newcite{eisner-2000-iwptbook}, this is analogous to dependency grammar's use of head automata to constrain a word's sequence of left and right children.
-For example, in {\em ``The boy that Jill skipped with fell down,''} the word {\em with} uses a lexical entry that requires it to link to a governing verb to the left, an extracted object farther to the left, and nothing to the right.
-Each entry has a hand-assigned cost in \{0,1,2\}, and the parser finds the parse of minimum total cost \cite{LINK-1999,vepstas-2014}.
-
-Given a link grammar parse, it would be straightforward to convert it to an acyclic dependency parse by orienting all edges rightward.
-However, the result may violate the {\sc Reachable} constraint.
-Instead we could orient all edges by depth-first search from the root node, which yields a DAG satisfying all our constraints.
-However, this might result in inconsistent annotation conventions, with some \texttt{S}-labeled links pointing from subject to verb and others from verb to subject.
-
-In the English link grammar, an \texttt{S} edge encodes a ``subject-verb'' relation {\em whose left word serves as the subject}.
-We would expect verbs to point to their subject arguments in dependency grammar, and so we surmise that all \texttt{S} links should be interpreted as pointing leftward (from verb to subject: {\em ``\mbox{Jack $\stackrel{\texttt{S}}{\leftarrow}$ is} falling''}).
-
-In general, we supposed that the link grammar lexicon designers actually had a {\em consistent} direction in mind for each {\em edge label}.
-This does not imply that English subjects must always appear to the left of the verb!
-The link grammar designers took care to use a distinct \texttt{SI} label in cases of subject-verb inversion, and we surmise that \texttt{SI} links are intended to point rightward (again from verb to subject: {\em ``\mbox{Is $\stackrel{\texttt{SI}}{\rightarrow}$ Jack} falling?''}).
-Similarly, different edge labels are used for English ``object-verb'' relations according to whether it is the left or the right word that serves as the object.
-These labels are presumably intended to encode different edge directions.
-
-% PREVIOUS ATTEMPT
-% This does not rule out variation in the word order used to express a semantic relation.
-For example, the English link grammar allows both SVO and VSO sentences, but with different edge labels.
-An \texttt{S} edge encodes a ``subject-verb'' relation whose the left word serves as the subject, while an \texttt{SI} edge encodes the same relation when the the right word serves as the subject.
-
-Our goal in this paper is to recover these implicit directions by global optimization.
-We seek a fixed mapping from labels to directions such that link grammar parses become directed dependency parses that satisfy all of our constraints.
-
-Our first thought was to seek a direction mapping such that no parsed word sequence allowed by the link grammar lexicon could possibly violate our constraints after directions were imposed.
-This is a well-defined constraint programming problem.
-For example, to prevent cyclicity, we would require (roughly speaking) that no word type in the lexicon could follow a sequence of directed rightward links through other word types and then a leftward link back to itself.
-
-% REMOVED AT REVIEWER SUGGESTION
-% However, working directly with the link grammar lexicon format is somewhat tricky. We also feared
-However, we feared that there would not be a feasible solution---because of errors in the lexicon or linguistically unnatural word sequences not anticipated by the grammar designers.
-In this case it would be unclear how to relax our constraints.
-
-Thus, rather than considering all theoretically possible word sequences, we chose to use a sample of {\em naturally occurring} sentences parsed by the link grammar, and to seek a direction mapping so that {\em these} parses would not violate our constraints after directions were imposed.
-If no such mapping exists, then we are willing to orient a few edge tokens in the wrong direction to ensure that the parses are still well-formed---but we minimize the number of such violations.
-In this way, the empirical distribution of sentences guides our assignment of directions.
-We are releasing the resulting multi-headed directed corpus via our personal websites.
-\noteJE{is this the best way?}
-
-% Link grammars are a grammatical system equivalent to context-free grammars that assign linking requirements to every given word.
-A link parser then tries to satisfy all of these requirements for every word of a sentence while still maintaining planarity.
-The resulting links describe the relationships between constituents in a parse.
-
-% The link grammar is based on a set of handwritten dictionaries.
-Instead of going through these dictionaries, we learned the grammar using an ILP.
-This approach also potentially allows us to analyze link grammar dictionaries other than English.
-\todo{explore other link grammar dictionaries}
-
-% \NoteJE{what statistics does it use?  what languages are available}
-
-% Link parsing in contrast produces a multi-headed planar graph with undirected edges, where every edge has a label describing the relationship between two constituents in a parse.
-In this paper we explore whether these relationships include dependencies, and whether the multi-headedness of the link grammar offers additional dependency relationships not found in other corpora.
-
-% Dependency parsing is the task of mapping a sentence to a projective (not always projective?) directed acyclic tree.
-Link parsing in contrast produces a multi-headed planar graph with undirected edges, where every edge has a label describing the relationship between two constituents in a parse.
-In this paper we explore whether these relationships include dependencies.
-To determine the directional dependencies within the link edge labels we will use integer linear programming, encoding the problem in the Zimpl little language \cite{Koch2004}.
-It turns out that the link parses roughly only match half of the conll dependency corpus.
-However this is because \todo{}.
-
-\section{Data Sets}
-We used the English sentences from the CoNLL 2007 Shared Task \cite{CONLL-SHARED-2007}---a subset of the Penn Treebank for which {\em single}-headed reference parses are available. 
-%We also used a prefix of the Russian National Corpus,\footnote{That is, a prefix when considering the files in Unix sort order by filename.
-We dropped sentences with $\leq 2$ words, as well as file 22592319, which appeared to contain charts.} which is unparsed.
-We also used a prefix of the Russian News Commentary data from the ACL 2013 Shared Task of Machine Translation,\footnote{\url{http://www.statmt.org/wmt13/training-monolingual-nc-v8.tgz}} which is unparsed.
-
-We generated link parses using the AbiWord/CMU link grammar parser version 5.0.8 \cite{LINKPARSER-2014}.
-The parser's coverage is less than ideal: we obtained connected parses for only \input{figure/conll_analysis_sentences_total.tex}(of \input{figure/conll_analysis_sentences_original.tex}$\!\!$) English sentences and only \input{figure/russian_sentence_count.tex}(of \mbox{\input{figure/russian_original_sentence_count.tex}$\!\!$)} Russian sentences, discarding the other sentences.\footnote{When the link parser fails, it outputs a disconnected graph representing its best effort parse within a time limit.
-We removed these sentences for fear that the parses would be unreliable.
-% REMARK BY JUNEKI, PROBABLY TOO DISTRACTING
-% our ILP encoding tries to satisfy the edges of an entire connected parse.
-Given a solution, it would be possible to return to these disconnected parses, directionalize the edges, and include them along with the rest of the corpus.
-It would also be possible to include these sentences for the ILP to solve by modifying the ILP to satisfy the edges of all disconnected components of a parse.
-}
-These two languages have the most mature lexicons at present, although lexicons for 7 other languages are available.
-
-On English, the link grammar parses have \input{figure/conll_analysis_edge_percent.tex}more edges overall, indicating that their directed versions will have a few multi-headed constructions per sentence.
-They
-do differ in style from the single-headed CoNLL parses of the same English sentences.
-Only \input{figure/conll_analysis_precision.tex}of the links match CoNLL arcs, and only \input{figure/conll_analysis_recall.tex}of the CoNLL arcs match links.
-% DATA FOR THE ABOVE
-% \begin{figure}[ht!]
-%   \centering
-%   \small
-%   \input{figure/conll_analysis_arcs.tex}
-%   \caption{Arcs that match links}
-% \end{figure}
-% 
-% OLD
-% Using the sentences of the dependency corpus as comparison, we find that the link parses do not wholly subsume dependency parses, and that the undirected links match roughly three fourths of the arcs in the conll dependency corpus.
-\todo{This is because...}
-
-% With the English sentences we compared our link parse results to the original dependency annotations, and with the Russian sentences we only produced link parses.
-
-%\subsection{Discarding Incomplete Link Parses}
-
-%\NoteJE{ideally we could cut this section given what I wrote above.
-However, you seem to be saying that we did train on sentences with dropped words as long as the parses were connected.
-Didn't we decide to drop those sentences too?}
-
-%Not all of the data was used for this work.
-From the given sentences, we discarded those that the link parser could not process and returned a parse with nodes disconnected from the root.
-\input{figure/conll_analysis_sentences_disconnected} of the English sentences were removed in this way.
-The ILP was run only on the connected parses to produce a corpus of directionalized link parses.
-
-%For our analysis of these parses, we further discarded those that had dropped words, where the link parser could not attach links to every word in the sentence.
-About \input{figure/conll_analysis_sentences_dropped} of the non-disconnected sentences were removed, leaving us working only with complete link parses.
-
-%\begin{figure}[ht!]
-%  \centering
-%  \small
-%  \input{figure/conll_analysis_sentences.tex}
-%  \caption{\small The English sentences used for this work.}
-%\end{figure}
-
-%\noteJH{The stability results come from the second row of this table. The analysis later in the paper uses the third row. Putting this section here might be confusing to readers?}
-
-\section{Integer Linear Programming Model}
-
-For each undirected labeled edge $ij$ in the link corpus, where $i,j$ denote tokens in the same sentence with $i < j$, we introduce nonnegative integer variables $x_{ij}$ and $x_{ji}$ with a constraint $x_{ij}+x_{ji}=1$.
-We interpret $x_{ij}=1$ or $x_{ji}=1$ to mean that the link has direction $i \rightarrow j$ or $i \leftarrow j$, respectively.\footnote{In practice we halve the number of variables by replacing $x_{ji}$ with $1-x_{ij}$ for $j > i$, but that obscures the exposition.}
-
-% OLD VERSION
-% we introduce a variable $x_{ij}$ that is 0 or 1 according to whether $e$ points left or right.  We abbreviate $1-x_{ij}$ as $\bar{x}_{ij}$.
-
-For each non-0 token $v$, we ensure that it has at least one parent by constraining\footnote{To denote two linked tokens, we use variables $i,j$ when $i$ is to the left of $j$, or variables $u,v$ when $u$ is the parent of $v$.}
-\begin{align}\label{eqn:oneparent}
-\sum_u x_{uv} & \geq 1
-\end{align}
-% OLD VERSION
-% $$\sum_{u < v} x_{uv} + \sum_{u > v} \bar{x}_{vu} \geq 1$$
-where $u$ ranges only over tokens such that the relevant variable exists.
-%
-To prevent cycles,\footnote{This also ensures {\sc reachable}, given \eqnref{oneparent}.} for each token $v$ we introduce a depth variable $d_v$ in the range $[0,n_v]$ (not constrained to be integer), where $n_v$ is the length of the sentence containing $v$.
-We require a child's depth to be at least 1 greater than each of its parents' depths---constraints that can be satisfied iff the sentence has no directed cycles:
-\begin{align}\label{eqn:nocycles}
-(\forall u)\; d_v + (1+n_v) \cdot (1-x_{uv}) & \geq 1+d_u
-% OLD VERSION
-% $$(\forall u < v) d_v + (1+n_v) \cdot \bar{x}_{uv} \geq 1+d_u$$
-% $$(\forall u > v) d_v + (1+n_v) \cdot x_{vu} \geq 1+d_u$$
-\end{align}
-The second summand ensures that \eqnref{nocycles} is trivially satisfied (hence has no effect) when $u$ is {\em not} the parent of $v$.  
-\NoteJE{in final version: figure out whether it is faster to require the 0 tokens to have depth 0, or add all the depths (perhaps multiplied by 0.001) to the objective function to break ties in simplex, or neither.
-See also other notes in tex file.}
-% As a speedup, we can require the 0 tokens to have depth 0. \noteJE{does this really still make a difference to speed?  
-% \noteJE{what's the point of the root\_depth constraint in the .zpl file?  Almost trivially satisfied.}
-
-Finally, we encourage all links with the same label \noteJE{explain that we mean capital letters?} to have the same direction.
-For each label $L$, we introduce binary variables $r_L$ and $\ell_L$, which say whether a link of type $L$ is ``allowed'' to point right or left, respectively.
-For each undirected edge $ij$ of label $L$, with $i < j$, we write
-\begin{align}
-x_{ij} &\leq r_L + s_{ij} & 
-x_{ji} &\leq \ell_L + s_{ij}
-\end{align}
-where $s_{ij} \geq 0$ is a slack variable that allows an edge token to point in a disallowed direction if needed to ensure \eqnref{oneparent}--\eqnref{nocycles}.
-
-Our objective tries to minimize the number of allowed directions (by link type---cf.\@ \newcite{ravi2009}) and the total slack (by link token):
-\begin{align}\label{eqn:obj}
-\min \left( \sum_{L} r_L + \ell_L \right) \cdot \frac{N_L}{4} + \sum_{ij} s_{ij}
-\end{align}
-where $N_L$ is the number of link tokens with label $L$.
-Objective \eqnref{obj} is willing to tolerate up to 1/4 of those link tokens' using a disallowed direction before it prefers to allow {\em both} directions.
-\noteJE{check!}
-% CUT FOR SPACE
-% (i.e., $r_L+\ell_L=2$).
-
-\NoteJE{It might be a good idea to have the second allowed direction still carry some lesser slack penalty, so that we break ties in favor of the preferred direction.
-The two bits for each label would then say ``which direction is preferred?'' and ``is the other direction allowed?''}
-
-% OLD
-% We formulate an ILP to find the minimal set of link type to directionality assignments that would directionalize all the links in a set of link grammar parses, while still mandating that all of the resulting dependency parses would be connected DAGs, and with all the nodes reachable from the root.
-Our task of finding the smallest set of assignments that would explain the parses is also called model minimization\cite{ravi2009}.
-\noteJE{worth citing, but not quite the same thing}
-% For every encountered label we generate two boolean label/direction variables (e.g.
-label/left, label/right).
-Setting one of these variables to \textsc{true} allows the label type to only go in the specified direction, while setting both would allow the label to go in either way.
-% The objective is to minimize the number of label/direction variables set to \textsc{true}, while still satisfying acyclicity and reachability constraints.
-The encoding was written in the Zimpl little language \cite{Koch2004} and solved using the SCIP Optimization Suite \cite{achterberg2009scip}.
-% \subsection{Slack}
-% We introduced slack such that the link types were allowed to deviate from the majority one percent of the time before the ILP would assign both label/direction variables to \textsc{true}.
-This allows for flexibility against noise in the link parser's label assignments, balanced with allowing for both directions to be assigned if necessary.
-
-% 
-% \begin{figure*}[ht!]
-%   \vspace{-6pt}
-%   \centering
-%   \input{figure/sample_parses.tikz}
-%   \vspace{-6pt}
-% \end{figure*}
-
-\section{Experiments and Results}
-
-%\hspace*{-1.5in}
-\begin{figure*}[ht!]
-
-  \vspace{-6pt}
-  \centering
-
-  %\input{figure/sentences.tikz}
-
-  \hspace*{-0.5in}
-  \begin{subfigure}[b]{0.49\textwidth}
-	\begin{dependency}
-	  \begin{deptext}
-		{\scriptsize DT} \& {\scriptsize NN} \& {\scriptsize MD} \& {\scriptsize RB} \& {\scriptsize RB} \& {\scriptsize VB} \& {\scriptsize VB} \& {\scriptsize IN} \& {\scriptsize NN} \& {\scriptsize .} \\
-		the \& matter \& may \& never \& even \& be \& tried \& in \& court \& . \\
-		- \& n-u \& v \& e \& e \& v \& v-d \& r \& n-u \& - \\
-	  \end{deptext}
-	  \deproot[edge above, edge style = {blue, dotted}]{3}{\small ROOT}
-	  \depedge[edge below, edge style = {red, ultra thick}]{2}{3}{S}
-	  \depedge[edge above, edge style = {blue, ultra thick}]{3}{2}{\small SBJ}
-	  \depedge[edge above, edge style = {blue, dotted}]{3}{4}{\small ADV}
-	  \depedge[edge above, edge style = {blue, dotted}]{3}{5}{\small ADV}
-	  \depedge[edge below, edge style = {red, thick}]{3}{6}{I}
-	  \depedge[edge above, edge style = {blue, thick}]{3}{6}{\small VC}
-	  \depedge[edge above, edge style = {blue, dotted}, edge unit distance =1.5ex]{3}{10}{\small P}
-	  \depedge[edge below, edge style = {red, thick}]{2}{1}{D}
-	  \depedge[edge above, edge style = {blue, thick}]{2}{1}{\small NMOD}
-	  \depedge[edge below, edge style = {red, thick}]{7}{8}{MV}
-	  \depedge[edge above, edge style = {blue, thick}]{7}{8}{\small ADV}
-	  \depedge[edge below, edge style = {orange, thick}]{6}{7}{P}
-	  \depedge[edge above, edge style = {blue, thick}]{6}{7}{\small VC}
-	  \depedge[edge below, edge style = {red, thick}]{8}{9}{J}
-	  \depedge[edge above, edge style = {blue, thick}]{8}{9}{\small PMOD}
-	  \deproot[edge below, edge style = {red, dotted}]{2}{W}
-	  \deproot[edge below, edge style = {orange, ultra thick, dotted}]{7}{WV}
-	  \deproot[edge below, edge style = {red, dotted}]{10}{X}
-	  \depedge[edge below, edge style = {red, dotted}]{6}{4}{E}
-	  \depedge[edge below, edge style = {red, dotted}]{6}{5}{E}
-	\end{dependency}
-  \end{subfigure}
-  \hspace*{0.5in}\begin{subfigure}[b]{0.49\textwidth}
-	\begin{dependency}
-	  \begin{deptext}
-		{\scriptsize DT} \& {\scriptsize NN} \& {\scriptsize NN} \& {\scriptsize VB} \& {\scriptsize RB} \& {\scriptsize VB} \& {\scriptsize DT} \& {\scriptsize NN} \& {\scriptsize .} \\
-		the \& Paris \& bank \& does \& n't \& publish \& that \& figure \& . \\
-		- \& b \& n \& v \& - \& v \& j-d \& n \& - \\
-	  \end{deptext}
-	  \deproot[edge above, edge style = {blue, dotted}]{4}{\small ROOT}
-	  \depedge[edge below, edge style = {red, thick}]{3}{1}{D}
-	  \depedge[edge above, edge style = {blue, thick}]{3}{1}{\small NMOD}
-	  \depedge[edge below, edge style = {red, thick}]{3}{2}{AN}
-	  \depedge[edge above, edge style = {blue, thick}]{3}{2}{\small NMOD}
-	  \depedge[edge below, edge style = {red, thick}]{6}{8}{O}
-	  \depedge[edge above, edge style = {blue, thick}]{6}{8}{\small OBJ}
-	  \depedge[edge below, edge style = {red, ultra thick}]{3}{4}{S}
-	  \depedge[edge above, edge style = {blue, ultra thick}]{4}{3}{\small SBJ}
-	  \depedge[edge below, edge style = {red, thick}]{4}{5}{N}
-	  \depedge[edge above, edge style = {blue, thick}]{4}{5}{\small VMOD}
-	  \depedge[edge below, edge style = {orange, thick}]{4}{6}{I}
-	  \depedge[edge above, edge style = {blue, thick}]{4}{6}{\small VC}
-	  \depedge[edge above, edge style = {blue, dotted}, edge unit distance = 1.7ex]{4}{9}{\small P}
-	  \depedge[edge below, edge style = {red, thick}]{8}{7}{D}
-	  \depedge[edge above, edge style = {blue, thick}]{8}{7}{\small NMOD}
-	  \deproot[edge below, edge style = {red, dotted}]{3}{W}
-	  \deproot[edge below, edge style = {orange, ultra thick, dotted}]{6}{WV}
-	  \deproot[edge below, edge style = {red, dotted}]{9}{X}
-	\end{dependency}
-  \end{subfigure}
-
-  \hspace*{-1in}
-  \input{figure/long_parses.tikz}
-
-  %\input{figure/long_parse2.tikz}
-  \caption{\small The blue upper edges show CoNLL gold dependency parses; the red lower edges show our oriented version of the link grammar parses.
-Edges are shown as dotted lines if they appear only in one parse.
-Edges are highlighted in orange if the child has multiple parents.
-Edges that appear in both parses are solid lines, drawn thicker if the directions do not match.
-Vertical edges have parent 0.
-For 100 example parses, see Appendix~\ref{app:parses} of the supplementary material.}
-  \label{fig:parses}
-  \vspace{-6pt}
-
-\end{figure*}
-
-\begin{figure}[t!]
-  \small
-  \centering
-  \includegraphics[width=0.7\textwidth, keepaspectratio=true]{figure/precision_recall.png}
-  \caption{Rapid convergence to the direction mapping obtained on the largest dataset.
-The direction mappings obtained on small datasets have high {\em precision} relative to the one obtained on the largest dataset.
-Their {\em recall} grows as more link types are seen and directionalized.}\label{fig:convergence}
-  \vspace{-6pt}
-\end{figure}
-
-\begin{table*}[ht]
-% \renewcommand{\arraystretch}{1.2}    % more space between rows
-  \hspace*{-1in}
-  \input{figure/sample_link_analysis_coarse_table.tex}
-  \caption{Our solution, i.e., our reconstruction of the ``intended'' direction for each link type in the English Link Grammar.
-We also indicate the extent to which each of these link types (1) has a single dominant direction, (2) participates in multi-headed constructions, and (3) corresponds to CoNLL links of a predictable direction and type.
-For space reasons, we show only the start of this table---the full table can be found in Appendix~\ref{app:linktypes} of the supplementary material.}\label{tab:linktypes}
-\end{table*}
-
-We solved our ILP problem using the SCIP Optimization Suite \cite{achterberg2009scip}, encoding it using the ZIMPL language \cite{Koch2004}.
-Our largest run took 1.5 hours.
-\noteJH{A run of 10,000 sentences takes 3142 seconds (update in final version!)}  On English, only \input{figure/allowed_labels_both.tex}of \input{figure/allowed_labels_total.tex}link {\em types} allowed both directions, and only $\sum_{ij} s_{ij} = \input{figure/label_token_disallowed}$ of \input{figure/label_token_total.tex}link {\em tokens} required a disallowed direction via slack.
-\input{figure/conll_analysis_sentences_multiheaded.tex}of the English sentences but (alas) only \input{figure/russian_multiheaded.tex}of the Russian ones had at least one multi-headed word.
-See Table~\ref{tab:linktypes} and
-Appendix~\ref{app:linktypes} for detailed results.
-%\noteJE{2 significant digits is enough}
-%\noteJE{update these numbers}
-
-\subsection{Stability of Results}\label{sec:stability}
-
-We worried that the direction mapping might be unstable and sensitive to the input corpus.
-Happily, Figure~\ref{fig:convergence} shows otherwise (for both English and Russian).
-\noteJE{IMPORTANT!
-current graph is with fine-grained tags; rerun with the coarse-grained tags as in the rest of the paper} Using even a small prefix of the data got very high-precision results, in the sense that nearly all $r_L$ or $\ell_L$ variables that were 1 in this lightly trained mapping were also 1 in our largest run.
-The only disadvantage to using small data is low recall relative to the large run---many of the labels $L$ are not observed yet and so we do not yet allow either direction ($r_L=\ell_L=0$).
-
-We used only coarse link tags as our labels, keeping only the capital letters of a tag (and merging all \texttt{ID} tags).
-This is because other characters in a tag indicate fine-grained features such as plurality that generally do not affect link direction.
-However, when we tried using fine tags as our labels instead, we found that all refinements of the same coarse tag would almost always spontaneously agree on their preferred direction.
-This indicates that there is indeed a ``natural'' direction for the coarse tag and that we can find it.
-
-% OLD
-% We report on the convergence of the results of our ILP with increasing data.
-With increasing trials of sentences, we measure how similar the subsequent direction mappings would be with each other.
-Taking the direction table of the largest run as the answer key, we compared how much the solutions of the smaller runs deviated from it.
-% We recorded the precision of whether the mappings in the smaller runs could be found in those of the largest run, and similarly the recall of whether the mappings in the largest run could be found in the smaller runs.
-
-% Because the precision values were high, we found that the solutions of the previous runs were consistent with the largest run, and as the data increased so did the recall values, and so we would continue to encounter novel link types and then incorporate them.
-
-\subsection{Linguistic Analysis}
-
-The resulting English corpus uses a syntactic annotation scheme that is somewhat different from the CoNLL annotations.  Differences are tabulated in
-Appendix~\ref{app:linktypes} of the supplementary material, while the actual parses are contrasted in Appendix~\ref{app:parses}.
-Fragments of these appendices are shown in Table~\ref{tab:linktypes} and Figure~\ref{fig:parses}.
-% By permission of the program chairs, these appendices are available at the anonymous URL \url{http://tidyurl.com/LinksToDAG_supplement}.
-
-The link grammar results in multi-headed treatments of infinitivals, compound determiners, relative clauses, and embedding.
-The other annotation differences are generally reasonable, e.g., letting {\em 's} be the head of a possessive, and different handling of punctuation and lists.
-Of course, a few of the differences are due to parser attachment errors.
-
-The main vexation is the handling of subject-verb links.
-Under the English link grammar, the verb (or 0) that governs a clause will link to both the clause's subject and its last (main) verb.
-This would permit our desired treatment of {\em ``Jill persuaded him to skip''}, in which {\em ``him''} has two parents.
-But the ILP solution generally treats subjects as {\em parents} of verbs (thus we get {\em him} $\rightarrow$ {\em to}).
-The reason for this is an inconsistency in the link grammar itself.\footnote{Specifically, the link grammar is inconsistent about how it handles clauses with missing subjects ({\em ``Jill wanted to skip''}).
-In this case, the governing verb links to the clause's first (tensed) verb in lieu of its subject.
-The problem is that it no longer also links to the clause's last (main) verb as it would if the subject were present.
-Hence our method concludes that a VP is always headed by its first verb ({\em to} $\rightarrow$ {\em skip}).
-That is a respectable convention on its own, but recall that unfortunately, the governing verb does {\em not} link to this first verb when the subject is present.
-As a result, the only remaining possible parent for the first verb is the subject ({\em him} $\rightarrow$ {\em to} $\rightarrow$ {\em skip}).
-This leads to the odd solution where subjects are treated as parents of verbs in general.
-
-Note that subjects are not {\em always} parents, due to another inconsistency in the link grammar.
-In the construction {\em ``It was impossible, they said''}, the subject ({\em ``they''}) is inconsistently not linked to 0 but only to the verb ({\em ``said''}) and so must be the verb's child, no other parent being available.}
-Fixing the link grammar would presumably correct the link direction.
-As noted in section~\ref{sec:motivation}, it is arguably a positive result that our method was able to detect a problem in the grammar engineering.
-
-% Multiheaded:
-% IV (890/891) - multiheaded treatment of infinitivals
-% I (1783/2969) - multiheaded treatment of infinitivals
-% B (795/842) - leftward extraction from relative clauses
-% CV (1588/1799) - multiheaded treatment of subordination
-% AL (39/40) - compound determiners
-% PP (448/689) - subject attaches to aux as well as main verb
-% HA (3/3) -
-% JQ (2/2) - 
-% SF (85/111) - pleonastic subject
-% AF (6/12) - 
-% C (244/1813) - 
-% OF (155/291)
-% P (651/2060) - 
-% Q (12/30)- 
-% RS (300/305) - 
-% S (4256/7434) -
-% SX (4/6) - 
-% W (2321/4721) - 
-% WV (1800/4138) -
-
-\NoteJE{what do we say about Russian?  ask Ryan for final version?}
-
-% DOESN'T SEEM NECESSARY FOR THE LANGUAGES WE TRIED
-% \section{Future Work}
-% \subsection{Slack Hierarchy}
-% We would like to introduce slack on the link grammar types such that link types with the same coarse grained label would try and align the same way as the majority in the group, where the preceding capital letters of the link type denote the coarse grained label, while the subscript letters denote further information.
-This slack would place a prior on rare or never-before seen link types to be assigned in the same way as other similar link types.
-We think that this will give the ILP better flexibility to handle noise and novel link types while still trying to learn the overall link grammar.
-
-\section{Conclusions}
-
-We have presented an automatic ILP-based method to ``orient'' link grammar parses in multiple languages, turning them into rooted connected DAGs.
-This improves their linguistic interpretability and provides new corpora for experimenting with multi-headed dependency parsers.
-
-ILP may {\em in general} be a valuable technology for enriching existing annotated corpora.
-For example, the Penn Treebank project \cite{TREEBANK-1993} deliberately omitted types of annotations that plausibly could be added automatically.
-ILP can help by leveraging unsupervised corpus-wide information \cite{ravi2009}, enforcing annotations that are simultaneously well-formed per sentence and consistent across sentences.
-
-% \subsection{Sample of English Parses}
-% 
-% \begin{figure*}[ht!]
-%   \vspace{-6pt}
-%   \centering
-%   \input{figure/sample_parses.tikz}
-%   \vspace{-6pt}
-% \end{figure*}
-
-%\twocolumn
-
-%\newpage
+%\end{abstract}
 
@@
 
-\section{English Link Types}\label{app:linktypes}
-
-Recall that \url{http://www.abisource.com/projects/link-grammar/dict/summarize-links.html} documents the 122 link types in the English link grammar lexicon.
-Below is the result of our largest English run, whose corpus uses \input{figure/allowed_labels_total.tex}of these types.
-\noteJH{automatically generate this number correctly: 107}
-
-Other researchers could postprocess link grammar parses by using the first two columns.  These indicate whether a given link label $L$ tended to be oriented rightward or leftward by our ILP.
-
-The labels that do not have a consistent direction are mainly those that connect a conjunction to its conjuncts.\footnote{See \url{http://www.abisource.com/projects/link-grammar/dict/coordination.html}.}
-Our coarsening of the labels (see section~\ref{sec:stability}) discarded the characters that distinguish left from right conjuncts.
-
-The third column indicates where multi-headedness happens, by showing how often the child of an $L$ link had additional parents as well.
-
-The remaining columns compare to the CoNLL parses.
-They show the fraction of $L$-labeled link tokens that were also present in the CoNLL reference parse---and on that fraction, whether the orientation matches CoNLL's and what the dominant CoNLL label was.
-
-%\NoteJE{improve table formatting.  Numeric justification; en dashes; longtable; remove most hlines.}
-
-%\begin{table*}[ht]
-%\renewcommand{\arraystretch}{1.2}    % more space between rows
-%\hspace*{-1in}
-\input{figure/link_analysis_coarse_table.tex}
-%\end{table*}
-
-\section{Sample English Parses}\label{app:parses}
-
-\NoteJE{show the fine-grained link labels in the parses, even though the ILP deals with coarse-grained.  Juneki tried already to hack this in but was getting double links.}
-
-Below we display the first 100 English sentences whose widths do not exceed the width of this page.
-The blue upper half shows the CoNLL tagging and dependency parse.
-The red lower half shows the link grammar's tagging and our orientation of the link parse.
-% The dashes in the link grammar part of speech tag annotations mean that the link parser did not return one.
-A vertical edge to word token $v$ denotes an edge $0 \rightarrow v$.
-Words that were unknown to the link parser are flagged with \texttt{[!]}.
-
-Edges are shown as solid lines if they appear in both parses.  They are drawn more thickly if the directions do not match.
-
-Edges are shown as dotted lines if they appear only in one parse and not in the other.
-
-Edges are highlighted in orange if the child has multiple parents.
-
-% OLD NERDVIEW EXPLANATION
-% For every arc:
-% \begin{enumerate}
-% \item If there is a corresponding link between the same nodes then a solid blue line is drawn.
-% \item Or else a dotted blue line is drawn. 
-% \end{enumerate}
-% The bottom red arcs represent the oriented link grammar parses. For every link edge:
-% \begin{enumerate}
-% \item If there is a corresponding CoNLL arc between the same nodes then a solid red line is drawn.
-% \item If the matching arc is facing the opposite direction, then the link edge is drawn thicker.
-% \item Of the remaining edges, if an edge is a multiheaded link, then it is drawn with thick orange.
-% \item The rest of the remaining edges are drawn with dotted red lines.
-% \end{enumerate}
-
-% Uncomment this to add in all of the parses. Warning: compilation will take a long time.
-%\input{figure/parses.tikz}
-
 \end{document}

### 2014-12-09 doc/slides/hong+eisner.TLT13.slides.tex
slides so far
addressed some of Jason's minor comments

@@
 
+%\setbeamerfont{page number in head/foot}{}
+%\setbeamertemplate{footline}[frame number]
+
 \begin{document}
@@
 \section{Motivation}
-\frame{\frametitle{Motivation}
-  \begin{itemize}
-    \item A multiheaded dependency corpus would be useful for testing new parsing algorithms
-    \item[]  
-    \item[]  
-    \item[]  
-  \end{itemize}
-}
 
-\frame{\frametitle{Motivation}
-  \begin{itemize}
-    \item A multiheaded dependency corpus would be useful for testing new parsing algorithms
-    \item Such a corpus could be automatically annotated using Integer Linear Programming
-    \item[]  
-    \item[]  
-  \end{itemize}
-}
-
-\frame{\frametitle{Motivation}
-  \begin{itemize}
-    \item A multiheaded dependency corpus would be useful for testing new parsing algorithms
-    \item Such a corpus could be automatically annotated using Integer Linear Programming
-    \item We explored whether the Link Grammar could be adapted for this purpose.
-    \item[]  
-  \end{itemize}
-}
-
-\frame{\frametitle{Motivation}
-  \begin{itemize}
-    \item A multiheaded dependency corpus would be useful for testing new parsing algorithms
-    \item Such a corpus could be automatically annotated using Integer Linear Programming
-    \item We explored whether the Link Grammar could be adapted for this purpose.
-    \item The results of this are mixed, but provides a good case study.
-  \end{itemize}
-}
-
-\section{Multi-headed}
-
+%\subsection{Multi-headed}
 \frame{\frametitle{Multi-headedness}
+  Multi-headedness Can Capture Additional Linguistic Phenomenon
   \begin{itemize}
@@
 
-%\frame{\frametitle{Multi-headedness}
-%  \begin{figure}
-%    \begin{dependency}
-%      \begin{deptext}
-%      I \& think \& the \& man \& who \& you \& saw \& wants \& to \& eat \& sushi \\
-%      \end{deptext}
-%      \deproot[edge above, thick]{2}{ROOT}
-%      \depedge[edge above, thick, hide label]{2}{1}{} % think -> I
-%      \depedge[edge above, thick, hide label]{4}{3}{} % man -> the
-%      \depedge[edge above, thick, hide label]{4}{5}{} % man -> who
-%      \depedge[edge above, thick, hide label]{7}{6}{} % saw -> you
-%      \depedge[edge above, thick, hide label]{8}{4}{} % wants -> man
-%      \depedge[edge above, thick, hide label]{4}{7}{} % man -> saw
-%      \depedge[edge above, thick, hide label]{2}{8}{} % think -> wants
-%      \depedge[edge above, thick, hide label]{8}{10}{} % wants -> eat
-%      \depedge[edge above, thick, hide label]{10}{9}{} % eat -> to
-%      \depedge[edge above, thick, hide label]{10}{11}{} % eat -> sushi
-%      \depedge[edge below, edge style = {purple}, hide label, edge unit distance = 1.5ex, ultra thick]{10}{4}{} % eat -> man
-%    \end{dependency}
-%    \caption*{Example of a Multiheaded Parse}
-%  \end{figure}
-%}
+\subsection{Motivation}
+\frame{\frametitle{Motivation}
+  \begin{itemize}
+    \item A multiheaded dependency corpus would be useful for testing new parsing algorithms
+    \item[]  
+    \item[]  
+    \item[]  
+  \end{itemize}
+}
+
+\frame{\frametitle{Motivation}
+  \begin{itemize}
+    \item A multiheaded dependency corpus would be useful for testing new parsing algorithms
+    \item Such a corpus could be automatically annotated using Integer Linear Programming
+    \item[]  
+    \item[]  
+  \end{itemize}
+}
+
+\frame{\frametitle{Motivation}
+  \begin{itemize}
+    \item A multiheaded dependency corpus would be useful for testing new parsing algorithms
+    \item Such a corpus could be automatically annotated using Integer Linear Programming
+    \item We explored whether the Link Grammar could be adapted for this purpose.
+    \item[]  
+  \end{itemize}
+}
+
+\frame{\frametitle{Motivation}
+  \begin{itemize}
+    \item A multiheaded dependency corpus would be useful for testing new parsing algorithms
+    \item Such a corpus could be automatically annotated using Integer Linear Programming
+    \item We explored whether the Link Grammar could be adapted for this purpose.
+    \item The results of this are mixed, but provides a good case study.
+  \end{itemize}
+}
 
@@
 	\begin{deptext}
+	  - \& n-u \& v \& e \& e \& v \& v-d \& r \& n-u \& - \\
 	  the \& matter \& may \& never \& even \& be \& tried \& in \& court \& . \\
-	  - \& n-u \& v \& e \& e \& v \& v-d \& r \& n-u \& - \\
 	\end{deptext}
@@
 	\begin{deptext}
+	  - \& n-u \& v \& e \& e \& v \& v-d \& r \& n-u \& - \\
+	  the \& matter \& may \& never \& even \& be \& tried \& in \& court \& . \\
 	  {\scriptsize DT} \& {\scriptsize NN} \& {\scriptsize MD} \& {\scriptsize RB} \& {\scriptsize RB} \& {\scriptsize VB} \& {\scriptsize VB} \& {\scriptsize IN} \& {\scriptsize NN} \& {\scriptsize .} \\
-	  the \& matter \& may \& never \& even \& be \& tried \& in \& court \& . \\
-	  - \& n-u \& v \& e \& e \& v \& v-d \& r \& n-u \& - \\
 	\end{deptext}
-	\deproot[edge above, thick, edge style = {blue}]{3}{\small ROOT}
-	\depedge[edge below, thick, edge style = {red}]{2}{3}{S}
-	\depedge[edge above, thick, edge style = {blue}]{3}{2}{\small SBJ}
-	\depedge[edge above, thick, edge style = {blue}]{3}{4}{\small ADV}
-	\depedge[edge above, thick, edge style = {blue}]{3}{5}{\small ADV}
-	\depedge[edge below, thick, edge style = {red}]{3}{6}{I}
-	\depedge[edge above, thick, edge style = {blue}]{3}{6}{\small VC}
-	\depedge[edge above, thick, edge style = {blue}, edge unit distance =1.5ex]{3}{10}{\small P}
-	\depedge[edge below, thick, edge style = {red}]{2}{1}{D}
-	\depedge[edge above, thick, edge style = {blue}]{2}{1}{\small NMOD}
-	\depedge[edge below, thick, edge style = {red}]{7}{8}{MV}
-	\depedge[edge above, thick, edge style = {blue}]{7}{8}{\small ADV}
-	\depedge[edge below, thick, edge style = {red}]{6}{7}{P}
-	\depedge[edge above, thick, edge style = {blue}]{6}{7}{\small VC}
-	\depedge[edge below, thick, edge style = {red}]{8}{9}{J}
-	\depedge[edge above, thick, edge style = {blue}]{8}{9}{\small PMOD}
-	\deproot[edge below, thick, edge style = {red}]{2}{W}
-	\deproot[edge below, thick, edge style = {red}]{7}{WV}
-	\deproot[edge below, thick, edge style = {red}]{10}{X}
-	\depedge[edge below, thick, edge style = {red}]{6}{4}{E}
-	\depedge[edge below, thick, edge style = {red}]{6}{5}{E}
+	\deproot[edge below, thick, edge style = {blue}]{3}{\small ROOT}
+	\depedge[edge above, thick, edge style = {red}]{2}{3}{S}
+	\depedge[edge below, thick, edge style = {blue}]{3}{2}{\small SBJ}
+	\depedge[edge below, thick, edge style = {blue}]{3}{4}{\small ADV}
+	\depedge[edge below, thick, edge style = {blue}]{3}{5}{\small ADV}
+	\depedge[edge above, thick, edge style = {red}]{3}{6}{I}
+	\depedge[edge below, thick, edge style = {blue}]{3}{6}{\small VC}
+	\depedge[edge below, thick, edge style = {blue}, edge unit distance =1.5ex]{3}{10}{\small P}
+	\depedge[edge above, thick, edge style = {red}]{2}{1}{D}
+	\depedge[edge below, thick, edge style = {blue}]{2}{1}{\small NMOD}
+	\depedge[edge above, thick, edge style = {red}]{7}{8}{MV}
+	\depedge[edge below, thick, edge style = {blue}]{7}{8}{\small ADV}
+	\depedge[edge above, thick, edge style = {red}]{6}{7}{P}
+	\depedge[edge below, thick, edge style = {blue}]{6}{7}{\small VC}
+	\depedge[edge above, thick, edge style = {red}]{8}{9}{J}
+	\depedge[edge below, thick, edge style = {blue}]{8}{9}{\small PMOD}
+	\deproot[edge above, thick, edge style = {red}]{2}{W}
+	\deproot[edge above, thick, edge style = {red}]{7}{WV}
+	\deproot[edge above, thick, edge style = {red}]{10}{X}
+	\depedge[edge above, thick, edge style = {red}]{6}{4}{E}
+	\depedge[edge above, thick, edge style = {red}]{6}{5}{E}
   \end{dependency}
@@
 	\begin{deptext}
+	  - \& n-u \& v \& e \& e \& v \& v-d \& r \& n-u \& - \\
+	  the \& matter \& may \& never \& even \& be \& tried \& in \& court \& . \\
 	  {\scriptsize DT} \& {\scriptsize NN} \& {\scriptsize MD} \& {\scriptsize RB} \& {\scriptsize RB} \& {\scriptsize VB} \& {\scriptsize VB} \& {\scriptsize IN} \& {\scriptsize NN} \& {\scriptsize .} \\
-	  the \& matter \& may \& never \& even \& be \& tried \& in \& court \& . \\
-	  - \& n-u \& v \& e \& e \& v \& v-d \& r \& n-u \& - \\
 	\end{deptext}
-	\deproot[edge above, edge style = {blue, dotted}]{3}{\small ROOT}
-	\depedge[edge below, edge style = {red, ultra thick}]{2}{3}{S}
-	\depedge[edge above, edge style = {blue, ultra thick}]{3}{2}{\small SBJ}
-	\depedge[edge above, edge style = {blue, dotted}]{3}{4}{\small ADV}
-	\depedge[edge above, edge style = {blue, dotted}]{3}{5}{\small ADV}
-	\depedge[edge below, edge style = {red, thick}]{3}{6}{I}
-	\depedge[edge above, edge style = {blue, thick}]{3}{6}{\small VC}
-	\depedge[edge above, edge style = {blue, dotted}, edge unit distance =1.5ex]{3}{10}{\small P}
-	\depedge[edge below, edge style = {red, thick}]{2}{1}{D}
-	\depedge[edge above, edge style = {blue, thick}]{2}{1}{\small NMOD}
-	\depedge[edge below, edge style = {red, thick}]{7}{8}{MV}
-	\depedge[edge above, edge style = {blue, thick}]{7}{8}{\small ADV}
-	\depedge[edge below, edge style = {orange, thick}]{6}{7}{P}
-	\depedge[edge above, edge style = {blue, thick}]{6}{7}{\small VC}
-	\depedge[edge below, edge style = {red, thick}]{8}{9}{J}
-	\depedge[edge above, edge style = {blue, thick}]{8}{9}{\small PMOD}
-	\deproot[edge below, edge style = {red, dotted}]{2}{W}
-	\deproot[edge below, edge style = {orange, ultra thick, dotted}]{7}{WV}
-	\deproot[edge below, edge style = {red, dotted}]{10}{X}
-	\depedge[edge below, edge style = {red, dotted}]{6}{4}{E}
-	\depedge[edge below, edge style = {red, dotted}]{6}{5}{E}
+	\deproot[edge below, edge style = {blue, dotted}]{3}{\small ROOT}
+	\depedge[edge above, edge style = {red, ultra thick}]{2}{3}{S}
+	\depedge[edge below, edge style = {blue, ultra thick}]{3}{2}{\small SBJ}
+	\depedge[edge below, edge style = {blue, dotted}]{3}{4}{\small ADV}
+	\depedge[edge below, edge style = {blue, dotted}]{3}{5}{\small ADV}
+	\depedge[edge above, edge style = {red, thick}]{3}{6}{I}
+	\depedge[edge below, edge style = {blue, thick}]{3}{6}{\small VC}
+	\depedge[edge below, edge style = {blue, dotted}, edge unit distance =1.5ex]{3}{10}{\small P}
+	\depedge[edge above, edge style = {red, thick}]{2}{1}{D}
+	\depedge[edge below, edge style = {blue, thick}]{2}{1}{\small NMOD}
+	\depedge[edge above, edge style = {red, thick}]{7}{8}{MV}
+	\depedge[edge below, edge style = {blue, thick}]{7}{8}{\small ADV}
+	\depedge[edge above, edge style = {orange, thick}]{6}{7}{P}
+	\depedge[edge below, edge style = {blue, thick}]{6}{7}{\small VC}
+	\depedge[edge above, edge style = {red, thick}]{8}{9}{J}
+	\depedge[edge below, edge style = {blue, thick}]{8}{9}{\small PMOD}
+	\deproot[edge above, edge style = {red, dotted}]{2}{W}
+	\deproot[edge above, edge style = {orange, ultra thick, dotted}]{7}{WV}
+	\deproot[edge above, edge style = {red, dotted}]{10}{X}
+	\depedge[edge above, edge style = {red, dotted}]{6}{4}{E}
+	\depedge[edge above, edge style = {red, dotted}]{6}{5}{E}
   \end{dependency}
@@
 	\begin{deptext}
+	  - \& n-u \& v \& e \& e \& v \& v-d \& r \& n-u \& - \\
+	  the \& matter \& may \& never \& even \& be \& tried \& in \& court \& . \\
 	  {\scriptsize DT} \& {\scriptsize NN} \& {\scriptsize MD} \& {\scriptsize RB} \& {\scriptsize RB} \& {\scriptsize VB} \& {\scriptsize VB} \& {\scriptsize IN} \& {\scriptsize NN} \& {\scriptsize .} \\
-	  the \& matter \& may \& never \& even \& be \& tried \& in \& court \& . \\
-	  - \& n-u \& v \& e \& e \& v \& v-d \& r \& n-u \& - \\
 	\end{deptext}
-	\deproot[edge above, edge style = {blue, dotted}]{3}{\small ROOT}
-	\depedge[edge below, edge style = {red, ultra thick}]{2}{3}{S}
-	\depedge[edge above, edge style = {blue, ultra thick}]{3}{2}{\small SBJ}
-	\depedge[edge above, edge style = {blue, dotted}]{3}{4}{\small ADV}
-	\depedge[edge above, edge style = {blue, dotted}]{3}{5}{\small ADV}
-	\depedge[edge below, edge style = {red, thick}]{3}{6}{I}
-	\depedge[edge above, edge style = {blue, thick}]{3}{6}{\small VC}
-	\depedge[edge above, edge style = {blue, dotted}, edge unit distance =1.5ex]{3}{10}{\small P}
-	\depedge[edge below, edge style = {red, thick}]{2}{1}{D}
-	\depedge[edge above, edge style = {blue, thick}]{2}{1}{\small NMOD}
-	\depedge[edge below, edge style = {red, thick}]{7}{8}{MV}
-	\depedge[edge above, edge style = {blue, thick}]{7}{8}{\small ADV}
-	\depedge[edge below, edge style = {orange, thick}]{6}{7}{P}
-	\depedge[edge above, edge style = {blue, thick}]{6}{7}{\small VC}
-	\depedge[edge below, edge style = {red, thick}]{8}{9}{J}
-	\depedge[edge above, edge style = {blue, thick}]{8}{9}{\small PMOD}
-	\deproot[edge below, edge style = {red, dotted}]{2}{W}
-	\deproot[edge below, edge style = {orange, ultra thick, dotted}]{7}{WV}
-	\deproot[edge below, edge style = {red, dotted}]{10}{X}
-	\depedge[edge below, edge style = {red, dotted}]{6}{4}{E}
-	\depedge[edge below, edge style = {red, dotted}]{6}{5}{E}
+	\deproot[edge below, edge style = {blue, dotted}]{3}{\small ROOT}
+	\depedge[edge above, edge style = {red, ultra thick}]{2}{3}{S}
+	\depedge[edge below, edge style = {blue, ultra thick}]{3}{2}{\small SBJ}
+	\depedge[edge below, edge style = {blue, dotted}]{3}{4}{\small ADV}
+	\depedge[edge below, edge style = {blue, dotted}]{3}{5}{\small ADV}
+	\depedge[edge above, edge style = {red, thick}]{3}{6}{I}
+	\depedge[edge below, edge style = {blue, thick}]{3}{6}{\small VC}
+	\depedge[edge below, edge style = {blue, dotted}, edge unit distance =1.5ex]{3}{10}{\small P}
+	\depedge[edge above, edge style = {red, thick}]{2}{1}{D}
+	\depedge[edge below, edge style = {blue, thick}]{2}{1}{\small NMOD}
+	\depedge[edge above, edge style = {red, thick}]{7}{8}{MV}
+	\depedge[edge below, edge style = {blue, thick}]{7}{8}{\small ADV}
+	\depedge[edge above, edge style = {orange, thick}]{6}{7}{P}
+	\depedge[edge below, edge style = {blue, thick}]{6}{7}{\small VC}
+	\depedge[edge above, edge style = {red, thick}]{8}{9}{J}
+	\depedge[edge below, edge style = {blue, thick}]{8}{9}{\small PMOD}
+	\deproot[edge above, edge style = {red, dotted}]{2}{W}
+	\deproot[edge above, edge style = {orange, ultra thick, dotted}]{7}{WV}
+	\deproot[edge above, edge style = {red, dotted}]{10}{X}
+	\depedge[edge above, edge style = {red, dotted}]{6}{4}{E}
+	\depedge[edge above, edge style = {red, dotted}]{6}{5}{E}
   \end{dependency}
 \end{figure}
-
-\begin{itemize}
-\item[] %This is probably due to an inconsistency of the Link Grammar, discovered by our method.
-\end{itemize}
 
@@
 	\begin{deptext}
+	  - \& n-u \& v \& e \& e \& v \& v-d \& r \& n-u \& - \\
+	  the \& matter \& may \& never \& even \& be \& tried \& in \& court \& . \\
 	  {\scriptsize DT} \& {\scriptsize NN} \& {\scriptsize MD} \& {\scriptsize RB} \& {\scriptsize RB} \& {\scriptsize VB} \& {\scriptsize VB} \& {\scriptsize IN} \& {\scriptsize NN} \& {\scriptsize .} \\
-	  the \& matter \& may \& never \& even \& be \& tried \& in \& court \& . \\
-	  - \& n-u \& v \& e \& e \& v \& v-d \& r \& n-u \& - \\
 	\end{deptext}
-	\deproot[edge above, edge style = {gray}, hide label]{3}{\small ROOT}
-	\depedge[edge below, edge style = {red, ultra thick}]{2}{3}{S}
-	\depedge[edge above, edge style = {blue, ultra thick}]{3}{2}{\small SBJ}
-	\depedge[edge above, edge style = {gray}, hide label]{3}{4}{\small ADV}
-	\depedge[edge above, edge style = {gray}, hide label]{3}{5}{\small ADV}
-	\depedge[edge below, edge style = {gray}, hide label]{3}{6}{I}
-	\depedge[edge above, edge style = {gray}, hide label]{3}{6}{\small VC}
-	\depedge[edge above, , edge style = {gray}, hide label, edge unit distance =1.5ex]{3}{10}{\small P}
-	\depedge[edge below, edge style = {gray}, hide label]{2}{1}{D}
-	\depedge[edge above, edge style = {gray}, hide label]{2}{1}{\small NMOD}
-	\depedge[edge below, edge style = {gray}, hide label]{7}{8}{MV}
-    \depedge[edge above, edge style = {gray}, hide label]{7}{8}{\small ADV}
-	\depedge[edge below, edge style = {gray}, hide label]{6}{7}{P}
-	\depedge[edge above, edge style = {gray}, hide label]{6}{7}{\small VC}
-	\depedge[edge below, edge style = {gray}, hide label]{8}{9}{J}
-	\depedge[edge above, edge style = {gray}, hide label]{8}{9}{\small PMOD}
-	\deproot[edge below, edge style = {gray}, hide label]{2}{W}
-    \deproot[edge below, edge style = {gray}, hide label]{7}{WV}
-	\deproot[edge below, edge style = {gray}, hide label]{10}{X}
-	\depedge[edge below, edge style = {gray}, hide label]{6}{4}{E}
-	\depedge[edge below, edge style = {gray}, hide label]{6}{5}{E}
+	\deproot[edge below, edge style = {gray}, hide label]{3}{\small ROOT}
+	\depedge[edge above, edge style = {red, ultra thick}]{2}{3}{S}
+	\depedge[edge below, edge style = {blue, ultra thick}]{3}{2}{\small SBJ}
+	\depedge[edge below, edge style = {gray}, hide label]{3}{4}{\small ADV}
+	\depedge[edge below, edge style = {gray}, hide label]{3}{5}{\small ADV}
+	\depedge[edge above, edge style = {gray}, hide label]{3}{6}{I}
+	\depedge[edge below, edge style = {gray}, hide label]{3}{6}{\small VC}
+	\depedge[edge below, , edge style = {gray}, hide label, edge unit distance =1.5ex]{3}{10}{\small P}
+	\depedge[edge above, edge style = {gray}, hide label]{2}{1}{D}
+	\depedge[edge below, edge style = {gray}, hide label]{2}{1}{\small NMOD}
+	\depedge[edge above, edge style = {gray}, hide label]{7}{8}{MV}
+    \depedge[edge below, edge style = {gray}, hide label]{7}{8}{\small ADV}
+	\depedge[edge above, edge style = {gray}, hide label]{6}{7}{P}
+	\depedge[edge below, edge style = {gray}, hide label]{6}{7}{\small VC}
+	\depedge[edge above, edge style = {gray}, hide label]{8}{9}{J}
+	\depedge[edge below, edge style = {gray}, hide label]{8}{9}{\small PMOD}
+	\deproot[edge above, edge style = {gray}, hide label]{2}{W}
+    \deproot[edge above, edge style = {gray}, hide label]{7}{WV}
+	\deproot[edge above, edge style = {gray}, hide label]{10}{X}
+	\depedge[edge above, edge style = {gray}, hide label]{6}{4}{E}
+	\depedge[edge above, edge style = {gray}, hide label]{6}{5}{E}
   \end{dependency}
 \end{figure}
+
+}
+
+\frame{\frametitle{Link Results: Subject-Verb links are backwards}
 
@@
 \end{itemize}
+
+\begin{figure}
+  \begin{dependency}
+	\begin{deptext}
+	  Jill \& wanted \& to \& skip \\
+	\end{deptext}
+    \deproot[edge above, edge unit distance=2ex]{1}{W}
+    \deproot[edge above, edge unit distance=2ex]{2}{WV}
+	\depedge[edge above]{2}{1}{S}
+	\depedge[edge above]{2}{3}{TO}
+	\depedge[edge above]{3}{4}{I}
+	\depedge[edge above]{2}{4}{IV}
+  \end{dependency}
+\end{figure}
+
+\begin{figure}
+  \begin{dependency}
+	\begin{deptext}
+	  Jill \& persuaded \& him \& to \& skip \\
+	\end{deptext}
+    \deproot[edge above, edge unit distance=2ex]{1}{W}
+    \deproot[edge above, edge unit distance=2ex]{2}{WV}
+	\depedge[edge above]{2}{1}{S}
+	\depedge[edge above]{2}{3}{O}
+	\depedge[edge above]{2}{4}{TO}
+	\depedge[edge above]{4}{5}{I}
+	\depedge[edge above]{2}{5}{IV}
+  \end{dependency}
+\end{figure}
 

### 2014-12-09 doc/slides/hong+eisner.TLT13.slides.tex
cleaned up the animations in the presentation

@@
   \begin{itemize}
-    \item A multiheaded dependency corpus would be useful for testing new parsing algorithms
-    \item[]  
-    \item[]  
-    \item[]  
-  \end{itemize}
-}
-
-\frame{\frametitle{Motivation}
-  \begin{itemize}
-    \item A multiheaded dependency corpus would be useful for testing new parsing algorithms
-    \item Such a corpus could be automatically annotated using Integer Linear Programming
-    \item[]  
-    \item[]  
-  \end{itemize}
-}
-
-\frame{\frametitle{Motivation}
-  \begin{itemize}
-    \item A multiheaded dependency corpus would be useful for testing new parsing algorithms
-    \item Such a corpus could be automatically annotated using Integer Linear Programming
-    \item We explored whether the Link Grammar could be adapted for this purpose.
-    \item[]  
-  \end{itemize}
-}
-
-\frame{\frametitle{Motivation}
-  \begin{itemize}
-    \item A multiheaded dependency corpus would be useful for testing new parsing algorithms
-    \item Such a corpus could be automatically annotated using Integer Linear Programming
-    \item We explored whether the Link Grammar could be adapted for this purpose.
-    \item The results of this are mixed, but provides a good case study.
+    \uncover<1->{\item A multiheaded dependency corpus would be useful for testing new parsing algorithms}
+    \uncover<2->{\item Such a corpus could be automatically annotated using Integer Linear Programming}
+    \uncover<3->{\item We explored whether the Link Grammar could be adapted for this purpose.}
+    \uncover<4->{\item The results of this are mixed, but provides a good case study.}
   \end{itemize}
@@
   \begin{itemize}
-  \item[]  
-  \end{itemize}
-}
-
-\frame{\frametitle{ILP Results}
-  \begin{itemize}
-    \item[] Link types that allowed both directions:
-      \begin{itemize}
-      \item[] 7 / 113 = 6.19\%
-      %\item[] $\frac{7}{113} = 6.19\%$ 
-      \end{itemize}
-    \item[] Link tokens that required disallowed direction via slack:
-      \begin{itemize}
-        %\item[] $\frac{4043}{195000} = 2.07\%$ 
-        \item[] 4043 / 195,000 = 2.07\%
-      \end{itemize}
-  \end{itemize}
-  \begin{itemize}
-  \item[] The link labels mostly have a consistent direction.
+    \uncover<1>{\item[]}
+    \uncover<2->{\item[] The link labels mostly have a consistent direction.}
   \end{itemize}

### 2014-12-09 doc/slides/hong+eisner.TLT13.slides.tex
improved animations

@@
 
-%\setbeamerfont{page number in head/foot}{}
-%\setbeamertemplate{footline}[frame number]
+\setbeamerfont{page number in head/foot}{}
+\setbeamertemplate{footline}[frame number]
 
@@
 \date{\today}
-\frame{\titlepage \NoteJH{Include pictures of us?}}
+\frame{\titlepage} %\NoteJH{Include pictures of us?}
 
@@
 
-\frame{\frametitle{Consistency of Directionalized Links}
+\frame{\frametitle{Consistency of Directionalized Links \onslide<2->{with Slack}}
 \begin{itemize}
@@
     \end{itemize}
+
     \begin{align}\label{direction+slack}
-      x_{ij} &\leq r_L &
-      x_{ji} &\leq \ell_L 
+      x_{ij} &\leq r_L \onslide<2->{ + s_{ij}} &
+      x_{ji} &\leq \ell_L \onslide<2->{ + s_{ij}}
     \end{align}
+
     \begin{align}\label{eqn:obj}
-      \min \left( \sum_L r_L + \ell_L \right)
+      \min \left( \sum_L r_L + \ell_L \right) \onslide<2->{\cdot \frac{N_L}{4} + \sum_{ij}s_{ij}}
     \end{align}
-    \begin{itemize}
-      \item[]  
-      \item[]  
-    \end{itemize}
 
-    \begin{itemize}
-      \item[]
-    \end{itemize}
-
-\end{itemize}
-}
-
-\frame{\frametitle{Consistency of Directionalized Links with Slack}
-\begin{itemize}
-  \item[] Consistency of Directionalized Links
-    \begin{itemize}
-      \item[] $r_L, \ell_L \in \{0,1\}$: whether links with label $L$ allowed left/right
-    \end{itemize}
-    \begin{align}\label{direction+slack}
-      x_{ij} &\leq r_L + s_{ij} &
-      x_{ji} &\leq \ell_L + s_{ij} 
-    \end{align}
-    \begin{align}\label{eqn:obj}
-      \min \left( \sum_L r_L + \ell_L \right) \cdot \frac{N_L}{4} + \sum_{ij}s_{ij}
-    \end{align}
-    \begin{itemize}
+    \onslide<2->{\begin{itemize}
       \item[] $s_{ij} \in \mathbb{R} \geq 0$: slack variable 
@@
     \end{itemize}
+    }
 

### 2014-12-10 doc/slides/tableprep.py
updated slides. Added link analysis table from the appendix (I had to cut it down). Added the stability-of-results graph.

@@ -1,0 +1,56 @@
+
+
+filename = "link_analysis_coarse_table.tex"
+newfilename = "link_table_filtered.tex"
+newfilename2 = "link_table_filtered_part2.tex"
+
+
+f = open(filename)
+
+new = open(newfilename, "w")
+new2 = open(newfilename2, "w")
+
+i = 0
+newlines = 0
+
+for line in f:
+    i += 1
+
+    if i < 12:
+        if "tabular" in line:
+            new.write("\\begin{tabular}{|l|l|l|l|l|l|}\n")
+            new2.write("\\begin{tabular}{|l|l|}\n")
+        elif "Label" in line:
+            new.write("Label & Rightward & Multiheaded & CoNLL Match & CoNLL Dir Match\\\\\\hline\n")
+            new2.write("Label & CoNLL Label\\\\\\hline\n")
+        else:
+            new.write(line)
+            new2.write(line)
+        continue
+
+    if not i < 123:
+        new.write(line)
+        new2.write(line)
+        continue
+
+
+
+    line = line.strip()
+    line = line.split("&")
+
+    total_num = int(line[2].split()[1].replace("(","").replace(")","").split("/")[1])
+
+    if total_num > 1450:
+        
+        
+        line2 = [line[0], line[-1]]
+        line = line[:-1]
+
+        line = "&".join(line) + " \\\\"
+        line2 = "&".join(line2)
+
+        print >> new, line
+        print >> new2, line2
+        newlines += 1
+        #print newlines, total_num, line
+        

### 2014-12-10 doc/slides/hong+eisner.TLT13.slides.tex
updated slides. Added link analysis table from the appendix (I had to cut it down). Added the stability-of-results graph.
cleaned up/improved the summarizing statistics slide.

@@
 \usepackage{subcaption}
+\usepackage{longtable}
 
@@
     \item[] Grammar-based formalism for projective dependency parsing with undirected links
-    %\item We set out to discover if the undirected links are actually directed relationships.
   \end{itemize}
@@
   \end{figure}
-
 }
@@
 %\end{itemize}
-  \begin{table}[h]
+
+  Data Sets taken from:
+  \begin{itemize}
+    \item[] CoNLL 2007 Shared Task (English)
+    \item[] ACL 2013 Shared Task of Machine Translation (Russian)
+  \end{itemize}
+
+  \begin{figure}[h]
+    \centering
     \begin{tabular}{|l|l|l|}
       \hline
-      & \# Sentences & \# Connected Parses \\ \hline
+       & Input Sentences & Output Connected Parses \\ \hline
        English  & 18,577        & 10,960               \\ \hline
-      Russian (unlabeled) & 18,577        & 4,913                \\ \hline
+       Russian & 18,577        & 4,913                \\ \hline
     \end{tabular}
-  \end{table}
+  \end{figure}
+}
 
-  On the English data set:
-
-  \begin{itemize}
-  \item[] Link Data has 8\% additional edges over the CoNLL.
-  \item[] 52\% of links match CoNLL arcs
-  \item[] 57\% of CoNLL arcs match links
-  \end{itemize}
-
-  \NoteJH{I skipped the ``Stability of Results'' section}
+\frame{\frametitle{Stability of Results}
+  \begin{figure}
+    \centering
+    \includegraphics[width=0.7\linewidth]{precision_recall}
+  \end{figure}
 }
@@
 
-\frame{\frametitle{ILP Results}
-%  \begin{table}[h]
-%    \begin{tabular}{|l|l||l|}
-%      \hline
-%      & & out of total \\ \hline
-%      Link types & & \\allowing both directions & 7 & 113 \\ \hline
-%      Link tokens & & \\requiring disallowed direction & & \\via slack & 4043 & 195,000 \\ \hline
-%    \end{tabular}
-%  \end{table}
+\frame{\frametitle{On the English Data Set:}
+
+  \uncover<2->{Multiheadedness}
   \begin{itemize}
-    \item[] Link types that allowed both directions:
+  \uncover<2->{\item[] Link Data has 8\% additional edges over the CoNLL.}
+  \end{itemize}
+
+  \uncover<3->{CoNLL Matches}
   \begin{itemize}
-        \item[] 7 / 113 = 6.19\%
+  \uncover<3->{\item[] 52\% of links match CoNLL arcs
+  \item[] 57\% of CoNLL arcs match links}
   \end{itemize}
-    \item[] Link tokens that required disallowed direction via slack:
+
+  \uncover<4->{Directionality}%\uncover<5->{ is Mostly Consistent!}
   \begin{itemize}
-      \item[] 4043 / 195,000 = 2.07\%
+  \uncover<4->{\item[] 6.19\% of link types allowed both directions %($\frac{7}{113}$)
+  \item[] 2.07\% of link tokens required disallowed direction via slack %($\frac{4043}{195,000}$)
+  }
   \end{itemize}
-  \end{itemize}
-  \begin{itemize}
-    \uncover<1>{\item[]}
-    \uncover<2->{\item[] The link labels mostly have a consistent direction.}
-  \end{itemize}
+
+  %\uncover<4->{The link labels mostly have a consistent direction.}
+  %\uncover<5->{\item[] The direction is often the same as CoNLL when there is a corresponding arc}
+
 }
+
+\frame{\frametitle{ILP Results: Top 25 Most Occuring Labels}
+
+  \begin{figure}
+    \centering
+    \input{link_table_filtered.tex}
+  \end{figure}
+}
+
+%\frame{\frametitle{ILP Results: Top 25 Most Occuring Labels}
+%  \input{link_table_filtered_part2.tex}
+%}
 
@@
 
-\NoteJH{Should I include a section of the appendix table? Which rows should I include?}
 }

### 2014-12-12 doc/slides/hong+eisner.TLT13.slides.tex
slides so far.
slides so far. Added a new intro slide with visualization. Still discussing with Jason what to do with the reversed subject arc.

@@
 \section{Motivation}
+
+\frame{\frametitle{Single-headedness}
+  Dependency parse treebanks today are single-headed.
+
+\begin{figure}
+  \begin{dependency}
+	\begin{deptext}
+	  {\scriptsize DT} \& {\scriptsize NN} \& {\scriptsize MD} \& {\scriptsize RB} \& {\scriptsize RB} \& {\scriptsize VB} \& {\scriptsize VB} \& {\scriptsize IN} \& {\scriptsize NN} \& {\scriptsize .} \\
+	  the \& matter \& may \& never \& even \& be \& tried \& in \& court \& . \\
+	\end{deptext}
+	\deproot[edge above, thick, edge style = {blue}]{3}{\small ROOT}
+	\depedge[edge above, thick, edge style = {blue}]{3}{2}{\small SBJ}
+	\depedge[edge above, thick, edge style = {blue}]{3}{4}{\small ADV}
+	\depedge[edge above, thick, edge style = {blue}]{3}{5}{\small ADV}
+	\depedge[edge above, thick, edge style = {blue}]{3}{6}{\small VC}
+	\depedge[edge above, thick, edge style = {blue}, edge unit distance =1.5ex]{3}{10}{\small P}
+	\depedge[edge above, thick, edge style = {blue}]{2}{1}{\small NMOD}
+	\depedge[edge above, thick, edge style = {blue}]{7}{8}{\small ADV}
+	\depedge[edge above, thick, edge style = {blue}]{6}{7}{\small VC}
+	\depedge[edge above, thick, edge style = {blue}]{8}{9}{\small PMOD}
+  \end{dependency}
+  \caption*{Some example dependency parse.}
+\end{figure}
+
+}
 
@@
 \begin{itemize}
-\item This is due to an inconsistency of the Link Grammar, discovered by our method. 
+\item In our paper we claimed this was due to an inconsistency of the Link Grammar, discovered by our method.
+
 \end{itemize}
@@
 
+%\begin{figure}
+%  \begin{dependency}
+%	\begin{deptext}
+%	  Jill \& persuaded \& him \& to \& skip \\
+%	\end{deptext}
+%    \deproot[edge above, edge unit distance=2ex]{1}{W}
+%    \deproot[edge above, edge unit distance=2ex]{2}{WV}
+%	\depedge[edge above]{2}{1}{S}
+%	\depedge[edge above]{2}{3}{O}
+%	\depedge[edge above]{2}{4}{TO}
+%	\depedge[edge above]{4}{5}{I}
+%	\depedge[edge above]{2}{5}{IV}
+%  \end{dependency}
+%\end{figure}
+
 \begin{figure}
@@
 	\begin{deptext}
-	  Jill \& persuaded \& him \& to \& skip \\
+	  Jill \& thinks \& he \& will \& skip \\
 	\end{deptext}
@@
     \deproot[edge above, edge unit distance=2ex]{2}{WV}
-	\depedge[edge above]{2}{1}{S}
-	\depedge[edge above]{2}{3}{O}
-	\depedge[edge above]{2}{4}{TO}
+	\depedge[edge above]{1}{2}{S}
+	\depedge[edge above]{2}{3}{C}
+	\depedge[edge above]{3}{4}{S}
 	\depedge[edge above]{4}{5}{I}
-	\depedge[edge above]{2}{5}{IV}
+	\depedge[edge above]{2}{5}{CV}
   \end{dependency}

### 2014-12-13 doc/slides/hong+eisner.TLT13.slides.tex
Changed the slide discussing the reversed subject links
Edited the slides with about half of Jasons suggestions. The other half still coming.

@@
 \usepackage{longtable}
+\usepackage{pdfpages}
 
@@
       \depedge[edge above, thick, hide label]{4}{5}{}
-      \depedge[edge below, ultra thick, hide label, edge style = {purple}, edge unit distance = 1.5ex]{3}{5}{}
+      \depedge[edge below, ultra thick, hide label, edge style = {purple}, edge unit distance = 1.5ex]{5}{3}{}
+      \depedge[edge below, ultra thick, hide label, edge style = {purple}, edge unit distance = 1.5ex]{2}{5}{}
     \end{dependency}
@@
   \item[] Grammar-based formalism for projective dependency parsing with undirected links
+  \item[] Orignal formalism and English Link Grammar created by Davy Temperley, Daniel Sleator, John Lafferty (1991)
   \end{itemize}
+}
+
+\subsection{Link Grammars Intro}
+\frame{\frametitle{Link Grammars}
+  \includegraphics[width=\linewidth]{linkgrammar_example1.png}\footnote{These figures were clipped from the orginal Link Grammar paper: \\``Parsing English with a Link Grammar'' by Sleator and Temperley}
+}
+
+\frame{\frametitle{Link Grammars}
+  \includegraphics[width=\linewidth]{linkgrammar_example2.png}
+}
+
+\frame{\frametitle{Link Grammars}
+  \includegraphics[width=\linewidth]{linkgrammar_example3.png}
 }
@@
 
+\subsection{What is ILP?}
+\frame{\frametitle{What is Integer Linear Programming?}
+\begin{itemize}
+  \item Its an optimization problem where some or all of the variables are integers. 
+  \item The objective function and constraints are linear.
+  \item Its NP-Hard! But good solvers exist that work well most of the time.
+  \item Our ILP is encoded as a ZIMPL program and solved using the SCIP Optimization Suite
+\end{itemize}
+}
+
+\subsection{Corpus Building}
+\frame{\frametitle{Corpus Building Strategy}
+\begin{itemize}
+  \item We start with a corpus of link grammar parses.
+  \item We try to directionalize this corpus, choosing a direction for every link in every parse.
+  \item We use an Integer Linear Program to find consistent directions for every link
+\end{itemize}
+}
+
+\subsection{ILP Model}
 \frame{\frametitle{Integer Linear Programming Model}
-
 Encoded Constraints:
 \begin{itemize}
-  \item Connectedness
-  \item Acyclicity
-  \item Consistency of Directionalized Links
+  \item Connectedness\uncover<2->{: Every word has a link attaching it}
+  \item Acyclicity\uncover<2->{: No cycles!}
+  \item Consistency of Directionalized Links\uncover<2->{:
+
+    Similar links oriented the same way}
 \end{itemize}
@@
 \item[] $x_{ij} + x_{ji} = 1$
+\end{itemize}
+\uncover<2->{\begin{center}
+    A link can only be oriented left or right
+\end{center}}
 
-%\item $x_{ij} + x_{ji} = 1$ (A link can only be oriented left or right)
-
-\end{itemize}
-
-%\NoteJH{TODO: Incomplete slide. Need to describe things like depth of the variable, label}
 }
@@
     \end{align}
+    \uncover<2->{\begin{center}
+        A word has at least 1 parent
+    \end{center}}
 
@@
     \end{align}
+    \uncover<2->{\begin{center}
+        The depth of a child is greater than the depth of the parent
+    \end{center}}
 \end{itemize}
@@
 \frame{\frametitle{Stability of Results}
+
+  \begin{itemize}
+  \item We worried that the recovered direction mapping might be unstable and sensitive to the input corpus. 
+  \item We compared the results of increasing runs of sentences.
+  \end{itemize}
+
   \begin{figure}
     \centering
-    \includegraphics[width=0.7\linewidth]{precision_recall}
+    \includegraphics[width=0.6\linewidth]{precision_recall}
   \end{figure}
@@
   \begin{itemize}
-  \uncover<2->{\item[] Link Data has 8\% additional edges over the CoNLL.}
+  \uncover<2->{\item[] Link Data has 8\% additional edges over the CoNLL.
+
+    (average about 2 multiheaded words per sentence)}
   \end{itemize}
@@
 \begin{itemize}
-\item In our paper we claimed this was due to an inconsistency of the Link Grammar, discovered by our method.
-
+\item This is due to a possible inconsistency of the Link Grammar, discovered by our method. 
 \end{itemize}
 
-\begin{figure}
-  \begin{dependency}
-	\begin{deptext}
-	  Jill \& wanted \& to \& skip \\
-	\end{deptext}
-    \deproot[edge above, edge unit distance=2ex]{1}{W}
-    \deproot[edge above, edge unit distance=2ex]{2}{WV}
-	\depedge[edge above]{2}{1}{S}
-	\depedge[edge above]{2}{3}{TO}
-	\depedge[edge above]{3}{4}{I}
-	\depedge[edge above]{2}{4}{IV}
-  \end{dependency}
-\end{figure}
-
-%\begin{figure}
-%  \begin{dependency}
-%	\begin{deptext}
-%	  Jill \& persuaded \& him \& to \& skip \\
-%	\end{deptext}
-%    \deproot[edge above, edge unit distance=2ex]{1}{W}
-%    \deproot[edge above, edge unit distance=2ex]{2}{WV}
-%	\depedge[edge above]{2}{1}{S}
-%	\depedge[edge above]{2}{3}{O}
-%	\depedge[edge above]{2}{4}{TO}
-%	\depedge[edge above]{4}{5}{I}
-%	\depedge[edge above]{2}{5}{IV}
-%  \end{dependency}
-%\end{figure}
+%    +-------WV-------+--------CV-------+
+%    +---Wd---+---Ss--+--Ce-+-Ss-+---I--+
+%    |        |       |     |    |      |
+%LEFT-WALL Jill.f thinks.v he will.v skip.v
 
 \begin{figure}
-  \begin{dependency}
+\vspace{-6pt}
+\centering
+  \begin{dependency}[edge style={-}]
 	\begin{deptext}
-	  Jill \& thinks \& he \& will \& skip \\
+	  Jill \& thinks \& he \& will \& skip\\
 	\end{deptext}
-    \deproot[edge above, edge unit distance=2ex]{1}{W}
-    \deproot[edge above, edge unit distance=2ex]{2}{WV}
+    %\deproot[edge above, edge unit distance=2ex]{1}{W}
+    %\deproot[edge above, edge unit distance=2ex]{2}{WV}
 	\depedge[edge above]{1}{2}{S}
@@
   \end{dependency}
+  %\caption*{Links with label I is forced to go right}
+\end{figure}
+
+%    +--------------WV-------------+
+%    +---Wd---+---Ss--+--MVi-+--I--+
+%    |        |       |      |     |
+%LEFT-WALL Jill.f thinks.v to.r skip.v
+
+\begin{figure}
+\vspace{-6pt}
+\centering
+  \begin{dependency}[edge style={-}]
+	\begin{deptext}
+	  Jill \& hopes \& to \& skip\\ % Originally: Jill thinks to skip. The sentence was altered to be clearer.
+	\end{deptext}
+    %\deproot[edge above, edge unit distance=2ex]{1}{W}
+    %\deproot[edge above, edge unit distance=2ex]{2}{WV}
+	\depedge[edge above]{1}{2}{S}
+	\depedge[edge above]{2}{3}{MV}
+	\depedge[edge above]{3}{4}{I}
+  \end{dependency}
+  %\caption*{Links with label I is forced to go right}
 \end{figure}
@@
 }
+
+%\frame{\frametitle{Possible simpler explanation}
+
+%linkparser> 
+%linkparser> Jill made him go
+%Found 5 linkages (3 had no P.P. violations)
+%Linkage 1, cost vector = (UNUSED=0 DIS= 0.00 LEN=6)
+%
+%    +-------WV-------+----I*j---+
+%    +---Wd---+---Ss--+--Ox-+    |
+%    |        |       |     |    |
+%LEFT-WALL Jill.f made.v-d him go.v
+
+%\begin{figure}
+%\vspace{-6pt}
+%\centering
+%  \begin{dependency}
+%	\begin{deptext}
+%	  Jill \& made \& him \& go \\
+%	\end{deptext}
+%    \deproot[edge above, edge unit distance=2ex]{1}{W}
+%    \deproot[edge above, edge unit distance=2ex]{2}{WV}
+%	\depedge[edge above]{1}{2}{S}
+%	\depedge[edge above]{2}{4}{I}
+%	\depedge[edge above]{2}{3}{O}
+%  \end{dependency}
+%  \caption*{Links with label I is forced to go right}
+%\end{figure}
+
+%\begin{figure}
+%  \vspace{-12pt}
+%  \begin{dependency}
+%	\begin{deptext}
+%	  Jill \& would \& go \\
+%	\end{deptext}
+%    \deproot[edge above, edge unit distance=2ex]{1}{W}
+%    \deproot[edge above, edge unit distance=2ex]{3}{WV}
+%	\depedge[edge above]{1}{2}{S}
+%	\depedge[edge above]{2}{3}{I}
+%  \end{dependency}
+%  \caption*{If I is forced to go right, then S is as well}
+%\end{figure}
+%}
 

### 2014-12-13 doc/slides/hong+eisner.TLT13.slides.tex
revamped intro slides as per Jasons suggestions.

@@
 
-\section{Motivation}
+\section{Introduction}
+\frame{\frametitle{Introduction}
+\begin{itemize}
+\item This talk is about automatically converting from one annotation style to another.
+\item The conversion could be hard, where information is fragmented or missing.
+
+%This is often useful.  You may want to do it yourself.  
+\item We use a general technique, Integer Linear Programming to recover this missing information.
+\end{itemize}
+}
+
+\frame{\frametitle{In Our Case: What We Started With}
+  \begin{figure}
+    \begin{dependency}[edge style={-}]
+	  \begin{deptext}
+	    - \& n-u \& v \& e \& e \& v \& v-d \& r \& n-u \& - \\
+	    the \& matter \& may \& never \& even \& be \& tried \& in \& court \& . \\
+	  \end{deptext}
+	  \depedge[edge above, thick, edge style = {red}]{2}{3}{S}
+	  \depedge[edge above, thick, edge style = {red}]{3}{6}{I}
+	  \depedge[edge above, thick, edge style = {red}]{2}{1}{D}
+	  \depedge[edge above, thick, edge style = {red}]{7}{8}{MV}
+	  \depedge[edge above, thick, edge style = {red}]{6}{7}{P}
+	  \depedge[edge above, thick, edge style = {red}]{8}{9}{J}
+	  \deproot[edge above, thick, edge style = {red}]{2}{W}
+	  \deproot[edge above, thick, edge style = {red}]{7}{WV}
+	  \deproot[edge above, thick, edge style = {red}]{10}{X}
+	  \depedge[edge above, thick, edge style = {red}]{6}{4}{E}
+	  \depedge[edge above, thick, edge style = {red}]{6}{5}{E}
+    \end{dependency}
+    \caption*{Link Grammar: Parse with undirected edges}
+  \end{figure}
+}
+
+\frame{\frametitle{What We Want:}
+  \begin{figure}
+    \begin{dependency}
+	  \begin{deptext}
+	    - \& n-u \& v \& e \& e \& v \& v-d \& r \& n-u \& - \\
+	    the \& matter \& may \& never \& even \& be \& tried \& in \& court \& . \\
+	  \end{deptext}
+	  \depedge[edge above, thick, edge style = {red}]{2}{3}{S}
+	  \depedge[edge above, thick, edge style = {red}]{3}{6}{I}
+	  \depedge[edge above, thick, edge style = {red}]{2}{1}{D}
+	  \depedge[edge above, thick, edge style = {red}]{7}{8}{MV}
+	  \depedge[edge above, thick, edge style = {red}]{6}{7}{P}
+	  \depedge[edge above, thick, edge style = {red}]{8}{9}{J}
+	  \deproot[edge above, thick, edge style = {red}]{2}{W}
+	  \deproot[edge above, thick, edge style = {red}]{7}{WV}
+	  \deproot[edge above, thick, edge style = {red}]{10}{X}
+	  \depedge[edge above, thick, edge style = {red}]{6}{4}{E}
+	  \depedge[edge above, thick, edge style = {red}]{6}{5}{E}
+    \end{dependency}
+    \caption*{Multiheaded parse with directionalized edges}
+  \end{figure}
+
+}
+
+\frame{\frametitle{Why We Want That}
+  \begin{itemize}
+    \item We want to develop parsing algorithms for parses that look like this
+    \item We couldn't figure out where to get the data to test them.
+  \end{itemize}
+}
 
 \frame{\frametitle{Single-headedness}
-  Dependency parse treebanks today are single-headed.
+  \begin{itemize}
+  \item Dependency parse treebanks today are either single-headed or not planar.
+  \item Stanford Dependencies are multiheaded but not planar
+  \end{itemize}
 
@@
 
+\uncover<2->{\begin{center}
+    Link Grammar is almost a multiheaded planar corpora! We just need to directionalize the links.
+\end{center}}
 }
 
+\section{Motivation}
 %\subsection{Multi-headed}
-\frame{\frametitle{Multi-headedness}
+\frame{\frametitle{Why Multi-headedness?}
   Multi-headedness Can Capture Additional Linguistic Phenomenon
@@
 
-\frame{\frametitle{Link Grammars: Example Parse}
+\frame{\frametitle{Link Grammars: Same Example Parse Again}
   \begin{figure}
@@
 
-\frame{\frametitle{Link Grammars: Converting into a Directed Acyclic Graph}
+\frame{\frametitle{Link Grammars: Converting into a Directed Acyclic Graph Again}
   \begin{figure}
@@
 
+\frame{\frametitle{Link Results: Subject-Verb links are backwards}
+\begin{itemize}
+      %\item Different dependency formalisms disagree on whether a clause is headed by the aux verb or the main verb.
+  \item The Link Grammar seems to be inconsistent about whether the auxiliary verb or the main verb is the head of a clause. 
+  \item Sometimes the governing verb links to the auxilliary, and sometimes to the main, depending on the type of clause. 
+  \item But the governing verb usually links to the subject when there is one.
+  \item So this makes the subject a consistent choice to make the head of a clause.
+\end{itemize}
+
+%governing verb -\> subject -\> auxilliary verb -\> main
+
+\begin{center}
+  To fix this, we could edit the link grammar, link parses, or the ILP. %\\ Our technique is only as useful as the input given.
+\end{center}
+
+    %Question: Which do we pick?
+    %Answer: Neither! We actually pick the subject as the head of the clause! 
+    %Question: Why??
+    %Answer:
+    %    The LG folks seem to be inconsistent within their grammar about whether the aux verb or the main verb is the head of a clause. 
+    %        Sometimes the governing verb links to aux, and sometimes to main, depending on the type of clause.
+    %    But the governing verb usually links to the subject when there is one!
+    %    Therefore, the subject ends up being a consistent choice for the head of the clause.  We get governing_verb --> subject --> aux --> main as our annotation scheme.
+    %Question: Couldn't this be fixed?
+    %Answer: Of course it could.  You could edit the link grammar or the link parses or the ILP.  Our technique is only as useful as the input you give it ...  
+}
+
 \section{Conclusions}

### 2014-12-13 doc/slides/hong+eisner.TLT13.slides.tex
Edited the slides some more.

@@
 \subsection{Link Grammars Intro}
-\frame{\frametitle{Link Grammars}
+\frame{\frametitle{Link Grammars: How They Work}
   \includegraphics[width=\linewidth]{linkgrammar_example1.png}\footnote{These figures were clipped from the orginal Link Grammar paper: \\``Parsing English with a Link Grammar'' by Sleator and Temperley}
@@
 
-\frame{\frametitle{Link Grammars}
+\frame{\frametitle{Link Grammars: How They Work}
   \includegraphics[width=\linewidth]{linkgrammar_example2.png}
@@
 
-\frame{\frametitle{Link Grammars}
+\frame{\frametitle{Link Grammars: How They Work}
   \includegraphics[width=\linewidth]{linkgrammar_example3.png}
@@
 
-\frame{\frametitle{Link Grammars: Same Example Parse Again}
+\frame{\frametitle{Link Grammars: Same Example Parse From Before Again}
   \begin{figure}
@@
 
-\frame{\frametitle{Link Grammars: Converting into a Directed Acyclic Graph Again}
+\frame{\frametitle{Link Grammars: Converting into a Directed Acyclic Graph}
   \begin{figure}
@@
       %\item Different dependency formalisms disagree on whether a clause is headed by the aux verb or the main verb.
-  \item The Link Grammar seems to be inconsistent about whether the auxiliary verb or the main verb is the head of a clause. 
-  \item Sometimes the governing verb links to the auxilliary, and sometimes to the main, depending on the type of clause. 
-  \item But the governing verb usually links to the subject when there is one.
-  \item So this makes the subject a consistent choice to make the head of a clause.
+  \uncover<1->{\item The Link Grammar seems to be inconsistent about whether the auxiliary verb or the main verb is the head of a clause. }
+  \uncover<2->{\item Sometimes the governing verb links to the auxilliary, and sometimes to the main, depending on the type of clause. }
+  \uncover<3->{\item But the governing verb usually links to the subject when there is one.}
+  \uncover<4->{\item So this makes the subject a consistent choice to make the head of a clause.}
 \end{itemize}
@@
 
+\uncover<5->{
 \begin{center}
@@
 \end{center}
-
+}
     %Question: Which do we pick?

### 2014-12-13 doc/slides/hong+eisner.TLT13.slides.tex
url link citation for SCIP

@@
   \item[] Grammar-based formalism for projective dependency parsing with undirected links
-  \item[] Orignal formalism and English Link Grammar created by Davy Temperley, Daniel Sleator, John Lafferty (1991)
+  \item[] Orignal formalism and English Link Grammar created by Davy Temperley, Daniel Sleator, and John Lafferty (1991)
   \end{itemize}
@@
   \item Its NP-Hard! But good solvers exist that work well most of the time.
-  \item Our ILP is encoded as a ZIMPL program and solved using the SCIP Optimization Suite
+  \item Our ILP is encoded as a ZIMPL program and solved using the SCIP Optimization Suite\footnote{\url{http://scip.zib.de/}} 
 \end{itemize}
@@
 
-\frame{\frametitle{Link Results: Subject-Verb links are backwards}
+\subsection{Subject-Verb Links}
+\frame{\frametitle{Link Results: Subject-Verb Links are Backwards}
 \begin{figure}
@@
 
-\frame{\frametitle{Link Results: Subject-Verb links are backwards}
+\frame{\frametitle{Link Results: Subject-Verb Links are Backwards}
 \begin{figure}
@@
 
-\frame{\frametitle{Link Results: Subject-Verb links are backwards}
+\frame{\frametitle{Link Results: Subject-Verb Links are Backwards}
 
@@
 
-\frame{\frametitle{Link Results: Subject-Verb links are backwards}
+\frame{\frametitle{Link Results: Subject-Verb Links are Backwards}
 \begin{itemize}
@@
   \item Link Grammar parses can be oriented into connected DAGs
-  \item New corpora for building multi-headed dependency parsers
-  \item ILP can be used to build or annotate corpora
+  \item new corpora for building multi-headed dependency parsers
+  \item ILP can be used to build or annotate corpora. 
   \end{itemize}

### 2015-01-07 doc/slides/hong+eisner.TLT13.slides.tex
small edits
Fixed some of Jason's minor comments of the slides.

@@
 
-\frame{\frametitle{What We Want:}
+\frame{\frametitle{What We Wanted:}
   \begin{figure}
@@
 
-\frame{\frametitle{Why We Want That}
+\frame{\frametitle{Why We Wanted That}
   \begin{itemize}
@@
   \item[] Grammar-based formalism for projective dependency parsing with undirected links
-  \item[] Orignal formalism and English Link Grammar created by Davy Temperley, Daniel Sleator, and John Lafferty (1991)
+  \item[] Original formalism and English Link Grammar created by Davy Temperley, Daniel Sleator, and John Lafferty (1991)
   \end{itemize}
@@
 \frame{\frametitle{Link Grammars: How They Work}
-  \includegraphics[width=\linewidth]{linkgrammar_example1.png}\footnote{These figures were clipped from the orginal Link Grammar paper: \\``Parsing English with a Link Grammar'' by Sleator and Temperley}
+  \includegraphics[width=\linewidth]{linkgrammar_example1.png}\footnote{These figures were clipped from the original Link Grammar paper: \\``Parsing English with a Link Grammar'' by Sleator and Temperley}
 }
@@
 
-  \caption*{Top half is CoNLL. Bottom half is the directionalized link parse.}
+  \caption*{Bottom half is CoNLL. Top half is the directionalized link parse.}
 \end{figure}
@@
 \begin{itemize}
-  \item Its an optimization problem where some or all of the variables are integers. 
+  \item An optimization problem where some or all of the variables are integers. 
   \item The objective function and constraints are linear.
-  \item Its NP-Hard! But good solvers exist that work well most of the time.
+  \item In general, it's NP-Hard! But good solvers exist that work well most of the time.
   \item Our ILP is encoded as a ZIMPL program and solved using the SCIP Optimization Suite\footnote{\url{http://scip.zib.de/}} 
@@
 
-\frame{\frametitle{ILP Results: Top 25 Most Occuring Labels}
+\frame{\frametitle{ILP Results: Top 25 Most Occurring Labels}
 

### 2015-01-28 doc/poster/hong+eisner.MASC-SLL.poster.tex
MASC poster version 1

\documentclass[a0,landscape]{a0poster}

\usepackage{hyperref}
\bibliographystyle{plain}
%\usepackage{acl2014}
\newcommand{\newcite}[1]{\cite{#1}}   % ACL's newcite is not available in plain bib style, so just use cite1

\usepackage{multicol} % This is so we can have multiple columns of text side-by-side
\columnsep=100pt % This is the amount of white space between the columns in the poster
\columnseprule=3pt % This is the thickness of the black line between the columns in the poster

\usepackage[svgnames]{xcolor} % Specify colors by their 'svgnames', for a full list of all colors available see here: http://www.latextemplates.com/svgnames-colors

\usepackage{times} % Use the times font
%\usepackage{palatino} % Uncomment to use the Palatino font

\usepackage{graphicx} % Required for including images
\usepackage{booktabs} % Top and bottom rules for table
\usepackage[font=small,labelfont=bf]{caption} % Required for specifying captions to tables and figures
\usepackage{amsfonts, amsmath, amsthm, amssymb} % For math fonts, symbols and environments
\usepackage{wrapfig} % Allows wrapping text around tables and figures

%\usepackage{xspace}
%\usepackage{enumitem}
%\usepackage[small]{caption}
%\usepackage{longtable}

%\usepackage{soul}

\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
%\usepackage{caption}
\usepackage{subcaption}
\usepackage{pdfpages}

\usepackage{tikz-dependency}

\newcommand{\eqnref}[1]{\eqref{eqn:#1}}

\usepackage[disable]{todonotes}   % insert [disable] to disable all notes
% Note that these macros accept optional arguments such as 
% [size=\small,bordercolor=red].
\newcommand{\Note}[4][]{\todo[author=#2,color=#3,fancyline,#1]{#4}}
\newcommand{\noteJH}[2][]{\Note[size=\small,#1]{JH}{blue!40}{#2}}
\newcommand{\noteJE}[2][]{\Note[size=\small,#1]{JE}{green!40}{#2}}   
\newcommand{\notewho}[3][]{\Note[size=\small,#1]{#2}{orange!40}{#3}}  % extra arg with miscellaneous author
\newcommand{\NoteJH}[2][]{\noteJH[inline,#1]{#2}}
\newcommand{\NoteJE}[2][]{\noteJE[inline,#1]{#2}}
\newcommand{\Notewho}[3][]{\notewho[inline,#1]{#2}{#3}}  % extra arg with miscellaneous author

\begin{document}

%----------------------------------------------------------------------------------------
%	POSTER HEADER 
%----------------------------------------------------------------------------------------

% The header is divided into three boxes:
% The first is 55% wide and houses the title, subtitle, names and university/organization
% The second is 25% wide and houses contact information
% The third is 19% wide and houses a logo for your university/organization or a photo of you
% The widths of these boxes can be easily edited to accommodate your content as you see fit
%
\begin{minipage}[b]{0.19\linewidth}
  \begin{flushleft}
    \includegraphics[width=15cm]{lti_logo.png} % Logo or a photo of you, adjust its dimensions here
  \end{flushleft}
\end{minipage}
%
\begin{minipage}[b]{0.55\linewidth}
  \centering \veryHuge \color{NavyBlue} \textbf{Deriving Multi-Headed Planar Dependency Parses from Link Grammar Parses} 
 \color{Black} \\
  %\Huge\textit{An Exploration of Complexity}\\[1cm] % Subtitle
\begin{minipage}{0.4\linewidth}
  \begin{flushleft}
  \centering \huge \textbf{Juneki Hong}\\
  \centering \huge Carnegie Mellon University 
  \end{flushleft}
\end{minipage}
\begin{minipage}{0.4\linewidth}
  \begin{flushright}
  \centering \huge \textbf{Jason Eisner}\\
  \centering \huge Johns Hopkins University
  \end{flushright}
\end{minipage}
\end{minipage}
\begin{minipage}[b]{0.19\linewidth}
  \begin{flushright}
    \includegraphics[width=10cm]{jhu_seal.png}
  \end{flushright}
\end{minipage}

\vspace{1cm} % A bit of extra whitespace between the header and poster content

%----------------------------------------------------------------------------------------

\begin{multicols}{4} % This is how many columns your poster will be broken into, a poster with many figures may benefit from less columns whereas a text-heavy poster benefits from more

\color{Navy} % Navy color for the summary

\section*{Summary}

\begin{itemize}
\item Multi-headed dependency corpora would allow for the development of richer syntactic formalisms. 
%However few exist, especially for the planar or projective case.
\item In order to produce projective multi-headed corpora, we consistently directionalize Link Grammar using Integer Linear Programming.
\item The resulting parses differ in style from CoNLL-style parses of the same sentences.
\end{itemize}
\color{DarkSlateGray} % DarkSlateGray color for the rest of the content

\vspace{-5mm}
\section*{Multi-Headed Dependency Parsing}
Relaxing single-headed constraints common in dependency parsing would allow for constructions such as Control, Relativization, and Conjunction.
%
\vspace{-5mm}
\subsection*{Control}
%
\begin{figure}[H]
  \centering
  %\begin{dependency}
  %  \begin{deptext}
  %    Jill \& likes \& to \& skip \\
  %  \end{deptext}
  %  \deproot[edge above, thick, hide label, edge unit distance = 1.5ex]{2}{}
  %  \depedge[edge above, thick, hide label]{2}{1}{}
  %  \depedge[edge below, ultra thick, hide label, edge style = {purple}, edge unit distance = 0.9ex]{4}{1}{}
  %  \depedge[edge above, thick, hide label]{2}{3}{}
  %  \depedge[edge above, thick, hide label]{3}{4}{}
  %\end{dependency}
  %\caption*{Jill is the subject of two verbs}
  \begin{dependency}
    \begin{deptext}
      Jill \& persuaded \& Jack \& to \& skip \\
    \end{deptext}
    \deproot[edge above, thick, hide label, edge unit distance = 1.5ex]{2}{}
    \depedge[edge above, thick, hide label]{2}{1}{}
    \depedge[edge above, thick, hide label]{2}{3}{}
    \depedge[edge above, thick, hide label]{3}{4}{}
    \depedge[edge above, thick, hide label]{4}{5}{}
    \depedge[edge below, ultra thick, hide label, edge style = {purple}, edge unit distance = 1.5ex]{5}{3}{}
    \depedge[edge below, ultra thick, hide label, edge style = {purple}, edge unit distance = 1.5ex]{2}{5}{}
  \end{dependency}
  \caption*{Jack is the object of one verb and the subject of another}
\end{figure}
%
\vspace{-5mm}
%
\subsection*{Relativization}
%
\begin{figure}[H]
  \centering
  \begin{dependency}
    \begin{deptext}
      The \& boy \& that \& Jill \& skipped \& with \& fell \& down \\
    \end{deptext}
    \deproot[edge above, thick, hide label, edge unit distance = 2ex]{7}{}
    \depedge[edge above, thick, hide label]{2}{1}{}
    \depedge[edge above, thick, hide label]{2}{3}{}
    \depedge[edge above, thick, hide label]{3}{5}{}
    \depedge[edge above, thick, hide label]{5}{4}{}
    \depedge[edge above, thick, hide label]{5}{6}{}
    \depedge[edge above, thick, hide label, edge unit distance = 1.4ex]{7}{2}{}
    \depedge[edge above, thick, hide label]{7}{8}{}
    \depedge[edge below, ultra thick, hide label, edge style = {purple}, edge unit distance = 1.0ex]{6}{2}{}
  \end{dependency}
  \caption*{The boy is the object of \textit{with} as well as the subject of \textit{fell}.}
\end{figure}
%
\vspace{-5mm}
%
\subsection*{Conjunction}
\begin{figure}[H]
  \centering
  \begin{dependency}
    \begin{deptext}
      Jack \& and \& Jill \& went \& up \& the \& hill \\
    \end{deptext}
    \deproot[edge above, thick, hide label, edge unit distance = 2ex]{4}{}
    \depedge[edge above, thick, hide label]{2}{1}{}
    \depedge[edge above, thick, hide label]{2}{3}{}
    \depedge[edge above, thick, hide label]{4}{2}{}
    \depedge[edge above, thick, hide label]{4}{5}{}
    \depedge[edge above, thick, hide label]{5}{7}{}
    \depedge[edge above, thick, hide label]{7}{6}{}
    \depedge[edge below, ultra thick, hide label, edge style = {purple}, edge unit distance = 1.3ex]{4}{1}{}
    \depedge[edge below, ultra thick, hide label, edge style = {purple}, edge unit distance = 1.5ex]{4}{3}{}
  \end{dependency}
  \caption*{Jack and Jill serve as the two arguments to \textit{and}, but are also subjects of \textit{went}.}
\end{figure}
%
\section*{Link Grammars}
\begin{itemize}
\item Grammar-based formalism for projective dependency parsing with undirected links. 
\item A label on the link describes the relationship between two words.
\item The original formalism and English Link Grammar was created by Davy Temperley, Daniel Sleator, and John Lafferty\cite{sleator-temperley-1991}.
\end{itemize}
%\vspace{-5mm}
\begin{figure}[H]
\centering
\includegraphics{linkgrammar_example1.png}
\caption*{Example visualization of a link grammar, taken from their original paper. Words have ``links'' that attach to others. These link attachments can either be optional or required.}
\end{figure}
\begin{figure}[H]
\centering
\includegraphics{linkgrammar_example2.png}
\caption*{A parse where all the link attachments have been satisfied. Link attachments must be projective. A parse cannot be completed until all words have a complete set of link attachments.}
\end{figure}
%\begin{figure}[H]
%\centering
%\includegraphics{linkgrammar_example3.png}
%\caption*{Link attachments must be projective. A parse cannot be completed until all words have a complete set of link attachments.}
%\end{figure}

\section*{Integer Linear Programming}
\begin{itemize}
\item ILP is an optimization problem where the objective function and constraints are linear, while some or all of the variables are integers. 
\item In general, it's NP-Hard, but good solvers exist that often work well.
\item Our ILP is encoded as a ZIMPL program and solved with SCIP Optimization Suite\cite{Koch2004, achterberg2009scip}
\end{itemize}

\subsection*{Corpus Building Strategy}
\begin{itemize}
\item Start with corpus of link grammar parses to directionalize. 
\item Use ILP to find the minimum set of directionality assignments that satisfy constraints.
\end{itemize}
% such as acyclicity, connectedness, and directional consistency.

\subsection*{ILP Link Orientation Variables}
\begin{itemize}
\item[] For each sentence, for each edge $i,j$, where $i < j$
\begin{figure}[H]
  \centering
  \begin{dependency}[edge style={-}]
    \begin{deptext}
      . \& . \& . \& $i$ \& . \& . \& . \& $j$ \& . \& . \& . \\
    \end{deptext}
    \depedge[edge above, thick, edge unit distance = 1.1ex]{4}{8}{L}
  \end{dependency}
\end{figure}
\item[] Orientation of each link can be represented by variables that can either be oriented left or oriented right:
$$x_{ij}, x_{ji} \in \mathbb{Z} \geq 0$$
$$x_{ij} + x_{ji} = 1$$
\end{itemize}

\vspace{-2cm}
\subsection*{ILP Constraints}
\subsubsection*{Acyclicity}

\begin{itemize}
\item[] Given that node $u$ is the parent of $v$
\item[] $n_v$: length of the sentence containing node $v$
\item[] $d_v \in [0, n_v]$: depth of the node from the root of the sentence
\item[] We enforce that the depth of a child is greater than that of the parent:
\begin{align}
  (\forall_u)\; d_v + (1 + n_v) \cdot (1 - x_{uv}) & \geq 1+d_u
\end{align}
\end{itemize}

\subsubsection*{Connectedness}
To ensure that every word is reachable from a root, a word must have at least one parent. Together with acyclicity, this enforces connectedness.
\begin{align}
  \sum_u x_{uv} & \geq 1 
\end{align}

\subsubsection*{Consistency of Directionalized Links}
Links with same label type are encouraged to be oriented in the same way. We introduce variables to represent whether links with label $L$ are allowed to go left or right.
$$r_L, \ell_L \in \{0,1\}$$
We introduce slack variables $s_{ij}$ to allow some links to go in disallowed directions with a penalty.
\begin{itemize}  
\item[] $s_{ij} \in \mathbb{R} \geq 0$
\item[] $N_L$: number of link tokens with label $L$
\end{itemize}
\begin{align}\label{direction+slack}
  x_{ij} &\leq r_L + s_{ij} &
  x_{ji} &\leq \ell_L + s_{ij}
\end{align}
\begin{align}\label{eqn:obj}
  objective = \min \left( \sum_L r_L + \ell_L \right) \frac{N_L}{4} + \sum_{ij}s_{ij}
\end{align}

\section*{Data Sets}
Data Sets taken from:
\begin{itemize}
\item[] CoNLL 2007 Shared Task (English)
\item[] ACL 2013 Shared Task of Machine Translation (Russian)
\end{itemize}
\begin{figure}[H]
  \centering
  \begin{tabular}{|l|l|l|}
    \hline
    & Input Sentences & Output Connected Parses \\ \hline
    English  & 18,577        & 10,960               \\ \hline
    Russian & 18,577        & 4,913                \\ \hline
  \end{tabular}
\end{figure}

\section{Experiments and Results}

\begin{figure}
  \centering
  \includegraphics[width=0.6\linewidth]{example_parse.png}
  \caption*{Bottom half is from the CoNLL 2007 shared task. Top half is our directionalized link parse. Edges are shown as solid lines if they appear in both parses.  They are drawn more thickly if the directions do not match, and they are shown as dotted lines if they appear only in one parse and not in the other. Edges are highlighted in orange if the child has multiple parents.}
\end{figure}

\vspace{-5mm}
\begin{figure}[H]
  \centering
  \begin{dependency}
	\begin{deptext}
	  - \& n-u \& v \& e \& e \& v \& v-d \& r \& n-u \& - \\
	  the \& matter \& may \& never \& even \& be \& tried \& in \& court \& . \\
	  {\scriptsize DT} \& {\scriptsize NN} \& {\scriptsize MD} \& {\scriptsize RB} \& {\scriptsize RB} \& {\scriptsize VB} \& {\scriptsize VB} \& {\scriptsize IN} \& {\scriptsize NN} \& {\scriptsize .} \\
	\end{deptext}
	\deproot[edge below, thick, edge style = {blue}]{3}{\small ROOT}
	\depedge[edge above, thick, edge style = {red}]{2}{3}{S}
	\depedge[edge below, thick, edge style = {blue}]{3}{2}{\small SBJ}
	\depedge[edge below, thick, edge style = {blue}]{3}{4}{\small ADV}
	\depedge[edge below, thick, edge style = {blue}]{3}{5}{\small ADV}
	\depedge[edge above, thick, edge style = {red}]{3}{6}{I}
	\depedge[edge below, thick, edge style = {blue}]{3}{6}{\small VC}
	\depedge[edge below, thick, edge style = {blue}, edge unit distance =1.5ex]{3}{10}{\small P}
	\depedge[edge above, thick, edge style = {red}]{2}{1}{D}
	\depedge[edge below, thick, edge style = {blue}]{2}{1}{\small NMOD}
	\depedge[edge above, thick, edge style = {red}]{7}{8}{MV}
	\depedge[edge below, thick, edge style = {blue}]{7}{8}{\small ADV}
	\depedge[edge above, thick, edge style = {red}]{6}{7}{P}
	\depedge[edge below, thick, edge style = {blue}]{6}{7}{\small VC}
	\depedge[edge above, thick, edge style = {red}]{8}{9}{J}
	\depedge[edge below, thick, edge style = {blue}]{8}{9}{\small PMOD}
	\deproot[edge above, thick, edge style = {red}]{2}{W}
	\deproot[edge above, thick, edge style = {red}]{7}{WV}
	\deproot[edge above, thick, edge style = {red}]{10}{X}
	\depedge[edge above, thick, edge style = {red}]{6}{4}{E}
	\depedge[edge above, thick, edge style = {red}]{6}{5}{E}
  \end{dependency}

  \caption*{Bottom half from CoNLL 2007 shared task. Top half is our directionalized link parse.}
\end{figure}

\subsubsection*{Multiheadedness}
On the English data Set, the link data has 8\% additional edges over the CoNLL. (average about 2 multiheaded words per sentence)

\subsubsection*{CoNLL Matches}
\begin{itemize}
\item[] 52\% of links match CoNLL arcs
\item[] 57\% of CoNLL arcs match links
\end{itemize}

\subsubsection*{Directionality}
\begin{itemize}
\item[] 6.19\% of link types allowed both directions %($\frac{7}{113}$)
\item[] 2.07\% of link tokens required disallowed direction via slack %($\frac{4043}{195,000}$)
\end{itemize}

\subsection*{Stability of Results}
To see whether the recovered direction mapping might be unstable and sensitive to the input corpus, we compared results of increasing runs of sentences. 
\begin{figure}[H]
  \centering
  \includegraphics[width=0.6\linewidth]{precision_recall.png}
  \caption*{The direction mappings obtained on small datasets have high {\em precision} relative to the one obtained on the largest dataset.  Their {\em recall} grows as more link types are seen and directionalized.}\label{fig:convergence}
\end{figure}  
%Rapid convergence to the direction mapping obtained on the largest dataset.

\vspace{-1cm}
\subsection*{Backwards Subject-Verb Links}
In our directionalized corpus subjects point to verbs instead of verbs pointing to subjects. This is due to a possible inconsistency of the Link Grammar discovered by our method.
\begin{itemize}
\item Link Grammar seems to be inconsistent about whether the auxiliary verb or the main verb is the head of a clause. 
\item Governing verb links to either auxilliary or main, depending on the clause type, but governing verbs usually link to subject when there is one. This makes subject a consistent choice to make head of a clause. To fix, we can edit the link grammar, link parses, or the ILP. 
\end{itemize}
%----------------------------------------------------------------------------------------
%	CONCLUSIONS
%----------------------------------------------------------------------------------------
\color{SaddleBrown} % SaddleBrown color for the conclusions to make them stand out
\vspace{-1cm}
\section*{Conclusions}
\begin{itemize}
\item Link Grammar parses can be oriented into connected DAGs
\item A new corpus for building multi-headed dependency parsers
\item ILP can be used to help annotate missing data in corpora. 
\end{itemize}

\color{DarkSlateGray} % Set the color back to DarkSlateGray for the rest of the content

%----------------------------------------------------------------------------------------
%	REFERENCES
%----------------------------------------------------------------------------------------

%\nocite{*} % Print all references regardless of whether they were cited in the poster or not
\begin{bibliography}{LinksToDAG_final}

\end{bibliography}

%----------------------------------------------------------------------------------------
%	ACKNOWLEDGEMENTS
%----------------------------------------------------------------------------------------

\subsubsection*{Acknowledgements}
This material is based upon work supported by the National Science Foundation under Grant No. 1423276.  The work was mainly conducted while the first author was at Johns Hopkins University.

%----------------------------------------------------------------------------------------

\end{multicols}
\end{document}

### 2015-08-20 doc/slides/hong+eisner.TLT13.slides.tex
updated slides

@@
 \title{Deriving Multi-Headed Planar Dependency Parses from Link Grammar Parses}
-\author{Juneki Hong, Jason Eisner}
-\date{\today}
+\author{Juneki Hong and Jason Eisner}
+\date{}%\today}
 \frame{\titlepage} %\NoteJH{Include pictures of us?}
@@
 \begin{itemize}
-\item This talk is about automatically converting from one annotation style to another.
-\item The conversion could be hard, where information is fragmented or missing.
+\item This talk is about converting from one annotation style to another.
+\item The conversion could be hard, where information is fragmented, missing, or ambiguous.
 
 %This is often useful.  You may want to do it yourself.  
-\item We use a general technique, Integer Linear Programming to recover this missing information.
+\item We use a general technique, Integer Linear Programming to help us do this conversion.
 \end{itemize}
@@
 
-\section{Motivation}
+\section{Motivation, Overview}
 %\subsection{Multi-headed}
@@
 
-\subsection{Control}
+\subsection{Control, Relativization, Conjunction}
 \frame{\frametitle{Control}
@@
 
-\subsection{Relativization}
+%\subsection{Relativization}
 \frame{\frametitle{Relativization}
@@
 
-\subsection{Conjunction}
+%\subsection{Conjunction}
 \frame{\frametitle{Conjunction}
@@
   \end{itemize}
+}
+
+\subsection{Corpus Building}
+\frame{\frametitle{Corpus Building Strategy}
+\begin{itemize}
+\item We start with some sentences and parse them with LG Parser
+\item We take the undirected parses and try to directionalize them.
+\item We use an ILP to assign consistent directions for each link type.
+\end{itemize}
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{images/corpusbuilding}
+\end{figure}
 }
@@
 
-\frame{\frametitle{Link Grammars: Converting into a Directed Acyclic Graph}
-  \begin{figure}
-    \begin{dependency}
-	  \begin{deptext}
-	    - \& n-u \& v \& e \& e \& v \& v-d \& r \& n-u \& - \\
-	    the \& matter \& may \& never \& even \& be \& tried \& in \& court \& . \\
-	  \end{deptext}
-	  \depedge[edge above, thick, edge style = {red}]{2}{3}{S}
-	  \depedge[edge above, thick, edge style = {red}]{3}{6}{I}
-	  \depedge[edge above, thick, edge style = {red}]{2}{1}{D}
-	  \depedge[edge above, thick, edge style = {red}]{7}{8}{MV}
-	  \depedge[edge above, thick, edge style = {red}]{6}{7}{P}
-	  \depedge[edge above, thick, edge style = {red}]{8}{9}{J}
-	  \deproot[edge above, thick, edge style = {red}]{2}{W}
-	  \deproot[edge above, thick, edge style = {red}]{7}{WV}
-	  \deproot[edge above, thick, edge style = {red}]{10}{X}
-	  \depedge[edge above, thick, edge style = {red}]{6}{4}{E}
-	  \depedge[edge above, thick, edge style = {red}]{6}{5}{E}
-    \end{dependency}
-
-    \caption*{Directionalize the edges}
-  \end{figure} 
-}
+%\frame{\frametitle{Link Grammars: Converting into a Directed Acyclic Graph}
+%  \begin{figure}
+%    \begin{dependency}
+%	  \begin{deptext}
+%	    - \& n-u \& v \& e \& e \& v \& v-d \& r \& n-u \& - \\
+%	    the \& matter \& may \& never \& even \& be \& tried \& in \& court \& . \\
+%	  \end{deptext}
+%	  \depedge[edge above, thick, edge style = {red}]{2}{3}{S}
+%	  \depedge[edge above, thick, edge style = {red}]{3}{6}{I}
+%	  \depedge[edge above, thick, edge style = {red}]{2}{1}{D}
+%	  \depedge[edge above, thick, edge style = {red}]{7}{8}{MV}
+%	  \depedge[edge above, thick, edge style = {red}]{6}{7}{P}
+%	  \depedge[edge above, thick, edge style = {red}]{8}{9}{J}
+%	  \deproot[edge above, thick, edge style = {red}]{2}{W}
+%	  \deproot[edge above, thick, edge style = {red}]{7}{WV}
+%	  \deproot[edge above, thick, edge style = {red}]{10}{X}
+%	  \depedge[edge above, thick, edge style = {red}]{6}{4}{E}
+%	  \depedge[edge above, thick, edge style = {red}]{6}{5}{E}
+%    \end{dependency}
+%    
+%    \caption*{Directionalize the edges}
+%  \end{figure} 
+%}
 
@@
 
-\subsection{Corpus Building}
-\frame{\frametitle{Corpus Building Strategy}
-\begin{itemize}
-  \item We start with a corpus of link grammar parses.
-  \item We try to directionalize this corpus, choosing a direction for every link in every parse.
-  \item We use an Integer Linear Program to find consistent directions for every link
-\end{itemize}
-}
-
 \subsection{ILP Model}
@@
 \begin{itemize}
-  \item Connectedness\uncover<2->{: Every word has a link attaching it}
-  \item Acyclicity\uncover<2->{: No cycles!}
-  \item Consistency of Directionalized Links\uncover<2->{:
+  \item Acyclicity\uncover<2->{: (\textbf{No cycles!})}
+  \item Connectedness\uncover<3->{: (\textbf{Every word is reachable from a root})}
+  \item Consistency of Directionalized Links\uncover<4->{:
 
-    Similar links oriented the same way}
+    (\textbf{Similar links oriented the same way})}
 \end{itemize}
+
+\begin{figure}
+  \begin{subfigure}[b]{0.3\textwidth}
+    \uncover<2->{\includegraphics[width=\linewidth]{images/acyclic2}}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.3\textwidth}
+    \uncover<3->{\includegraphics[width=\linewidth]{images/connectedness}}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.3\textwidth}
+    \uncover<4->{\includegraphics[width=\linewidth]{images/consistency}}
+  \end{subfigure}
+
+\end{figure}
+
 }
@@
 \uncover<2->{\begin{center}
-    A link can only be oriented left or right
+    \textbf{An individual link token can either be oriented left or oriented right}
 \end{center}}
@@
 
-\frame{\frametitle{Connectedness, Acyclicity}
+\frame{\frametitle{Acyclicity, Connectedness}
 \begin{itemize}
-  \item[] Connectedness
-    \begin{align}
-      \sum_u x_{uv} & \geq 1 
-    \end{align}
-    \uncover<2->{\begin{center}
-        A word has at least 1 parent
-    \end{center}}
-
   \item[] Acyclicity
@@
     \uncover<2->{\begin{center}
-        The depth of a child is greater than the depth of the parent
+        \textbf{The depth of a child is greater than the depth of the parent}
+    \end{center}}
+
+  \item[] Connectedness
+    \begin{align}
+      \sum_u x_{uv} & \geq 1 
+    \end{align}
+    \uncover<3->{\begin{center}
+        \textbf{A word has at least 1 parent}
     \end{center}}
 \end{itemize}
-
 }
@@
     \begin{itemize}
-      \item[] $r_L, \ell_L \in \{0,1\}$: whether links with label $L$ allowed left/right
+      \item[] $r_L, \ell_L \in \{0,1\}$: whether all links with label $L$ allowed left/right
     \end{itemize}
@@
 
+    Objective Function:
     \begin{align}\label{eqn:obj}
@@
     \end{itemize}
-
+    }
+    \onslide<3->{
     \begin{itemize}
-      \item[] Slack allows a few links with label $L$ in disallowed directions
+      \item[] \textbf{Slack allows a few links with label $L$ in disallowed directions}
     \end{itemize}
@@
   \begin{itemize}
-  \item We worried that the recovered direction mapping might be unstable and sensitive to the input corpus. 
+  \item We were worried that the recovered direction mapping might be unstable and sensitive to the input corpus. 
   \item We compared the results of increasing runs of sentences.
@@
     \input{link_table_filtered.tex}
+  \end{figure}
+}
+
+\frame{\frametitle{ILP Results: Top 25 Most Occurring Labels}
+  \begin{figure}
+    \centering
+    \input{link_table_filtered_highlighted.tex}
+  \end{figure}
+}
+
+\frame{\frametitle{ILP Results: Top 25 Most Occurring Labels}
+  \begin{figure}
+    \centering
+    \input{link_table_filtered_highlighted_explanation.tex}
   \end{figure}
@@
   \item Link Grammar parses can be oriented into connected DAGs
-  \item new corpora for building multi-headed dependency parsers
-  \item ILP can be used to build or annotate corpora. 
+  \item A new corpus available for building multi-headed dependency parsers
+  \item ILP can be used to help annotate incomplete or missing data in corpora. 
   \end{itemize}

### 2015-08-20 doc/slides/link_table_filtered_highlighted_explanation.tex
updated slides

\begin{tiny}
\centering
%\begin{longtable}[width=\textwidth]{|l|l|l|l|l|l|}
\begin{tabular}{|l|l|l|l|l|l|}
  \hline
  Label & Rightward & Multiheaded & CoNLL Match & CoNLL Dir Match\\\hline
  %\endhead

  \textbf{B} & \textbf{100\% (1514/1515)} & \textbf{61\% (919/1515)} & \textbf{53\% (806/1515)} & \textbf{84\% (678/806)} \\
  \hline
\end{tabular}
\end{tiny}

%\begin{small}
\begin{itemize}
  \uncover<2->{
  \item[] ``B'' link relative clauses
  }
\end{itemize}
%\end{small}

\begin{small}
\uncover<2->{
  \begin{subfigure}[b]{0.48\textwidth}
    \centering
  \begin{dependency}[edge style={-}]
    \begin{deptext}
      The \& dog \& I \& had \& chased \& was \& green \\
    \end{deptext}
    \depedge[edge above, edge style = {red, thick}]{2}{3}{R}
    \depedge[edge above, edge style = {red, thick}]{3}{4}{S}
    \depedge[edge above, edge style = {red, thick}]{4}{5}{PP}
    \depedge[edge above, edge style = {orange, ultra thick}]{2}{5}{B}
  \end{dependency}
\end{subfigure}
}
\uncover<3->{
  \begin{subfigure}[b]{0.48\textwidth}
    \centering
  \begin{dependency}[edge style={-}]
    \begin{deptext}
      I \& told \& him \& I \& had \& oranges \\
    \end{deptext}
    \depedge[edge above, edge style = {red, thick}, edge unit distance = 1.5ex]{2}{4}{CE}
    \depedge[edge above, edge style = {red, thick}]{4}{5}{S}
    \depedge[edge above, edge style = {orange, ultra thick}]{2}{5}{CV}
  \end{dependency}
\end{subfigure}
}

\end{small}

\begin{tiny}
\begin{tabular}{|l|l|l|l|l|l|}
  \hline
  Label & Rightward & Multiheaded & CoNLL Match & CoNLL Dir Match\\\hline
  \textbf{CV} & \textbf{100\% (3237/3237)} & \textbf{100\% (3237/3237)} & \textbf{56\% (1827/3237)} & \textbf{28\% (512/1827)}  \\
  \hline
  %\end{longtable}
\end{tabular}
\end{tiny}

%\begin{small}
\begin{itemize}
  \uncover<3->{
  \item[] ``CV'' link conjunctions to main verbs of clauses.
  }
\end{itemize}

%\end{small}
